### 2013-04-11 globaldsl13/conclusion.tex

\section{Conclusion} % and Future Work}
\label{s:conclusion}

In this paper, we described how extensible parsing in Wyvern makes for
a solid platform to support whitespace-delimited, type-directed embedded DSLs. In the
future, we aim to implement a wide variety of DSLs in Wyvern tweaking
our approach and implementation thereof to provide a comprehensive example of
supporting multiple interacting DSLs in a safe and easy-to-use manner.

% \todo{tie features to goals}

% \todo{implementation and validation plans}