### 2013-04-11 globaldsl13/globaldsl13.tex

\documentclass[english]{sigplanconf}
\usepackage{palatino}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{listings}
\usepackage{url}
\usepackage{babel}
\usepackage{graphicx}
%\usepackage{microtype}

%\lstset{
%  numbers=left,
%  basicstyle=\ttfamily,
%  tabsize=4,
%  frame=single,
%}

\lstset{tabsize=3, basicstyle=\ttfamily\small, commentstyle=\itshape\rmfamily, numbers=left, numberstyle=\tiny, language=java,moredelim=[il][\sffamily]{?},mathescape=true,showspaces=false,showstringspaces=false,columns=fullflexible,xleftmargin=15pt,escapeinside={(@}{@)}, morekeywords=[1]{application,policy,external,component,connects,to,meth,let,in,val,where,return,group,by,within,connect,with,attr,html,head,title,style,body,div,keyword,location,schema,must,salt,select,text}}
\lstloadlanguages{Java,VBScript,XML,HTML,Python}

\newcommand{\todo}[1]{\textbf{[TODO: #1]}}
\newcommand{\keyw}[1]{\texttt{\textbf{#1}}}

\begin{document}

\title{Whitespace-Delimited, Type-Directed Parsing for Embedded DSLs}

\authorinfo{Cyrus Omar, Benjamin Chung$^{1}$, Darya Kurilova, Alex Potanin$^{2}$,
  and Jonathan Aldrich}
{Carnegie Mellon University}
{\{comar, darya, aldrich\}@cs.cmu.edu, bwchung@andrew.cmu.edu$^{1}$, and alex@ecs.vuw.ac.nz$^{2}$}
% Maybe just keep my cs account active after I leave for a few years?
% Otherwise: alex@ecs.vuw.ac.nz is my email...
\maketitle

\begin{abstract}

Domain-specific languages enhance ease of use, expressiveness and
verifiability, 
but defining and using different 
DSLs within a single application remains difficult.  
We present an approach to embedded DSLs where:
1) a whitespace-delimited lexer is shared across DSLs, and 
2) the parsing and type checking phases occur in tandem so that types can define domain-specific parsing strategies governing whitespace-delimited blocks.
We argue that this approach occupies a sweet spot with both 
high expressiveness and ease of use. We outline a design
and ongoing implementation of this strategy
and examine the relationship between the type-directed and keyword-directed strategies for parsing of embedded DSLs.

\end{abstract}

\begin{sloppypar}

\input{intro}
\input{motivation}
\input{approach}
\input{implementation}
\input{related}
\input{conclusion}

%TODO TODO TODO TODO TODO TODO TODO TODO add acknowledgments here!!!!!!  ARO

\end{sloppypar}

\bibliographystyle{plain}
\bibliography{biblio}
\end{document}