### 2013-04-11 globaldsl13/intro.tex

\section{Introduction}
\label{s:intro}

Domain-specific languages (DSLs) \cite{shellchapter, coldfusion,
  sinatra}
% have been widely-studied because they 
enable the expression of  
domain-specific ideas naturally and permit domain-specific
verification and compilation strategies.  For DSLs to reach their full
potential, however, it must be possible to easily use multiple
DSLs at once within a host general-purpose language (GPL), such that DSL code
and GPL glue code can inter-operate to form a complete application. More specifically, we propose the following essential design criteria:

\begin{itemize}

\item \emph {Composability}: It should be possible to use multiple DSLs and a GPL
%, in addition to a GPL,
within a single program unit.  %Here 
Within the text file-based paradigm familiar to most developers, this 
means including multiple DSLs within a single file.
 Moreover, it should be possible to embed code written in one DSL
  within another DSL without them requiring knowledge of each other or interfering with one another. So, more specifically, DSLs should be \emph{safely composable}.

\item \emph{Interoperability}: DSLs should be able to \emph{interoperate}. A necessary condition for interoperability is that it should be
  possible to pass and operate on values 
%such as functions or data structures
  defined in foreign DSLs. Additional requirements, such as the ability to do so naturally (rather than requiring a large amount of ``glue code'') and safely may also be relevant in many settings. 
%Moreover,
  Minimally, however, if one DSL defines a function or data structure satisfying an interface specified in
  a common interface description language (such as the type system of
  a host GPL), code written in another DSL should be able to use those
  values according to that interface, without requiring that the client DSL
  have knowledge of the details of the provider DSL or \emph{vice versa}.

\end{itemize}

%\item type system: one DSL depends on types defined by another DSL,
%  can use objects of that type in special ways [this goes beyond the
%    scope - save for a later paper]

In addition to composability and interoperability, we believe that to be most useful, a system supporting DSLs should have the following characteristics:

\begin{itemize}

\item \emph{Flexibility}: Have flexible enough DSL syntax to support a wide variety of text-based notations;

\item \emph{Identifiability}: Make it easy for programmers to identify which code is written in which DSL;

\item \emph{Consistency}: Encourage DSLs written in similar styles in order to enhance readability and learnability of each DSL;

\item \emph{Simplicity}: Keep the cost of defining a DSL low.

%\item Share conventions between DSLs and a host language, making each DSL easier for programmers to learn, helping programmers to identify which code is written in which DSL, and avoiding unintended conflicts between DSLs. \todo{Avoid conflicts (visual and real), enhance learnability}

%\item Reuse low-level mechanisms and design decisions from a host language, thereby reducing the cost of defining DSLs.  \todo{Easy to define DSLs}

\end{itemize}

We are developing a comprehensive approach that has the potential to meet these design requirements well, based on defining DSLs within a language specifically designed for extensibility from within.
The approach inherently supports multiple DSLs in the same file, and
supports interoperability because the DSLs are defined by translation
into a common internal language.  One key aspect of our proposed solution, and one of the focal points of this paper, is
that we restrict the syntax of DSLs in a few key ways that greatly simplify
the mechanism without significantly decreasing expressiveness:
\begin{itemize}

\item The host language and its DSLs share a tokenization and lexing 
  strategy, standardizing conventions for identifiers, operators,
  constants, and comments.  This avoids the cost of defining lexing
  within each DSL, avoids many kinds of low-level clashes between
  languages, and makes the composed language more readable. Note that this does not limit the ability of a DSL to define new keywords and other constructs.

\item Constructs in the host language and its DSLs are
  \emph{whitespace-delimited} at the top level, according to a particular strategy that we will describe.  Various forms of parentheses 
  can also serve as delimiters.  Thus, indentation levels or
  parenthesized expressions clearly delimit blocks that are governed by a particular 
  language.  This makes the boundaries of each DSL clear to
  the programmer and the compiler, enhancing usability and guaranteeing that 
  subtle conflicts cannot arise.

\end{itemize}

Within this syntactic framework, we then propose a novel syntax extension
mechanism in which the \emph{expected type of an expression}, rather than an explicit keyword, 
%, rather than a keyword, 
signals the switch to a new DSL that parses the tokenized and lexed code within a subsequent delimited block according to a type-associated grammar. The  more common keyword-directed strategy arises as a special case of this type-directed strategy. 
We will begin with some motivating examples (\S\ref{s:motivation}), describe our approach (\S\ref{s:approach}) and implementation (\S\ref{s:implementation}), and conclude with related work (\S\ref{s:related}).

% keep discussion of type-based parsing brief - active typing for parsing (only)
% avoids conflict with Cyrus' paper

%The rest of the paper is organized as follows.  The next section
%illustrates our approach by example, discussing the components of the
%solution in more detail. Section~\ref{s:approach} outlines our
%approach, shows a wider variety of examples and discusses variations
%of our approach.  Our in-progress implementation of the proposal is
%described in Section~\ref{s:implementation}.  Section~\ref{s:related}
%compares to related work, and Section~\ref{s:conclusion} concludes.

% with a discussion of future work.