### 2013-06-14 globaldsl13/conclusion.tex

% !TEX root = globaldsl13.tex
%\section{Conclusion} % and Future Work}
%\label{s:conclusion}
%
%In this paper, we described how extensible parsing in Wyvern makes for
%a solid platform to support whitespace-delimited, type-directed embedded DSLs. In the
%future, we aim to implement a wide variety of DSLs in Wyvern tweaking
%our approach and implementation thereof to provide a comprehensive example of
%supporting multiple interacting DSLs in a safe and easy-to-use manner.
%
% \todo{tie features to goals}

% \todo{implementation and validation plans}

\section*{Acknowledgements}
We thank the anonymous reviewers for helpful comments, and acknowledge the support of the Department of Defense and the Air Force Research Laboratory. CO is supported by the NSF Graduate Research Fellowship.