### 2013-06-14 globaldsl13/globaldsl13.tex

%\documentclass[english]{sigplanconf}
\documentclass[english]{sig-alternate}
\usepackage{times}
%\usepackage[T1]{fontenc}
%\usepackage[latin9]{inputenc}
\usepackage{listings}
\usepackage{url}
\usepackage{babel}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\renewcommand{\ttdefault}{pcr}
%\usepackage{microtype}

%\lstset{
%  numbers=left,
%  basicstyle=\ttfamily,
%  tabsize=4,
%  frame=single,
%}

\lstset{tabsize=2, basicstyle=\ttfamily\small, commentstyle=\itshape\rmfamily, numbers=left, numberstyle=\color{gray}\ttfamily\tiny, language=java,moredelim=[il][\sffamily]{?},mathescape=true,showspaces=false,showstringspaces=false,columns=fullflexible,xleftmargin=15pt,escapeinside={(@}{@)}, morekeywords=[1]{application,policy,external,component,connects,to,meth,let,in,val,where,return,group,by,within,connect,with,attr,html,head,title,style,body,div,keyword,location,schema,must,salt,select,text,type,then}}
\lstloadlanguages{Java,VBScript,XML,HTML,Python}

\newcommand{\todo}[1]{\textbf{[TODO: #1]}}
\newcommand{\keyw}[1]{\texttt{\textbf{#1}}}

   \conferenceinfo{GlobalDSL}{'13, Montpellier, France}
   \CopyrightYear{2013}
   \crdata{978-1-4503-2043-6}

\begin{document}

\title{Type-Directed, Whitespace-Delimited Parsing for Embedded DSLs}
\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
% 1st. author
\alignauthor
Cyrus Omar, Benjamin Chung, Darya Kurilova, Alex Potanin$^{1}$ and Jonathan Aldrich\\
       \affaddr{Carnegie Mellon University}\\
%       \affaddr{1932 Wallamaloo Lane}\\
%       \affaddr{Wallamaloo, New Zealand}\\
       \email{\{comar, bwchung, darya, aldrich\}@cs.cmu.edu, and alex@ecs.vuw.ac.nz$^{1}$}
% 2nd. author
%\alignauthor
%G.K.M. Tobin\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{P.O. Box 1212}\\
%       \affaddr{Dublin, Ohio 43017-6221}\\
%       \email{webmaster@marysville-ohio.com}
% 3rd. author
%\alignauthor Lars Th{\o}rv{\"a}ld\\
%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
%       \affaddr{Hekla, Iceland}\\
%       \email{larst@affiliation.org}
}

%\authorinfo{Cyrus Omar, Benjamin Chung, Darya Kurilova, Alex Potanin$^{1}$ 
%  and Jonathan Aldrich}
%{Carnegie Mellon University}
%{\{comar, bwchung, darya, aldrich\}@cs.cmu.edu, and alex@ecs.vuw.ac.nz$^{1}$}
% Maybe just keep my cs account active after I leave for a few years?
% Otherwise: alex@ecs.vuw.ac.nz is my email...
\maketitle

\begin{abstract}

Domain-specific languages improve ease-of-use, expressiveness and
verifiability, 
but defining and using different 
DSLs within a single application remains difficult.  
We introduce an approach for embedded DSLs where 1) whitespace delimits DSL-governed blocks, and 2) the parsing and type checking phases occur in tandem so that the expected type of the block determines which domain-specific parser governs that block.
We argue that this approach occupies a sweet spot, providing   
high expressiveness and ease-of-use while maintaining safe composability. We introduce the design, provide examples and describe an ongoing implementation of this strategy in the Wyvern programming language. We also discuss how a more conventional keyword-directed strategy for parsing of DSLs can arise as a special case of this type-directed strategy.
\end{abstract}
\begin{sloppypar}

\input{intro}
\input{motivation}
\input{approach}
\input{implementation}
\input{related}
\input{conclusion}

%TODO TODO TODO TODO TODO TODO TODO TODO add acknowledgments here!!!!!!  ARO

\bibliographystyle{abbrv}
\bibliography{biblio}
\end{sloppypar}

\end{document}