### 2013-08-26 thesis-proposal/conclusion.tex
Updated abstract, most of introduction, beginning of approach

% !TEX root = omar-thesis-proposal.tex
\section{Conclusion}

Monolithic programming systems enforce a dichotomy between \emph{built-in constructs}, which enjoy support throughout the programming system but must be designed and implemented by the system designer in advance, and \emph{user-defined constructs}, which can be distributed as user libraries, but must creatively combine and repurpose the small set of available built-in constructs to express all desired run-time, compile-time and edit-time behaviors. This dichotomy has forced researchers and domain experts who want to significantly advance the state-of-the-art in programming systems, particularly in compiled, statically-typed systems where this dichotomy is most salient, to design and develop new languages. But this then couples their core innovations with a collection of unrelated design decisions, requires more effort both for the experts developing the innovation and its targeted user community, and leads to intrinsic problems for users developing applications consisting of components written in different languages.
\newpage
\listoftodos
\newpage