### 2013-09-10 thesis-proposal/acc.tex
updated proposal

% !TEX root = omar-thesis-proposal.tex
\section{Active Code Completion}