### 2013-09-17 thesis-proposal/acc.tex
Updated beginning of proposal

% !TEX root = omar-thesis-proposal.tex
\section{Active Code Completion}\label{acc}