### 2013-09-17 thesis-proposal/conclusion.tex

% !TEX root = omar-thesis-proposal.tex
\section{Conclusion}

\newpage
\listoftodos
\newpage