### 2014-01-09 thesis-proposal/acc.tex
continuing work on proposal

% !TEX root = omar-thesis-proposal.tex
\section{Active Code Completion}\label{acc}
Murphy-Hill and Murphy, CSCW paper on peer interaction to enable discovery of tools

Improving program navigation with an active help system. CASCON 2010.

Improving software engineer's fluency by recommending dev environment commands. FSE 2012.