### 2014-01-16 thesis-proposal/acc.tex
Full rough draft.

% !TEX root = omar-thesis-proposal.tex
\section{Active Code Completion}\label{acc}
Software developers today make heavy use of the code completion support found in modern source code editors  \cite{murphy_how_2006}. Most editors provide code completion in the form of a floating menu containing  contextually-relevant variables, fields, methods, types and other code snippets. By navigating and selecting from this menu, developers are able to avoid many common spelling and logic errors, eliminate unnecessary keystrokes and explore unfamiliar APIs without incurring the mental overhead associated with switching to an external documentation tool or API browser.
% The way code completion menus brought documentation systems into the editor, active code completion interfaces bring external tools into the editor.

Several refinements and additions to the code completion menu have previously been suggested in the literature. These have focused on leveraging additional sources of information, such as databases of usage history \cite{robbes_how_2008}\cite{HouPletcher2011}, inheritance information \cite{HouPletcher2011}, API-specific information \cite{HouPletcher2011}\cite{Lee+2008}, partial abbreviations \cite{Han+2009}, examples extracted from code repositories \cite{bruch_learning_2009}\cite{Brandt+2010} and crowdsourced information \cite{mooty_calcite:_2010}\cite{SnipMatch}, to increase the relevance and sophistication of the featured menu items. As with the standard form of code completion, many of these sources of data can also be utilized via external tools (e.g. Calcite \cite{mooty_calcite:_2010} uses information that could already be accessed using the Jadeite \cite{conf/vl/StylosFYM09} tool). Empirical evidence presented in these studies, however, suggests that directly integrating these kinds of tools into the editor is particularly effective. For example, users of the Calcite tool completed 40\% more tasks in a lab study (unfortunately, a Jadeite control group was not included.)

In all of these systems, the code completion interface has remained primarily menu-based. When an item is selected, code is inserted immediately, without further input from the developer. These systems are also difficult to extend: a fixed strategy determines the completions that are available, so library providers cannot directly specify new domain-specific or contextually-relevant logic. 
In this paper we propose a technique called {\it active code completion} that eliminates these restrictions\footnote{Portions of this work previously appeared in a poster abstract \cite{ACC_VLHCC}.}. This  makes developing and integrating a broad array of highly-specialized developer tools directly into the editor, via the familiar code completion command, significantly simpler. This technique is motivated by the evidence discussed above and further evidence provided in this paper that developers prefer, and make more effective use of, tools that do not require leaving the immediate editing environment.

In this paper, we discuss active code completion in the context of object construction. For example, consider the following Java code fragment:

\begin{lstlisting}
  public Color getDefaultColor() \{
      return \textvisiblespace
\end{lstlisting}

\begin{figure*}\label{color}
\begin{center}
\includegraphics[width=40pc]{color_palette.png}\end{center}
\caption{(a) An example code completion palette associated with the \texttt{Color} class. (b) The source code generated by this palette.}
\end{figure*}

If the developer invokes the code completion command at the indicated cursor position (\textvisiblespace), the editor  looks  for a {\it palette definition} associated with the {\it type} of the expression being entered, which in this case is  \verb|Color|. If an associated palette is found, a menu item briefly describing this palette is added to the standard code completion menu. When selected, the corresponding palette is shown, replacing the standard code completion menu. Figure 1a gives an example of a simple palette that may be associated with the \verb|Color| class\footnote{A video demonstrating this process is available at \\\url{http://www.cs.cmu.edu/~NatProg/graphite.html}.}.

The developer can interact with such palettes to provide parameters and other information related to her intent, and receive immediate feedback about the effect these choices will have on the behavior of the object being constructed. When this interaction is complete, the palette generates appropriate source code for insertion at the cursor. Figure 1b shows the inserted code after the user presses \textsc{enter}.

In accordance with best practices, we sought to address the following questions before designing and implementing our active code completion system:

\begin{itemize}
\item What {\it specific} use cases exist for this form of active code completion in a professional development setting? 
\item What {\it general} criteria are common to types that would and would not benefit from an associated palette?
\item What are some relevant usability and design criteria for palettes designed to address such use cases?
\item What capabilities must the underlying active code completion system provide to enable these use cases and user interface designs?
\end{itemize}

To help us answer these questions, we conducted a survey of 473 professional developers (Section II). Their responses, along with information gathered from informal interviews and code corpus analyses, revealed a number of non-trivial functional requirements for palette interfaces as well as the underlying active code completion architecture (Section III). Participants also suggested a large number of use cases, demonstrating the broad applicability of this technique. We organize these into several broad categories (Section IV).

Next, we describe Graphite, an Eclipse plug-in that implements the active code completion architecture for the Java programming language (Section V), allowing Java library developers to associate custom palettes with their own classes. We describe several design choices that we made to satisfy the requirements discovered in our preliminary investigations and briefly examine necessary trade-offs.

Finally, we conducted a pilot lab study with a more complex palette, implemented using Graphite, that assists developers as they write regular expressions (Section VI). The study provides specific evidence in support of the broader claim that highly-specialized tools that are integrated directly with the editing environment are particularly useful. We conclude that active code completion systems like Graphite are useful because they make developing, deploying and discovering such tools fundamentally simpler.

\subsection{Remaining Tasks}
\begin{itemize}
\item Related work: Murphy-Hill and Murphy, CSCW paper on peer interaction to enable discovery of tools; 
Improving program navigation with an active help system. CASCON 2010.; 
Improving software engineer's fluency by recommending dev environment commands. FSE 2012.
\item Discuss safety more specifically.
\end{itemize}