### 2014-01-16 thesis-proposal/timeline.tex

% !TEX root = omar-thesis-proposal.tex
\section{Timeline}
\begin{itemize}
\item @$\lambda$ - ICFP - Mar. 1
\item Wyvern - ECOOP or OOPSLA - late March
\item Ace - PLDI or OOPSLA - late March
\item Then write everything up.
\end{itemize}