### 2014-02-11 thesis-proposal/conclusion.tex

% !TEX root = omar-thesis-proposal.tex
%\section{Conclusion}

%\newpage
%\listoftodos
%\newpage