### 2014-02-11 thesis-proposal/timeline.tex

% !TEX root = omar-thesis-proposal.tex
\section{Timeline}
To summarize, we plan on completing the work in this thesis per the following schedule:
\begin{itemize}
\item The work on @$\lambda$ will be submitted on Mar. 1 to ICFP. 
\item The work on Ace will be submitted on Mar. 25 to OOPSLA.
\item The work on Wyvern will, pending likely final acceptance, be written up in final form by May 12 for ECOOP.
\item The work on Graphite has been completed and published at ICSE.
\end{itemize}