### 2014-06-04 atlang-gpce14/atlang-gpce14.tex
atlang-gpce14 submission

%\documentclass[12pt]{article}
\documentclass[9pt,preprint]{sigplanconf}
\newcommand{\sigmat}[1]{\sigma_{\text{#1}}}

\newcommand{\F}[1]{{\sf #1}~}
\newcommand{\FF}[1]{{\sf #1}}
\newcommand{\Q}{\FF{Arg}}
\usepackage{sfmath}
\usepackage{times}
%\setlength{\parskip}{0pt}
%\setlength{\parsep}{0pt}
\setlength{\headsep}{10pt}
%\setlength{\topskip}{0pt}
%\setlength{\topmargin}{0pt}
%\setlength{\topsep}{0pt}
%\setlength{\partopsep}{0pt}

\usepackage{subfigure}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb} 
%\usepackage{amsthm}
%\newtheorem{corollary}{Corollary}
\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
 \usepackage{longtable}
 \usepackage{booktabs}
 \usepackage{pdflscape}
 \usepackage{colortbl}%
 \newcommand{\myrowcolour}{\rowcolor[gray]{0.925}}
 \usepackage{wasysym}
\lstset{
  language=Python,
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  commentstyle=\itshape\color{light-gray},
  basicstyle=\ttfamily\scriptsize,
  morekeywords={lambda, self, assert, cls, @classmethod, with},
  numbers=left,
  numberstyle=\scriptsize\color{light-gray}\textsf,
  numbersep=0.5em,
  xleftmargin=1.5em,
  stringstyle=\color{mauve},
  escapeinside={^}{^},
  aboveskip=1pt,
  belowskip=1pt
}
\lstdefinestyle{Bash}{
    language={}, 
    numbers=left,
    numberstyle=\scriptsize\color{light-gray}\textsf,
    moredelim=**[is][\color{blue}\bf\ttfamily]{`}{`}
}
\lstdefinestyle{OpenCL}{
  language=C++,
  morekeywords={kernel, __kernel, global, __global, size_t, get_global_id, sin, printf, int2}
}

\newcommand{\lstinlinep}[1]{\lstinline[language=Python,basicstyle=\ttfamily\small]{#1}}

\usepackage{float}
\floatstyle{ruled}
\newfloat{codelisting}{tp}{lop}
\floatname{codelisting}{Listing}
\setlength{\floatsep}{5pt}
\setlength{\textfloatsep}{6pt}
%\setlength{\intextsep}{0pt}
\setlength{\belowcaptionskip}{0pt}

\definecolor{very-light-gray}{gray}{0.9}

\usepackage{url}

\usepackage{todonotes}

\usepackage{placeins}

\usepackage{textpos}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\lefthyphenmin=62
\sloppy
%\makeatletter
%\renewcommand\section{\@startsection{section}{1}{\z@}%
%                       {-18\p@ \@plus -4\p@ \@minus -4\p@}%
%                       {8\p@ \@plus 4\p@ \@minus 4\p@}%
%                       {\normalfont\large\bfseries\boldmath
%                        \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
%\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
%                       {-18\p@ \@plus -4\p@ \@minus -4\p@}%
%                       {8\p@ \@plus 4\p@ \@minus 4\p@}%
%                       {\normalfont\normalsize\bfseries\boldmath
%                        \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
%\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
%                       {-5\p@ \@plus -4\p@ \@minus -4\p@}%
%                       {-0.5em \@plus -0.22em \@minus -0.1em}%
%                       {\normalfont\normalsize\bfseries\boldmath}}
%\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
%                       {-4\p@ \@plus -4\p@ \@minus -4\p@}%
%                       {-0.5em \@plus -0.22em \@minus -0.1em}%
%                       {\normalfont\normalsize\itshape}}
%\makeatother

    \renewcommand{\topfraction}{0.9}  % max fraction of floats at top
    \renewcommand{\bottomfraction}{0.8} % max fraction of floats at bottom
\newcommand{\rulename}[1]{{\textsc{\textbf{#1}}}}
\def \TirNameStyle #1{\small\rulename{#1}}
\begin{document}

\conferenceinfo{-}{-} 
\copyrightyear{-} 
\copyrightdata{[to be supplied]}

\title{Active Type Constructors}
\subtitle{Unambiguously Growing a Type System Inside a Python}
%\titlebanner{{\tt \textcolor{Red}{{\small Under Review -- distribute within CMU only.}}}}        % These are ignored unless
%\preprintfooter{Distribute within CMU only.}   % 'preprint' option specified.

%\authorinfo{~}{~}{~}
\authorinfo{Cyrus Omar \and Jonathan Aldrich}
          {          Carnegie Mellon University}
          {\{comar, aldrich\}@cs.cmu.edu}

\maketitle
\begin{abstract}
%Evidence suggests that programmers are reluctant to adopt new languages to
%gain access to new abstractions, even when they agree that these abstractions
%could be valuable. This suggests a need for languages that satisfy two key
%growth criteria: compatibility with existing programming ecosystems; and
%internal extensibility, so that new abstractions do not require new
%languages.
%
%We introduce @\texttt{lang}, a language compatible with the popular Python ecosystem.
%While Python, like other similar languages, is satisfactory for simple
%scripting, @\texttt{lang} is designed for more complex situations where static
%typechecking is beneficial. Unlike most statically-typed languages, @\texttt{lang}'s type
%system can be extended from within, by associating compile-time functions with
%type definitions. The compiler selectively invokes these according to a
%type-directed protocol, avoiding interference problems. Despite these design
%constraints, active types in @\texttt{lang} can express the static and dynamic semantics
%of a range of functional, object-oriented, parallel and domain-specific
%abstractions, all within libraries.
Programmers justifiably prefer abstractions available as libraries for an existing language over those requiring a switch to a new language, but library providers cannot extend these language's type systems directly, so powerful statically typed abstractions often remain obscure. This suggests a need for a language that is \emph{backwards compatible} with the libraries, tools and infrastructure of a widely adopted language but that has an  \emph{internally extensible type system}, organized around a mechanism that guarantees that any combination of extensions can be imported  without  ambiguity.

Our aim in this paper is thus to 1) introduce such a mechanism and 2) show how it can be implemented in a backwards compatible manner within an existing widely used language. We do so with two artifacts: @\texttt{lang}, an  extensible statically typed language implemented as a library inside Python, and @$\lambda$, a minimal calculus. Both @\texttt{lang} and @$\lambda$ have a fixed syntax but  determine the semantics of each term constructor by delegating control, according to a novel bidirectional typechecking and translation protocol, to methods associated with a relevant \emph{active type constructor}, defined in a library. The type constructor delegated control over a term is stable even when additional extensions are imported, so we avoid ambiguities by construction. % By flexibly  repurposing a fixed syntax in this type-directed manner, we avoid ambiguities by construction. %In @\texttt{lang}, types and functions are defined using Python's syntax, distributed using Python's package system and can call into Python libraries, so @\texttt{lang} is uniquely backwards compatible. We show that active type constructors as implemented in @\texttt{lang} permit practical library implementations of a variety of type system fragments that would normally need to be built into a language. %We also show how to construct safe, natural and extensible foreign function interfaces using a novel form of ``staged'' type propagation from Python into @\texttt{lang} and then through to OpenCL. % can be used to enable type system extensions and demonstrates how type safety can be maintained by lifting a technique from the typed compilation literature into the language.%To show that the method scales, we give a full implementation of the OpenCL programming language as a library. %We briefly examine the type theoretic foundations of this approach and give modular criteria that guarantee that an extension is safe and modular. %@\texttt{lang} can be used from the shell or interactively from within Python\todo{take this sentence out?}.
%Researchers developing languages and abstractions for high-performance computing must consider a number of design criteria, including performance, verifiability, portability and ease-of-use. Despite the deficiencies of legacy tools and the availability of seemingly superior options, end-users have been reluctant to adopt new language-based abstractions. We argue that this can be largely attributed to a failure to consider three additional criteria: continuity, extensibility\- and interoperability. This paper introduces @\texttt{lang}, a language that aims to satisfy this more comprehensive set of design criteria. To do so, @\texttt{lang} introduces several novel compile-time mechanisms, makes principled design choices, and builds upon existing standards in HPC, particularly Python and OpenCL. OpenCL support, rather than being built into the language, is implemented atop an extensibility mechanism that also admits abstractions drawn from other seemingly disparate paradigms. The core innovation underlying this and other features of @\texttt{lang} is a novel reification of types as first-class objects at compile-time, representing a refinement to the concept of active libraries that we call \emph{active types}. We validate our overall design by considering a case study of a simulation framework enabling the modular specification and efficient execution of ensembles of neural simulations across a cluster of GPUs.
\end{abstract}

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Asking programmers to import a library is simpler than asking them to adopt a new programming language. Indeed, recent empirical studies underscore the difficulties of driving new languages into adoption, finding that extrinsic factors like compatibility with large existing code bases, library availability, familiarity and tool support are at least as important as intrinsic features of the language \cite{Meyerovich:2013:EAP:2509136.2509515,chen05}.

Unfortunately, researchers and domain experts designing new abstractions can  find it difficult to soundly and naturally implement them as libraries, particularly when these require strengthening the language's static type system. In these situations, abstraction providers generally develop a new language dialect, increasingly with the aid of tools like language workbenches \cite{erdweg2013state}. Unfortunately, this \emph{language-oriented approach} \cite{journals/stp/Ward94} is problematic at language boundaries: working with components written in a foreign language requires programming against (and thus revealing) its internal implementation. This is brittle, unnatural and unsafe and can cause performance issues due to additional necessary checks and conversions at language boundaries. %To avoid this, it is necessary to combine the different languages and their associated tools, but this often requires significant effort and expertise (we will discuss this further in Sec. \ref{related-work}).

These issues can prevent even dialects introducing only a few new constructs from being adopted. A study comparing a Java dialect, Habanero-Java (HJ), with a comparable library, \lstinlinep{java.util.concurrent}, for example, found that the language-based abstractions in HJ were easier to use and provided useful static guarantees \cite{cave2010comparing}. Nevertheless, it concluded that the library-based abstractions remained more practical outside the classroom because using HJ, as a distinct dialect of Java with its own type system, would be problematic in settings where some developers had not adopted it, but needed to interface with code that had. It would also be difficult to use it in combination with other abstractions also implemented as dialects of Java. Moreover, its tool support is more limited. 
This speaks to a widespread problem: programmers and development teams often cannot use  the typed abstractions they prefer because they are only available bundled with specialized languages they cannot  adopt \cite{Meyerovich:2012:SDR:2414721.2414724,Meyerovich:2013:EAP:2509136.2509515}. %These languages become, in essence, merely vehicles for communicating ideas, rather than useful tools.

 %This issue was perhaps most succinctly expressed by a participant in a recent study by Basili et al. \cite{basili2008understanding} who stated ``I hate MPI, I hate C++. [But] if I had to choose again, I would probably choose the same.''
%Unfortunately, researchers and domain experts (collectively, \emph{providers}) who design potentially useful new abstractions often find it difficult to implement them in terms of the general-purpose abstraction mechanisms available in existing languages. 
%Using a foreign construct can be  unnatural and, if the languages have  different type systems, can also cause safety problems.

Internally extensible languages have promised to reduce the problems associated with distinct dialects by giving abstraction providers more direct control over a language's syntax and semantics from within libraries. %In   an internally extensible language, programmers  can \emph{granularly} import the primitive constructs best suited to each component of their application or library. Providers can thus more easily develop, deploy and evaluate new abstractions in the context of existing codebases, narrowing one of the gaps between research and practice \cite{basili2008understanding}. 
%Indeed, abstractions could compete on their individual merit, rather than  by the serious adoption barriers mentioned above.
Unfortunately, the mechanisms available today have several problems. First, they are themselves often available only within a distinct dialect of an existing language and so face the classic ``chicken-and-egg'' problem: a language like SugarJ \cite{erdweg2011sugarj} (which permits syntax extensions over a fixed semantics) must overcome the same extrinsic barriers as a language like HJ. Second, giving abstraction providers too much control over the language can permit the definition of ambiguous extensions, a problem that is detected ``too late'': when extensions are combined by abstraction clients, who are not necessarily  capable of understanding and fixing the ambiguity.% Too many constraints can leave it difficult to express real-world abstractions.%It is difficult to guarantee that extensions are sound in isolation and more so to guarantee that they can be {safely} combined. %This requires further research before it would be appropriate to widely rely on them (which we will discuss as we go on).

\paragraph{Terminology}
A type system is typically defined in fragments organized around (and often identified by) a \emph{type constructor} (e.g. a fragment defining binary product types, organized around the type constructor $\fvar{prod}$). The syntax of each fragment can be characterized by its contributions to the \emph{concrete syntax} (e.g. expression forms \lstinlinep{(e, e)}, \lstinlinep{e.l} and \lstinlinep{e.r} and type form \lstinlinep{T * T}) and the \emph{abstract syntax}, where terms are written uniformly by applying a \emph{term constructor} to an \emph{index} and zero or more \emph{arguments}, e.g. $\FF{pair}[\FF{triv}](e_1; e_2)$ or $\FF{prj}[\FF{left}](e)$. Term indices are evaluated, if necessary, in an earlier \emph{phase} than term arguments and trivial indices are often omitted for clarity of presentation. The indices of expression terms,  like these examples,  are \emph{static terms}, $\sigma$. \emph{Types}, $\tau$, can be seen as static terms constructed by static term constructor $\FF{ty}$, indexed by a type constructor and having another (possibly trivial) static term as an argument, called (by convention) the \emph{type index}, e.g. $\FF{ty}[\fvar{prod}](\FF{spair}(\tau_1; \tau_2))$, often  abbreviated $\fvar{prod}\texttt{[}\tau_1; \tau_2\texttt{]}$ (%or stylized $\tau_1 \times \tau_2$. %Static terms are also called \emph{type-level terms}.  
 often, the static terms are not made explicit but rather assumed to be in the ambient metatheory).%Types classify expressions according to the rules comprising the fragment's type system. A type safe language is one where evaluation of a well-typed expression cannot lead to undefined behavior or result in a value of a different type, greatly simplifying static reasoning about program behavior and eliminating the need for many run-time checks.

%In a monolithic language design, decisions about which fragments are to be combined to constitute a complete  language are made centrally, so the designer can ensure that the concrete syntax is unambiguous and that other metatheoretic concerns, like guarantees of type safety, are maintained. To decentralize control over these decisions, so that fragments can be defined as libraries and combined arbitrarily, these guarantees must be shown to hold for all possible combinations of fragments deemed individually valid. This can quickly lead to difficulties.

  Each fragment also defines rules that contribute to the \emph{static semantics} (specifying how types are assigned to expressions at ``compile-time'') and the \emph{dynamic semantics} (specifying the behavior of well-typed expressions at ``run-time'') \cite{pfpl}. The dynamic semantics of full-scale languages are generally implemented (or even directly specified, as in the Harper-Stone semantics for Standard ML \cite{Harper00atype-theoretic}) by elaboration to a simpler \emph{internal language}.% In some languages, static terms themselves may have their own static and dynamic semantics (the \emph{static statics} and the \emph{static dynamics}).

The most apparently difficult problems related to combining fragments to form a typed programming language arise when attempting to enable the decentralized introduction of new concrete forms or term constructors. Syntactic ambiguities are difficult to modularly preclude and many tools operate by exhaustive case analysis over the concrete or abstract syntax of the terms of the language they work with (e.g. syntax highlighters, editor modes, pretty-printers, style checkers, documentation generators). Indeed, enabling the decentralized introduction of new term constructors is the standard example used when describing the more general problem of enabling the extension of a sum type with new cases as well as new  functions over its cases  \cite{Reynolds75}, prompting Wadler to dub it the \emph{expression problem} \cite{wadler1998expression}.

\paragraph{Approach} Fewer tools operate by exhaustive case analysis over type constructors. In a conventionally formulated programming language, however, it is difficult to introduce a new type system fragment without adding new concrete syntax or term constructors to accompany the new type constructors. In this paper, we aim to show that this need not be the case. We introduce a method that allows library providers to introduce new type system fragments by defining new type constructors in libraries while leaving the concrete syntax, the term constructors and the internal language fixed and enriching the static language so that it has its own dynamic semantics (the \emph{static dynamics}). The key idea is that we determine the semantics of each term not based directly on the identity of its term constructor (the usual ``syntax-directed'' approach), but rather by delegating to a static function associated with a relevant type constructor. The term constructor's identity only serves to determine how the delegate is assigned. We call user-defined type constructors equipped with static functions and distributed in libraries \emph{active type constructors}.

As a simple example, consider the concrete term \lstinlinep{e.r} (where \lstinlinep{e} is an expression and \lstinlinep{r} is a static label) corresponding to the now generalized abstract term  $\FF{attr}[\text{``r''}](e)$. Our semantics delegate responsibility for this term to a static function associated with the type constructor of the type recursively synthesized by the ``target'', $e$. As a result, tuple projection, as described above, can be introduced without the need to consider that it might syntactically coincide with projection out of a functional record or an object with a field labeled \lstinlinep{r} (both examples we will discuss). Each of these abstractions can use the same form and term constructor. As we will detail, terms for which there is no clear ``target'' from which to determine a delegate (e.g. number or list literals) instead either use the type they are being analyzed against or, if not available, a ``default'' delegate determined by the type constructor governing the function in which they appear. Symmetric binary terms (e.g. addition) use a third protocol, which we will discuss. There is always exactly one active type constructor delegated  responsibility over a term and this delegate is stable with respect to additional imports, so semantic ambiguities cannot arise by construction. %This is a generalization of a previous mechanism that we have developed, which deals only with extending the elaboration, but not the typechecking, of the bodies of delimited literals \cite{omar-ecoop14}.

%The approach we describe here is general with respect to the fixed syntax of the language it is introduced within. To implement it, we must only 1) define the delegation protocol that determines how a type constructor is delegated responsibility over each form, basing it on a bidirectional type system that distinguishes locations where an expression must \emph{synthesize} a type (e.g. to the right of bindings) from those where it can be \emph{analyzed} against a known type (this has also been called ``local type inference'' \cite{Pierce:2000:LTI:345099.345100}); 2) expose a fixed static language, which is used to implement active type constructors, construct types by applying type constructors to static indices and  import libraries; and 3) expose a fixed internal target language to the static language (fragment providers  implement the dynamic semantics of their fragments by translation to this language).

\paragraph{Specific Aims} Our specific aims in this paper are to introduce active type constructors and demonstrate that they are expressive, practically realizable within existing languages and theoretically well-founded. Our target audience is typed language fragment providers (researchers or domain experts) and language designers interested in extensibility mechanisms. We do not aim to evaluate the claim that our approach will ultimately be of use to fragment clients (i.e. normal developers) in writing and reasoning about a variety of real-world programs (this first requires adoption by our target audience). We describe two artifacts in pursuit of these aims:
\begin{enumerate}
\item Because we are committed to a fixed syntax, but aim to show that our mechanism does not critically rely on any particular syntactic choices, it is pragmatic to use an existing syntax for which tooling already exists. In this paper, we choose Python's syntax. Moreover, we use the Python language itself as both the static language and the internal language. This makes it easy for us to implement our language, called @\texttt{lang}, itself as a library inside Python, using its quasiquotation and reflection facilities.\footnote{The name was chosen as an initialism for ``actively  typed language'' and because Python marks function decorators, which we use extensively for this purpose, using the @ symbol.} Python is also an interesting starting point  because it  can be thought of as beginning with a degenerate static type system providing just one trivially indexed type constructor, \texttt{dyn} \cite{pfpl}. Taken together, @\texttt{lang} is almost completely backwards compatible with the Python ecosystem with remarkably little effort: we inherit a large number of  existing tools for working with Python files, Python's substantial packaging infrastructure and access to all existing Python libraries. We introduce @\texttt{lang} with practical examples of fragments as libraries in Sec. \ref{usage}.
\item In Sec. \ref{theory}, we describe active type constructors in their essential form, developing a minimal calculus called @$\lambda$. The syntax provides only variables, variable binding constructs, type ascription and two generic term constructors. Perhaps surprisingly, we show how a conventional syntax like Python's can be recovered by a purely syntactic desugaring to these term constructors. The semantics for @$\lambda$ take the form of a \emph{bidirectionally typed elaboration semantics} where the available type constructors are tracked by a \emph{fragment context}. We use a simple lambda calculus as both our static and internal language to emphasize the fundamental character of the approach.
\end{enumerate}
%@\texttt{lang} can be used from the shell, producing source files that can be further compiled or executed by external means. @\texttt{lang} functions targeting a language with Python bindings (e.g. Python itself, as well as the others just described) can also be compiled and invoked interactively from Python. 
%@\texttt{lang} supports a form of staged \emph{ad hoc} polymorphism (based on singleton types) wherein the ``dynamic type'' of a Python value can determine a static type and trigger active typechecking and translation ``just-in-time'' when Python code calls a generic @\texttt{lang} function. We show how this can significantly streamline interactive workflows (e.g. in scientific computing) that use Python for orchestration and exploratory tasks (e.g. plotting) but call into a statically-typed  language via a foreign-function interface (FFI) for performance- and safety-critical portions. We provide a complete implementation of the OpenCL type system (a language for low-level data parallel programming on GPUs), show several higher-level extensions to it, and discuss bindings that bridge between the type systems of @\texttt{lang}, OpenCL and Python's standard numerics package, \lstinlinep{numpy}, all as a library (contribution 3, Sec. \ref{targets}).%We will discuss in detail as our primary realistic case study.

%The remainder of the paper is organized as follows: in Sec. \ref{usage}, we describe the basic structure and usage of @\texttt{lang} with an example that uses several interesting @\texttt{lang} type libraries. In Sec. \ref{att}, we detail how active typing and translation delegate control, at compile-time, to active types and bases, and discuss how this represents a form of extensibility that side-steps the expression problem. In Sec. \ref{theory}, we reduce this mechanism to  a core lambda calculus, $\lam@\texttt{lang}$,  clarify the relationship to   type-level computation and show how type safety can be maintained. In Sec. \ref{targets}, we discuss generic functions, FFIs and staging via a complete implementation of the OpenCL type system (and extensions to it) as a library.

In Sec.  \ref{related}, we describe how active type constructors relate to several threads of related work. We conclude in Sec. \ref{discussion} by summarizing the key features needed by a host language to practically support active type constructors, and delineating present limitations and promising directions for future work. %The appendix has details omitted for concision as well as several more examples, including functional datatypes, flexible complex numbers and a previously little-explored type system for tracking string invariants using regular expressions statically (which we show being used in Listing \ref{example}, amongst  other examples).

\section{Active Type Constructors in @\texttt{lang}}\label{usage}

%
Listing \ref{example} shows an example of a well-typed @\texttt{lang} program for which strong correctness guarantees have been statically maintained using several type constructors defined in libraries. As suggested by line \ref{hello-ct}, the top level of every @\texttt{lang} file can be seen as a \emph{compilation script} written directly in Python, the static language. @\texttt{lang} requires no modifications to the language (version 2.6+ or 3.0+) or features specific to the primary implementation, CPython (so @\texttt{lang} supports alternative implementations like Jython and PyPy). This choice pays off immediately on the first line: @\texttt{lang}'s import mechanism is Python's import mechanism, so Python's build tools (e.g. \lstinlinep{pip}) and package repostories (e.g. \lstinlinep{PyPI}) are directly available for distributing @\texttt{lang} libraries, including those defining type constructors. Here, \lstinlinep{atlib} is the ``standard library'' but does not benefit from special support in \lstinlinep{atlang} itself.% Nothing in \lstinlinep{atlib} benefits from special support from \lstinlinep{atlang}, the core language package. We will see \lstinlinep{atlang} being used directly below. %@\texttt{lang} can be installed with a single command: \emph{omitted for review}. %@\texttt{lang} is agnostic about the particular design choices that we made in our examples.

%The top-level statements in an @\texttt{lang} file, like the \lstinlinep{print} statement on line 10, are executed to control the compile-time behavior, rather than the run-time behavior, of the program. %That is, Python serves as the \emph{compile-time metalanguage} (and, as we will see shortly, the \emph{type-level language}) of @\texttt{lang}. %For readers familiar with C/C++, Python can be thought of as serving a role similar to (but more general than) its preprocessor and template system (as we will see).
\subsection{Active Typechecking}

\paragraph{Types}
\begin{codelisting}[t]
\begin{lstlisting}
^\label{imports-start}^from atlib import record, decimal, dyn, string, 
  string_in, proto, fn, printf^\label{imports-end}^

^\label{hello-ct}^print "Hello, compile-time world!"

^\label{decl-Details}^Details = record[
  'amount' : decimal[2], 
  'memo' : dyn]
^\label{decl-Account-start}^Account = record[
  'name' : string, 
  'account_num' : string_in[r'\d{2}-\d{8}']]^\label{decl-Account-end}^
^\label{decl-Transfer}^Transfer = proto[Details, Account]

^\label{log-transfer-start}^@fn[Transfer, ...]
def log_transfer(t):
  """Logs a transfer to the console."""
  printf("Transferring %s to %s.", 
    t.amount.to_string, t.name)^\label{log-transfer-end}^

^\label{main-start}^@fn[...]
def __main__():
  ^\label{account-start}^account = {name: "Annie Ace", 
    account_num: "00-00000001"} [:Account]^\label{account-end}^
  log_transfer(({amount: 5.50, memo: None}, account))
  log_transfer(({amount: 15.00, memo: "Rent payment"}, 
    account))^\label{main-end}^

print "Goodbye, compile-time world!"
\end{lstlisting}
\caption{[\texttt{listing\ref{example}.py}] An @\texttt{lang} compilation script.}
\label{example}
\end{codelisting}
Types are constructed programmatically in the static language by applying a type constructor to an index:  \lstinlinep{tycon[idx]}. This is in contrast to many statically-typed languages where types (e.g. datatypes, interfaces) cannot be constructed computationally, but are only declared ``literally''. %Put another way, 
%@\texttt{lang} supports \emph{static computation at the level of types} and Python is its type-level language. The static language subsumes th%The compilation script can assign shorter names to types for convenience (removing the need for  facilities like \lstinlinep{typedef} in C or \lstinlinep{type} in Haskell). 
In this example, we see several useful types being constructed:
\begin{enumerate}
\item On line \ref{decl-Details}, we construct a functional record type with two (immutable) fields, assigning it to the identifier \lstinlinep{Details}. The field \lstinlinep{amount} has type \lstinlinep{decimal[2]}, which classifies decimal numbers having two decimal places, and \lstinlinep{memo} has type \lstinlinep{dyn}, classifying dynamic Python values.  %Here, \lstinlinep{record} and \lstinlinep{decimal} are {indexed type constructors}. 
Syntactically, the index for \lstinlinep{record} is provided using borrowed Python  syntax for array slices (e.g. \lstinlinep{a[min:max]}) to approximate conventional notation for type annotations. The field names are simply static strings (and could be computed rather than written literally, e.g. \lstinline{'a' + 'mount'}, to support features like \emph{type providers} \cite{Syme:2013:TIF:2429376.2429378}). 
\item On line \ref{decl-Account-start}, we construct another record type. The field \lstinlinep{name} has type \lstinlinep{string} and the field \lstinlinep{account_num} has a type classifying strings guaranteed statically to be in the regular language specified  by the  regular expression pattern provided as the index (written here as a \emph{raw string literal} \cite{python}). This  fragment, based on recent work, statically tracks the language an immutable string is in, providing operations like concatenation, demonstrating the expressive power of our approach \cite{fulton-thesis}. %To our knowledge, this type system is not  expressible in other simply typed languages (we do not consider dependently-typed logics without a distinct phase separation).
\item 
%Prototypic inheritance  is a feature of some dynamically-typed languages (most notably, Self \cite{Ungar:Smith:oopsla:1987} and Javascript). 
On line \ref{decl-Transfer}, we construct a simple \emph{prototypic object type} \cite{Lie86}, \lstinlinep{Transfer}, classifying values consisting of a \emph{fore} of type \lstinlinep{Details} and a \emph{prototype} of type \lstinlinep{Account}. %\footnote{We encourage the convention of naming types that are semantic in camelcase and types and type constructors that are structural in nature, like \texttt{dyn} or \texttt{record}, in lowercase.} 
 If a field cannot be found in the fore, the type system will delegate (statically) to the type of the prototype. This makes it easy to share the values of the fields of a common prototype amongst many fores, here to allow us to describe multiple transfers to the same account.% differing only in the transaction  details.
\end{enumerate}

\paragraph{Incomplete Types} Incomplete types arise from the application of a type constructor to an index containing an ellipsis (a valid array slice form in Python). The static terms \lstinlinep{fn[Transfer, ...]} and \lstinlinep{fn[...]} seen on lines \ref{log-transfer-start} and \ref{main-start} are incomplete types. We will discuss how the elided portion of the index (e.g. the return type) is determined below.

\paragraph{Active Type Constructors} Type constructors are implemented in @\texttt{lang} libraries as Python classes inheriting from \lstinlinep{atlang.Type} (classes here serving as Python's form of open sums). We show portions of \lstinlinep{atlib.fn} in Fig. \ref{fn-top}. Types are instances of these classes. Incomplete types are instead instances of \lstinlinep{atlang.IncompleteType}. The compiler controls instantiation by overloading the subscript operator on the \lstinlinep{atlang.Type} class object (using Python's \emph{metaclass} mechanism \cite{python}).

\paragraph{Function Definitions}
\begin{codelisting}[t]
\begin{lstlisting}
class fn(atlang.Type):
  ^\label{fn_init}^def __init__(self, idx): 
    # called by atlang.Type via the [] operator 
    atlang.Type.__init__(self, 
      self._check_and_norm(idx))

  @classmethod
  ^\label{syn_idx_FunctionDef}^def syn_idx_FunctionDef(cls, ctx, inc_idx, node):
    arg_types = cls._check_and_norm_inc(inc_idx)
    if not hasattr(ctx, 'assn_ctx'): 
      ctx.assn_ctx = { }
    ctx.assn_ctx.update(zip(node.args, arg_types))
    for stmt in node.body: ctx.check(stmt)
    last_stmt = node.body[-1]
    if isinstance(last_stmt, ast.Expr): 
      rty = last_stmt.value.ty
    else: rty = unit
    return (arg_types, rty) # fully specified index

  ^\label{ana_FunctionDef}^def ana_FunctionDef(self, ctx, node):
    if not hasattr(ctx, 'assn_ctx'): 
      ctx.assn_ctx = { } # top-level functions
    ctx.assn_ctx[node.name] = self # recursion
    ctx.assn_ctx.update(zip(node.args, self.idx[0]))
    # all but last
    for stmt in node.body[0:-1]: ctx.check(stmt)
    last_stmt = node.body[-1]
    if isinstance(last_stmt, ast.Expr): 
      ctx.ana(last_stmt.value, self.idx[1])
    elif self.idx[1] == unit or self.idx[1] == dyn: 
      ctx.check(last_stmt)
    else: raise atlang.TypeError("...", last_stmt)
\end{lstlisting}
\caption{A portion of the type constructor \texttt{atlib.fn}.}
\label{fn-top}
\end{codelisting}
 To be typechecked by @\texttt{lang} and define run-time behavior, function definitions must have an \emph{ascription}, provided by \emph{decorating} them with either a type or, here, an incomplete type (see \cite{python} for a discussion of Python decorators; we again use Python's metaclasses and operator overloading to support the use of a class as a decorator). Ascribed function definitions do not share Python's semantics and indeed the underlying Python function is discarded immediately after its abstract syntax and static environment (its closure and the global dictionary of the Python module it was defined in) have been extracted by the compiler using Python's built-in reflection capabilities and \lstinlinep{inspect} and \lstinlinep{ast} packages \cite{python}. The \lstinlinep{ast} package defines Python's term constructors. For function literals (term constructor \lstinlinep{FunctionDef}), the @$\texttt{lang}$ compiler delegates to the type constructor of the ascription in one of two ways, depending on the ascription.

If the ascription is an incomplete type, the compiler invokes the class method \lstinlinep{syn_idx_FunctionDef}, seen on line \ref{syn_idx_FunctionDef}, with the \emph{compiler context} (an object that provides hooks into the compiler and a storage location for accumulating information during typechecking), the incomplete index and the syntax tree of the function definition. Here, this method 1) checks and normalizes the incomplete index (here, checking that the argument types are indeed types and turning empty and single arguments into a 0- or 1-tuples for uniformity; not shown); 2) adds the argument names and types to a field of the context that tracks the types of local assignments (initializing it first if it is a top-level function, as in our example); 3) asks the compiler, via the method \lstinlinep{ctx.check}, to typecheck each statement in the body (discussed below); 4) if the term constructor of the last statement is \lstinlinep{ast.Expr} (a top-level expression), then the type of this expression is the return type, otherwise it is \lstinlinep{unit}; and finally, 5) a fully specified index (and thus a type) is synthesized for the function as a whole, here equivalent to \lstinlinep{fn[Transfer, atlib.unit]}. Note that this choice of using the last expression as the implicit return value (and considering any statement-level term constructor other than \lstinlinep{Expr} to have a trivial value) is made in a library, not by the language itself.% A different function type constructor could be written to support \lstinlinep{return} statements (omitted for concision).% The choices is made by clients .

Were the ascription a (complete) type, the compiler would delegate to \lstinlinep{fn} by calling an instance method on it, \lstinlinep{ana_FunctionDef}, seen on line \ref{ana_FunctionDef}. This method proceeds similarly, but does not need to perform the first step and last steps (and does not need to be a class method) because the index was already determined when the type was constructed, so it can be accessed via \lstinlinep{self.idx}. If the final statement is an expression, is analyzed against the provided return type (see below). Otherwise, only \lstinlinep{unit} or \lstinlinep{dyn} are valid return types (again, a choice made in a library)%. If it is, the expression . % Note that in this example, recursive functions are only supported when the ascription is a type, to avoid complications related to synthesizing a type for a recursive function call.
%On lines \ref{log-transfer-start}-\ref{log-transfer-end}, we see the function \lstinlinep{log_transfer}. We follow Python conventions by starting with a documentation string  that informally specifies the behavior of the function (so standard documentation generators can be used). Before moving into the body, however, we  also give an \emph{argument signature} (line \ref{arg-sig}) stating that the type of \lstinlinep{t} is \lstinlinep{Transfer}, the prototypic object type defined above.

%This again repurposes Python's array slice syntax to approximate conventional notation for type annotations. % (and following conventions for documentation strings, which also describe ``types'' at the end). % The ``keyword'' \lstinlinep{type} was chosen because it is the name of a built-in function in Python (and thus would not be used in the way shown for any other purpose).

\paragraph{Statements}
\begin{codelisting}[t]
\begin{lstlisting}
#class fn(atlang.Type): (cont'd) 
  @classmethod
  def check_Assign_Name(cls, ctx, stmt):
    x, e = stmt.target.id, stmt.value # see ast
    if x in ctx.assn_ctx: ctx.ana(e, ctx.assn_ctx[x])
    else: ctx.assn_ctx[x] = ctx.syn(e)

  @classmethod
  ^\label{syn_Name}^def syn_Name(cls, ctx, e):
    try: return ctx.assn_ctx[e.id]
    except KeyError:
      try: return self._syn_lift(ctx.static_env[e.id])
      except KeyError: raise atlang.TypeError("...", e)

  ^\label{default_asc}^@classmethod
  def syn_default_asc_Str(cls, ctx, e): return dyn
\end{lstlisting}
\caption{Some forms in the body of a function delegate to the type constructor of the function they are defined within (via class methods during typechecking).}
\label{fn-stmts}
\end{codelisting}
The \lstinlinep{ctx.check} method is defined by the compiler. As is the pattern in this paper, the compiler delegates to an active type constructor based on the term constructor of the statement being checked. Most statement-level term constructors other than \lstinlinep{Expr} simply delegate to the type constructor of the function they are defined within by calling class methods of the form \lstinlinep{check_TermCon}, where \lstinlinep{TermCon} is a term constructor or combination of term constructors in a few cases where a finer distinction than what Python itself made was necessary, as defined by \lstinlinep{ast}.

For example, the class method \lstinlinep{check_Assign_Name}, seen in Listing \ref{fn-stmts}, is called for statements of the form \lstinlinep{name = expr}, as on line \ref{account-start} of Listing \ref{example}. In this examples, the assignables context (the \lstinlinep{assn_ctx} field of the compiler context) is consulted to determine whether the name being assigned to already has a type due to a prior assignment, in which case the expression is analyzed against that type using \lstinlinep{ctx.ana}. If not, the expression must synthesize a type, using \lstinlinep{ctx.syn}, and a binding is added. %This interpretation of the assignment form cannot cause the function it appears in to return so \lstinlinep{None} is returned (Python returns \lstinlinep{None} by default if nothing is explicitly returned) and the subsumption rule applies trivially, so we do not need to define \lstinlinep{ana_rty_Assign_Name} separately.

%Return statements, on the other, require separately defining return type synthesis and analysis logic. The class method \lstinlinep{syn_rty_Return} is called for return statements when a return type is not yet known. It must synthesize a return type based on the value being returned (defaulting to \lstinlinep{unit}). Subsequent return statements are processed by \lstinlinep{ana_rty_Return}, which checks the provided value against the known return type. If no value is provided, then the return type must be \lstinlinep{unit}. Note that this function does not need to return a value; if a \lstinlinep{atlang.TypeError} is not raised, then analysis succeeds.

\paragraph{Bidirectional Typechecking of Expressions} The methods \lstinlinep{ctx.ana} and \lstinlinep{ctx.syn} are also defined by the compiler and can be called by type constructors to request type analysis (when the expected type is known) and synthesis (when the type is an ``output''), respectively, for an expression. Once again, the compiler delegates to a type constructor by a protocol that depends on the term constructor, invoking methods of the form \lstinlinep{ana_TermCon} or \lstinlinep{syn_TermCon}. This represents a form of bidirectional type system (sometimes also called a \emph{local type inference system}) \cite{Pierce:2000:LTI:345099.345100}, and the standard subsumption principle applies: if analysis is requested and an \lstinlinep{ana_TermCon} method is not defined but a \lstinlinep{syn_TermCon} method is, then synthesis proceeds and then the synthesized type is checked for equality against the type provided for analysis. Type equality is defined by checking that the two type constructors are identical and that their indices are equal. Two types with different type constructors are never equal, to ensure that the burden of ensuring that typing respects type equality is local to a single type constructor.  No form of subtyping or implicit coercion is supported for the purposes of this paper (though we have designed a mechanism that similarly localizes reasoning to a single type constructor, we do not have room to precisely describe it here.)

\paragraph{Names} If the term constructor of an expression is \lstinlinep{ast.Name}, as when referring to a bound variable or an assignable location, then the type constructor governing the function the term appears in is delegated to via the class method \lstinlinep{syn_Name} (names, when used as expressions, must synthesize a type). We see the definition of this method for \lstinlinep{fn} starting on line \ref{syn_Name} in Listing \ref{fn-stmts}. A name synthesizes a type if it is either in the assignables context or if it is in the static environment. In the former case, the type that was synthesized when the assignment occurred (by \lstinlinep{syn_Assign_Name}) is returned. In the latter case, we must lift the static value to run-time. For the purposes of this paper, we support only other typed @\texttt{lang} functions, Python functions and classes (which are given type \lstinlinep{dyn}, and can only be called with values of type \lstinlinep{dyn}) and modules (which are given a \emph{singleton type} -- a type indexed by the module reference itself -- from which @\text{lang} functions and Python functions and classes can be accessed as attributes in the usual way, see below).%In the future, we plan to detail how the interface between Python and @\texttt{lang} can be enriched to support moving typed values in and out of untyped Python functions safely (by inserting checks, defined again by a type constructor, at the interface between typed and untyped code), but we do not have enough room to adequately detail this mechanism here (and it is tangential to our specific aims).
\paragraph{Literal Expressions and Ascriptions}
\begin{codelisting}[t]
\begin{lstlisting}
class record(atlang.Type):
  def __init__(self, idx): 
    # Sig is an unordered mapping from fields to types
    atlang.Type.__init__(self, Sig.from_slices(idx))

  ^\label{syn_idx_Dict}^@classmethod
  def syn_idx_Dict(self, ctx, partial_idx, e):
    if partial_idx != Ellipsis: 
      raise atlang.TypeError("...bad index...", e)
    idx = []
    for f, v in zip(e.keys, e.values):
      if isinstance(f, ast.Name): 
        idx.append(slice(f.id, ctx.syn(v)))
      else: raise atlang.TypeError("...", f)
    return idx

  ^\label{ana_Dict}^def ana_Dict(self, ctx, e):
    for f, v in zip(e.keys, e.values):
      if isinstance(f, ast.Name):
        if f.id in self.idx.fields: 
          ctx.ana(v, self.idx[f.id])
        else: raise atlang.TypeError("...", f)
      else: raise atlang.TypeError("...", f)
    if len(self.idx.fields) != len(e.keys): 
      raise atlang.TypeError("...missing field...", e)

  ^\label{syn_Attribute}^def syn_Attribute(self, ctx, e):
    if e.attr in self.idx: return self.idx[e.attr]
    else: raise atlang.TypeError("...", e)
\end{lstlisting}
%
%  def trans_type(self, target):
%    if target.supports(examples.fp.EagerNAryProdLR):
%      return target.EagerNAryProdLR()
%    else if isinstance(target, examples.py.Dyn):
%      return target.Dyn()
\caption{A portion of the \texttt{atlib.record} type constructor.}
\label{record}
\end{codelisting}
Python designates literal forms for dictionaries, tuples, lists, strings, numbers and lambda functions. In @\texttt{lang}, the type constructor delegated control over terms arising from any of these forms is a function of how the typechecker encounters the term.

If the term appears in an analytic position, the type constructor of the type it is being analyzed against is delegated control. We see this on line \ref{account-start} of Listing \ref{example}: the dictionary literal form appears in an analytic context, here because an explicit type ascription, \lstinlinep{[:Account]}, was provided. Note that the ascription again repurposes Python's array slice syntax (the start is, conveniently,  optional). The compiler invokes the \lstinlinep{ana_Dict} method on the type the literal is being analyzed against, which is defined by its type constructor, here \lstinlinep{atlib.record}, shown on line \ref{ana_Dict} of Listing \ref{record}. This method analyzes each field value in the literal against the type of the field, extracted from the type index (an unordered mapping from field names to their types, so that type equality is up to reordering). The various literal forms inside the outermost form thus do not need an ascription because they appear in positions where the type they will be analyzed against is provided by \lstinlinep{record}. For example, the value of the field \lstinlinep{account_num} delegates to \lstinlinep{string_in} via the \lstinlinep{ana_Str} method (not shown).

An ascription directly on a literal can also be an incomplete type. For example, we can write \lstinline[basicstyle=\ttfamily\small]{[x, y] [:matrix[...]]} or more concisely \lstinline[basicstyle=\ttfamily\small]{[x, y] [:matrix]} instead of \lstinline[basicstyle=\ttfamily\small]{[x, y] [:matrix[i32]]} when we know \lstinline[basicstyle=\ttfamily\small]{x} and \lstinline[basicstyle=\ttfamily\small]{y} synthesize the type \lstinline[basicstyle=\ttfamily\small]{i32}, because the type constructor can use this information to synthesize the appropriate index. The decorator \lstinline[basicstyle=\ttfamily\small]{@fn[Transfer, ...]}, discussed previously, can be seen as a partial type ascription on a statement-level function literal.\footnote{In Python 3.0+, syntax for annotating function definitions directly with type-like annotations was introduced, so the entire index can be synthesized.} Lambda expressions support the same ascriptions:
\vspace{3px}
\begin{lstlisting}[numbers=none, basicstyle=\ttfamily\scriptsize]
(lambda x, y: x + y) [:fn[(i32, i32), ...]]
\end{lstlisting}
%\vspace{3px}
\noindent
Records support partial type ascription as well, e.g.: 
\vspace{3px}
\begin{lstlisting}[numbers=none,basicstyle=\ttfamily\scriptsize]
{name: "Deepak Dynamic", age: "27"} [:record]
\end{lstlisting}
%\vspace{3px}
\noindent
The compiler delegates to the type constructor by a class method in these cases (just as we saw above with \lstinlinep{fn}). In this case, the class method \lstinlinep{syn_idx_Dict}, shown on line \ref{syn_idx_Dict} of Listing \ref{record}, would be called. Because no record index was provided, an index must be synthesized from the literal itself, and thus the values are not analyzed against a type, but must each synthesize a type.

This brings us to the situation where a literal appears in a synthetic position, as the two string literals above do. In such a situation, the compiler delegates responsibility to the type constructor of the function that the literal appears in, asking it to provide a ``default ascription'' by calling the class method \lstinlinep{syn_default_asc_Str}, shown on line \ref{default_asc} of Listing \ref{fn-stmts}. In this case, we simply return \lstinlinep{dyn}, so the type of the expression above has type \lstinline[basicstyle=\ttfamily\small]{record['name': dyn, 'age' : dyn]}. A different function type constructor might make more precise choices. Indeed, a benefit of using Python as our static language is that it is relatively straightforward to provide a type constructor that allows us to provide different defaults, and control choices like the return semantics, as block-scoped ''pragmas'' in the static language using Python's \lstinlinep{with} statement \cite{python} (details omitted), e.g. 
\begin{codelisting}[h]
\begin{lstlisting}[numbers=none]
with fn.default_asc(Num=i64, Str=string, Dict=record):
  @fn[...] # : fn[(), record["a": i64, "b": string]]
  def test(): {a: 1, b: "no ascriptions needed!"} 
\end{lstlisting}
\caption{Block-scoped settings can be seen by type constructors.}
\label{defaultasc}
\end{codelisting}
\vspace{-10px}
%{In Python 3, syntax for annotating functions (but unfortunately not lambdas) with argument and return type annotations directly  (purely as metadata) was introduced \cite{pep3107}:
%\begin{lstlisting}[numbers=none]
%@fn
%def f(x : T) -> T: # ...
%\end{lstlisting}
%\vspace{-3px}
%This can also be thought of as an incomplete type ascription, which we could write more explicitly \lstinlinep{@fn[...]},  because the literal itself contains the necessary information to synthesize the full index. We show the more universally available notation here because Python 2.\textit{x} is much more widely adopted as of this writing.

%\paragraph{Compound Expressions} 
%For expressions other than names and literals, the type synthesized by one or more subexpressions determines the type constructor delegated responsibility for synthesis and analysis.%The sub-expression that determines this is a function of the term constructor.

\paragraph{Targeted Expressions} Expression forms having exactly one sub-expression, like \lstinlinep{-e} (term constructor \lstinlinep{UnaryOp_USub}) or \lstinlinep{e.attr} (term constructor \lstinlinep{Attribute}), or where there may be multiple sub-expressions but the left-most one is the only one required, like \lstinlinep{e[slices]} (term constructor  \lstinlinep{Subscript}) or \lstinlinep{e(args)} (term constructor \lstinlinep{Call}), are called \emph{targeted expressions}. For these, the compiler first synthesizes a type for the left-most sub-expression, then calls either the \lstinlinep{ana_TermCon} or \lstinlinep{syn_TermCon} methods on that type. For example, we see type synthesis for the field projection operation on records defined starting on line \ref{syn_Attribute} of Listing \ref{record}.% We omit the implementation for concision, but \lstinlinep{fn} defines \lstinlinep{syn_Call} as one would expect, by simply checking each argument against the corresponding argument in the index.

\paragraph{Binary Expressions} \begin{codelisting}[t]
\begin{lstlisting}
class string_in(atlang.Type):
  def __init__(self, idx):
    atlang.Type.__init__(self, self._rlang_of_rx(idx))

  # treats string as string_in[".*"]
  ^\label{handles_Add_with}^handles_Add_with = set([string]) 
  ^\label{syn_BinOp_Add}^@classmethod
  def syn_BinOp_Add(cls, ctx, e): 
    rlang_left = self._rlang_from_ty(ctx.syn(e.left)) 
    rlang_right = self._rlang_from_ty(ctx.syn(e.right))
    return string_in[self._concatenate_langs(
      rlang_left, rlang_right)]
\end{lstlisting}
\caption{Binary operations in \texttt{atlib.string\_in}.}
\label{string_in}
\end{codelisting}
\begin{codelisting}[t]
\begin{lstlisting}
def check_tycon(tycon):
  for other_tycon in tycon.handles_Add_with: 
    if tycon in other_tycon.handles_Add_with:
      raise atlang.AmbiguousTyconError("...", 
        tycon, other_tycon) # (other binops analagous)
\end{lstlisting}
\caption{For each type constructor definition and binary operator, \texttt{atlang} runs a modular handle set check to preclude ambiguity.}
\label{check_tycon}
\end{codelisting}
The major remaining expression forms are the binary operations, e.g. \lstinlinep{e + e} or \lstinlinep{e < e}. These are notionally symmetric, so it would not be appropriate to simply give the left-most one precedence. Instead, we begin by attempting to synthesize a type for both subexpressions. If both synthesize a type with the same type constructor, or only one synthesizes a type, that type constructor is delegated responsibility via a class method, e.g. \lstinlinep{syn_BinOp_Add} on line \ref{syn_BinOp_Add} of Listing \ref{string_in}.

If both synthesize a type but with different type constructors (e.g. we want to concatenate a \lstinlinep{string} and a \lstinlinep{string_in[r]}), then we consult the \emph{handle sets} associated as a class attribute with each type constructor, e.g. \lstinlinep{handles_Add_with} on line \ref{handles_Add_with} of Listing \ref{string_in}. This is a set of other type constructors that the type constructor defining the handle set may potentially support binary operations with. When a type constructor is defined, the language checks that if \lstinlinep{tycon2} appears in \lstinlinep{tycon1}'s handler set, then \lstinlinep{tycon1} does \emph{not} appear in \lstinlinep{tycon2}'s handler set. This is a very simple modular analysis (rather than one that can only be performed at ``link-time''), shown in Listing \ref{check_tycon}, that ensures that the type constructor delegated control over each binary expression is deterministic and  unambiguous without arbitrarily preferring one subexpression position over another. This check is performed  by using a metaclass hook (technically, this can be disabled; we assume that clients are not importing potentially adversarial extensions in this work, though lifting this assumption raises quite interesting questions that we hope will be addressed by future work).

\subsection{Active Translation}
Once typechecking is complete, the compiler enters the translation phase. %The compiler creates an instance of a class inheriting from \lstinlinep{ace.Target} for use during this phase via the \lstinlinep{init_target} method (line \ref{pybase}.54). This object provides methods for code generation and supports features like fresh variable generation, adding imports and so on. It's interface is not constrained by @\texttt{lang} (we will see what @\texttt{lang} requires of a target below) and the mechanics of code generation are orthogonal to the  focus of this paper, so we will discuss it relatively abstractly. The simplest API would be string-based code generation. For the Python target we use here, we generate ASTs. The API is based directly on the \lstinlinep{ast} library, with a few additional conveniences just mentioned.
This phase follows the same delegation protocol as the typechecking phase. Each \lstinlinep{check_}/\lstinlinep{syn_}/\lstinlinep{ana_TermCon} method has a corresponding \lstinlinep{trans_TermCon} method. These are all now instance methods, because all indices have been fully determined.

Examples of translation methods for the \lstinlinep{fn} and \lstinlinep{record} type constructors are shown in Listing \ref{trans}. The output of translation on our example is discussed in the next subsection. Translation methods have access to the context and node, as during typechecking, and must return a translation, which for our purposes, is simply another Python syntax tree (in practice, we offer some additional conveniences beyond \lstinlinep{ast}, such as fresh variable generation and lifting of imports and statements inside expressions to appropriate positions, in the module \lstinlinep{astx} distributed with the language). Translation methods can assume that the term is well-typed and the typechecking phase saves the type that was delegated control, along with the type that was assigned, as attributes of each term processed by the compiler. Note that not all terms need to have been processed by the compiler if they were otherwise reinterpreted by a type constructor given control over a parent term (e.g. the field names in a record literal are never treated as expressions, while they would be for a dictionary literal). %This protocol can be seen on lines \ref{context}.39-\ref{context}.47. 
\begin{codelisting}[t]
\begin{lstlisting}
#class fn(atlang.Type): (cont'd)
  def trans_FunctionDef(self, ctx, node):
    x_body = [ctx.trans(stmt) for stmt in node.body]
    x_fun = astx.copy_with(node, body=x_body)
    if node.name == "__main__": 
      x_invoker = ast.parse(
        'if __name__ == "__main__": __main__()')
      return ast.Suite([x_fun, x_invoker])
    else: return x_fun

  def trans_Assign_Name(self, ctx, stmt):
    return astx.copy_with(stmt, 
      target=ctx.trans(stmt.target), 
      value=ctx.trans(stmt.value))

  def trans_Name(self, ctx, e):
    if e.id in ctx.assn_ctx: return astx.copy(e)
    else: return self._trans_lift(
      ctx.static_env[e.id])

#class record(atlang.Type): (cont'd)
  def trans_Dict(self, ctx, e): 
    if len(self.idx.fields) == 1: 
      return ctx.trans(e.values[0])
    ast_dict = dict(zip(e.keys, e.values))
    return target.Tuple(ctx.trans(target, ast_dict[f])
      for f, ty in self.idx)

  def trans_Attribute(self, ctx, e):
    if len(self.idx.fields) == 1: return ctx.trans(e)
    else: return ast.Subscript(ctx.trans(e.value), 
      ast.Index(ast.Num(self.idx.position_of(e.attr))))
\end{lstlisting}
\caption{Translation methods for the types defined above.}
\label{trans}
\end{codelisting}

%Because we are targeting Python directly, most of our \lstinlinep{trans} methods are direct translations (factored out into a helper function). The main methods of interest that are not entirely trivial are \lstinlinep{trans_FunctionDef_outer} on lines \ref{pybase}.33-\ref{pybase}-37 and the two translation methods in Listing \ref{record}, which implement the logic described in Sec. \ref{compilation}. 
%
%Each type constructor must also specify a \lstinlinep{trans_type} method. This is important when targeting a typed language (so that the translations of type annotations can be generated, for example). As we will see in the next two sections, this is also critical to ensuring type safety. The language \emph{checks} not only that translations having a given type are well-typed, but that they have the type specified by this method (requiring that the target provide a \lstinlinep{is_of_type} method). Here, because we are simply targeting Python directly, the \lstinlinep{trans_type} method on line \ref{record}.49-\ref{record}.50 simply generates \lstinlinep{target.dyn}. 
%
%When this phase is complete, each node processed by the context will have a translation, available via the \lstinlinep{trans} attribute. In particular, each typed function has a translation. Note that some nodes are never processed by the context because they were reinterpreted by the delegate (e.g. the field names in a record literal), so they do not have translations (as expected, given our discussion above).
%
%To support external compilation, the target must have an \lstinlinep{emit} method that takes the compilation script's file name and a reference to a string generator (an instance of \lstinlinep{ace.util.CG}) and emits source code. The string generator we provide can track indentation levels (to support Python code generation and make generation for other languages more readable, for the purposes of debugging). It allows non-local string generation via the concept of user-defined \emph{locations}.  Each file that needs to be generated is a location and there can also be locations within a file (e.g. the imports vs. the top-level code), specified by a target the first time it finds that a necessary location is not defined. A generated entity (e.g. an import, class definition or function definition) can only be added once at a location. We saw this in Listing \ref{example-out} for the decimal-to-string conversion. The API will be discussed further in the appendix.

\subsection{Standalone Compilation}\label{compilation} 
\begin{codelisting}[t]
\begin{lstlisting}[style=Bash]
$ `@ listing1.py`
Hello, compile-time world!
Goodbye, compile-time world!
[@] _atout_listing1.py successfully generated.
$ `python _atout_listing1.py`
Transferring 5.50 to Annie Ace.
Transferring 15.00 to Annie Ace.
\end{lstlisting}
\caption{Compiling \texttt{listing\ref{example}.py} using the \texttt{@} script.}
\label{external-compilation}
\end{codelisting}
\begin{codelisting}[t]
\begin{lstlisting}
import atlib.runtime as _at_i0_

def log_transfer(t):
  _at_i0_.print("Transferring %s to %s." % 
    (_at_i0_.dec_to_str(t[0][0], 2), t[1][0]))

def __main__():
  account = ("Annie Ace", "00-00000001")
  log_transfer((((5, 50), None), account))
  log_transfer((((15, 0), "Rent payment"), account))
if __name__ == "__main__": __main__()
\end{lstlisting}
\caption{[\texttt{\_atout\_listing\ref{example}.py}] The file generated in Listing \ref{external-compilation}.}
\label{example-out}
\end{codelisting}

Listing \ref{external-compilation} shows how to invoke the \lstinlinep{@} compiler at the shell to typecheck and translate then execute \lstinlinep{listing1.py}. %Note that the {\texttt{print}} statements at the top  of the  compilation script were evaluated during compilation only. %These two steps can be combined by running \lstinlinep{ace listing1.py} (the intermediate file is not generated unless explicitly requested in this case).
The @\texttt{lang} compiler is itself a Python library and \lstinlinep{@} is a simple Python script that invokes it in two steps:
\begin{enumerate}
\item It evaluates the compilation script to completion.
\item For any top-level bindings in the environment that are @\texttt{lang} functions, it initiates typehecking and translation as described above. If no type errors are discovered, the ordered set of translations are collected (obeying order dependencies) and emitted. If a type error is discovered, an error is displayed.
\end{enumerate}

In our example, there are no type errors, so the file shown in Listing \ref{example-out}  is generated. This file is meant only to be executed, not edited or imported directly. The invariants necessary to ensure that execution does not ``go wrong'', assuming the extensions were implemented correctly, were checked statically and entities having no bearing on execution, like field names and types themselves, were erased. The dynamic semantics of the type constructors used in the program were implemented by translation to Python:% In particular, note: 
\begin{enumerate}
\item \lstinlinep{fn} recognized the function name \lstinlinep{__main__} as special, inserting the standard Python code to invoke it if the file is run directly.
\item Records translated to tuples (the field names were not needed).
\item Decimals translated to pairs of integers. String conversion is defined in a ``runtime'' package with a ``fresh'' name.
\item Terms of type \lstinlinep{string_in[r"..."]} translated to strings. Checks could here be performed  statically.% Note that there are situations (e.g. if a string is read in from the console) that would necessitate an initial run-time check, but no further checks in downstream functions.
\item Prototypic objects translated to pairs consisting of the fore and the prototype. Dispatch to the appropriate record based on the field name was static (line 5).
\end{enumerate}

\paragraph{Type Errors} Listing \ref{oops} shows an example of code containing several type errors. If analagous code were written in Python itself, these could only be found if the code was executed on the first day of the month (and not all of the issues would  immediately be caught as run-time exceptions). In @\texttt{lang}, these can be found without executing the code (i.e. true static typechecking).

\begin{codelisting}[t]
\begin{lstlisting}
from listing1 import fn, dyn, Account, Details, 
  log_transfer
import datetime
@fn[dyn, dyn]
def pay_oopsie(memo):
  if datetime.date.today().day == 1: # @lang to Python
^\label{error-start}^    account = {nome: "Oopsie Daisy", 
      account_num: "0-00000000"} [:Account] # (format)
    details = {amount: None, memo: memo} [:Details]
    details.amount = 10 # (immutable)
    log_transfer((account, details)) # (backwards) ^\label{error-end}^
print "Today is day " + str(datetime.date.today())
pay_oopsie("Hope this works..") # Python to @lang
print "All done."
\end{lstlisting}
\caption{[\texttt{listing\ref{oops}.py}] Lines \ref{error-start}-\ref{error-end} each have a type error.}
\label{oops}
\end{codelisting}
\begin{codelisting}[t]
\begin{lstlisting}[style=Bash]
$ `python listing^\color{blue}\ref{oops}^.py`
Today is day 2
Traceback (most recent call last):
  File "listing^\ref{oops}^", line 9, in <module>
atlang.TypeError: 
    File "listing^\ref{oops}^.py", line ^\ref{error-start}^, col 12, in <module>: 
      [record] Invalid field name: nome
\end{lstlisting}
\caption{Execution never proceeds into a function with a type error.}
\label{oops-exec}
\end{codelisting}

\subsection{Interactive Invocation} @\texttt{lang} functions over values of type \lstinlinep{dyn} can be invoked directly from Python. Typechecking and translation occurs upon first invocation (subsequent calls are cached; we assume that the static environment is immutable). We see this occurring when we execute the code in Listing \ref{oops} using the Python interpreter in Listing \ref{oops-exec}.

In future work, we plan to detail how to insert dynamic checks and wrapper objects, defined by active type constructors in a manner similar to how static checks are defined here, so that values that are not of type \lstinlinep{dyn} can be passed in and out of untyped Python code. Because these can be seen as implicit coercions to and from \lstinlinep{dyn}, and we do not aim to introduce this feature in this paper, we omit discussion of this topic. Explicit coercions can be implemented using the mechanisms described above. For example, \lstinlinep{string_in} provides an introductory form that checks a provided \lstinlinep{string}  or \lstinlinep{dyn} value dynamically, raising an exception in the case of failure: \lstinlinep{[raw_input()] : string_in[r"\\d+"]}.

%.; for example, the unexpected use of \lstinlinep{None}\footnote{The notation \texttt{+T} constructs a \texttt{None}-able option type; see appendix.}). Static type checking allows us to find these errors during compilation. Listing \ref{oops-compilation} shows the result of attempting to compile this code. The compilation script completes (so that functions can refer to each other mutually recursively), then the typed functions in the top-level environment are typechecked. The typechecker raises an exception statically at the first error, and \lstinlinep{acec} prints it to the console as shown.

%When encountering a compound term (e.g. \lstinlinep{t.amount} in Listing \ref{example}), the compiler defers control over type  and translation to the type of a designated subexpression (e.g. \lstinlinep{t}) according to a fixed fixed \emph{dispatch protocol}. Below are examples of the choices made in @\texttt{lang}.\todo{full in appendix?}
%\begin{itemize}
%\item Responsibility over {\bf attribute access} (\texttt{e.attr}), {\bf subscripting} (\texttt{e[e1]}) and \textbf{calls} (\lstinlinep{e(e1, ..., en)}) and {\bf unary operations} (e.g. \lstinlinep{-e}) is handed to the type recursively assigned to \texttt{e}.
%\item Responsibility over {\bf binary operations} (e.g. \lstinlinep{e1 + e2}) is first handed to the type assigned to the left operand. If it indicates that it cannot handle the operation, the type assigned to the right operand is handed responsibility. {Note that this operates like the corresponding rule in Python's \emph{dynamic} operator overloading mechanism; see Sec. \ref{related} for a discussion.}
%\item Responsibility over \textbf{constructor calls} (\lstinlinep{[t](e1, ..., en)}), where \lstinlinep{t} is a \emph{compile-time Python expression} evaluating to an active type, is handed to that type. If \lstinlinep{t} evaluates to a \emph{family} of types, like \lstinlinep{clx.Cplx}, the active type is first generated via a class method, as discussed below.
%%\item Responsibility over {\bf simple assignment statements}, is handed to the type of the variable on the left (which, as we will see below, is determined by the active base). If this type does not provide special assignment semantics, the base must handle it.%Destructuring assignment is also supported by a somewhat more complex protocol.\todo{clarify}
%\end{itemize}
%%
%%\todo{To be revised from here on down}The core of @\texttt{lang} consists of about 1500 lines of Python code implementing its primary concepts: generic functions, concrete functions, active types, active bases and active targets.  The latter three comprise @\texttt{lang}'s extension mechanism. Extensions provide semantics to, and govern the compilation of, @\texttt{lang} functions, rather than logic in @\texttt{lang}'s core. %Indeed, the name ``@\texttt{lang}'' might itself be an acronym: it is an ``active compilation environment''.

% As a result, execution is essentially as fast as it could be given the target language chosen (Python). We will target an even faster language, \lstinlinep{OpenCL}, in Sec. \ref{targets}.

%Because \lstinlinep{t} will be statically guaranteed to have type \lstinlinep{Transfer},

\section{@$\lambda$: Active Type Constructors, Minimally}\label{theory}
\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
\newcommand{\tri}{i}
\newcommand{\trt}{\cev\tau}
\renewcommand{\tnil}[1]{[]}
\newcommand{\ltxt}[1]{\ell_{\text{#1}}}
\newcommand{\abslbl}[1]{\FF{lbl}[{\texttt{#1}}]}
\begin{figure*}[t!]
\small
$
\begin{array}{ll}
\F{import} \Phi_\text{nat}, \Phi_\text{lprod} & \FF{import}[\Phi_\text{nat}, \Phi_\text{lprod}](\\
\F{static~let} \tvar{nat} = \fvar{nat}\texttt{\textbf{[}}\FF{nil}\texttt{\textbf{]}} & \FF{slet}[\FF{ty}[\fvar{nat}](\FF{nil})](\tvar{nat}.\\
\F{let} zero = \texttt{0} : \tvar{nat} & \FF{let}(\FF{asc}[\tvar{nat}](\FF{lit}[\FF{lbl}[\texttt{0}]](\cdot)); zero.\\
\F{let} one = zero.{\texttt{s}} & \FF{let}(\FF{targ}[\abslbl{s}](zero; \cdot); one.\\
\F{let} plus = (\lambda x{:}\tvar{nat}. \lambda y{:}\tvar{nat}. &  \FF{let}(\FF{asc}[\FF{incty}[\fvar{fn}](\tvar{nat})](\lambda(x.\FF{asc}[\FF{incty}[\fvar{fn}](\tvar{nat})](\lambda(y.\\
\quad x\cdot{\texttt{rec}}(y; \lambda p.\lambda r. r\cdot{\texttt{s}})) & \quad \FF{targ}[\abslbl{rec}](x; \lambda(p.\lambda(r.\FF{targ}[\abslbl{s}](r; \cdot)))))))); plus.\\
\F{let} two = plus~one~one & \FF{let}(\FF{targ}[\FF{nil}](\FF{targ}[\FF{nil}](plus; one); one); two.\\
\{{\texttt{one}} {=} one; {\texttt{two}} {=} two\} : \fvar{lprod}\texttt{[}\cdots\texttt{]}~~~~~~& \FF{asc}[\FF{incty}[\fvar{lprod}](\FF{nil})](\FF{lit}[\FF{cons}(\abslbl{one}; \FF{cons}(\abslbl{two}; \FF{nil}))](one; two)))))))\\
% & \quad \\
%\quad\quad \\
%\quad\quad\quad & \quad\quad\quad \\
%\{{\texttt{one}} {=} one; {\texttt{two}} {=} plus(one)(one)\} : \fvar{lprod} ~~~~& \FF{asc}[\FF{incty}[\fvar{lprod}](\FF{nil})](\\&
%\quad 
%one; \\&
%\quad \FF{targ}[\FF{nil}](\FF{targ}[\FF{nil}](plus; one); one))))))))
\end{array}
$~~
%$\begin{array}{rcl}
%\fvar{tycon}\texttt{[}\sigma\texttt{]} & := & \FF{ty}[\fvar{tycon}](\sigma)\\
%\fvar{tycon}\texttt{[}\cdots\sigma\cdots\texttt{]} & := & \FF{incty}[\fvar{tycon}](\sigma)\end{array}$
%~~~~~~~~$
%\begin{array}{rcl}
%\fvar{tycon} & := & \FF{incty}[\fvar{tycon}](\FF{nil})\\
%\fvar{tycon}\texttt{[}\cdots\texttt{]} & := & \FF{incty}[\fvar{tycon}](\FF{nil})\\
%\end{array}$\\
%~\\
%\hspace{-50px}$
%\tiny
%\begin{array}{rcl}
%\lambda x{:}\sigma.e & := & \FF{asc}[\FF{incty}[\fvar{fn}](\sigma)](\lambda(x.e))\\
%(e_1, ..., e_n) & := & \FF{lit}[\FF{nil}](e_1; \ldots; e_n)\\
%\{{\texttt{lbl}}_1=e_1, \ldots, {\texttt{lbl}}_n=e_n\} & := & \FF{lit}[\FF{cons}(\FF{lbl}[{\texttt{lbl}}_1]; \ldots\FF{cons}(\FF{lbl}[{\texttt{lbl}}_n]; \FF{nil}))](e_1; \ldots; e_n)\\
%\text{(any numeric label)}~~~~{\texttt{n}} & := & \FF{lit}[\FF{lbl}[\texttt{n}]]()\\
%e(e_1) & := & \FF{targ}[\FF{nil}](e; e_1)\\
%e\cdot{\texttt{lbl}} & := & \FF{targ}[\FF{lbl}[{\texttt{lbl}}]](e; \cdot)\\
%e\cdot{\texttt{lbl}}(e_1, ..., e_n) & := & \FF{targ}[\FF{lbl}[{\texttt{lbl}}]](e; e_1; \ldots; e_n)
%%%(e_1, ..., e_n) & := & intro[()](e_1; ...; e_n)
%%%n &  := & intro[n]()
%%%s & := & intro[s]()
%%%
%%%e.l & := & elim[l](e)
%%%e[e_1; ...; e_n] & := & elim[()](e_1, ..., e_n)
%%%e.l(e_1; ...; e_n) & := & elim[l]
%\end{array}
%$
%
\caption{A program written using conventional concrete syntax, left, syntactically desugared to the abstract syntax on the right.}
\label{atlam-example}
\end{figure*}
\begin{figure}[t!]
\newcommand{\eqval}[1]{#1}
\newcommand{\noneqval}[1]{#1}
\small
\[
\begin{array}{rlcl}
\text{programs} & \rho & ::= & \FF{import}[\Phi](e)\\
\text{fragments} & \Phi & ::= & \emptyset \pipe \Phi, \fvar{tycon}=\{ \delta \}\\
\text{tycon defs} & \delta & ::= & \FF{analit}=\sigma; ~\FF{synidxlit}=\sigma; \\
& & & \FF{anatarg}=\sigma;~\FF{syntarg}=\sigma\\
\text{expressions} & e & ::= & x \pipe \FF{let}(e; x.e) \pipe   \FF{slet}[\sigma](\tvar{x}.e) \pipe \FF{asc}[\sigma](e) \\
& & \pipe & \lambda(x.e) \pipe \FF{lit}[\sigma](\bar{e}) \pipe \FF{targ}[\sigma](e; \overline{e}) \\
\text{static terms} & \sigma & ::= & \tvar{x}  \pipe \noneqval{\lambda(\tvar{x}.\sigma)} \pipe\FF{ap}(\sigma; \sigma) \pipe \FF{fail}\\
 & & \pipe & \eqval{\FF{ty}[\fvar{tycon}](\sigma)} \pipe \FF{tycase}[\fvar{tycon}](\sigma; \tvar{x}.\sigma; \sigma) \\
& & \pipe & \eqval{\FF{incty}[\fvar{tycon}](\sigma)}  \\
& & \pipe & \eqval{\FF{lbl}[{\texttt{lbl}}]} \pipe \FF{lbleq}(\sigma; \sigma; \sigma; \sigma) \\
& & \pipe & \eqval{\FF{nil}} \pipe \eqval{\FF{cons}(\sigma; \sigma)} \pipe  \FF{listrec}(\sigma; \sigma; \tvar{x}.\tvar{y}.\sigma) \\
 & & \pipe & \noneqval{\FF{arg}[e]} \pipe \FF{ana}(\sigma; \sigma; \tvar{x}.\sigma) \pipe \FF{syn}(\sigma; \tvar{x}.\tvar{y}.\sigma)\\
 & & \pipe &  \noneqval{\titerm{\iota}}\\
\text{internal terms} & \iota & ::= & \iup{\sigma} \pipe x \pipe \lambda(x.\iota) \pipe \FF{iap}(\iota; \iota) \\
& & \pipe  &  \FF{inil} \pipe \FF{icons}(\iota; \iota) \pipe \FF{ilistrec}(\iota; \iota; x.y.\iota)

\end{array}
\]
%\vspace{-15px}
\caption{Abstract syntax of @$\lambda$. Metavariable $\fvar{tycon}$ ranges over type constructor names (assumed globally unique), ${\texttt{lbl}}$ over static labels, $x, y$ over expression variables and $\tvar{x}, \tvar{y}$ over static variables. We indicate that variables or static variables are bound within a term or static term by separating them with a dot, e.g. $\tvar{x}.\tvar{y}.e$, and abbreviate a sequence of zero or more expressions as $\overline{e}$. %Static values are shown annotated with either $\color{gray}=$ or $\color{gray}\neq$. Values having and containing only sub-terms with $\color{gray}=$ annotations are those for which syntactic and semantic equality coincide, and only such values are valid type indices, captured in the judgement $\sigma~\mathtt{eqval}_\Phi$.   % Highlighted terms are exclusively used by fragment providers and will be described further below; $\FF{arg}[e]$ is introduced only by the semantics and is a value, and the IL quotation form $\titerm{\iota}$ is a value if it does not contain any unquote forms, $\iup{\sigma}$.
}
\label{syntax-clients}
\vspace{5px}
\end{figure}
%\begin{figure}[t]
%\small
%$\begin{array}{rcl}
%\fvar{tycon}\texttt{[}\sigma\texttt{]} & := & \FF{ty}[\fvar{tycon}](\sigma)\\
%\fvar{tycon}\texttt{[}\cdots\sigma\cdots\texttt{]} & := & \FF{incty}[\fvar{tycon}](\sigma)\end{array}$
%~~~~~~~~$
%\begin{array}{rcl}
%\fvar{tycon} & := & \FF{incty}[\fvar{tycon}](\FF{nil})\\
%\fvar{tycon}\texttt{[}\cdots\texttt{]} & := & \FF{incty}[\fvar{tycon}](\FF{nil})\\
%\end{array}$\\
%~\\
%$
%\begin{array}{rcl}
%\lambda x{:}\sigma.e & := & \FF{asc}[\FF{incty}[\fvar{fn}](\sigma)](\lambda(x.e))\\
%(e_1, ..., e_n) & := & \FF{lit}[\FF{nil}](e_1; \ldots; e_n)\\
%\{{\texttt{lbl}}_1=e_1, \ldots, {\texttt{lbl}}_n=e_n\} & := & \FF{lit}[\FF{cons}(\FF{lbl}[{\texttt{lbl}}_1]; \ldots\FF{cons}(\FF{lbl}[{\texttt{lbl}}_n]; \FF{nil}))](e_1; \ldots; e_n)\\
%\text{(any numeric label)}~~~~{\texttt{n}} & := & \FF{lit}[\FF{lbl}[\texttt{n}]]()\\
%e(e_1) & := & \FF{targ}[\FF{nil}](e; e_1)\\
%e\cdot{\texttt{lbl}} & := & \FF{targ}[\FF{lbl}[{\texttt{lbl}}]](e; \cdot)\\
%e\cdot{\texttt{lbl}}(e_1, ..., e_n) & := & \FF{targ}[\FF{lbl}[{\texttt{lbl}}]](e; e_1; \ldots; e_n)
%%%(e_1, ..., e_n) & := & intro[()](e_1; ...; e_n)
%%%n &  := & intro[n]()
%%%s & := & intro[s]()
%%%
%%%e.l & := & elim[l](e)
%%%e[e_1; ...; e_n] & := & elim[()](e_1, ..., e_n)
%%%e.l(e_1; ...; e_n) & := & elim[l]
%\end{array}
%$
%\vspace{-5px}
%\caption{\small Selected desugarings from a stylized conventional concrete syntax to abstract syntax.}
%\label{desugaring}
%\vspace{10px}
%\end{figure}

%

We now turn our attention to a type theoretic formulation of the key mechanisms described above atop a minimal abstract syntax, shown in Fig. \ref{syntax-clients}. This syntax supports a \emph{purely syntactic desugaring} from a conventional concrete syntax, shown by example in Fig. \ref{atlam-example}.

\paragraph{Fragment Client Perspective} A program, $\rho$, consists of a series of fragment imports, $\Phi$, defining active type constructors for use in an expression, $e$. As described in the introduction, expressions are indexed by static terms, $\sigma$, and indices that can only be trivial are omitted. The abstract syntax of $e$ provides $\FF{let}$ binding of variables and for convenience, static values can also be bound to a static variable, $\tvar{x}$ (distinguished in bold for clarity), using $\FF{slet}$. Static terms have a \emph{static dynamics} and evaluate to \emph{static values} or fail.

\subparagraph{Types and Ascriptions} An expression can be ascribed a {type} or an {incomplete type}, both  {static values} constructed, as in the introduction, by naming an imported type constructor, $\fvar{tycon}$, and providing a type index, another static value. The static language also includes lists and atomic \emph{labels} for use in compound indices. %The remaining static term constructors are of use mainly to fragment providers (see below).

\subparagraph{Literal Desugaring} All concrete literals (other than lambda expressions, which are built in)  desugar to an abstract term of the form $\FF{lit}[\sigma](\overline{e})$, where $\sigma$ captures all static portions of the literal (e.g. a list of the labels used as field names in the labeled product literal in Fig. \ref{atlam-example}, or the numeric label used for the natural number zero) and $\overline{e}$ is a list of sub-expressions (e.g. the field values). % is only one term constructor corresponding to each basic method by which a type constructor is delegated responsibility over a term.

\subparagraph{Targeted Expression Desugaring} All non-introductory operations go through a targeted expression form (e.g. $e(e)$, or $e\cdot\texttt{lbl}$, or $e\cdot\texttt{lbl}(\overline{e})$; see previous section), which all desugar to an abstract term of the form $\FF{targ}[\sigma](e; \overline{e})$ where $\sigma$ again captures all static portions of the term (e.g. the label naming an operation, e.g. $\texttt{s}$ or $\texttt{rec}$ on natural numbers \cite{pfpl}), $e$ is the  target (e.g. the natural number being operated on, the function being called, or the record we wish to project out of) and $\overline{e}$ are all other arguments (e.g. the base and recursive cases of the recursor, or the argument being applied).% We conjecture that formalizing these semantics would not present a substantial challenge.

\paragraph{Bidirectional Active Typechecking and Translation}
\newcommand{\atjsynX}[3]{\Gamma \vdash_\fvalCtx #1 \Rightarrow #2 \leadsto #3}
\newcommand{\atjanaX}[3]{\Gamma \vdash_\fvalCtx #1 \Leftarrow #2 \leadsto #3}
\newcommand{\atjerrX}[1]{\Gamma \vdash_\fvalCtx #1~ \mathtt{error}}
\begin{figure*}[t!]
\small
$\fbox{\inferrule{}{\atjsynX{e}{\sigma}{\iota}}}$~~
$\fbox{\inferrule{}{\atjanaX{e}{\sigma}{\iota}}}$~~
$\Gamma ::= \emptyset \pipe \Gamma, x \Rightarrow \sigma$
%$\fbox{\inferrule{}{\atjerrX{e}}}$
\renewcommand{\MathparLineskip}{\lineskiplimit=.6\baselineskip\lineskip=.6\baselineskip plus .2\baselineskip}
\begin{mathpar}
\inferrule[att-subsume]{
  \atjsynX{e}{\sigma}{\iota}
}{
  \atjanaX{e}{\sigma}{\iota}
}

\inferrule[att-var]{
  x \Rightarrow \sigma \in \Gamma
}{
  \atjsynX{x}{\sigma}{x}
}

\inferrule[att-ana-let]{
  \atjsynX{e_1}{\sigma_1}{\iota_1}\\\\
  \Gamma, x \Rightarrow \sigma_1 \vdash_\fvalCtx e_2 \Leftarrow \sigma_2 \leadsto \iota_2
}{
  \atjanaX{\FF{let}(e_1; x.e_2)}{\sigma_2}{\FF{iap}(\lambda(x.\iota_2); \iota_1)}
}

\inferrule[att-syn-let]{
  \atjsynX{e_1}{\sigma_1}{\iota_1}\\\\
  \Gamma, x \Rightarrow \sigma_1 \vdash_\fvalCtx e_2 \Rightarrow \sigma_2 \leadsto \iota_2
}{
  \atjsynX{\FF{let}(e_1; x.e_2)}{\sigma_2}{\FF{iap}(\lambda(x.\iota_2); \iota_1)}
}

\inferrule[att-syn-asc]{
    \sigma \Downarrow_{\emptyset; \emptyset} \sigma'\\
     \sigma'~\mathtt{ty}_\fvalCtx\\
  \atjanaX{e}{\sigma'}{\iota}
}{
  \atjsynX{\FF{asc}[\sigma](e)}{\sigma'}{\iota}
}

\inferrule[att-ana-slet]{
   \sigma_1 \Downarrow_{\emptyset; \emptyset} \sigma_1'\\
    \atjanaX{[\sigma_1'/\tvar{x}]e}{\sigma_2}{\iota}
}{
  \atjanaX{\FF{slet}[\sigma_1](\tvar{x}.e)}{\sigma_2}{\iota}
}

\inferrule[att-syn-slet]{
   \sigma_1 \Downarrow_{\emptyset; \emptyset} \sigma_1'\\
    \atjsynX{[\sigma_1'/\tvar{x}]e}{\sigma_2}{\iota}
}{
  \atjsynX{\FF{slet}[\sigma_1](\tvar{x}.e)}{\sigma_2}{\iota}
}

\inferrule[att-ana-lam]{
    \sigma_1~\mathtt{ty}_\fvalCtx\\\sigma_2~\mathtt{ty}_\fvalCtx\\
  \Gamma, x \Rightarrow \sigma_1 \vdash_\fvalCtx e \Leftarrow \sigma_2 \leadsto \iota
}{
  \atjanaX{\lambda x.e}{\FF{ty}[\fvar{fn}](\FF{cons}(\sigma_1; \FF{cons}(\sigma_2; \FF{nil})))}{\lambda(x.\iota)}
}

\inferrule[att-syn-idx-lam]{
    \sigma_1~\mathtt{ty}_\fvalCtx\\
  \Gamma, x \Rightarrow \sigma_1 \vdash_\fvalCtx e \Rightarrow \sigma_2 \leadsto \iota
}{
  \atjsynX{\FF{asc}[\FF{incty}[\fvar{fn}](\sigma_1)](\lambda(x.e))}{\FF{ty}[\fvar{fn}](\FF{cons}(\sigma_1; \FF{cons}(\sigma_2; \FF{nil})))}{\lambda(x.\iota)}
}
%
%\inferrule[att-lam]{
% \tau_1 \Downarrow_{\Gamma;\fvalCtx} \tau_1'\\
%   \trepof{\tau_1'} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\\\
% \Gamma, x \Rightarrow \tau_1' \vdash_\fvalCtx e \Rightarrow \tau_2 \leadsto \iota
%%  \iota \hookrightarrow_\fvalCtx \iota'\\
%%    \sigma \hookrightarrow_\fvalCtx \sigma'
%%  \ddbar{\fvar{Arrow}}{\fvalCtx}{\trepof{\tau_1'}}{\sbar_1}\\
% %\delfromtau{$\Xi_0$}{\fvalCtx}{\tau_1'}{\sabs}\\\\
%}{
% \atjsynX{\elam{x}{\tau_1}{e}}{\ttype{arrow}{(\tau_1', \tau_2)}}{\ilam{x}{\sigma'}{\iota}}
%}

\inferrule[att-ana-lit]{
  \vdash_\fvalCtx \FF{analit}(\fvar{tycon})=\sigmat{def}\\
  \mathtt{args}(\overline{e})=\sigmat{args}\\\\
  \sigmat{def} ~\sigmat{tyidx} ~\sigmat{tmidx} ~\sigmat{args} \Downarrow_{\Gamma; \fvalCtx} \titerm{\iota}
% \FF{ap}(\FF{ap}(\FF{ap}(\sigmat{def}; \sigmat{tyidx}); \sigmat{tmidx}); \sigmat{args}) \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota}
% \trepof{\ttype{tycon}{\tauidx'}} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\
}{
  \atjanaX{\FF{lit}[\sigmat{tmidx}](\overline{e})}{\FF{ty}[\fvar{tycon}](\sigmat{tyidx})}{\iota}
}

\inferrule[att-syn-idx-lit]{
  \vdash_\fvalCtx \FF{synidxlit}(\fvar{tycon})=\sigmat{def}\\
  \mathtt{args}(\overline{e})=\sigmat{args}\\\\
  \sigmat{def} ~\sigmat{incidx} ~\sigmat{tmidx} ~\sigmat{args} \Downarrow_{\Gamma;\fvalCtx} \FF{cons}(\sigmat{tyidx}; \FF{cons}(\titerm{\iota}; \FF{nil}))\\
% \FF{ty}[\fvar{tycon}](\sigmat{tyidx})~\mathtt{ty}_\fvalCtx
% \FF{ap}(\FF{ap}(\FF{ap}(\sigmat{def}; \sigmat{incidx}); \sigmat{tmidx}); \sigmat{args}) \Downarrow_{\Gamma;\fvalCtx}\FF{cons}(\sigmat{tyidx}; \FF{cons}(\titerm{\iota}; \FF{nil}))
}{
  \atjsynX{\FF{asc}[\FF{incty}[\fvar{tycon}](\sigmat{incidx})](\FF{lit}[\sigmat{tmidx}](\overline{e}))}{\FF{ty}[\fvar{tycon}](\sigmat{tyidx})}{\iota}
}
%
%\inferrule[att-i-asc-ty]{
% \atjanaX{I}{\tau}{\iota}
%}{
% \atjsynX{I : \tau}{\tau}{\iota}
%}
%
%\inferrule[att-i-asc-tycon]{
% \vdash_\fvalCtx \FF{isyn}(\fvar{tycon})=\taudef\\\\
% \taudef~\tauidx~((\Gamma; e_1)? {::}{\ldots}{::}(\Gamma; e_n)? {::} []) \Downarrow_{\Gamma;\fvalCtx} \tden{\titerm{\iota}}{\ttype{tycon}{\tauidx'}}\\
%     \trepof{\tau_1'} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\
% \Gamma \vdash_\fvalCtx \iota : \sigma
%}{
% \atjsynX{\FF{intro}[\tauidx](e_1; \ldots; e_n)] :: \fvar{tycon}}{\ttype{tycon}{\tauidx'}}{\iota'}
%}

\inferrule[att-ana-targ]{
  \atjsynX{e_\text{targ}}{\FF{ty}[\fvar{tycon}](\sigmat{tyidx})}{\iota_\text{targ}}\\\\
  \vdash_\fvalCtx \FF{anatarg}(\fvar{tycon})=\sigmat{def}\\
  \mathtt{args}(\overline{e})=\sigmat{args}\\\\
  \sigmat{def} ~\sigmat{tyidx} ~\titerm{\iota_\text{targ}} ~\sigmat{ty} ~\sigmat{tmidx} ~\sigmat{args} \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota}
% \FF{ap}(\FF{ap}(\FF{ap}(\FF{ap}(\FF{ap}(\sigmat{def}; \sigmat{tyidx}); \titerm{\iota}); \sigma); \sigmat{tmidx}); \sigmat{args}) \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota'}
}{
  \atjanaX{\FF{targ}[\sigmat{tmidx}](e_\text{targ}; \overline{e})}{\sigmat{ty}}{\iota}
}

\inferrule[att-syn-targ]{
  \atjsynX{e_\text{targ}}{\FF{ty}[\fvar{tycon}](\sigmat{tyidx})}{\iota_\text{targ}}\\\\
  \vdash_\fvalCtx \FF{syntarg}(\fvar{tycon})=\sigmat{def}\\
  \mathtt{args}(\overline{e})=\sigmat{args}\\\\
  \sigmat{def} ~\sigmat{tyidx} ~\titerm{\iota_\text{targ}} ~\sigmat{tmidx} ~\sigmat{args} \Downarrow_{\Gamma;\fvalCtx} \FF{cons}(\sigmat{ty}; \FF{cons}(\titerm{\iota}; \FF{nil}))\\
  \sigmat{ty}~\mathtt{ty}_\fvalCtx
}{
  \atjsynX{\FF{targ}[\sigmat{tmidx}](e_\text{targ}; \overline{e})}{\sigmat{ty}}{\iota'}
}
\end{mathpar}
\caption{Bidirectional active typechecking and translation. For concision, we use standard functional notation for static function application.}
\label{atj}
\end{figure*}
%In this section, we will describe the foundations of \emph{active typechecking and translation} by constructing a core lambda calculus, @$\lambda$. 
%A \emph{program} in @$\lambda$, $\rho$, consists of a series of fragment definitions, $\phi$, for use by an external term, $e$. An example of an @$\lambda$ program, $\rho_{\text{example}}$, that uses {fragment definitions} $\phi_{\texttt{nat}}$ and $\phi_{\texttt{lprod}}$, defining primitive natural numbers and labeled products (i.e. ordered records), is shown in Fig. \ref{nat-sugared}. In @\texttt{lang},  fragment definitions are packaged separately; we do not include this in the core calculus for simplicity.
%
%In this example, we see the introduction and elimination forms for natural numbers, functions and labeled tuples all being used. However, in the core syntax, shown (from the perspective of fragment clients) in Fig. \ref{client-syntax}, there are only two external introductory forms, $\lambda x.e$ and $\FF{intro}[\sigma](\overline{e})$, and one external elimination form, $e\cdot\FF{elim}[\sigma](\overline{e})$, where $\overline{e}$ is shorthand for zero or more semicolon-separated external terms, called the \emph{arguments} of the introduction or elimination operation. The forms used in Fig. \ref{nat-sugared} are \emph{derived forms}, defined in terms of these core forms in Fig. \ref{derived}. 
%
%Both fragments and external terms can contain \emph{static terms}, $\sigma$, which are evaluated statically to \emph{static values}, $\hat\sigma$. The static language is itself a simply-typed lambda calculus. We call the types of static terms \emph{kinds}, $\kappa$, by convention. Static variables are written in bold font, $\tvar{x}$, to emphasize that they are distinct from variables bound by external terms, written $x$. For example, in the example in Figure \ref{nat-sugared}, $\tvar{nat}$ is a static variable bound by the fragment $\phi_{\texttt{nat}}$ to the natural number type. Types have kind $\kTypeBlur$ and are introduced by naming a \emph{type constructor}, written in small-caps, e.g. $\fvar{nat}$ and $\fvar{lprod}$, and providing an \emph{index}, which is a static term of the \emph{index kind} of the type constructor.  For example, $\fvar{nat}$ is indexed by $\kunit$ (because there is only one natural number type), so $\tvar{nat}$ is defined to be $\ttype{nat}{\tunit}$, and $\fvar{lprod}$ is indexed by $\klist{\kpair{\FF{L}}{\kTypeBlur}}$, classifying static lists pairing \emph{labels} with types. Labels are static values of kind $\FF{L}$, written abstractly using the metavariables containing $\ell$ in our examples. 
%
%Before typechecking the external term in a program, the program is kind checked and its static terms are normalized to static values.  We write static values using the hatted metavariable $\hat \sigma$. Other syntactic metavariables in our system also have hatted forms if they can contain static terms. The desugared, statically normalized version of the example in Figure \ref{nat-sugared}, $\hat\rho_{\text{example}}$, is shown in Figure \ref{nat-normal}. %The syntax of @$\lambda$ from the perspective of a fragment client is given in Fig. \ref{atlam-syntax}. We will see a few additions relevant only to fragment providers when we discuss defining fragments below.
%
%\subsection{External Terms}
% a program consists of first \emph{kind checking} its static terms, then normalizing its static terms, then typechecking the external term and,  simultaneously, \emph{translating} it to a term, $\iota$, in the \emph{typed internal language}. The dynamic behavior of an external term is determined entirely by its translation to this language. Internal types are written $\tau$. The internal language is only exposed to fragment providers, not to clients (i.e. normal developers).

%We will now give a core typed lambda calculus, @$\lambda$, that captures the semantics described in the previous sections. It is intended to make precise how active typechecking and translation works and how our mechanism relates to existing work on bidirectional typechecking, type-level computation and typed compilation while abstracting away from the details of Python's syntax and imposing a stronger type-level semantics that will allow us to state metatheoretic properties of interest. We will assume a fixed base for functions (providing the standard semantics of lambda functions) and target language, which we here call the \emph{internal language}.
 % program written in core @$\lambda$, without syntactic sugar and with all static terms normalized, is shown in Fig. \ref{nat-desugared}. The syntax of core @$\lambda$ is shown in Fig. \ref{grammar} and the syntactic desugarings are shown in Fig. \ref{desugaring}. The definition of $\phi_{\texttt{nat}}$ is shown in Fig. \ref{nat-atfrag}. We will discuss these in the following sections.

%In this section, we will develop an ``actively typed'' version of the simply-typed lambda calculus with simply-kinded type-level computation called $\lam@\texttt{lang}$. More specifically, the level of types, $\tau$, will itself form a simply-typed lambda calculus. \emph{Kinds} classify type-level terms in the same way that types conventionally classify expressions. Types become just one kind  of type-level value (which we will write $\kTypeBlur$, though it is also variously written $\star$, \lstinlinep{T} and \lstinlinep{Type} in various settings). Rather than there being a fixed set of type and operator constructors, we allow the programmer to declare new  constructors, and give their static and dynamic semantics by writing type-level functions. The kind system combined with techniques borrowed from the typed compilation literature and a form of type abstraction will allow us to prove strong type safety, decidability and conservativity theorems.
%
%The syntax of Core $\lam@\texttt{lang}$ is given in Fig. \ref{grammar}. An example of a program defining type and operator constructors that can be used to construct an active embedding of G\"odel's \textbf{T} into $\lam@\texttt{lang}$ is given in Fig. \ref{nat}. We will discuss its semantics and how precisely the embedding, seen being used starting on line 15 to ultimately compute the sum of two and two, works as we go on. Natural numbers can, of course, be isomorphically embedded in existing languages, with a similar usage and asymptotic performance profile (up to function call overhead as an abstract type, for example). We will provide more sophisticated examples where this is less feasible later on (and note that type abstraction is an orthogonal mechanism).
%a type constructor declaration, $\fvar{nat}$, indexed trivially, together with three operator constructors, also all indexed trivially, that implement the standard introductory forms for natural numbers as well as the recursor operator (as in G\"odel's T \cite{pfpl}). Following the type constructor declaration, we apply $\fvar{nat}$ with the trivial index, $\tunit$, to form the type $\tvar{nat}$. Finally, we write an external term that uses the operators associated with $\fvar{nat}$ and the built-in constructor $\fvar{Parr}$, governing partial functions, to define an addition function and compute the addition of the natural numbers  two and two. We will introduce a more convenient concrete syntax in later portions of this thesis; for now we will restrict ourselves to the abstract syntax so that this example can directly aid in understanding the semantics.

%\begin{figure}[t]
%\small
%$$\begin{array}{rrcl}  
%%\textbf{programs} & 
%\text{programs} & \rho & ::= & \F{using}\phi~\F{in}e \\
%\text{fragments} & \phi &  ::= & \F{tycon}\fvar{tycon}~\F{of}\kappat{tyidx}~{\{}\F{iana} \tau_1; \F{esyn} \tau_2; \F{rep} \tau_3{\}} %\pipe \phi; \phi %\\
%%\\
%%& \pipe & 
%\pipe \F{def}\tvar{t} : \kappa = \tau \pipe \phi; \phi%\pdef{t}{\kappa}{\tau}{\progsort} \pipe 
%\\ %\pfam{\familyDf}{\progsort}
%%\\&  \pipe & \pdef{t}{\kappa}{\tau}{\progsort}  \pipe e
%%\pipe \pdef{t}{\kappa}{\tau}{\progsort} 
%\text{external terms}& e & ::= & \evar{x} \pipe \F{let} x = e_1~\F{in} e_2 \pipe \lambda x.e \pipe e : \tau\\
%&& \pipe & \FF{intro}[\taut{opidx}](\splat{e}{1}{n})\pipe e\cdot\FF{elim}[\taut{opidx}](\splat{e}{1}{n}) \\
%
%\text{type-level terms} & \tau   & ::=   &   \tvar{t} \pipe \tifeq{\tau_{1}}{\tau_{2}}{\kappa}{\tau_{3}}{\tau_{4}} \pipe \terr  \\
% && \pipe & 
% %\F{let}\tvar{t}{:}\kappa=\tau_1~\F{in} \tau_2 \pipe 
%                           \tlam{t}{\kappa}{\tau} \pipe 
%                           \tapp{\tau_1}{\tau_2} \\
%&&\pipe&                     
%                           \tnil{\kappa} \pipe \tcons{\tau_1}{\tau_2} \pipe 
%                                      \tfold{\tau_1}{\tau_2}{\tvar{t}_{hd}}{\tvar{t}_{tl}}{\tvar{t}_{rec}}{\tau_3}
%                           \\
%     && \pipe  &     \ell  \\
%     && \pipe &  
%   \tunit \pipe 
%                           \tpair{\tau_{1}}{\tau_{2}} \pipe 
%                           \tfst{\tau} \pipe 
%                           \tsnd{\tau} 
%                           \\  
%     && \pipe & \tinl{\kappa_1, \kappa_2}{\tau_1} \pipe \tinr{\kappa_1, \kappa_2}{\tau_2} \pipe \tsumcase{\tau}{t}{\tau_1}{t}{\tau_2}\\
%&&\pipe  &   \ttype{tycon}{\taut{tyidx}} \pipe  \tfamcase{\tau}{tycon}{t}{\tau_1}{\tau_2}\\
%%                            \\                          %       & & \pipe & \tfamcase{\tau}{Fam}{x}{\tau_1}{\tau_2}\\
%%                                                
%%\\     &  \pipe & \tden{\tau_2}{\tau_1} \pipe \terr %\pipe \tden{\ibar}{\tau}^{\checkmark} 
%%\pipe \ttypeof{\tau} \pipe \itransof{\tau} \\% \tdencase{\tau}{x}{t}{\tau_1}{\tau_2}\\
%% %& & \pipe & 
%%%                           \tdencase{\tau}{y}{x}{\tau_1}{\tau_2}
%%%                            \\
%%
% && \pipe & \titerm{\iota}
%     \pipe \titype{\sigma} \pipe \trepof{\tau} \\
%   && \pipe &  (\Gamma; e)? \pipe \FF{syn}(\tau_1; \tvar{t}_{ty},\tvar{t}_{trans}.\tau_2) \pipe \FF{ana}(\tau_1; \tau_2; \tvar{t}_{trans}.\tau_3)\\
%%\textbf{external terms}         & e & ::= & \evar{x} \pipe 
%%%                           \efix{x}{\tau}{e} \pipe 
%%                            \elam{\evar{x}}{\tau}{e} \pipe 
%%                            \eop{tycon}{op}{
%%                              \tauidx
%%                            }{
%%                                  \splat{e}{1}{n}
%%                            } \\
%%                  &     &   &   \\
%%
%%
%%                        
%%%\text{deabstracted}& \iota & ::= & \mathcal{G}[\iota, \sigma]\\
%%                        
%%\\
%%\\
%%              
%%% &  & \pipe & \tvalof{\tau_1}{\tau_2} \pipe \iup{\tau} \\
%%%                         \trepof{\tau} \pipe \dup{\tau}\\
%%\text{translational IL} & \bar{\iota} & ::= & x \pipe \ifix{x}{\bar \sigma}{\bar \iota} 
%%  %\pipe \ilam{x}{\bar \sigma}{\bar \iota} \pipe \iapp{\bar \iota_1}{\bar \iota_2} 
%%  \pipe \cdots \pipe \itransof{\tau} \\           
%% & \bar{\sigma} & ::= & \darrow{\bar \sigma_1}{\bar \sigma_2} \pipe \cdots \pipe \dup{\tau} \pipe \trepof{\tau} \\
%%%\text{abstracted} & \sabs & ::= & \darrow{\sabs_1}{\sabs_2} \pipe \cdots \pipe \sabsrep{\tau}
%%                      \\
%\text{kinds} & \kappa  & ::= & \karrow{\kappa_1}{\kappa_2} \pipe \klist{\kappa} \pipe \FF{L} \pipe \kunit \pipe \kpair{\kappa_1}{\kappa_2} \pipe \ksum{\kappa_1}{\kappa_2} \pipe \kTypeBlur \pipe \kITerm \pipe \kIType \pipe \Q
%%                          \klabel \pipe
%%                          \klist{\kappa} \pipe
%%                        \kunit \pipe 
%%                        \kpair{\kappa_{1}}{\kappa_{2}} \\
%%                        &&\pipe&
%%                        \ksum{\kappa_1}{\kappa_2} \pipe
%%                        \kTypeBlur \pipe \kDen \pipe 
%%                        \kIType               
%%%\textbf{ops signature}     & \Theta  & ::= & \kOpEmpty \pipe \kOp{\Theta}{op}{\kappai}\\
%%%                                   &   &   & \\
%\\
%\text{typing contexts} & \Gamma & ::= & \emptyset \pipe \Gamma, x \Rightarrow \tau\\
%\\
%\text{internal terms} & \iota  & ::= & \evar{x} \pipe 
%                       \ifix{\evar{x}}{\sigma}{\iota} \pipe
%                       \ilam{\evar{x}}{\sigma}{\iota} \pipe 
%                       \iapp{\iota_{1}}{\iota_{2}} \pipe n \pipe \iota_1 \pm \iota_2 \pipe \FF{if0}(\iota; \iota_1; \iota_2)\\
%   && \pipe &  \iunit \pipe \ipair{\iota_1}{\iota_2} \pipe \ifst{\iota} \pipe \isnd{\iota} \pipe \iup{\tau} 
%%\\&\pipe&               \cdots \pipe  
%%                        n \pipe \iop{\iota_{1}}{\iota_{2}} \pipe \iIfEq{\iota_{1}}{\iota_{2}}{\dint}{\iota_{3}}{\iota_{4}} \\
%%& \pipe & 
%%                        \iunit \pipe
%%                        \ipair{\iota_{1}}{\iota_{2}} \pipe 
%%                        \ifst{\iota} \pipe
%%                        \isnd{\iota} 
%%\\&\pipe&                       
%%                         \iinl{\sigma_2}{\iota_1} \pipe \iinr{\sigma_1}{\iota_2} \\&\pipe & \icase{e}{x}{e_1}{x}{e_2} \\
%\\ \text{internal types} & \sigma  & ::= &    \darrow{\sigma_1}{\sigma_2} \pipe \dint \pipe \dunit \pipe \dpair{\sigma_1}{\sigma_2} %\pipe \dint \pipe \dunit \pipe \dpair{\sigma_1}{\sigma_2} \pipe \dsum{\sigma_1}{\sigma_2}  
% \pipe \dup{\tau}
%\end{array}$$
%%\vspace{-10pt}
%\caption{\small Syntax of Core @$\lambda$. Here, $x$ ranges over external and internal language variables, $\tvar{t}$ ranges over type-level variables, $\fvar{tycon}$ ranges over type constructor names, $\ell$ ranges over labels, $n$ ranges over integers and $\pm$ ranges over standard binary operations over integers. The introductory form for arguments, $(\Gamma; e)?$, should only be constructed by the compiler, not by type constructor providers.
%\label{grammar}}
%\end{figure}
%\begin{figure*}[t]
%\small
%\begin{tabular}{r l l l}
% & \textbf{initial} & \textbf{statically normal}\\
%\textbf{programs} & $\rho ::=$ & $\hat\rho ::=$\\
% & ~~$\FF{using}~{\phi}~\FF{in}~e$ & ~~$\FF{using}~{\hat\phi}~\FF{in}~\hat e$\\
%\textbf{fragment decls.} & $\phi ::=$ & $\hat\phi ::=$\\
%tycon decls. & ~~$\FF{tycon}~\fvar{tycon} : \Theta = \theta$ & ~~$\FF{tycon}~\fvar{tycon} : \Theta = \hat\theta$\\
%static binding & ~~$\FF{def}~\tvar{x} : \kappa = \sigma$ & \\
% & ~~$\phi; \phi$ & ~~$\hat\phi; \hat\phi$\\
%\textbf{static terms} & $\sigma ::=$ & $\hat\sigma ::= $ &\textbf{kinds}~~~  $\kappa ::= $\\
%static values & ~~$\hat \sigma$\\
%static variables & ~~$\tvar{x}$ & ~ & ~\\
%types & ~~$\ttype{tycon}{\sigma}$ & ~~$\ttype{tycon}{\hat\sigma}$ & ~~$\kTypeBlur$\\
%%static products & ~~$\tunit$ & ~~$\tunit$ & ~~$\kunit$\\
%%& ~~$(\sigma, \sigma)$ & ~~$(\hat\sigma, \hat\sigma)$ & ~~$\kpair{\kappa}{\kappa}$\\
%%static sums & ~~$\FF{in}[\ell](\sigma)$ & ~~$\FF{in}[\ell](\hat\sigma)$ & ~~$\Sigma[\{\overline{\ell \hookrightarrow \kappa}\}]$\\
%static lists & ~~$[]_\kappa$ & ~~$[]_\kappa$ & ~~$\klist{\kappa}$\\
% & ~~$\sigma :: \sigma$ & ~~$\hat\sigma :: \hat\sigma$ & \\
%static labels & ~~$\ell$ & ~~$\ell$ & ~~$\FF{L}$\\
%\\
%\textbf{external terms} & $e ::= $ & $\hat e ::=$ \\
%variables & ~~$x$ & ~~$x$\\
%binding & ~~$\FF{let}~x=e~\FF{in}~e$ & ~~$\FF{let}~x=\hat{e}~\FF{in}~\hat{e}$\\
%fixpoints & ~~$\FF{fix}~x{:}\sigma~\FF{is}~e$ & ~~$\FF{fix}~x{:}\hat\sigma~\FF{is}~\hat e$\\
%ascription & ~~$e : \sigma$ & ~~$\hat{e} : \hat\sigma$\\
%%introductory forms & ~~$\nu$ & ~~$\hat\nu$\\
%%eliminatory forms & ~~$\eta$ & ~~$\hat\eta$\\
%% & $\nu ::= $ & $\hat \nu ::=$\\
%analytic lambda & ~~$\lambda x.e$ & ~~$\lambda x.\hat{e}$\\
%%synthetic lambda & ~~$\lambda x {:} \sigma.e$ & ~~$\lambda x{:}\hat\sigma.\hat e$\\
%analytic intro & ~~$\FF{intro}[\sigma](\overline{e})$ & ~~$\FF{intro}[\hat\sigma](\overline{\hat e})$\\
%%synthetic intro & ~~$\fvar{tycon}\cdot\FF{intro}[\sigma](\overline{e})$ & ~~$\fvar{tycon}\cdot\FF{intro}[\hat \sigma](\overline{\hat e})$\\
%% & $\eta ::= $ & $\hat\eta ::=$\\
%synthetic elim & ~~$e\cdot\FF{elim}[\sigma](\overline{e})$ & ~~$\hat e\cdot\FF{elim}[\hat \sigma](\overline{\hat e})$\\
%\end{tabular}
%\caption{Syntax from the perspective of fragment clients (i.e. normal developers)}
%\end{figure*}
%\begin{figure*}[t]
%\small
%\hspace{-5px}
%$
%\begin{array}{rclr}
%%\FF{intro}(e_1; \ldots; e_n) & := & \FF{intro}[()](e_1; \ldots; e_n)\\
%\sigma\langle e_1; \ldots; e_n \rangle & := & \FF{intro}[\tfst{\sigma}](e_1; \ldots; e_n) : \tsnd{\sigma} & \text{assisted intro}\\
%%\langle\sigmaidx\rangle(e_1; \ldots; e_n) & := & \FF{intro}[\sigmaidx](e_1; \ldots; e_n)\\
%%[\sigmaidx](e_1; \ldots; e_n) & := & \FF{intro}[\sigmaidx](e_1; \ldots; e_n)\\
%\{ \ell_1{=}~e_1, \ldots, \ell_n{=}~e_n \} & := & \FF{intro}[\ell_1 :: \ldots :: \ell_n :: []_{\FF{L}}](e_1; \ldots; e_n) & \text{labeled seq.}  \\
%(e_1, ..., e_n) & := & \FF{intro}[()](e_1; \ldots; e_n) & \text{unlabeled seq.}\\
%%n & := & \FF{intro}[n]() & \text{number literal} \\
%%s & := & \FF{intro}[s]() & \text{string literal}\\
%%e\cdot\FF{elim}(e_1; \ldots; e_n) & := & e\cdot\FF{elim}[()](e_1; \ldots; e_n)\\
%e~e_1 & := & e\cdot\FF{elim}[()](e_1) & \text{application}\\
%\sigma~e~\langle e_1; \ldots; e_n \rangle  & := & e\cdot\FF{elim}[\sigma](e_1; \ldots; e_n) & \text{assisted elim}\\
%e.\ell & := & e\cdot\FF{elim}[\ell]() & \text{projection}\\
%%\F{case}e~\{\ell_1\langle x_1 \rangle \Rightarrow e_1~|~\ldots~|~\ell_n\langle x_n \rangle \Rightarrow e_n\} & := & e\cdot\FF{elim}[\ell_1 :: \ldots :: \ell_n :: []_{\FF{L}}]( & \text{case analysis}\\
%% & & \quad \lambda x_1.e_1; \ldots; \lambda x_n.e_n)\\
%\sigma_1 \rightharpoonup \sigma_2 & := & \ttype{parr}{(\sigma_1, \sigma_2)} & \text{arrow types}
%%!TEX encoding = UTF-8 Unicode
%\begin{figure*}
%\newcommand{\tcl}[1]{\multicolumn{2}{l}{#1}}
%\small
%\begin{tabular}{r l l l l}
% & \tcl{\textbf{initial}} & \textbf{statically normal} \\
%\textbf{tycon definitions} & \tcl{$\delta ::=$}  & $\hat\delta ::=$ \\
% & \tcl{~~$\FF{iana}~\sigma; \FF{esyn}~\sigma; \FF{rep}~\sigma$} & ~~$\FF{iana}~\hat\sigma; \FF{esyn}~\hat\sigma; \FF{rep}~\hat\sigma$\\
% \\
%\textbf{(static language, cont.)} & \tcl{$\sigma ::= ...$} & $\hat\sigma ::= ...$ & $\kappa ::= ...$\\
%%type elim & \tcl{~~$\tfamcase{\sigma}{tycon}{x}{\sigma}{\sigma}$} & \\
%%static product elim & \tcl{~~$\sigma.\FF{prl}$} &\\
%% & \tcl{~~$\sigma.\FF{prr}$} &\\
%%static sum elim & \tcl{~~$\FF{case}~\sigma~\{\overline{l\langle\tvar{x}\rangle\Rightarrow\sigma}\}$} &\\
%%static list elim & \tcl{~~$\FF{listrec}(\sigma; \sigma; \tvar{h}, \tvar{t}, \tvar{r}.\sigma)$} & ~\\
%decidable equality & \tcl{~~$\tifeq{\hat\sigma}{\hat\sigma}{\kappa}{\hat\sigma}{\hat\sigma}$} & ~ & ~\\
%type error & \tcl{~~$\FF{tyerr}$} & ~ \\
%%static functions & \tcl{~~$\tlam{x}{\kappa}{\sigma}$} & ~~$\tlam{x}{\kappa}{\sigma}$ & ~~$\karrow{\kappa}{\kappa}$\\
%% & \tcl{~~$\sigma~\sigma$} & ~ & ~\\
%term translations & \tcl{~~$\titerm{\tri}$} & ~~$\titerm{\iota}$ & ~~$\kITerm$\\
%type translations & \tcl{~~$\titype{\trt}$} & ~~$\titype{\tau}$ & ~~$\kIType$\\
% & \tcl{~~$\trepof{\sigma}$} & ~\\
%arguments & \multicolumn{3}{c}{\color{gray}(introduced only by typechecker)} & ~~$\Q$\\
% & \tcl{~~$\FF{ana}(\sigma; \sigma; \tvar{x}.\sigma)$} & \\
% & \tcl{~~$\FF{syn}(\sigma; \tvar{x}, \tvar{t}.\sigma)$} & \\
%\\
%\textbf{internal language} & $\tri ::= $ & $\trt ::= $ & $\iota ::= $ & $\tau ::=$\\
%splice & ~~$\iup{\sigma}$ & ~~$\dup{\sigma}$ & \\
%variables & ~~$x$ &  & ~~$x$ & \\
%%fixpoints & ~~$\FF{fix}~x{:}\trt~\FF{is}~\tri$ & & ~~$\FF{fix}~x{:}\tau~\FF{is}~\iota$\\
%%functions & ~~$\lambda x{:}\trt.\tri$ & ~~$\trt \rightharpoonup \trt$ & ~~$\lambda x{:}\tau.\iota$ & ~~$\tau \rightharpoonup \tau$\\
%% & ~~$\tri~\tri$ & ~ & ~~$\iota~\iota$\\
%%products & ~~$()$ & ~~$\dunit$ & ~~$()$ & ~~$\dunit$\\
%% & ~~$(\tri, \tri)$ & ~~$\dpair{\trt}{\trt}$ & ~~$(\iota, \iota)$ & ~~$\dpair{\tau}{\tau}$\\
%% & ~~$\tri.\FF{prl}$ & ~ & ~~$\tri.\FF{prl}$ & ~\\
%%  & ~~$\tri.\FF{prr}$ & ~ & ~~$\tri.\FF{prr}$ & ~\\
%%%products & ~~$(\overline{l=\tri})$ & ~~$\Pi[\overline{(l, \trt)}]$ & ~~$(\overline{l = \iota})$ & ~~$\Pi[\overline{(l, \tau)}]$\\
%%% & ~~$\tri\cdot l$ & ~ & ~~$\iota\cdot l$ & ~\\
%%sums & ~~$\FF{in}[\ell](\tri)$ & ~~$\Sigma[\{\overline{\ell \hookrightarrow \trt}\}]$ & ~~$\FF{in}[\ell](\iota)$ & ~~$\Sigma[\{\overline{\ell \hookrightarrow \tau}\}]$\\
%% & ~~$\FF{case}~ \tri ~\{\overline{\ell(x) \Rightarrow \tri}\}$ & ~ & ~~$\FF{case}~\iota~\{\overline{\ell(x)\Rightarrow \iota}\}$ & ~\\
%%integers & ~~$\FF{int}[z]$ & ~~$\FF{int}$ & ~~$\FF{int}[z]$ & ~~$\FF{int}$\\
%% & ~~$\tri \pm \tri$ & ~ & ~~$\iota \pm \iota$ & ~\\
%% & ~~$\FF{if0}(\tri; \tri; \tri; \tri)$ & ~ & ~~$\FF{if0}(\iota; \iota; \iota; \iota)$ & ~
%% %& ~~$
%\end{tabular}
%\caption{Syntax from the perspective of fragment providers also includes an internal language and static language.}
%\end{figure*}
%\begin{figure*}
%\small
%\begin{tabular}{r l l}
%external typing context & $\Gamma ::= \emptyset \pipe \Gamma, x \Rightarrow \ttype{tycon}{\hat\sigma}$\\
%\\
% & \textbf{initial} & \textbf{statically normal} \\
%\textbf{static language, cont.} & $\sigma ::= ...$ & $\hat\sigma ::= ...$\\
%argument intro & ~~$\FF{arg}(\Gamma; \hat e)$ & ~~$\FF{arg}(\Gamma; \hat e)$
%\end{tabular}
%\caption{Syntax from the perspective of the semantics / compiler.}
%\end{figure*}
\newcommand{\fkctx}{\mathcal{F}}
\newcommand{\ok}[1]{#1~\mathtt{ok}}
\newcommand{\normalize}[3]{#1 \Downarrow_{#2} #3}
\newcommand{\normalizeX}[2]{\normalize{#1}{\Phi}{#2}}
\newcommand{\serr}[1]{#1~\mathtt{tyerr}}
%\begin{figure*}
%\small
%\begin{tabular}{r l l}
%\textbf{context} & \textbf{syntax} & \textbf{well-formedness}\\
%fragment kind context & $\fkctx ::= \fkctx_0 \pipe \fkctx, \fvar{tycon}[\kappa]~\{\FF{intro}[\kappa]; \FF{elim}[\kappa]\}$ & $\vdash \fkctx$\\
%kinding context & $\Delta ::= \Delta \pipe \Delta, \tvar{x} : \kappa$\\
%fragment context & $\Phi ::= \Phi_0 \pipe \Phi, \fvar{tycon}~\{\hat\delta\}$ & $\vdash \Phi \sim \fkctx$\\
%external typing context & $\Gamma ::= \emptyset \pipe \Gamma, x \Rightarrow \hat\sigma$ & $\vdash_\Phi \Gamma$\\
%internal typing context & $\Omega ::= \emptyset \pipe \Omega, x : \tau$\\
%~
%\end{tabular}
%\\
%\\
%\begin{tabular}{l l l l l}
%\textbf{kinding}& \multicolumn{2}{l}{\textbf{static normalization}} & \multicolumn{2}{l}{\textbf{program compilation}}\\
%$\ok{\rho}$ & $\normalize{\rho}{}{\hat\rho}$ & $\serr{\rho}$ & $\hat\rho \leadsto \iota$\\
%$\Delta \vdash_\fkctx {\phi}\sim\fkctx$ & $\normalizeX{\phi}{\hat\phi}$ & $\serr{\phi}$ & $\hat\phi \sim \Phi$\\
%$\Delta \vdash_\fkctx \delta\sim\{\kappa; \kappa\}$ & $\normalizeX{\delta}{\hat\delta}$ & $\serr{\delta}$ & \multicolumn{2}{l}{\textbf{active typechecking and translation}}\\
%$\Delta \vdash_\fkctx \sigma : \kappa$ & $\normalizeX{\sigma}{\hat \sigma}$ & $\serr{\sigma}$ & $\Gamma \vdash_\Phi \hat e \Leftarrow \hat\sigma \leadsto \iota$ & $\Gamma \vdash_\Phi \hat e \Rightarrow \hat\sigma \leadsto \iota$\\
%$\Delta \vdash_\fkctx \ok{e}$ & $\normalizeX{e}{\hat e}$ & $\serr{e}$ & $\vdash_\Phi \hat\sigma \leadsto \tau$ & $\vdash_\Phi \Gamma \leadsto \Omega$\\
%$\Delta \vdash_\fkctx \ok{\tri}$ & $\normalizeX{\tri}{\iota}$ & $\serr{\tri}$ & \multicolumn{1}{l}{\textbf{internal statics}} & \textbf{internal dynamics}\\
%$\Delta \vdash_\fkctx \ok{\trt}$ & $\normalizeX{\trt}{\tau}$ & $\serr{\trt}$ & $\Omega \vdash \iota : \tau$ & $\iota \mapsto \iota$~~~~~~~~$\iota~\mathtt{val}$
%\end{tabular}
%\caption{Summary of judgements}
%\end{figure*}

%
%\begin{figure*}
%\small
%$
%\begin{array}{lcl}
%\delta_{\texttt{nat}} & := & \quad \FF{iana}~\tlam{opidx}{\kunit+\kunit}{\tlam{tyidx}{\kunit}{\tlam{args}{\klist{\FF{Arg}}}{
% \tsumcase{\tvar{opidx}\\
%&&\quad\quad}{\_}{\tvar{isnil}~\tvar{args}~\titerm{0}\\
%&&\quad\quad}{\_}{\tvar{decons1}~\tvar{args}~\tlam{a}{\FF{Arg}}{\FF{ana}(\tvar{a}; \ttype{nat}{\tunit}; \tvar{x}.\titerm{\iup{\tvar{x}}+1})}}
%}}}\\
%&&\quad \FF{esyn}~\tlam{opidx}{\kunit}{\tlam{tyidx}{\kunit}{\tlam{x}{\kITerm}{\tlam{args}{\klist{\FF{Arg}}}{\tvar{decons2}~\tvar{args}~\tlam{a1}{\FF{Arg}}{\tlam{a2}{\FF{Arg}}{\\
%&&\quad\quad \FF{syn}(\tvar{a1}; \tvar{t1}, \tvar{i1}.\\
%&&\quad\quad\quad \F{let}\tvar{t2} : \kTypeBlur = \ttype{arrow}{(\ttype{nat}{\tunit}, \ttype{arrow}{(\tvar{t1}, \tvar{t1})})}~\F{in} \\
%&&\quad\quad\quad \FF{ana}(\tvar{a2}; \tvar{t2}; \tvar{i2}.(\tvar{t1},\\
%&&\quad\quad\quad\quad \titerm{\iapp{(\ifix{f}{\darrow{\dint}{\dup{\trepof{\tvar{t2}}}}}{\ilam{x}{\dint}{\\&&\quad\quad\quad\quad\quad\quad \FF{if0}(x; \iup{\tvar{i1}}; {
%   %\\&\quad\quad\quad\quad\quad\quad
%   \iapp{\iup{\tvar{i2}}~(x-1)}{(\iapp{f}{(x-1)})}})}}\\
%   &&\quad\quad\quad\quad\quad)}{\iup{\tvar{x}}}})))}}}}}}\\
%&&\quad \FF{rep}~\tlam{tyidx}{\kunit}{\titype{\dint}}\\
%\end{array}
%$
%\caption{\small The definition of the nat type constronstructor.}
%\end{figure*}
%\begin{figure}
%\small
%\begin{flalign}
%& \F{tycon}\fvar{record}~\F{of}\klist{\kpair{\FF{L}}{\kTypeBlur}}~\{\\
%& \quad \FF{iana}~\{\tlam{i}{\kpair{\klist{\kpair{\FF{L}}{\kTypeBlur}}}{\klist{\FF{L}}}}{\tlam{a}{\klist{\Q}}{\\& \quad\quad\tfold{\tvar{zip3}~\tfst{\tvar{i}}~\tsnd{\tvar{i}}~\tvar{a}}
%  {\titerm{\tunit}}{h}{t}{r}{
% \\&\quad\quad\quad\tifeq{\tfst{\tvar{first}~\tvar{h}}}{\tvar{second}~\tvar{h}}{\FF{L}}{
% \\&\quad\quad\quad\quad
% \F{let}~\tvar{x}:\kITerm~=~\FF{ana}(\tvar{third}~\tvar{h}; \tsnd{\tvar{first}~\tvar{h}})
%~\FF{in}
%\\ & \quad\quad\quad\quad \tfold{\tvar{t}}{\tvar{x}
%}{\_}{\_}{\_}{\titerm{(\iup{\tvar{x}},\iup{\tvar{r}})}}  }{\terr}
%%     \F{let}~\tvar{tt}:\FF{TT}=~\FF{syn}(\tsnd{\tsnd{\tvar{h}}}) \F{in}\\
%%\\     & \quad\quad\quad 
%}
%}}\}\\
%& \quad \FF{isyn}~\{\tlam{i}{\klist{\FF{L}}}{\tlam{a}{\klist{\Q}}{
%\\&\quad\quad \tfold{\tvar{zip2}~\tvar{i}~\tvar{a}}{\tden{\titerm{()}}{\ttype{record}{\tnil{{\kpair{\FF{L}}{\kTypeBlur}}}}}}{h}{t}{r}{
%\\&\quad\quad\quad \F{let}~\tvar{htt}:\FF{TT}~=~\FF{syn}(\tsnd{\tvar{h}})
%~\FF{in}
%\\&\quad\quad\quad \F{let}~\tvar{hty}:\kTypeBlur~=~\ttypeof{\tvar{htt}}
%\\&\quad\quad\quad \F{let}~\tvar{ty}:\kTypeBlur~=~\ttype{record}{(\tfst{\tvar{h}}, \tvar{hty})::\tnil{{\kpair{\FF{L}}{\kTypeBlur}}}}
%~\FF{in}
%\\&\quad\quad\quad \tfold{\tvar t}{\tden{\itransof{\tvar{htt}}}{\tvar{ty}}}{\_}{\_}{\_}{
%\\&\quad\quad\quad\quad \tden{\titerm{(\iup{\itransof{\tvar{htt}}},\iup{\itransof{\tvar{r}}})}}{\tvar{ty}}}}
%}}\}\\
%& \quad \FF{esyn}~\{\tlam{i}{\FF{L}}{\tlam{a}{\klist{\Q}}{
% \tvar{decons1}~\tvar{a}~\tlam{ty}{\kTypeBlur}{\tlam{x}{\kITerm}{
%\\&\quad\quad    \tfamcase{\tvar{ty}}{record}{sig}{
%\\&\quad\quad\quad \tfold{\tvar{sig}}{\terr}{h}{t}{r}{
%\\&\quad\quad\quad\quad \tifeq{\tfst{\tvar{h}}}{\tvar{i}}{\FF{L}}{\tfold{\tvar{t}}{\tden{\tvar{x}}{\tsnd{\tvar{h}}}}{\_}{t'}{\_}{
%\\&\quad\quad\quad\quad\quad \tfold{\tvar{t'}}{\tden{\titerm{\ifst{\iup{\tvar{x}}}}}{\tsnd{\tvar{h}}}}{\_}{\_}{\tvar{r'}}{
%\\&\quad\quad\quad\quad\quad\quad \tden{\titerm{\isnd{\iup{{\itransof{\tvar r}}}}}}{\tsnd{\tvar{h}}}}}\\&\quad\quad\quad\quad\hspace{-3px}}{\tvar{r}}
%}\\&\quad\quad\hspace{-3px}}{\terr}
% }}
%}}\\
%& \quad \FF{rep}~\{\tlam{i}{\klist{\kpair{\FF{L}}{\kTypeBlur}}}{   \\& \quad \quad \tfold{\tvar{i}}{\titype{\dunit}}{s}{j}{r}{
%     \tfold{\tvar{j}}{\trepof{\tsnd{\tvar{s}}}}{\_}{\_}{\_}{\\
%   &\quad\quad\quad
%     \titype{\dpair{\dup{\trepof{\tsnd{\tvar{s}}}}}{\dup{\tvar{r}}}}
%     }
%   }}\}\\
%& \}; \\
%& \F{let}\tvar{R} : \kTypeBlur = \ttype{record}{(\ell_1, \ttype{record}{\tnil{{\FF{L}\times\kTypeBlur}}}) :: \tnil{{\FF{L}\times\kTypeBlur}}}~\F{in}\\
%&\F{let}id = \elam{x}{\tvar{R}}{\{\ell_1=~x\cdot\ell_1\} :: \fvar{record}}\\
%&\F{let}triv = id[\{\ell_1 = \{\}\}]\cdot\ell_1
%%&\F{let}one = \FF{intro}[\ell_\text{succ}]
%%&\F{let}plus = \elam{x}{\tvar{nat}}{\elam{y}{\tvar{nat}}{\FF{elim}[()](x; y; \\
%%& \quad \elam{r}{\tvar{nat}}{intro[\ell_\text{succ}](r) : \tvar{nat}}})}
%\end{flalign}
%\caption{The definition of the record type in $\lam@\texttt{lang}$ using the desugarings in Figure \ref{desugaring} and with the addition of simple let bindings for both type and term variables (not shown). Some type-level helper functions also omitted for concision.}
%\label{record-theory}
%\end{figure}

%These correspond to the premises of the \emph{central compilation judgement} $\pkcompiles{\rho}{\iota}$:
%\[
%\inferrule[p-compiles]{
% \emptyset \vdash_{\fvalCtx_0} \rho\\
%%  \progOK{\emptyset}{\fvalCtx_0}{\rho}\\
% \pcompiles{\fvalCtx_0}{\rho}{\iota}
%}{\pkcompiles{\rho}{\iota}}
%\]
%This anchors our exposition; we will describe how it is derived (i.e. how to write a compiler for $\lam@\texttt{lang}$) in the following sections. 
The static semantics are specified by the \emph{bidirectional active typechecking and translation judgements} shown in Fig. \ref{atj}, which relate an expression, $e$, to a {type}, $\sigma$, and a \emph{translation}, $\iota$, under \emph{typing context} $\Gamma$ using imported {fragments} $\Phi$. The judgement form $\atjsynX{e}{\sigma}{\iota}$ specifies \emph{type synthesis} (the type is an ``output''), whereas $\atjanaX{e}{\sigma}{\iota}$ specifies \emph{type analysis} (the type is an ``input''). This can be seen as combining a bidirectional type system (in the style of Pierce and Turner \cite{Pierce:2000:LTI:345099.345100} and a number of subsequent formalisms and languages, e.g. Scala) with an elaboration semantics in the style of the Harper-Stone semantics for Standard ML \cite{Harper00atype-theoretic}. Our language of \emph{internal terms}, $\iota$, includes only functions and lists for simplicity. The form $\iup{\sigma}$ is used as an ``unquote'' operator, and will appear only in intermediate portions of a typing derivation, not in a translation (discussed below).

The first two rows of rules in Fig. \ref{atj} are essentially standard. \rulename{att-subsume} specifies the subsumption principle described in the previous section: if a type can be synthesized, then the term can also be analyzed against that type. We decide type equality purely syntactically here. \rulename{att-var} specifies that variables always synthesize types and elaborate identically. The typing context, $\Gamma$, maps variables to types in essentially the conventional way \cite{pfpl}. The rules $\rulename{att-ana-let}$ and $\rulename{att-syn-let}$ first synthesize a type for the bound value, then add this binding to the context and analyze or synthesize the body of the binding. The translation is to an internal function application, in the conventional manner. $\rulename{att-asc}$ begins by normalizing the provided index and checking that it is a type (Fig. \ref{tleval}, top). If so, it analyzes the ascribed expression against that type. The rules $\rulename{att-ana-slet}$ and $\rulename{att-syn-slet}$ eagerly evaluate the provided static term to a static value, then immediately perform the substitution (demonstrating the phase separation) in the expression before analysis or synthesis proceeds.%The normalization judgement for type-level terms is written $\tau \Downarrow_{\Gamma;\fvalCtx} \tau'$. The kinding rules (not shown here) guarantee that normalization of type-level terms cannot go wrong (we will refine what precisely this means later).

\paragraph{Lambdas}
The rule \rulename{att-ana-lam} performs type analysis on a lambda abstraction, $\lambda(x.e)$. If it succeeds, the translation is the corresponding lambda in the internal language. The type constructor $\textsc{fn}$ is included implicitly in $\Phi$ and must have a type index consisting of a pair of types (pairs are here just lists of length 2 for simplicity). Because both the argument type and return type are known, the body of the lambda is analyzed against the return type after extending the context with the argument type. This is thus the usual type analysis rule for functions in a bidirectional setting.

The rule \rulename{att-syn-idx-lam} covers the case where a lambda abstraction has an incomplete type ascription   providing only the argument type. This corresponds to the concrete syntax seen in the definition of $plus$ in Fig. \ref{atlam-example}. Here, the return type must be synthesized by the body.

These two rules can be compared to the rules in Listing \ref{fn-top}. The main difference is that in @$\lambda$, the language itself manages variables and contexts, rather than the type constructor. This is largely for simplicity, though it does limit us in that we cannot define function type constructors that require alternative or additional contexts. Addressing this in the theory is one avenue for future work.

Function application can be defined directly as a targeted expression, as seen in the example, which we will discuss below.
\paragraph{Fragment Provider Perspective}
\begin{figure}[t]
\small
\centering
$
\begin{array}{l}
\Phi_\text{fn} := \fvar{fn}=\{\FF{analit}=\FF{nil}; \FF{synidxlit}=\FF{nil}; \FF{anatarg}=\FF{nil}; \\
~\quad \FF{syntarg}=\lambda \tvar{tyidx}.\lambda \tvar{ifn}.\lambda \tvar{tmidx}.\lambda \tvar{args}.\\ 
\quad\quad  \tvar{isnil}~\tvar{tmidx}~(\tvar{decons1}~\tvar{args}~\lambda \tvar{arg}. \tvar{decons2} ~\tvar{tyidx}~\lambda \tvar{inty}.\lambda \tvar{rty}.\\
\quad\quad\quad  \FF{ana}(\tvar{arg}; \tvar{inty}; \tvar{ia}.\tvar{pair}~\tvar{rty}~\titerm{\FF{iap}(\iup{\tvar{ifn}};\iup{\tvar{ia}})}))\}
\end{array}
$
\caption{The \textsc{fn} fragment defines function application.}
\label{fn-tycon}
\vspace{10px}
\end{figure}
\begin{figure}[t]
\small
\centering
$
\begin{array}{l}
\Phi_\text{nat} := \fvar{nat}=\{\FF{analit}=\lambda \tvar{tyidx}.\lambda \tvar{tmidx}.\lambda \tvar{args}.\\ 
\quad\quad\tvar{isnil}~\tvar{tyidx}~(\tvar{lbleq}~\tvar{tmidx}~\tvar{lbl}[\texttt{0}]~(\tvar{isnil}~\tvar{args}~ \titerm{\tvar{inil}}));\\
\quad  \FF{synidxlit}=\lambda \tvar{incidx}.\lambda \tvar{tmidx}.\lambda \tvar{args}.\\
\quad\quad \tvar{isnil}~\tvar{incidx}~(\tvar{lbleq}~\tvar{tmidx}~\tvar{lbl}[\texttt{0}]~(\tvar{isnil} ~\tvar{args}~(\tvar{pair}~\FF{nil}~\titerm{\tvar{inil}})));\\
\quad  \FF{anatarg}=\lambda \tvar{tyidx}.\lambda \tvar{i1}.\lambda \tvar{ty}.\lambda \tvar{tmidx}.\lambda \tvar{args}.\\
  \quad\quad \tvar{lbleq}~\tvar{tmidx}~\FF{lbl}[\texttt{rec}]~(\tvar{decons2}~\tvar{args}~
    \lambda \tvar{arg1}.\lambda \tvar{arg2}.\\
    \quad \quad\quad  \FF{ana}(\tvar{arg1}; \tvar{ty}; \tvar{i2}.\FF{ana}(\tvar{arg2}; \tvar{fn2ty}~\FF{ty}[\fvar{nat}](\FF{nil})~\tvar{ty}~\tvar{ty}; \tvar{i3}.\\
    \quad\quad\quad\quad \titerm{\FF{ilistrec}(\iup{\tvar{i1}}; \iup{\tvar{i2}}; x, y.\FF{iap}(\FF{iap}(\iup{\tvar{i3}}; x); y))})));\\
\quad \FF{syntarg} = \lambda \tvar{tyidx}.\lambda \tvar{i1}.\lambda \tvar{tmidx}.\lambda \tvar{args}.\\
\quad\quad \tvar{lbleq}~\tvar{tmidx}~\FF{lbl}[\texttt{s}]~(\tvar{isnil}~\tvar{args}~(\\
\quad\quad\quad \tvar{pair}~\FF{ty}[\fvar{nat}](\tvar{nil})~\titerm{\FF{icons}(\FF{inil};\iup{\tvar{i1}})}))\}
\end{array}
$
\caption{The $\fvar{nat}$ fragment, based on G\"odel's \textbf{T} \cite{pfpl}.}
\vspace{10px}
\label{nat-tycon}
\end{figure}
\begin{figure}[t]
\small
\centering
$
\begin{array}{l}
\Phi_\text{lprod} := \fvar{lprod}=\{\FF{analit} = \lambda \tvar{tyidx}.\lambda \tvar{tmidx}.\lambda \tvar{args}. \\
\quad\quad \FF{listrec}(\tvar{zipexact3}~\tvar{tyidx}~\tvar{tmidx}~\tvar{args}; \titerm{\FF{inil}}; \tvar{h}. \tvar{ri}.\\
\quad\quad\quad \tvar{decons3} ~\tvar{h} ~\lambda \tvar{idxitem}.\lambda \tvar{lbl}.\lambda \tvar{e}.\tvar{decons2}~\tvar{idxitem}~\lambda \tvar{lblidx}. \lambda \tvar{tyidx}. \\
\quad\quad\quad\quad \tvar{lbleq}~\tvar{lbl}~\tvar{lblidx}~(\FF{ana}(\tvar{e}; \tvar{tyidx}; \tvar{i}.\titerm{\FF{icons}(\iup{\tvar{i}}, \iup{\tvar{ri}})})));\\
\quad\FF{synidxlit} = \lambda \tvar{incidx}.\lambda \tvar{tmidx}.\lambda \tvar{args}.\\
\quad\quad \FF{listrec}(\tvar{zipexact2}~\tvar{tmidx}~\tvar{args}; \tvar{pair} ~\FF{nil}~ \titerm{\FF{inil}}; \tvar{h}. \tvar{r}.\\
\quad\quad\quad \tvar{decons2} ~\tvar{h}~\lambda \tvar{lbl}.\lambda \tvar{e}.\tvar{decons2}~\tvar{h}~\lambda \tvar{ridx}.\lambda \tvar{ri}.\FF{syn}(\tvar{e};  \tvar{ty}. \tvar{i}. \\
\quad\quad\quad\quad \tvar{pair}~\FF{cons}(\tvar{pair}~\tvar{lbl}~ \tvar{ty}; \tvar{ridx})~\titerm{\FF{icons}(\iup{\tvar{i}}; \iup{\tvar{ri}})}));\\
\quad \FF{anatarg}=\FF{nil}; \text{\color{gray} (destructuring let could be implemented here)}\\ % 
\quad \FF{syntarg} = \lambda \tvar{tyidx}.\lambda \tvar{i}.\lambda \tvar{lbl}.\lambda \tvar{args}.\\
\quad\quad \tvar{isnil}~\tvar{args}~(\tvar{pair} ~(\tvar{lookup}~\tvar{lbl}~\tvar{tyidx})~\\
\quad\quad\quad \titerm{\FF{iap}(\FF{iap}(nth;\iup{\tvar{i}}); \iup{\tvar{itermofn} ~(\tvar{posof}~\tvar{lbl}~\tvar{tyidx})})})
\end{array}
$
\caption{The $\fvar{lprod}$ fragment (labeled products are like records, but the field order matters; cf. Listing \ref{record}).}
\label{lprod-tycon}
\vspace{10px}
\end{figure}
\begin{figure}[t]
\small
$\fbox{\inferrule{}{\sigma~\mathtt{ty}_\fvalCtx}}$
\vspace{-20px}
\renewcommand{\MathparLineskip}{\lineskiplimit=.6\baselineskip\lineskip=.6\baselineskip plus .2\baselineskip}
\begin{mathpar}
\inferrule[ty]{
  \FF{ty}[\fvar{tycon}](\sigma) \Downarrow_{\emptyctx; \fvalCtx} \FF{ty}[\fvar{tycon}](\sigma)
}{\FF{ty}[\fvar{tycon}](\sigma)~\mathtt{ty}_\fvalCtx}
\end{mathpar}
$\fbox{\inferrule{}{\sigma \Downarrow_{\Gamma;\fvalCtx} \sigma}}$~
%$\fbox{\inferrule{}{\sigma \Downarrow_{\Gamma;\fvalCtx}^= \sigma}}$~
\vspace{-15px}
\begin{mathpar}
%\inferrule[repof]{
% \tau \Downarrow_{\Gamma;\fvalCtx} \ttype{tycon}{\tauidx}\\
% \vdash_\fvalCtx \FF{rep}(\fvar{tycon}) = \taurep\\
% \taurep~\tauidx \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}
%}{
% \FF{repof}(\tau) \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}
%}
%
\inferrule[n-ty]{
  \fvar{tycon} \in \text{dom}(\fvalCtx)\\
  \sigma \Downarrow_{\Gamma;\fvalCtx} \sigma'
}{\FF{ty}[\fvar{tycon}](\sigma)~\Downarrow_{\Gamma;\fvalCtx} \FF{ty}[\fvar{tycon}](\sigma')}
\\
\inferrule[n-arg]{ }{\FF{arg}[e]~\Downarrow_{\Gamma;\fvalCtx} \FF{arg}[e]}

\inferrule[n-ana]{
  \sigma_1 \Downarrow_{\Gamma;\fvalCtx} \FF{arg}[e]\\
  \sigma_2 \Downarrow_{\Gamma;\fvalCtx} \sigmat{ty}\\
  \sigmat{ty}~\mathtt{ty}_\fvalCtx\\\\
  \atjanaX{e}{\sigmat{ty}}{\iota}\\
  [\titerm{\iota}/\tvar{x}]\sigma_3 \Downarrow_{\Gamma;\fvalCtx} \sigma_3'
}{
  \FF{ana}(\sigma_1; \sigma_2; \tvar{x}.\sigma_3) \Downarrow_{\Gamma;\fvalCtx} \sigma_3'
}

\inferrule[n-syn]{
    \sigma_1 \Downarrow_{\Gamma;\fvalCtx} \FF{arg}[e]\\
  \atjsynX{e}{\sigma}{\iota}\\\\
    [\sigma/\tvar{x}_{1}, \titerm{\iota}/\tvar{x}_{2}]\sigma_2 \Downarrow_{\Gamma;\fvalCtx} \sigma_2'
}{
  \FF{syn}(\sigma_1; \tvar{x}_{1}.\tvar{x}_{2}.\sigma_2) \Downarrow_{\Gamma;\fvalCtx} \sigma_2'
}

\inferrule[n-quote]{
  \iota \curlyveedownarrow_{\Gamma;\fvalCtx} \iota'
}{\titerm{\iota} \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota'}}
\end{mathpar}
$\fbox{\inferrule{}{\iota \curlyveedownarrow_{\Gamma;\fvalCtx} \iota}}$
\vspace{-15px}
\begin{mathpar}

\inferrule[q-uq]{\sigma \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota}}{\iup{\sigma} \curlyveedownarrow_{\Gamma;\fvalCtx} \iota}

\inferrule[q-x]{ }{x \curlyveedownarrow_{\Gamma;\fvalCtx} x}

\inferrule[q-lam]{\iota \curlyveedownarrow_{\Gamma;\fvalCtx} \iota'}
{\lambda(x.\iota) \curlyveedownarrow_{\Gamma;\fvalCtx} \lambda (x.\iota')}
%
%\inferrule[q-ap]{\iota_1 \curlyveedownarrow_{\Gamma;\fvalCtx} \iota_1'\\
%\iota_2 \curlyveedownarrow_{\Gamma;\fvalCtx} \iota_2'}{
%\FF{iap}(\iota_1; \iota_2) \curlyveedownarrow_{\Gamma;\fvalCtx} \FF{iap}(\iota_1'; \iota_2')}
\end{mathpar}
\caption{Selected normalization rules for the static language.}
\vspace{10px}
\label{tleval}
\end{figure}

Fragment providers define type constructors by defining ``methods'', $\delta$, that control analysis and synthesis literals and targeted expressions. These are static functions invoked by the final four rules in Fig. \ref{atj}, which we will describe next. The type constructors $\fvar{fn}$, $\fvar{nat}$ and $\fvar{lprod}$ are shown in Figs. \ref{fn-tycon}-\ref{lprod-tycon}, respectively. They use helper functions for working with labels ($\tvar{lbleq}$), lists ($\tvar{isnil}$, $\tvar{decons1}$, $\tvar{decons2}$, $\tvar{decons3}$, $\tvar{zipexact2}$, $\tvar{zipexact3}$, and $\tvar{pair}$), lists of pairs interpreted as finite mappings ($\tvar{lookup}$ and $\tvar{posof}$) and translating a static representation of a number to an internal representation of that number ($\tvar{itermofn}$). We also assume an internal function $nth$ that retrieves the $n$th element of a list. All of these are standard or straightforward and omitted for concision. Failure cases for the static helper functions evaluate to $\FF{fail}$. Derivation of the typing judgement does not continue if a $\FF{fail}$ occurs (corresponding to an \lstinlinep{atlang.TypeError} propagated directly to the compiler).

\subparagraph{Literals} The rule $\rulename{att-ana-lit}$, invokes the $\FF{analit}$ method of the type constructor of the type the literal is being analyzed against, asking it to return a translation (a value of the form $\titerm{\iota}$). The type and term index are provided, as well as a list of \emph{reified arguments}: static values of the form $\FF{arg}[e]$. The $\fvar{fn}$ type constructor does not implement this (functions are introduced only via lambdas). The $\fvar{nat}$ type constructor implements this by checking that the term index was the label corresponding to $\texttt{0}$ and no arguments were provided. The $\fvar{lprod}$ type constructor is more interesting: it folds over each corresponding item in the type index (a pair consisting of a label and a type), the term index (a label) and the argument list, checking that the labels match and programmatically analyzing the argument against the type it should have. A reified argument $\sigma_1$ against a type $\sigma_2$, binding the translation to $\tvar{x}$ in $\sigma_3$ if successful (and failing otherwise) with the static term $\FF{ana(\sigma_1; \sigma_2; \tvar{x}.\sigma_3)}$, the semantics of which are in Fig. \ref{tleval}. Labeled products translate to lists by recursively composing the translations of the field values using the ``unquote'' form, $\iup{\sigma}$, which is eliminated during normalization (also seen in Fig. \ref{tleval}). This function can be compared to the method \lstinlinep{ana_Dict} in Listing \ref{record}.

Literals with an incomplete type ascription can synthesize a type by rule \rulename{att-syn-idx-lit}. The $\FF{synidxlit}$ method of type constructor of the partial ascription is called with the incomplete type index, the term index and the arguments as above, and must return a pair consisting of the complete type index and the translation. Again, $\fvar{fn}$ does not implement this. The $\fvar{nat}$ type constructor supports it, though it is not particularly interesting, as $\fvar{nat}$ is always indexed trivially, so it follows essentially the same logic as in $\FF{analit}$. The $\fvar{lprod}$ type constructor is more interesting: in this case, when the incomplete type index is trivial (as in the example in Fig. \ref{atlam-example}), the list of pairs of labels and types must be synthesized from the literal itself. This is done by programatically synthesizing a type and translation for a reified argument using $\FF{syn(\sigma_1; \tvar{x}.\tvar{y}.\sigma_2)}$, also specified in Fig. \ref{tleval}. The type index and translation are recursively formed. This can be compared to the class method \lstinlinep{syn_idx_Dict} in Listing \ref{record}.

\subparagraph{Targeted Terms} Targeted terms are written $\FF{targ}[\sigmat{tmidx}](e_\text{targ}; \overline{e})$. The type constructor of the type recursively synthesized by the \emph{target}, $e_\text{targ}$, is delegated control over analysis and synthesis via the methods $\FF{anatarg}$ and $\FF{syntarg}$, respectively, as seen in rules \rulename{att-ana-targ} and \rulename{att-syn-targ}. Both receive the type index, the translation of the target, the term index and the reified arguments. The former also receives the type being analyzed and only needs to produce a translation. The latter must produce a pair consisting of a type and a translation.

The $\fvar{fn}$ type constructor defines function application by straightforwardly implementing $\FF{syntarg}$.  Because of subsumption, $\FF{anatarg}$ need not be separately defined.

The $\fvar{nat}$ type constructor defines the successor operation synthetically and the recursor operation analytically (because it has two branches that must have the same type). The latter analyzes the second argument against a function type, avoiding the need to handle binding itself. Natural numbers translate to lists, so the $\FF{nat}$ recursor can be implemented straightforwardly using the list recursor. In a practical implementation, we might translate natural numbers to integers and use a fixpoint computation instead.

The $\fvar{lprod}$ type constructor defines the projection operation synthetically, using the helper functions mentioned above to lookup the appropriate item in the type index. Note that one might also define an analytic targeted operation on labeled products corresponding to pattern matching, e.g. $\FF{let}~\{\texttt{a}{=}x, \texttt{b}{=}y\}=r~\FF{in}~e$. @$\texttt{lang}$ supports this using Python's syntax for destructuring assignment, but we must omit the details.

\paragraph{Metatheory} The formulation shown here guarantees that type synthesis actually produces a type, given well-formed contexts. The definitions are straightforward and the proof is a simple induction with simple lemmas about the normalization semantics. %We plan to prove the full proofs in a companion tech report (available upon request).

\begin{theorem}[Synthesis]
If $\fvalCtx~\mathtt{fragment}$ and $\Gamma~\mathtt{ctx}_\fvalCtx$ and $\atjsynX{e}{\sigma}{\iota}$ then $\sigma~\mathtt{ty}_\fvalCtx$.
\end{theorem}

We can also prove that importing additional type constructors cannot change the semantics of a previously well-typed term, assuming that naming conflicts have been resolved by some extrinsic mechanism. Indeed, the proof is an essentially trivial induction because of the way we have structured our mechanism. The type constructor delegated responsibility over a term is deterministically determined irrespective of the structure of $\fvalCtx$. We plan to provide full proofs of these theorems in a technical report.

\begin{theorem}[Unambiguous Extension]
If $\fvalCtx~\mathtt{fragment}$ and $\Gamma~\mathtt{ctx}_\fvalCtx$ and $\atjsynX{e}{\sigma}{\iota}$ and $\fvalCtx'~\mathtt{fragment}$ and $\text{dom}(\fvalCtx) \cap \text{dom}(\fvalCtx') = \emptyset$ then ${\Gamma}\vdash_{\fvalCtx;\fvalCtx'}{e}\Rightarrow{\sigma}\leadsto{\iota}$.
\end{theorem}
%Metatheoretic guarantees about the translation cannot be provided in the formulation as given. However, inserting straightforward checks to guarantee that the translation is a closed term, and provide simple mechanisms for hygiene, would be simple, but are omitted.
%\noindent
%\subsubsection{Rules} We briefly describe each rule and relevant additional judgements below:
%
%\noindent
%The rule $\rulename{att-lam}$ says that lambda terms can only be analyzed against arrow types and translate to lambda terms in the internal language.
%
%The rule $\textsc{\textbf{att-intro-ana}}$ says that generalized introductory forms can only be analyzed against a type. That type constructor is consulted to extract its \emph{introductory operator definition}, written (e.g. in Figure \ref{nat-atfrag}) after the $\FF{iana}$ keyword as a type-level function. This function is given the provided operator index, the type's type index and a list encapsulating the operator's  \emph{arguments}. An argument is reified as a type-level term of the form $(\Gamma, e)?$ and has kind $\FF{Arg}$. Note that only the compiler should construct such a term (in practice, this could be enforced purely syntactically, so we do not enforce this judgmentally here). The operator definition is responsible for producing a translation, or evaluating to $\terr$ if this is not possible. A translation is simply a quoted internal term, written $\titerm{\iota}$, with kind $\kITerm$. Like quoted internal types, quoted internal terms support an unquote form, written $\iup{\tau}$, which is normalized away (in a capture-avoiding manner, not shown).
%
%Natural numbers have two introductory forms, each indexed trivially. Because there is only one generalized introductory form, we instead index the operator by a simple sum kind with two trivial cases, $\kunit + \kunit$. The first case corresponds to the zero operator and the second to the successor. In the introductory operator definition, we see in the first case that the translation $0$ is produced after checking that no arguments were provided (using a simple helper function, $\tvar{isnil}$, not shown). In the second case, we first extract the single argument (using the helper function $\tvar{decons1}$, not shown). We then analyze it against the type $\ttype{nat}{\tunit}$ using one of the elimination forms for arguments, $\FF{ana}(\tau_1; \tau_2; \tvar{t}_{trans}.\tau_3)$. If it succeeds, we construct the appropriate translation based on the translation of the argument. The normalization rule for successful analysis, $\textsc{ana}$, is shown in Figure \ref{tleval}. 
%
%Note also that lambda is technically the introductory form for the $\fvar{arrow}$ type constructor, but because it requires manipulating the context, it must be built in to @$\lambda$. We do not currently support type system fragments that require adding or manipulating existing contexts.
%
%The rule $\rulename{att-elim-syn}$ says that elimination forms synthesize a type. This is done again by consulting a type-level function, called the \emph{eliminatory operator definition}, associated with a type constructor, here the type constructor of the type  recursively synthesized for the primary operand, $e$. This definition is invoked with the type index, the operator index, the translation of the primary operand and a list of arguments. Because elimination forms synthesize a type, this definition must produce both a type and a translation. Representational consistency is then checked.
%
%Note that the elimination form for the $\fvar{arrow}$ type constructor can be defined in this manner (because it does not require binding variables), so application is simply syntactic sugar, rather than a built-in operator in the external language.
%
%The constructor context, $\Phi$, tracks user-defined type  constructors introduced in the ``imported'' fragments. %Note that fragments can also export type-level values bound to type-level variables (written in bold font, e.g. $\tvar{nat}$, $\tvar{s}$, $\tvar{z}$, etc.), but substitution will be performed prior to typechecking, so they need not be tracked by $\Phi$. % Weakening and exchange of the constructor context closely related to the issue of conservativity that we will return to. Each constructor is identified by name, so there is no analog to contraction. 
% This form of semantics can be seen as lifting into the language specification the first stage of a type-directed compiler like the TIL compiler for Standard ML \cite{tarditi+:til-OLD} and has some parallels to the Harper-Stone semantics for Standard ML \cite{Harper00atype-theoretic}. There, as in Wyvern, external terms were given meaning by elaboration from the EL to an IL having the same type system. Here the two languages have different type systems, so we call it a translation rather than an elaboration (arranging the judgement form slightly differently to emphasize this distinction, cf. above).

%In @$\lambda$, the internal language (IL) provides partial functions (via the generic fixpoint operator of Plotkin's PCF), simple product  types and integers for the sake of our example (and as a nod toward practicality on contemporary machines). In practice, the internal language could be any typed  language with a specification for which type safety and decidability of typechecking have been satisfyingly determined. In Sec. \ref{ace}, we will see how the internal language can itself be made user-definable. The internal type system serves as a ``floor'': guarantees that must hold for terms of any type (e.g. that out-of-bounds access to memory never occurs) must be maintained by the internal type system. User-defined constructors can enforce invariants stronger than those the internal type system maintains at particular types, however. Performance is also ultimately limited by the internal language and downstream compilation stages that we do not here consider (safe compiler extension has been discussed in previous work, e.g. \cite{conf/pldi/TatlockL10}).
%
%The external language has a fixed syntax with six forms: variables, $\FF{let}$-bindings, lambda terms, type ascription and generalized introductory and elimination forms. As we will see, the generalized introductory form is given meaning by the type it is analyzed against (similar to the protocol for TSLs in Wyvern), and the elimination form is given meaning by the type of the external term being eliminated ($e$). This represents an internalization into the language of Gentzen's inversion principle \cite{gentzen}\todo{what to cite for this?}. Fig. \ref{desugaring} shows how to recover more conventional introductory and elimination forms by a purely syntactic desugaring.%\footnote{Note that for even more flexibility, we could also include TSLs, but we choose to avoid that for simplicity (and to show that, strictly speaking, one need not have an extensible syntax to have an extensible type system, and \textit{vice versa}).}
%
%\subsection{Types and Type-Level Computation}\label{types}
%@$\lambda$ supports, and makes extensive use of, simply-kinded type-level computation. Specifically, type-level terms, $\tau$, themselves form a typed lambda calculus. The classifiers of type-level terms are called \emph{kinds}, $\kappa$, to distinguish them from  \emph{types}, which are  type-level values of kind $\kTypeBlur$. %In @\texttt{lang}, the type-level language (together with the level of programs) is written in Python, which can be thought of as having a rather limited kind system (with one kind, \lstinlinep{dyn}). Here, we are able to more precisely discuss the kinds of values in the type-level language. 
%As in Sec. \ref{att}, types are formed by applying a \emph{type constructor} to an \emph{index}. User-defined type constructors are declared in fragment definitions using $\FF{tycon}$. Each constructor in the program must have a unique name, written e.g. \fvar{nat} or \fvar{lprod}. %\footnote{We assume naming conflicts can be avoided by some extrinsic mechanism.} 
%A type constructor must also declare an \emph{index kind}, $\kappat{tyidx}$. A type is introduced by applying a type constructor to an index of this kind, written $\ttype{Tycon}{\taut{tyidx}}$.  For example, the type of natural numbers is indexed trivially (i.e. by kind $\kunit$), so it is written $\ttype{Nat}{\tunit}$.
%
% To permit the embedding of interesting type systems, the type-level language includes several kinds other than $\kTypeBlur$. We lift several functional data structures to the type level: here, only unit ($\kunit$), binary products ($\kpair{\kappa_1}{\kappa_2}$), binary sums ($\ksum{\kappa_1}{\kappa_2}$) and lists ($\klist{\kappa}$), in addition to labels (introduced as $\ell$, possibly with a subscript, having kind \FF{L}).  The type constructor $\fvar{nat}$ is indexed trivially because there is only one natural number type, but $\fvar{lprod}$ would be indexed by a list of pairs of {labels}  and types. The type constructor $\fvar{arrow}$ is included in the initial constructor context, $\fvalCtx_0$, and has index kind $\kpair{\kTypeBlur}{\kTypeBlur}$.
% As with the internal language, in practice, one could include a richer programming language and retain the spirit of the calculus, as long as it does not introduce general recursion at the type level. For example, our desugarings add support for number literals and string literals, which require adding numbers and strings to the type-level language (and providing a means for lifting them from the type-level language to the internal language, as we will discuss).
%  
% %We see a record type, abbreviated $\tvar{R}$, constructed on line 26. 
%
%The kind $\kTypeBlur$ also has an elimination form, $\tfamcase{\tau}{Tycon}{x}{\tau_1}{\tau_2}$ allowing the extraction of a type index by case analysis against a contextually-available type constructor. To a first approximation, one might think of type constructors as constructors of a built-in open datatype \cite{conf/ppdp/LohH06}, $\kTypeBlur$, at the type-level. Like open datatypes, there is no notion of exhaustiveness so the default case is required for totality. %We will see where this is used shortly.
%
%
%
%%We will write the kind of types as $\kTypeBlur$, though it is also written $\star$ or \lstinlinep{Type} in various similarly structured languages (see Sec. \ref{related-work}). 
%% Rather than there being a fixed set of type constructors, we allow the programmer to declare new type  constructors, and give the static and dynamic semantics of their associated operators, by writing type-level functions. In the semantics for this calculus, our kind system combined with techniques borrowed from the typed compilation literature and a form of type abstraction allow us to prove strong type safety, decidability and conservativity theorems.
%
%
%
%% and integers ($\dint$). We also include labels ($\klabel$), written in a slanted font, e.g. $\tlabel{myLabel}$, which are string-like values that only support comparison and play a distinguished role in the expanded syntax, as we will later discuss. Our first example, $\fvar{nat}$, is indexed trivially, i.e. by unit kind, $\kunit$, so there is only one natural number type, $\ttype{nat}{\tunit}$, but we will show examples of type constructors that are indexed in more interesting ways in later portions of this work. For example, $\fvar{LabeledTuple}$ has index kind $\klist{\kpair{\klabel}{\kTypeBlur}}$. 
% 
% Type constructors are not first-class; they do not themselves have arrow kind as in some kind systems (e.g.  \cite{watkins2008specifying}; Ch. 22 of \emph{PFPL} describes a related system \cite{pfpl}). The type-level language does, however, include total functions of arrow kind, written $\karrow{\kappa_1}{\kappa_2}$. Type constructor application can be wrapped in a type-level function to emulate a first-class or uncurried version of a type constructor for convenience.% (indeed, such a wrapper could be generated automatically, though we do not do so). 
%
%Two type-level terms of kind $\kTypeBlur$ are equivalent if they apply the same constructor, identified by name, to equivalent indices. Going further, we ensure that deciding type equivalence requires only checking for syntactic equality after normalization by imposing the restriction that equivalence at a type constructor's index kind must be decidable in this way. Our treatment of equivalence in the type-level language is thus quite similar to the treatment of term-level equality using ``equality types'' in a language like Standard ML.
%% A kind $\kappa$ is an  \emph{equality kind} if $\kEq{\kappa}$ can be derived (see appendix). 
%Conditional branching on the basis of equality at an equality kind can be performed in the type-level language. Equivalence at arrow kind is not decidable by our criteria, so type-level functions cannot appear within type indices. This also prevents general recursion from arising at the type level. Without this restriction, a type-level function taking a type as an argument could ``smuggle in'' a self reference as a type index, extracting it via case analysis (continuing our analogy to open datatypes, this is closely related to the positivity condition for inductive datatypes in total functional languages like Coq).% as maintaining the metatheoretic guarantee that typing respects type equivalence would impose a substantial burden in such a setting.% (a na\"ive approach to this would impose non-trivial extrinsic proof obligations onto extension developers that, unlike in others in this thesis, could threaten type safety).

%Every type constructor also defines type-level functions called $\FF{iana}$ and $\FF{esyn}$, which we will describe below, and a \emph{representation schema}, a type-level function that associates with every type an internal type. We will return to this after introducing operators.
%
%\subsection{Core External Forms and Desugaring}\label{opcons}
%The syntax for external terms (Figure \ref{grammar}) contains variables, $\lambda$ terms, three generalized introductory forms and a single generalized elimination form. The introductory forms are either unascribed, ascribed with a type or ascribed with a type constructor, as in our discussion of @\texttt{lang}. These generalized forms take a single a type-level value as an {index} and $n \geq 0$ arguments, which are other external terms. To better motivate this choice, we can give a purely syntactic desugaring of a Python-like syntax with labels to these forms, shown in Figure \ref{desugaring}. It is instructive to rewrite lines 27-28 of Figure \ref{record-theory} using these desugarings.% In @\texttt{lang}, desugarings can be user-defined 
%
%User-defined operator constructors are declared using \textsf{opcon}.  For reasons that we will discuss, our calculus associates every operator  constructor with a type constructor. The \emph{fully-qualified name} of every operator constructor, e.g. $\fvar{nat}.\opvar{z}$, must be unique. Operator constructors, like type constructors, declare an index kind, $\kappaidx$. In our first example, all the operator constructors are indexed trivially (by index kind $\kunit$), but other examples use more interesting indices. For example, in SML, the projection operator \lstinlinep{#3} can be applied to an $n$-tuple, $e$, iff $n \geq 3$. Note that it thus cannot be a function with a standard arrow type. Notionally, \lstinlinep{#} is an operator constructor and \lstinlinep{3} is its index. In an active embedding of $n$-tuples into $\lam@\texttt{lang}$, this would be written $\eop{Tuple}{prj}{3}{e}$ (we will nearly recover ML's syntax later). $\fvar{LabeledTuple}.\opvar{prj}$ is the operator constructor used to access a field of a labeled tuple, so it has index kind $\klabel$. An operator itself is, notionally, selected by indexing an operator constructor, e.g. $\fvar{nat}.\opvar{s}\langle \tunit \rangle$, but technically neither operator constructors nor operators are first-class at any level (additional machinery would be needed, e.g. an \textsf{Op} kind, but this is not fundamental to our calculus). 
%Instead, in the external language, an operator constructor is applied by simultaneously providing an index and  $n \geq 0$ \emph{arguments}, written $\eop{Tycon}{op}{\tauidx}{\splat{e}{1}{n}}$\footnote{It may be helpful to distinguish between type/operator constructors and \emph{term formers}. There are term formers at all levels in the calculus. For example, operator constructor application and $\lambda$ are  external term formers, and type constructor application is a type-level term former. We might write these following Harper's conventions for abstract syntax to highlight this distinction \cite{pfpl}: $\mathtt{lam}[\tau](x.e)$, $\mathtt{ocapp}[\fvar{Tycon}, \opvar{op}, \tauidx](\splat{e}{1}{n})$ and $\texttt{tcapp}[\fvar{Tycon}](\tau)$.}. For example, on line 18 of Fig. \ref{nat}, we see the operator constructors $\fvar{nat}.\opvar{z}$ and $\fvar{nat}.\opvar{s}$ being applied to compute $two$. %\footnote{Although our focus here is entirely on semantics, a brief note on syntax: in the expanded syntax, the trivial indices and empty argument lists can be omitted, so we could write \texttt{Nat.s(Nat.s(Nat.z))}. With the ability to ``open'' a type's operators into the context, we could shorten this still to \texttt{s(s(z))}. Alternatively, with the ability to define a TSL in a manner similar to that in Sec. \ref{aparsing}, we might instead just write \texttt{2}.}

%The expressive and metatheoretic power of the calculus arises from how the rules for the active typing judgement handle these  generalized forms. Rather than fixing the specification of a finite collection of operator constructors and tasking the \emph{compiler} with deciding a typing derivation on its basis, the specification instead delegates to a type-level function associated with a type constructor. In the core calculus, this is one of three functions, called \FF{iana}, \FF{isyn} and \FF{esyn} in the grammar.\footnote{This means that any particular type constructor only supports a single introduction and elimination desugaring. This is a minor inconvenience in some cases that is resolved in @\texttt{lang} by the use of methods.} We see how this is done with the rules for the active typing judgements, given in Figure \ref{atj}.

%The rule $\rulename{att-lam}$ must take into account the fact that, because we support type-level computation, the type annotation on the argument may not be in normal form. Thus, we evaluate it to normal form. Note that the kinding rules (not shown) will guarantee that, because $\tau_1$ is of kind $\kTypeBlur$, its normal form, $\tau'_1$, is of the form $\ttype{tycon}{\tauidx}$ for some normal $\tauidx$. To generate an appropriate internal type annotation in the translation, we need to compute the representation type of $\tau_1'$. This involves calling the representation function associated with its type constructor (rule \fvar{repof} in Figure \ref{tleval}). The form $\titerm{\sigma}$ is a \emph{quoted internal type}. Note in the syntax that there is an unquote form, $\dup{\tau}$, which allows us to compose internal types compositionally without needing to expose an elimination form. Indeed, it is an interesting facet of our calculus that we never need to examine syntax trees directly to implement extensions. The evaluation semantics remove quotations, so the normal form of a quoted internal type contains an internal type with no quotations.
%
%Lambda functions are the only introductory form requiring special support in the calculus (because they need to  manipulate the context; see Discussion). The next three rules show how any other abstractions that we define (e.g. records, decimals, etc.) make use of a generalized introductory form. 
%
%The rule  $\rulename{att-i-unasc}$ shows that unascribed introductory forms can only be analyzed against a type. Given such a type, the rule extracts a definition, named $\FF{iana}$, from the type constructor (cf. the \lstinlinep{ana_Dict} and \lstinlinep{trans_Dict} methods earlier). It calls this function with a pair containing the operator and type index and a list of \emph{reified arguments}, which have kind $\Q$ and introductory form $(\Gamma; e)?$. These are only constructed by the compiler (there would be no corresponding form in the concrete syntax). Their purpose is to allow the definition to programatically invoke synthesis and analysis as needed using the $\FF{ana}$ and $\FF{syn}$ operators. We can see the $\fvar{record}$ type constructor doing so on line 5 of Figure \ref{record-theory} to ensure that the field value provided as an argument has the same type as the corresponding label (accomplished by simultaneously folding over all three pieces of input data). The rule for performing analysis, $\fvar{ana}$, is in Figure \ref{tleval}. If it succeeds, analysis returns a \emph{translation}, which is a quoted internal term with kind $\kITerm$ and introductory form $\titerm{\iota}$. Like quoted internal types, there is an unquote form that is eliminated during evaluation. The operator definition uses this to construct a translation for the record. As before, empty records translate to units and records with a single field are unadorned. In this example, records with two or more fields translate to nested tuples. 
%
%Because we have a representation type associated with the type, we can check that the translation is \emph{representationally consistent} (the final premise). As we will discuss, representational consistency combined with type safety of the internal language implies type safety overall (it arises as a strengthening of the inductive hypothesis needed to prove type safety, and is closely related to work on \emph{type-preserving compilation} in the TIL compiler for Standard ML, \cite{tarditi+:til-OLD}).
%
%The next rule, $\rulename{att-i-asc-ty}$ states that introductory forms ascribed with a type are analyzed against that type.
%
%Introductory forms ascribed with a type constructor can, according to the final introductory rule, $\rulename{att-i-asc-tycon}$, synthesize a type via the definition $\FF{isyn}$. It is passed the operator index (there is no type index, since we only have a type constructor) and a list of arguments, as before. In this case, it must return not just a translation, but also a type. We use the form $\tden{\tau_2}{\tau_1}$ for such a pairing, which has kind $\FF{TT}$. The type and translation can be extracted from it using the appropriately named elimination forms (indeed, as given, it is merely a pair, but we give it special syntax for clarity -- it looks like the conclusion of the typing judgement -- and for other reasons that will become clear in future work). To synthesize a type from a list of labels and arguments, we must be able to synthesize types from the arguments. The $\FF{syn}(\tau)$ operator permits this, per rule $\fvar{syn}$, seen being used in Listing \ref{record-theory}. We check representational consistency of the result, as before.
%
%Finally, we show the rule for elimination forms. It operates by first synthesizing a type for the ``primary'' subterm, then extracting the $\FF{esyn}$ definition from its type constructor. As before, it is called with the arguments and representational consistency is checked. Note that the primary subterm is itself the first argument (though we have already synthesized a type for it, it is more uniform and clear to allow the operator  to do so again, which is done in our example by the $\tvar{decons1}$ helper function).

%\subsection{Representational Consistency Implies Type Safety}\label{repcon}
%A key invariant that our operator definitions maintain is that well-typed external terms of type $\ttype{Nat}{\tunit}$ always translate to well-typed internal terms of internal type $\dint$, the representation type of $\ttype{Nat}{\tunit}$. Verifying this for the zero case is simple. For the translation produced by the successor case to be of internal type $\dint$ requires that $\iup{\tvar{x}}$ be of internal type $\dint$. Because it is the result of analyzing an argument against $\ttype{Nat}{\tunit}$, and the only other introductory form is the zero case, this holds inductively. We will discuss the recursor later, but it also maintains this invariant inductively.
%
%This invariant would not hold if, for example, we allowed the type and translation: $$({{\ttype{Nat}{\tunit},\titerm{(0, ())}}})$$
%\noindent
%In this case, there would be two different internal types, $\dint$ and $\dpair{\dint}{\dunit}$,  associated with a single external type, $\ttype{Nat}{\tunit}$. This would make it impossible to reason \emph{compositionally} about the translation of an external term of type $\ttype{Nat}{\tunit}$, so our implementation of the successor would produce ill-typed translations in some cases but not others. Similarly, we wouldn't be able to write functions over all natural numbers because there would not be a well-typed translation to give to such a function. This violates type safety: there are now well-typed external terms, according to the active typechecking and translation judgement, for which evaluating the corresponding translation would ``go wrong''. 
%
%To reason compositionally about the semantics of well-typed external terms when they are given meaning by translation to a typed internal language, the system must maintain the following property: for every  type, $\tau$, there must exist an internal type, $\sigma$, called its \emph{representation type}, such that the translation of every external term of type $\tau$ has internal type $\sigma$. This principle of \emph{representational consistency} arises essentially as a strengthening of the inductive hypothesis necessary to prove that all well-typed external terms translate to well-typed internal terms because $\lambda$ and successor are defined compositionally. It is closely related to the concept of \emph{type-preserving compilation} developed by Morrisett et al. for the TIL compiler for SML \cite{tarditi+:til-OLD}, here lifted into the language. Our judgements check this extension correctness property directly by typechecking each translation produced by a user extension (the translations of, for example, lambda terms will be inductively representationally consistent, so no additional check is needed).%That is, instead of being used as a necessary condition for compiler correctness, we are using it as a sufficient condition for type safety. 
%%We emphasize the distinction between translation (which in our calculus endows terms with a dynamic semantics) and compilation (which must preserve the dynamic semantics of terms). %In the presence of extensions, checking for representational consistency becomes subtle if we wish to guarantee conservativity, as we will discuss in the next subsection.
%
%%For the semantics to ensure that representational consistency is maintained by all operator definitions, we require, as briefly mentioned in Sec. \ref{types}, that each type constructor declare a \emph{representation schema} after the keyword \textsf{schema}. This is a type-level function of kind $\karrow{\kappaidx}{\kIType}$, where $\kappaidx$ is the index kind of the type constructor. The kind $\kIType$ has introductory form $\titype{\bar \sigma}$ and no elimination form, and is similar to the kind $\kITerm$ introduced above. There are two forms of \emph{translational internal type}, $\bar \sigma$, that do not correspond to forms in $\sigma$. As with translational internal terms, these are included to allow an internal type to be formed compositionally:
%%\begin{enumerate}
%%\item $\dup{\tau}$ splices in another translational internal type $\tau$
%%\item $\trepof{\tau}$ refers to the representation type of type $\tau$
%%\end{enumerate}
%%
%%When the semantics (i.e. the compiler) needs to determine the representation type of the type $\ttype{Tycon}{\tauidx}$ it simply applies the representation schema of $\fvar{Tycon}$ to ${\tauidx}$. Note that the representation type of a type cannot be extracted directly from within the type-level language, again for reasons that we will discuss below. 
%%
%%These additional forms are not needed by the representation schema of $\fvar{nat}$  because it is trivially indexed. In Fig. \ref{tuple}, we will discuss an example of the type constructor $\fvar{Tuple}$, implementing the semantics of $n$-tuples by translation to nested binary products. Here, the representation schema refers to the representations of the tuple's constituent types and is computed by folding over a list, so both of these are used\todo{add this example from appendix of ESOP}. 
%
%Translational internal terms can contain translational internal types. We see this in the definition of the recursor on natural numbers. The type assignment is the arbitrary type, $\tvar{t2}$. We cannot know what the representation type of $\tvar{t2}$ is, so we refer to it abstractly using $\trepof{\tvar{t2}}$. Luckily, the proof of representational consistency of this operator is parametric over the representation type of $\tvar{t2}$.

%\subsection{Metatheory}\label{metatheory}
%\newcommand{\F}[1]{\textsf{#1}~}
%\newcommand{\FF}[1]{\textsf{#1}}
%\newcommand{\Q}{\FF{Arg}}
%\renewcommand{\tnil}[1]{[]}
%\begin{figure}[t]
%\small
%$$\begin{array}{rcl} 
%%\textbf{programs} & 
%\rho & ::= & \F{tycon}\fvar{tycon}~\F{of}\kappaidx~{\{}TC{\}}; \rho \pipe e\\ %\pfam{\familyDf}{\progsort}
%%\\&  \pipe & \pdef{t}{\kappa}{\tau}{\progsort}  \pipe e
%%\pipe \pdef{t}{\kappa}{\tau}{\progsort} 
% TC & ::= & \F{iana} \{\tau\}; \F{isyn} \{\tau\}; \F{esyn} \{\tau\}; \F{rep} \{\tau\}\\
%\\ 
%e & ::= & \evar{x} \pipe \elam{\evar{x}}{\tau}{e} \pipe I \pipe I : \tau \pipe I :: \fvar{tycon} \pipe E\\
%I & ::= &  \FF{intro}[\tauidx](\splat{e}{1}{n})\\
%E & ::= & e\cdot\FF{elim}[\tauidx](\splat{e}{1}{n})\\
%\\
% \tau  & ::=   &   \tvar{t} \pipe 
%                           \tlam{t}{\kappa}{\tau} \pipe 
%                           \tapp{\tau_1}{\tau_2}  
%\\&\pipe&                      
%                           \tnil{\kappa} \pipe \tcons{\tau_1}{\tau_2} \pipe 
%                                      \tfold{\tau_1}{\tau_2}{h}{t}{r}{\tau_3}
%                           \\
%     & \pipe &     \ell \pipe 
%   \tunit \pipe 
%                           \tpair{\tau_{1}}{\tau_{2}} \pipe 
%                           \tfst{\tau} \pipe 
%                           \tsnd{\tau} \pipe \cdots 
%                           \\  
%%     & \pipe & \tinl{\kappa_2}{\tau_1} \pipe \tinr{\kappa_1}{\tau_2} \\
%%     &\pipe& \tsumcase{\tau}{t}{\tau_1}{t}{\tau_2} \\
%&\pipe &   \ttype{tycon}{\tauidx} \\&\pipe& \tfamcase{\tau}{tycon}{x}{\tau_1}{\tau_2}\\
%&\pipe &           \tifeq{\tau_{1}}{\tau_{2}}{\kappa}{\tau_{3}}{\tau_{4}} 
%%                            \\                          %       & & \pipe & \tfamcase{\tau}{Fam}{x}{\tau_1}{\tau_2}\\
%%                                                
%\\      &  \pipe & \tden{\tau_2}{\tau_1} \pipe \terr %\pipe \tden{\ibar}{\tau}^{\checkmark} 
%\pipe \ttypeof{\tau} \pipe \itransof{\tau} \\% \tdencase{\tau}{x}{t}{\tau_1}{\tau_2}\\
%% %& & \pipe & 
%%%                           \tdencase{\tau}{y}{x}{\tau_1}{\tau_2}
%%%                            \\
%%
% & \pipe & \titerm{\iota}
%     \pipe \titype{\sigma} \pipe \trepof{\tau} \\
%   & \pipe &  (\Gamma; e)? \pipe \FF{syn}(\tau) \pipe \FF{ana}(\tau; \tau')\\
%%\textbf{external terms}         & e & ::= & \evar{x} \pipe 
%%%                           \efix{x}{\tau}{e} \pipe 
%%                            \elam{\evar{x}}{\tau}{e} \pipe 
%%                            \eop{tycon}{op}{
%%                              \tauidx
%%                            }{
%%                                  \splat{e}{1}{n}
%%                            } \\
%%                  &     &   &   \\
%%
%%
%%                        
%%%\text{deabstracted}& \iota & ::= & \mathcal{G}[\iota, \sigma]\\
%%                        
%%\\
%%\\
%%              
%%% &  & \pipe & \tvalof{\tau_1}{\tau_2} \pipe \iup{\tau} \\
%%%                         \trepof{\tau} \pipe \dup{\tau}\\
%%\text{translational IL} & \bar{\iota} & ::= & x \pipe \ifix{x}{\bar \sigma}{\bar \iota} 
%%  %\pipe \ilam{x}{\bar \sigma}{\bar \iota} \pipe \iapp{\bar \iota_1}{\bar \iota_2} 
%%  \pipe \cdots \pipe \itransof{\tau} \\           
%% & \bar{\sigma} & ::= & \darrow{\bar \sigma_1}{\bar \sigma_2} \pipe \cdots \pipe \dup{\tau} \pipe \trepof{\tau} \\
%%%\text{abstracted} & \sabs & ::= & \darrow{\sabs_1}{\sabs_2} \pipe \cdots \pipe \sabsrep{\tau}
%%                      \\
%\kappa & ::= & \karrow{\kappa_1}{\kappa_2} \pipe \klist{\kappa} \pipe \FF{L} \pipe \kunit \pipe \kpair{\kappa_1}{\kappa_2} \pipe \cdots \\
% & \pipe & \kTypeBlur \pipe \FF{TT} \pipe \kITerm \pipe \kIType \pipe \Q
%%                          \klabel \pipe
%%                          \klist{\kappa} \pipe
%%                        \kunit \pipe 
%%                        \kpair{\kappa_{1}}{\kappa_{2}} \\
%%                        &&\pipe&
%%                        \ksum{\kappa_1}{\kappa_2} \pipe
%%                        \kTypeBlur \pipe \kDen \pipe 
%%                        \kIType               
%%%\textbf{ops signature}     & \Theta  & ::= & \kOpEmpty \pipe \kOp{\Theta}{op}{\kappai}\\
%%%                                   &   &   & \\
%\\
%\\
%\iota  & ::= & \evar{x} \pipe 
%                       \ifix{\evar{x}}{\sigma}{\iota} \pipe
%                       \ilam{\evar{x}}{\sigma}{\iota} \pipe 
%                       \iapp{\iota_{1}}{\iota_{2}} 
%                       \\ & \pipe & \iunit \pipe \ipair{\iota_1}{\iota_2} \pipe \ifst{\iota} \pipe \isnd{\iota} \pipe \cdots \pipe \iup{\tau}
%%\\&\pipe&               \cdots \pipe  
%%                        n \pipe \iop{\iota_{1}}{\iota_{2}} \pipe \iIfEq{\iota_{1}}{\iota_{2}}{\dint}{\iota_{3}}{\iota_{4}} \\
%%& \pipe & 
%%                        \iunit \pipe
%%                        \ipair{\iota_{1}}{\iota_{2}} \pipe 
%%                        \ifst{\iota} \pipe
%%                        \isnd{\iota} 
%%\\&\pipe&                       
%%                         \iinl{\sigma_2}{\iota_1} \pipe \iinr{\sigma_1}{\iota_2} \\&\pipe & \icase{e}{x}{e_1}{x}{e_2} \\
%\\ \sigma  & ::= &    \darrow{\sigma_1}{\sigma_2} \pipe \dunit \pipe \dpair{\sigma_1}{\sigma_2} %\pipe \dint \pipe \dunit \pipe \dpair{\sigma_1}{\sigma_2} \pipe \dsum{\sigma_1}{\sigma_2}  
%\pipe \cdots \pipe \dup{\tau}
%\end{array}$$
%%\vspace{-10pt}
%\caption{\small Syntax of Core $\lam@\texttt{lang}$. Here, $x$ ranges over external and internal language variables, $\tvar{t}$ ranges over type-level variables, $\fvar{tycon}$ ranges over type constructor names and $\ell$ ranges over labels.
%\label{grammar}}
%\end{figure}
%\begin{figure}
%\small
%\begin{flalign}
%& \F{tycon}\fvar{record}~\F{of}\klist{\kpair{\FF{L}}{\kTypeBlur}}~\{\\
%& \quad \FF{iana}~\{\tlam{i}{\kpair{\klist{\kpair{\FF{L}}{\kTypeBlur}}}{\klist{\FF{L}}}}{\tlam{a}{\klist{\Q}}{\\& \quad\quad\tfold{\tvar{zip3}~\tfst{\tvar{i}}~\tsnd{\tvar{i}}~\tvar{a}}
%  {\titerm{\tunit}}{h}{t}{r}{
% \\&\quad\quad\quad\tifeq{\tfst{\tvar{first}~\tvar{h}}}{\tvar{second}~\tvar{h}}{\FF{L}}{
% \\&\quad\quad\quad\quad
% \F{let}~\tvar{x}:\kITerm~=~\FF{ana}(\tvar{third}~\tvar{h}; \tsnd{\tvar{first}~\tvar{h}})
%~\FF{in}
%\\ & \quad\quad\quad\quad \tfold{\tvar{t}}{\tvar{x}
%}{\_}{\_}{\_}{\titerm{(\iup{\tvar{x}},\iup{\tvar{r}})}}  }{\terr}
%%     \F{let}~\tvar{tt}:\FF{TT}=~\FF{syn}(\tsnd{\tsnd{\tvar{h}}}) \F{in}\\
%%\\     & \quad\quad\quad 
%}
%}}\}\\
%%& \quad \FF{isyn}~\{\tlam{i}{\klist{\FF{L}}}{\tlam{a}{\klist{\Q}}{
%%\\&\quad\quad \tfold{\tvar{zip2}~\tvar{i}~\tvar{a}}{\tden{\titerm{()}}{\ttype{record}{\tnil{{\kpair{\FF{L}}{\kTypeBlur}}}}}}{h}{t}{r}{
%%\\&\quad\quad\quad \F{let}~\tvar{htt}:\FF{TT}~=~\FF{syn}(\tsnd{\tvar{h}})
%%~\FF{in}
%%\\&\quad\quad\quad \F{let}~\tvar{hty}:\kTypeBlur~=~\ttypeof{\tvar{htt}}
%%\\&\quad\quad\quad \F{let}~\tvar{ty}:\kTypeBlur~=~\ttype{record}{(\tfst{\tvar{h}}, \tvar{hty})::\tnil{{\kpair{\FF{L}}{\kTypeBlur}}}}
%%~\FF{in}
%%\\&\quad\quad\quad \tfold{\tvar t}{\tden{\itransof{\tvar{htt}}}{\tvar{ty}}}{\_}{\_}{\_}{
%%\\&\quad\quad\quad\quad \tden{\titerm{(\iup{\itransof{\tvar{htt}}},\iup{\itransof{\tvar{r}}})}}{\tvar{ty}}}}
%%}}\}\\
%%& \quad \FF{esyn}~\{\tlam{i}{\FF{L}}{\tlam{a}{\klist{\Q}}{
%%  \tvar{decons1}~\tvar{a}~\tlam{ty}{\kTypeBlur}{\tlam{x}{\kITerm}{
%%\\&\quad\quad   \tfamcase{\tvar{ty}}{record}{sig}{
%%\\&\quad\quad\quad \tfold{\tvar{sig}}{\terr}{h}{t}{r}{
%%\\&\quad\quad\quad\quad \tifeq{\tfst{\tvar{h}}}{\tvar{i}}{\FF{L}}{\tfold{\tvar{t}}{\tden{\tvar{x}}{\tsnd{\tvar{h}}}}{\_}{t'}{\_}{
%%\\&\quad\quad\quad\quad\quad \tfold{\tvar{t'}}{\tden{\titerm{\ifst{\iup{\tvar{x}}}}}{\tsnd{\tvar{h}}}}{\_}{\_}{\tvar{r'}}{
%%\\&\quad\quad\quad\quad\quad\quad \tden{\titerm{\isnd{\iup{{\itransof{\tvar r}}}}}}{\tsnd{\tvar{h}}}}}\\&\quad\quad\quad\quad\hspace{-3px}}{\tvar{r}}
%%}\\&\quad\quad\hspace{-3px}}{\terr}
%%  }}
%%}}\\
%%& \quad \FF{rep}~\{\tlam{i}{\klist{\kpair{\FF{L}}{\kTypeBlur}}}{  \\& \quad \quad \tfold{\tvar{i}}{\titype{\dunit}}{s}{j}{r}{
%%    \tfold{\tvar{j}}{\trepof{\tsnd{\tvar{s}}}}{\_}{\_}{\_}{\\
%%    &\quad\quad\quad
%%    \titype{\dpair{\dup{\trepof{\tsnd{\tvar{s}}}}}{\dup{\tvar{r}}}}
%%    }
%%  }}\}\\
%%& \}; \\
%%& \F{let}\tvar{R} : \kTypeBlur = \ttype{record}{(\ell_1, \ttype{record}{\tnil{{\FF{L}\times\kTypeBlur}}}) :: \tnil{{\FF{L}\times\kTypeBlur}}}~\F{in}\\
%%&\F{let}id = \elam{x}{\tvar{R}}{\{\ell_1=~x\cdot\ell_1\} :: \fvar{record}}\\
%%&\F{let}triv = id[\{\ell_1 = \{\}\}]\cdot\ell_1
%%%&\F{let}one = \FF{intro}[\ell_\text{succ}]
%%%&\F{let}plus = \elam{x}{\tvar{nat}}{\elam{y}{\tvar{nat}}{\FF{elim}[()](x; y; \\
%%%& \quad \elam{r}{\tvar{nat}}{intro[\ell_\text{succ}](r) : \tvar{nat}}})}
%\end{flalign}
%\caption{The definition of the record type in  using the desugarings in Figure \ref{desugaring} and with the addition of simple let bindings for both type and term variables (not shown). Some type-level helper functions also omitted for concision.}
%
%\label{record-theory}
%\end{figure}
%We will now give a core typed lambda calculus, $\lam@\texttt{lang}$, that captures the semantics described in the previous sections. It is intended to make precise how active typechecking and translation works and how our mechanism relates to existing work on bidirectional typechecking, type-level computation and typed compilation while abstracting away from the details of Python's syntax and imposing a stronger type-level semantics that will allow us to state metatheoretic properties of interest. We will assume a fixed base for functions (providing the standard semantics of lambda functions) and target language, which we here call the \emph{internal language}.
%
%The syntax of $\lam@\texttt{lang}$ is shown in Figure \ref{grammar} and an example \emph{program} that defines a type constructor, $\fvar{record}$, with a semantics similar to that given in Listing \ref{record}\footnote{Technically, this defines labeled tuples, because the order of labels matters.}, using it to write an identity function and compute the value $triv$ (the empty record), is shown in Figure \ref{record-theory}.
%
%%In this section, we will develop an ``actively typed'' version of the simply-typed lambda calculus with simply-kinded type-level computation called $\lam@\texttt{lang}$. More specifically, the level of types, $\tau$, will itself form a simply-typed lambda calculus. \emph{Kinds} classify type-level terms in the same way that types conventionally classify expressions. Types become just one kind  of type-level value (which we will write $\kTypeBlur$, though it is also variously written $\star$, \lstinlinep{T} and \lstinlinep{Type} in various settings). Rather than there being a fixed set of type and operator constructors, we allow the programmer to declare new  constructors, and give their static and dynamic semantics by writing type-level functions. The kind system combined with techniques borrowed from the typed compilation literature and a form of type abstraction will allow us to prove strong type safety, decidability and conservativity theorems.
%%
%%The syntax of Core $\lam@\texttt{lang}$ is given in Fig. \ref{grammar}. An example of a program defining type and operator constructors that can be used to construct an active embedding of G\"odel's \textbf{T} into $\lam@\texttt{lang}$ is given in Fig. \ref{nat}. We will discuss its semantics and how precisely the embedding, seen being used starting on line 15 to ultimately compute the sum of two and two, works as we go on. Natural numbers can, of course, be isomorphically embedded in existing languages, with a similar usage and asymptotic performance profile (up to function call overhead as an abstract type, for example). We will provide more sophisticated examples where this is less feasible later on (and note that type abstraction is an orthogonal mechanism).
%%a type constructor declaration, $\fvar{nat}$, indexed trivially, together with three operator constructors, also all indexed trivially, that implement the standard introductory forms for natural numbers as well as the recursor operator (as in G\"odel's T \cite{pfpl}). Following the type constructor declaration, we apply $\fvar{nat}$ with the trivial index, $\tunit$, to form the type $\tvar{nat}$. Finally, we write an external term that uses the operators associated with $\fvar{nat}$ and the built-in constructor $\fvar{Parr}$, governing partial functions, to define an addition function and compute the addition of the natural numbers  two and two. We will introduce a more convenient concrete syntax in later portions of this thesis; for now we will restrict ourselves to the abstract syntax so that this example can directly aid in understanding the semantics.
%
%\subsection{Overview}\label{programs}
%
%A \emph{program}, $\rho$, consists of a series of constructor declarations followed by an external term, $e$.  Compiling a program consists of first \emph{kind checking} it (see below), then typechecking the external term and simultaneously {translating} it to a term, $\iota$, in the {typed internal language}. 
%%These correspond to the premises of the \emph{central compilation judgement} $\pkcompiles{\rho}{\iota}$:
%%\[
%%\inferrule[p-compiles]{
%%  \emptyset \vdash_{\fvalCtx_0} \rho\\
%%% \progOK{\emptyset}{\fvalCtx_0}{\rho}\\
%%  \pcompiles{\fvalCtx_0}{\rho}{\iota}
%%}{\pkcompiles{\rho}{\iota}}
%%\]
%%This anchors our exposition; we will describe how it is derived (i.e. how to write a compiler for $\lam@\texttt{lang}$) in the following sections. 
%The key judgements are the \emph{bidirectional active typing judgements}  (Fig. \ref{att}, which we describe starting in Sec. \ref{opcons}). They relate an external term, $e$, to a {type}, $\tau$, called its \emph{type assignment}, and an internal term, $\iota$, called its \emph{translation}, under \emph{typing context} $\Gamma$ and \emph{constructor context} $\Phi$. The first is synthesis, the second analysis.
%\[\Gamma \vdash_\fvalCtx e \Rightarrow \tau \leadsto \iota
%~~~~~~~
%\Gamma \vdash_\fvalCtx e \Leftarrow \tau \leadsto \iota\]
%%\[\ecompilesAX{e}{\tau}{\iota}\]
%
%The typing context, $\Gamma$, maps variables to types in essentially the conventional way (\cite{pfpl} contains the necessary background for this section). The constructor context, $\Phi$, tracks user-defined type  constructors. % Weakening and exchange of the constructor context closely related to the issue of conservativity that we will return to. Each constructor is identified by name, so there is no analog to contraction.
%
%The dynamic behavior of an external term is determined entirely by its translation to the internal language, which has a conventional operational semantics. This form of semantics can be seen as lifting into the language specification the first stage of a type-directed compiler like the TIL compiler for Standard ML \cite{tarditi+:til-OLD} and has some parallels to the Harper-Stone semantics for Standard ML, where external terms were also given meaning by elaboration from the EL to an IL \cite{Harper00atype-theoretic}. %
%
%In $\lam@\texttt{lang}$, the internal language (IL) provides partial functions (via the generic fixpoint operator of Plotkin's PCF) and simple product  types for the sake of our example. In practice, the internal language could be any typed  language with a specification for which type safety and decidability of typechecking have been satisfyingly determined. The internal type system serves as a ``floor'': guarantees that must hold for terms of any type (e.g. that out-of-bounds access to memory never occurs) must be maintained by the internal type system. User-defined constructors can enforce invariants stronger than those the internal type system maintains at particular types, however. Performance is also ultimately limited by the internal language and downstream compilation stages that we do not here consider (safe compiler extension has been discussed in previous work, e.g. \cite{conf/pldi/TatlockL10}).
%
%
%\subsection{Types and Type-Level Computation}\label{types}
%$\lam@\texttt{lang}$ supports, and makes extensive use of, simply-kinded type-level computation. Specifically, type-level terms, $\tau$, themselves form a typed lambda calculus. The classifiers of type-level terms are called \emph{kinds}, $\kappa$, to distinguish them from  \emph{types}. Types are  type-level values of kind $\kTypeBlur$. In @\texttt{lang}, the type-level language (together with the level of programs) is written in Python, which can be thought of as having a rather limited kind system (with one kind, \lstinlinep{dyn}). Here, we are able to more precisely discuss the kinds of values in the type-level language. 
%
%In most languages, types are formed by applying one of a collection of \emph{type constructors} to zero or more \emph{indices}. In $\lam@\texttt{lang}$, the situation is notionally similar. User-defined type constructors can be declared at the top of a program (or lifted to the top, in practice) using \textsf{tycon}. Each constructor in the program must have a unique name, written e.g. \fvar{record}.\footnote{We assume naming conflicts can be avoided by some extrinsic mechanism.} A type constructor must also declare an \emph{index kind}, $\kappaidx$. 
%
% To permit the embedding of interesting type systems, the type-level language includes several kinds other than $\kTypeBlur$. We lift several functional data structures to the type level: here, only unit ($\kunit$), binary products ($\kpair{\kappa_1}{\kappa_2}$) and lists ($\klist{\kappa}$), in addition to labels (introduced as $\ell$, possibly with a subscript, having kind \FF{L}).  Our record type constructor is indexed by a list of pairs of {labels}  and types (a signature, in essence; line 1). The type constructor $\fvar{arrow}$ is included in the initial constructor context, $\fvalCtx_0$ and has index kind $\kpair{\kTypeBlur}{\kTypeBlur}$.
% In practice, one could include a richer functional programming language and retain the spirit of the calculus, as long as it does not introduce general recursion at the type level. 
%  
%A type is introduced by applying a type constructor to an index of this kind, written $\ttype{Tycon}{\tauidx}$.  For example, the type of natural numbers is indexed trivially, so it is written $\ttype{Nat}{\tunit}$. We see a record type, abbreviated $\tvar{R}$, constructed on line 26. 
%
%The kind $\kTypeBlur$ also has an elimination form, $$\tfamcase{\tau}{Tycon}{x}{\tau_1}{\tau_2}$$ allowing the extraction of a type index by case analysis against a contextually-available type constructor. To a first approximation, one might think of type constructors as constructors of a built-in open \cite{conf/ppdp/LohH06}, $\kTypeBlur$, at the type-level. Like open datatypes, there is no notion of exhaustiveness so the default case is required for totality. We will see where this is used shortly.
%
%
%
%%We will write the kind of types as $\kTypeBlur$, though it is also written $\star$ or \lstinlinep{Type} in various similarly structured languages (see Sec. \ref{related-work}). 
%% Rather than there being a fixed set of type constructors, we allow the programmer to declare new type  constructors, and give the static and dynamic semantics of their associated operators, by writing type-level functions. In the semantics for this calculus, our kind system combined with techniques borrowed from the typed compilation literature and a form of type abstraction allow us to prove strong type safety, decidability and conservativity theorems.
%
%
%
%% and integers ($\dint$). We also include labels ($\klabel$), written in a slanted font, e.g. $\tlabel{myLabel}$, which are string-like values that only support comparison and play a distinguished role in the expanded syntax, as we will later discuss. Our first example, $\fvar{nat}$, is indexed trivially, i.e. by unit kind, $\kunit$, so there is only one natural number type, $\ttype{nat}{\tunit}$, but we will show examples of type constructors that are indexed in more interesting ways in later portions of this work. For example, $\fvar{LabeledTuple}$ has index kind $\klist{\kpair{\klabel}{\kTypeBlur}}$. 
% 
% Type constructors are not first-class; they do not themselves have arrow kind as in some kind systems (e.g.  \cite{watkins2008specifying}; Ch. 22 of \emph{PFPL} describes a related system \cite{pfpl}). The type-level language does, however, include total functions of arrow kind, written $\karrow{\kappa_1}{\kappa_2}$. Type constructor application can be wrapped in a type-level function to emulate a first-class or uncurried version of a type constructor for convenience (indeed, such a wrapper could be generated automatically, though we do not do so). 
%
%Two type-level terms of kind $\kTypeBlur$ are equivalent if they apply the same constructor, identified by name, to equivalent indices. Going further, we ensure that deciding type equivalence requires only checking for syntactic equality after normalization by imposing the restriction that equivalence at a type constructor's index kind must be decidable in this way. Our treatment of equivalence in the type-level language is thus quite similar to the treatment of term-level equality using ``equality types'' in a language like Standard ML.
%% A kind $\kappa$ is an  \emph{equality kind} if $\kEq{\kappa}$ can be derived (see appendix). 
%Conditional branching on the basis of equality at an equality kind can be performed in the type-level language. Equivalence at arrow kind is not decidable by our criteria, so type-level functions cannot appear within type indices. This also prevents general recursion from arising at the type level. Without this restriction, a type-level function taking a type as an argument could ``smuggle in'' a self reference as a type index, extracting it via case analysis (continuing our analogy to open datatypes, this is closely related to the positivity condition for inductive datatypes in total functional languages like Coq).% as maintaining the metatheoretic guarantee that typing respects type equivalence would impose a substantial burden in such a setting.% (a na\"ive approach to this would impose non-trivial extrinsic proof obligations onto extension developers that, unlike in others in this thesis, could threaten type safety).
%
%Every type constructor also defines a \emph{representation}, a type-level function that associates with every type an internal \emph{representation type} (analagous to \lstinlinep{trans_type} above). We will return to this after introducing the external forms and operator definitions.
%
%\subsection{Core External Forms and Desugaring}\label{opcons}
%The syntax for external terms (Figure \ref{grammar}) contains variables, $\lambda$ terms, three generalized introductory forms and a single generalized elimination form. The introductory forms are either unascribed, ascribed with a type or ascribed with a type constructor, as in our discussion of @\texttt{lang}. These generalized forms take a single a type-level value as an {index} and $n \geq 0$ arguments, which are other external terms. To better motivate this choice, we can give a purely syntactic desugaring of a Python-like syntax with labels to these forms, shown in Figure \ref{desugaring}. It is instructive to rewrite lines 27-28 of Figure \ref{record-theory} using these desugarings:% In @\texttt{lang}, desugarings can be user-defined 
%\noindent
%\vspace{-5px}
%$$
%\begin{array}{l}
%\FF{let}~id=\lambda x:\tvar{R}.\FF{intro}[\ell_1 :: []](x\cdot\FF{elim}[\ell_1]()) :: \fvar{record}\\
%\FF{let}~triv=id\cdot\FF{elim}[()](\FF{intro}[\ell_1 :: []](\FF{intro}[[]]()))\cdot\FF{elim}[\ell_1]()
%\end{array}
%$$
%\vspace{-5px}
%\begin{figure}[t]
%\small
%\[
%\begin{array}{rcl}
%\{ \ell_1{=}~e_1, \ldots, \ell_n{=}~e_n \} & := & \FF{intro}[\ell_1 :: \ldots :: \ell_n :: []](e_1; \ldots; e_n)\\
%(e_1, ..., e_n) & := & \FF{intro}[()](e_1; \ldots; e_n)\\
%\#n & := & \FF{intro}[n]()\\
%"s" & := & \FF{intro}[s]()\\
%e.\ell & := & e\cdot\FF{elim}[\ell]()\\
%e[e_1; \ldots; e_n] & := & e\cdot\FF{elim}[()](e_1; \ldots; e_n)\\
%e.\ell(e_1; ...; e_n) & := & e\cdot\FF{elim}[\ell](e_1; \ldots; e_n)
%%(e_1, ..., e_n) & := & intro[()](e_1; ...; e_n)
%%n &  := & intro[n]()
%%s & := & intro[s]()
%%
%%e.l & := & elim[l](e)
%%e[e_1; ...; e_n] & := & elim[()](e_1, ..., e_n)
%%e.l(e_1; ...; e_n) & := & elim[l]
%\end{array}
%\]
%\caption{Desugaring from conventional syntax to core forms. We assume that the type-level language has numbers, $n$, and strings, $s$ (not shown for concision).}
%\label{desugaring}
%\end{figure}
%
%%User-defined operator constructors are declared using \textsf{opcon}.  For reasons that we will discuss, our calculus associates every operator  constructor with a type constructor. The \emph{fully-qualified name} of every operator constructor, e.g. $\fvar{nat}.\opvar{z}$, must be unique. Operator constructors, like type constructors, declare an index kind, $\kappaidx$. In our first example, all the operator constructors are indexed trivially (by index kind $\kunit$), but other examples use more interesting indices. For example, in SML, the projection operator \lstinlinep{#3} can be applied to an $n$-tuple, $e$, iff $n \geq 3$. Note that it thus cannot be a function with a standard arrow type. Notionally, \lstinlinep{#} is an operator constructor and \lstinlinep{3} is its index. In an active embedding of $n$-tuples into $\lam@\texttt{lang}$, this would be written $\eop{Tuple}{prj}{3}{e}$ (we will nearly recover ML's syntax later). $\fvar{LabeledTuple}.\opvar{prj}$ is the operator constructor used to access a field of a labeled tuple, so it has index kind $\klabel$. An operator itself is, notionally, selected by indexing an operator constructor, e.g. $\fvar{nat}.\opvar{s}\langle \tunit \rangle$, but technically neither operator constructors nor operators are first-class at any level (additional machinery would be needed, e.g. an \textsf{Op} kind, but this is not fundamental to our calculus). 
%%Instead, in the external language, an operator constructor is applied by simultaneously providing an index and  $n \geq 0$ \emph{arguments}, written $\eop{Tycon}{op}{\tauidx}{\splat{e}{1}{n}}$\footnote{It may be helpful to distinguish between type/operator constructors and \emph{term formers}. There are term formers at all levels in the calculus. For example, operator constructor application and $\lambda$ are  external term formers, and type constructor application is a type-level term former. We might write these following Harper's conventions for abstract syntax to highlight this distinction \cite{pfpl}: $\mathtt{lam}[\tau](x.e)$, $\mathtt{ocapp}[\fvar{Tycon}, \opvar{op}, \tauidx](\splat{e}{1}{n})$ and $\texttt{tcapp}[\fvar{Tycon}](\tau)$.}. For example, on line 18 of Fig. \ref{nat}, we see the operator constructors $\fvar{nat}.\opvar{z}$ and $\fvar{nat}.\opvar{s}$ being applied to compute $two$. %\footnote{Although our focus here is entirely on semantics, a brief note on syntax: in the expanded syntax, the trivial indices and empty argument lists can be omitted, so we could write \texttt{Nat.s(Nat.s(Nat.z))}. With the ability to ``open'' a type's operators into the context, we could shorten this still to \texttt{s(s(z))}. Alternatively, with the ability to define a TSL in a manner similar to that in Sec. \ref{aparsing}, we might instead just write \texttt{2}.}
%
%\subsection{Operator Definitions and Representational Consistency}
%
%%
%\newcommand{\atjsynX}[3]{\Gamma \vdash_\fvalCtx #1 \Rightarrow #2 \leadsto #3}
%\newcommand{\atjanaX}[3]{\Gamma \vdash_\fvalCtx #1 \Leftarrow #2 \leadsto #3}
%\begin{figure}[t]
%\small
%$\fbox{\inferrule{}{\ecompilesAX{e}{\tau}{\iota}}}$
%~~~~$\Gamma ::= \emptyset \pipe \Gamma, x \Rightarrow \tau$
%\begin{mathpar}
%\inferrule[att-var]{
% x \Rightarrow \tau \in \Gamma
%}{
% \atjsynX{x}{\tau}{x}
%}
%
%\inferrule[att-syn-to-ana]{
% \atjsynX{e}{\tau}{\iota}
%}{
% \atjanaX{e}{\tau}{\iota}
%}
%
%\inferrule[att-lam]{
% \tau_1 \Downarrow_{\Gamma;\fvalCtx} \tau_1'\\
%   \trepof{\tau_1'} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\\\
% \Gamma, x \Rightarrow \tau_1' \vdash_\fvalCtx e \Rightarrow \tau_2 \leadsto \iota
%%  \iota \hookrightarrow_\fvalCtx \iota'\\
%%    \sigma \hookrightarrow_\fvalCtx \sigma'
%%  \ddbar{\fvar{Arrow}}{\fvalCtx}{\trepof{\tau_1'}}{\sbar_1}\\
% %\delfromtau{$\Xi_0$}{\fvalCtx}{\tau_1'}{\sabs}\\\\
%}{
% \atjsynX{\elam{x}{\tau_1}{e}}{\ttype{arrow}{(\tau_1', \tau_2)}}{\ilam{x}{\sigma'}{\iota}}
%}
%
%\inferrule[att-i-unasc]{
% \vdash_\fvalCtx \FF{iana}(\fvar{tycon})=\taudef\\
% \taudef~(\tauidx, \tauidx')~((\Gamma; e_1)? :: \ldots :: (\Gamma; e_n)? :: []) \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota}\\
%%  \trepof{\ttype{tycon}{\tauidx'}} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\
%   \trepof{\ttype{tycon}{\tauidx'}} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\
%   \Gamma \vdash_\fvalCtx \iota : \sigma
%}{
% \atjanaX{\FF{intro}[\tauidx](e_1; \ldots; e_n)}{\ttype{tycon}{\tauidx'}}{\iota}
%}
%
%\inferrule[att-i-asc-ty]{
% \atjanaX{I}{\tau}{\iota}
%}{
% \atjsynX{I : \tau}{\tau}{\iota}
%}
%
%\inferrule[att-i-asc-tycon]{
% \vdash_\fvalCtx \FF{isyn}(\fvar{tycon})=\taudef\\\\
% \taudef~\tauidx~((\Gamma; e_1)? {::}{\ldots}{::}(\Gamma; e_n)? {::} []) \Downarrow_{\Gamma;\fvalCtx} \tden{\titerm{\iota}}{\ttype{tycon}{\tauidx'}}\\
%     \trepof{\tau_1'} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\
% \Gamma \vdash_\fvalCtx \iota : \sigma
%}{
% \atjsynX{\FF{intro}[\tauidx](e_1; \ldots; e_n)] :: \fvar{tycon}}{\tau}{\iota'}
%}
%
%\inferrule[att-elim]{
% \atjsynX{e}{\ttype{tycon}{-}}{-}\\
% \vdash_\fvalCtx \FF{esyn}(\fvar{tycon})=\taudef\\
% \taudef~((\Gamma; e)? :: (\Gamma; e_1)? :: \ldots :: (\Gamma; e_n)? :: []) \Downarrow_{\Gamma;\fvalCtx} \tden{\tau}{\iota}\\
%     \trepof{\tau} \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}\\
% \Gamma \vdash_\fvalCtx \iota : \sigma
%}{
% \atjsynX{e\cdot\FF{elim}[\tauidx](e_1; \ldots; e_n)}{\tau}{\iota}
%}
%\end{mathpar}
%\caption{\small The active typing judgement. The normalization judgement for type-level terms ($\Downarrow$) and the representational consistency check will be in an appendix.}
%\label{atj}
%\end{figure}
%\begin{figure}
%\small
%\begin{mathpar}
%\inferrule[repof]{
% \tau \Downarrow_{\Gamma;\fvalCtx} \ttype{tycon}{\tauidx}\\
% \vdash_\fvalCtx \FF{rep}(\fvar{tycon}) = \taurep\\
% \taurep \tauidx \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}
%}{
% \FF{repof}(\tau) \Downarrow_{\Gamma;\fvalCtx} \titype{\sigma}
%}
%
%\inferrule[syn]{
%    \tau \Downarrow_{\Gamma;\fvalCtx} (\Gamma; e)?\\
% \atjsynX{e}{\tau}{\iota}
%}{
% \FF{syn}(\tau) \Downarrow_{\Gamma;\fvalCtx} \tden{\titerm{\iota}}{\tau}
%}
%
%\inferrule[ana]{
% \tau_1 \Downarrow_{\Gamma;\fvalCtx} (\Gamma; e)?\\
% \tau_2 \Downarrow_{\Gamma;\fvalCtx} \ttype{tycon}{\tauidx}\\\\
% \atjanaX{e}{\ttype{tycon}{\tauidx}}{\iota}\\
%}{
% \FF{ana}(\tau_1; \tau_2) \Downarrow_{\Gamma;\fvalCtx} \titerm{\iota}
%}
%\end{mathpar}
%\caption{\small Evaluation semantics for the type-level language. Missing rules (including error propagation rules, which immediately cause failure of typechecking) are unsurprising and will be given in an appendix.}
%\label{tleval}
%\end{figure}
%The expressive and metatheoretic power of the calculus arises from how the rules for the active typing judgement handle these  generalized forms. Rather than fixing the specification of a finite collection of operator constructors and tasking the \emph{compiler} with deciding a typing derivation on its basis, the specification instead delegates to a type-level function associated with a type constructor. In the core calculus, this is one of three functions, called \FF{iana}, \FF{isyn} and \FF{esyn} in the grammar.\footnote{This means that any particular type constructor only supports a single introduction and elimination desugaring. This is a minor inconvenience in some cases that is resolved in @\texttt{lang} by the use of methods.} We see how this is done with the rules for the active typing judgements, given in Figure \ref{atj}.
%
%The first two rules are standard in bidirectional type systems: variables synthesize types and if a term synthesizes a type, it can be analyzed against a type (cf. Listing \ref{context}). Variables translate to variables, and if we are simply converting from synthesis to analysis, translation is not affected.
%
%The rule $\rulename{att-lam}$ must take into account the fact that, because we support type-level computation, the type annotation on the argument may not be in normal form. Thus, we evaluate it to normal form. Note that the kinding rules (not shown) will guarantee that, because $\tau_1$ is of kind $\kTypeBlur$, its normal form, $\tau'_1$, is of the form $\ttype{tycon}{\tauidx}$ for some normal $\tauidx$. To generate an appropriate internal type annotation in the translation, we need to compute the representation type of $\tau_1'$. This involves calling the representation function associated with its type constructor (rule \fvar{repof} in Figure \ref{tleval}). The form $\titerm{\sigma}$ is a \emph{quoted internal type}. Note in the syntax that there is an unquote form, $\dup{\tau}$, which allows us to compose internal types compositionally without needing to expose an elimination form. Indeed, it is an interesting facet of our calculus that we never need to examine syntax trees directly to implement extensions. The evaluation semantics remove quotations, so the normal form of a quoted internal type contains an internal type with no quotations.
%
%Lambda functions are the only introductory form requiring special support in the calculus (because they need to  manipulate the context; see Discussion). The next three rules show how any other abstractions that we define (e.g. records, decimals, etc.) make use of a generalized introductory form. 
%
%The rule  $\rulename{att-i-unasc}$ shows that unascribed introductory forms can only be analyzed against a type. Given such a type, the rule extracts a definition, named $\FF{iana}$, from the type constructor (cf. the \lstinlinep{ana_Dict} and \lstinlinep{trans_Dict} methods earlier). It calls this function with a pair containing the operator and type index and a list of \emph{reified arguments}, which have kind $\Q$ and introductory form $(\Gamma; e)?$. These are only constructed by the compiler (there would be no corresponding form in the concrete syntax). Their purpose is to allow the definition to programatically invoke synthesis and analysis as needed using the $\FF{ana}$ and $\FF{syn}$ operators. We can see the $\fvar{record}$ type constructor doing so on line 5 of Figure \ref{record-theory} to ensure that the field value provided as an argument has the same type as the corresponding label (accomplished by simultaneously folding over all three pieces of input data). The rule for performing analysis, $\fvar{ana}$, is in Figure \ref{tleval}. If it succeeds, analysis returns a \emph{translation}, which is a quoted internal term with kind $\kITerm$ and introductory form $\titerm{\iota}$. Like quoted internal types, there is an unquote form that is eliminated during evaluation. The operator definition uses this to construct a translation for the record. As before, empty records translate to units and records with a single field are unadorned. In this example, records with two or more fields translate to nested tuples. 
%
%Because we have a representation type associated with the type, we can check that the translation is \emph{representationally consistent} (the final premise). As we will discuss, representational consistency combined with type safety of the internal language implies type safety overall (it arises as a strengthening of the inductive hypothesis needed to prove type safety, and is closely related to work on \emph{type-preserving compilation} in the TIL compiler for Standard ML, \cite{tarditi+:til-OLD}).
%
%The next rule, $\rulename{att-i-asc-ty}$ states that introductory forms ascribed with a type are analyzed against that type.
%
%Introductory forms ascribed with a type constructor can, according to the final introductory rule, $\rulename{att-i-asc-tycon}$, synthesize a type via the definition $\FF{isyn}$. It is passed the operator index (there is no type index, since we only have a type constructor) and a list of arguments, as before. In this case, it must return not just a translation, but also a type. We use the form $\tden{\tau_2}{\tau_1}$ for such a pairing, which has kind $\FF{TT}$. The type and translation can be extracted from it using the appropriately named elimination forms (indeed, as given, it is merely a pair, but we give it special syntax for clarity -- it looks like the conclusion of the typing judgement -- and for other reasons that will become clear in future work). To synthesize a type from a list of labels and arguments, we must be able to synthesize types from the arguments. The $\FF{syn}(\tau)$ operator permits this, per rule $\fvar{syn}$, seen being used in Listing \ref{record-theory}. We check representational consistency of the result, as before.
%
%Finally, we show the rule for elimination forms. It operates by first synthesizing a type for the ``primary'' subterm, then extracting the $\FF{esyn}$ definition from its type constructor. As before, it is called with the arguments and representational consistency is checked. Note that the primary subterm is itself the first argument (though we have already synthesized a type for it, it is more uniform and clear to allow the operator  to do so again, which is done in our example by the $\tvar{decons1}$ helper function).
%
%\subsubsection{Representational Consistency}
%The representational consistency lemmas says that active typechecking and translation of an external term, if it succeeds (producing a type, $\hat\sigma$ and a translation $\iota$), always produces a well-typed translation. More specifically, the translation's type is the representation type of $\hat\sigma$.
%
%\begin{theorem}[Representational Consistency]
%Given well-kinded contexts and an external term that is both well-kinded and well-typed:
%\begin{enumerate}
%\item $\vdash \Phi$ 
%\item $\vdash_\Phi \Gamma$
%\item $\vdash_\Phi e$
%\item Either:
%\begin{enumerate}
%\item[a.] $\Gamma \vdash_\Phi e \Leftarrow \hat{\sigma} \leadsto \iota$
%\item[b.] $\Gamma \vdash_\Phi e \Rightarrow \hat{\sigma} \leadsto \iota$
%\end{enumerate}
%\end{enumerate}
%there exists an internal typing context, $\Omega$, and an internal type, $\tau$, such that  $\vdash_\Phi \Gamma \leadsto \Omega$ and $\FF{rep}(\hat\sigma) \Downarrow_\Phi \titype{\tau}$ and $\Omega \vdash \iota : \tau$.
%\end{theorem}
%\begin{proof}
%The proof proceeds by induction on the typing derivation (4a or 4b). Because the representational consistency is explicitly checked in the rules for intro and elim forms, these cases follow directly. The remaining cases are standard and follow inductively, given standard lemmas about valid contexts and a lemma about type safety of the static language. This is a fixed, standard functional language with only a few simple additions (TODO detail this).
%\end{proof}
%
%\subsubsection{Type Safety}
%Representational safety implies that well-typed external terms produce well-typed internal terms. If the internal language is type safe, then we know that well-typed external terms do not ``go wrong''. The internal language is simply PCF with sums and products, and a base type of integers, so this is a completely standard result \cite{pfpl}.

%\subsubsection{Decidability}
%The metatheoretic properties of interest are: type safety and termination of the type-level language (guaranteed by our kind system, though the details will need to be provided in an appendix; termination is non-trivial), type safety and decidability of the internal language (it i.s a standard variant of PCF, so this is trivial) and representational consistency. Type safety for the language as a whole comes as a corollary of these lemmas. We plan to provide detailed proofs of these, but for now, they should be treated as conjectures and our formulation above as expository. 
%We believe that despite this, the formalization is surprisingly elegant and concise, given its expressive power. It is a useful exercise to implement natural numbers, ala G\"odel's \textbf{T}, using this calculus, assuming that the type-level and internal languages include integers and one can lift them from the former into the latter.
%%
%
%For the synthetic functions, it must decide the type assignment and the translation on the basis of the index and arguments, or decide that this is not possible. An operator constructor with index kind $\kappaidx$ must have a definition of kind \[\kappaidx \rightarrow \klist{\kDen} \rightarrow (\kDen + \kunit)\] 
%As this suggests, the active typing judgement  will provide the operator index and a list of {recursively determined \emph{derivates}}, which have kind $\kDen$, for the arguments and ask the definition to return a value of ``option kind'', $\ksum{\kDen}{\kunit}$, where the trivial case indicates that a derivate cannot be constructed due to an invalid index, an incorrect number of arguments or an argument of an invalid type.\footnote{In practice, we would require operator constructor providers to report information about the precise location of the error (e.g. which argument was of incorrect type) and provide an appropriate error message and other metadata, but we omit this in the semantics.}
%
%
%A \emph{program}, $\rho$, consists of a series of type constructor declarations followed by an external term, $e$. The syntax for external terms contains six forms: variables, $\lambda$ terms, a generalized unascribed introductory form, a generalized introductory form ascribed with a type, a generalized introductory form ascribed with a type constructor and a generalized elimination form. We will discuss how a more natural syntax  like Python's can be added by simple desugaring.\todo{syntax}
%
%
%
%Compiling a program consists of first \emph{kind checking} it (typechecking the type-level terms, which has no analog in @\texttt{lang} because the type-level language is Python; discussed further below), then typechecking the external term and {translating} it to a term, $\iota$, in the {typed internal language}. %These correspond to the premises of the \emph{central compilation judgement} $\pkcompiles{\rho}{\iota}$:
%%\[
%%\inferrule[p-compiles]{%
%%  \emptyset \vdash_{\fvalCtx_0} \rho\\
%%  \progOK{\emptyset}{\fvalCtx_0}{\rho}\\
%%  \pcompiles{\fvalCtx_0}{\rho}{\iota}
%%}{\pkcompiles{\rho}{\iota}}
%%\]
%%his anchors our exposition; we will describe how it is derived (i.e. how to write a compiler for $\lam@\texttt{lang}$) in the following sections. 
%
%The key judgement in the calculus is the \emph{active typing judgement}  (Fig. \ref{att}, which we describe starting in Sec. \ref{opcons}). It relates an external term, $e$, to a {type}, $\tau$, called its \emph{type assignment}, and an internal term, $\iota$, called its \emph{translation}, under \emph{typing context} $\Gamma$ and \emph{constructor context} $\Phi$: 
%\[\ecompilesX{e}{\tau}{\iota}\]
%
%The typing context $\Gamma$ maps variables to types in essentially the conventional way (\cite{pfpl} contains the necessary background for this paper). The constructor context tracks user-defined type and operator constructors. % Weakening and exchange of the constructor context closely related to the issue of conservativity that we will return to. Each constructor is identified by name, so there is no analog to contraction.
%There is no separate operational semantics for the external language. Instead, the dynamic behavior of an external term is determined by its translation to the internal language, which has a more conventional operational semantics. This form of semantics can be seen as lifting into the language specification the first stage of a type-directed compiler like the TIL compiler for Standard ML \cite{tarditi+:til-OLD} and has some parallels to the Harper-Stone semantics for Standard ML, where external terms were also given meaning by elaboration from the EL to an IL \cite{Harper00atype-theoretic}. %
%
%In $\lam@\texttt{lang}$, the internal language (IL) provides partial functions (via the generic fixpoint operator of Plotkin's PCF), simple product and sum types and a base type of integers (to make our example interesting and as a nod toward speed on contemporary machines). In practice, the internal language could be any typed  language with a specification for which type safety and decidability of typechecking have been satisfyingly determined. The internal type system serves as a ``floor'': guarantees that must hold for \emph{all} well-typed terms, independent of the constructor context (e.g. that out-of-bounds access to memory never occurs), must be maintained by the internal type system. User-defined constructors can (only) enforce invariants stronger than those the internal type system maintains. Performance is also ultimately limited by the internal language and downstream compilation stages that we do not here consider (safe compiler extension has been discussed in previous work, e.g. \cite{conf/pldi/TatlockL10}).
%

%\begin{figure}
%\small
%\begin{flalign}
%\small
% & \family{Tuple}{\klist{\kTypeBlur}}{\\
% & \quad \tops{new}{\kunit}{\_}{a}{\tlam{\_}{\kunit}{\tlam{a}{\klist{\kDen}}{\\
% & \quad\quad
%   \tfold{\tvar{a}}{\tden{{()}}{\ttype{Tuple}{\tnil{\kTypeBlur}}}}{d}{b}{r}{\\
%     & \quad\quad\quad \tfamcase{\ttypeof{\tvar{r}}}{Ntuple}{i}{\\
%     & \quad\quad\quad\quad \tfold{\tvar{b}}{\tden{\itransof{\tvar{d}}}{\ttype{Ntuple}{\tcons{\ttypeof{\tvar{d}}}{\tvar{i}}}}\\& \quad\quad\quad\quad}{\_}{\_}{\_}{\tden{\\& \quad\quad\quad\quad\quad {(\itransof{\tvar{d}},\itransof{\tvar{r}})}}{\ttype{Ntuple}{\tcons{{\ttypeof{\tvar{d}}}}{\tvar{i}}}}}
%     }{\terr}
% }
% }}};\\
% & \quad \tops{prj}{\kint}{i}{a}{\lambda \tvar{i}:\kint.\lambda \tvar{a}:\klist{\kDen}. 
%   \tapp{\tvar{decons1}}{\tapp{\tvar{a}}{\tlam{d}{\kDen}{\\
% & \quad\quad\tfamcase{\ttypeof{\tvar{d}}}{Ntuple}{nl}{\\
% & \quad\quad \tfold{\tvar{nl}}{\terr}{t1}{j}{\_}{\\
% & \quad\quad \tfold{\tvar{j}}{\tifeq{\tvar{i}}{1}{\dint}{\tden{\itransof{d}}{\tvar{t1}}}{\terr}\\&\quad\quad\quad}{\_}{\_}{\_}{\\
% & \quad\quad\quad (\tlam{p}{\kpair{\kDen}{\dint}}{\tifeq{\tsnd{\tvar{p}}}{\tvar{i}}{\dint}{\tfst{\tvar{p}}}{\terr}})\\
% & \quad\quad\quad (\tapp{\tvar{foldl}}{\tapp{\tvar{nl}}{\tapp{(\tden{\tvar{x}}{\tvar{nt}},0)}{
%   \tlam{r}{\kpair{\kDen}{\dint}}{\tlam{t}{\kTypeBlur}{\tlam{ts}{\klist{\kTypeBlur}}{\\
% & \quad\quad\quad\quad \tifeq{\tvar{i}}{\tsnd{\tvar{r}}}{\dint}{\tvar{r}}{\tdencase{\tfst{\tvar{r}}}{rx}{\_}{\\
% & \quad\quad\quad\quad\quad \tifeq{\tvar{i}}{\tsnd{\tvar{r}}+1}{\dint}{\\
% & \quad\quad\quad\quad\quad\quad \tifeq{\tvar{ts}}{\tnil{\kTypeBlur}}{\klist{\kTypeBlur}}{(\tden{\tvar{rx}}{\tvar{t}},\tvar{i})\\
% & \quad\quad\quad\quad\quad\quad}{(\tden{\titerm{\ifst{\iup{\tvar{rx}}}}}{\tvar{t}},\tvar{i})}\\  
% & \quad\quad\quad\quad\quad}{(\tden{\titerm{\isnd{\iup{\tvar{rx}}}}}{\tvar{t}},\tsnd{\tvar{r}}+1)}}{\terr}}}}}
% }}})
% }}{\terr}}}}}
% }\\ & 
%%& \quad \tops{pr}a}{b}{c}{d}
% }{d}{i}{(\tlam{i}{\klist{\kTypeBlur}}{
%   \\& \quad \quad \tfold{\tvar{i}}{\titype{\dunit}}{s}{j}{r}{
%     \tfold{\tvar{j}}{\titype{\trepof{\tvar{s}}}}{\_}{\_}{\_}{\\
%   &\quad\quad\quad
%     \titype{\dpair{\trepof{\tvar{s}}}{\dup{\tvar{r}}}}
%     }
%   }
% })}
%\end{flalign}
%\caption{ABC}
%\label{tuple}
%\end{figure}
%\begin{figure}[t]
%\small
%$\fbox{\inferrule{}{\progOKX{\progsort}}}$
%~~~$\tvarCtx ::= \emptyctx \pipe \tvarCtxX{t}{\kappa}$
%~~~$\fvalCtx ::= \emptyctx \pipe \fvalCtxX{\fvalDf}$
%%$\fCtx ::= \Sigma_0  \pipe \fvalCtxX$
%\begin{mathpar}
%\inferrule[k-tycon]{
% \fvar{tycon} \notin \text{dom}(\fvalCtx)\\
% \kEq{\kappaidx}\\
% \tKind{\tvarCtx}{\fvalCtx}{\taurep}{\karrow{\kappaidx}{\kIType}}\\
% \opType{\tvarCtx}{\fvalCtxX{\fvalDf}}{\theta}{\Theta}\\
% \progOK{\tvarCtx}{\fvalCtxX{\fvalDf}}{\rho}
%}{
% \progOKX{\pfam{\familyDf}{\rho}}
%}
%
%%\inferrule[def-kinding]{
%%  \tKindX{\tau}{\kappa}\\
%%  \progOK{\tvarCtxX{t}{\kappa}}{\fvalCtx}{\rho}
%%}{
%%  \progOKX{\pdef{t}{\kappa}{\tau}{\rho}}
%%}
%%
%\inferrule[k-e-prog]{
% \exprOK{\tvarCtx}{\fvalCtx}{e}
%}{
% \progOKX{e}
%}
%\end{mathpar}
%$\fbox{$\opType{\tvarCtx}{\fvalCtx}{\theta}{\Theta}$}$
%\begin{mathpar}
%\inferrule[k-opcon]{
% \tKind{\tvarCtx}{\fvalCtx}{\taudef}{\karrow{\kappaidx}{\karrow{\klist{\kDen}}{(\ksum{\kDen}{\kunit})}}}
%}{
% \opType{\tvarCtx}{\fvalCtx}{{\tops{op}{\kappaidx}{i}{a}{\taudef}}}{
% \kOpS{op}{\kappaidx}}
%}
%
%\inferrule[k-opcons]{
% \opType{\tvarCtx}{\fvalCtx}{\theta_1}{\Theta_1}\\
% \opType{\tvarCtx}{\fvalCtx}{\theta_2}{\Theta_2}\\\\
% \text{dom}(\theta_1) \cap \text{dom}(\theta_2) = \emptyset
%}{
% \opType{\tvarCtx}{\fvalCtx}{\theta_1; \theta_2}{\Theta_1, \Theta_2}
%}
%\end{mathpar}
%$\fbox{$\exprOKX{e}$}$
%\begin{mathpar}
%\inferrule[k-e-var]{ }{
% \exprOK{\tvarCtx}{\fvalCtx}{\evar{x}}
%}
%
%\inferrule[k-e-lam]{
% \tKindX{\tau}{\kTypeBlur}\\
% \exprOK{\tvarCtx}{\fvalCtx}{e}
%}{
% \exprOKX{\elam{\evar{x}}{\tau}{e}}
%}
%
%\inferrule[k-e-op]{
% \fval{tycon}{-}{-}{\theta} \in \fvalCtx\\
% \tops{op}{\kappaidx}{i}{a}{-} \in \theta\\
% \tKindX{\tauidx}{\kappaidx}\\
% \exprOKX{e_1}\\
% \cdots\\
% \exprOKX{e_n}
%}{
% \exprOKX{\eop{tycon}{op}{\tauidx}{\splat{e}{1}{n}}}
%}
%\end{mathpar}
%$\fbox{$\kEq{\kappa}$}$
%\begin{mathpar}
%\inferrule[t-eq]{ }{
% \kEq{\kTypeBlur}
%}
%
%\inferrule[i-eq]{ }{
% \kEq{\kint}
%}
%
%\inferrule[l-eq]{ }{
% \kEq{\klabel}
%}
%
%\inferrule[list-eq]{
% \kEq{\kappa}
%}{
% \kEq{\klist{\kappa}}
%}
%\\
%\inferrule[u-eq]{ }{
% \kEq{\kunit}
%}
%
%\inferrule[p-eq]{
% \kEq{\kappa_1}\\
% \kEq{\kappa_2}
%}{
% \kEq{\kpair{\kappa_1}{\kappa_2}}
%}
%
%\inferrule[s-eq]{
% \kEq{\kappa_1}\\
% \kEq{\kappa_2}
%}{
% \kEq{\ksum{\kappa_1}{\kappa_2}}
%}
%\end{mathpar}
%\caption{\small Kinding for type and operator constructors and external terms, and equality kinds (see text).  Type-level terms stored in the constructor context are not needed during kinding, indicated using a dash.}
%\label{kindprog}
%\vspace{-10pt}
%\end{figure}

%\section{Interactive Compilation and Invocation}\label{targets}
%\subsection{OpenCL as an Active Library}
%The code in this section uses \lstinlinep{clx}, an example library implementing the semantics of the OpenCL programming language and extending it with some additional useful types, which we will discuss shortly. @\texttt{lang} itself has no built-in support for OpenCL.
%
%To briefly review, OpenCL provides a data-parallel SPMD programming model where developers define functions, called {\em kernels}, for execution on \emph{compute devices} like GPUs or multi-core CPUs \cite{opencl11}. Each thread executes the same kernel but has access to a unique index, called its \emph{global ID}. Kernel code is written in a variant of C99 extended with some new primitive types and operators, which we will introduce  as needed in our examples below.
%
%\subsection{Generic Functions}\label{genfn}
%%\begin{codelisting}[t]
%%\lstinputlisting[commentstyle=\color{mauve}]{listing7.py}
%%\caption{[\texttt{listing\ref{metaprogramming}.py}] Metaprogramming with @\texttt{lang}, showing how to construct generic functions from abstract syntax trees.}
%%\label{metaprogramming}
%%\end{codelisting}
%
%%\begin{codelisting}
%%\lstinputlisting[linebackgroundcolor={\btLstHL{5-8}}]{listing3.py}
%%\caption{[\texttt{listing\ref{map}.py}] A generic data-parallel higher-order map function targeting OpenCL.}
%%\label{map}
%%\end{codelisting}
%%\begin{codelisting}[t]
%%\lstinputlisting[commentstyle=\color{mauve}]{listing7.py}
%%\caption{[\texttt{listing\ref{metaprogramming}.py}] Metaprogramming with @\texttt{lang}, showing how to construct generic functions from abstract syntax trees.}
%%\label{metaprogramming}
%%\end{codelisting}
%Lines 3-4 introduce \lstinlinep{map}, an @\texttt{lang} function of three arguments that is governed by the \emph{active base} referred to by \lstinlinep{clx.base} and targeting the \emph{active target} referred to by \lstinlinep{clx.opencl}. The active target determines which language the function will compile to (here, the OpenCL kernel language) and mediates code generation. 
%
%The body of this function, highlighted in grey for emphasis, does not have Python's semantics. Instead, it will be governed by the active base together with the active types used within it. No such types have been provided explicitly, however. Because our type system is extensible, the code inside could be meaningful for many different assignments of types to the arguments. We call functions awaiting types \emph{generic functions}. Once types have been assigned, they are called \emph{concrete functions}.
%
%Generic functions are represented at compile-time as instances of \lstinlinep{ace.GenericFn} and consist of an abstract syntax tree, an {active base} and an {active target}. The purpose of the \emph{decorator} on line 3 is to replace the Python function on lines 4-8 with an @\texttt{lang} generic function having the same syntax tree and the provided active base and active target. 
%Decorators in Python are simply syntactic sugar for applying the decorator function directly to the function  being decorated \cite{python}. In other words, line 3 could be replaced by inserting the following  statement on line 9:
%\vspace{-0.28cm}
%\begin{verbbox}
%map = ace.fn(clx.base, clx.opencl)(map)
%\end{verbbox}
%\begin{figure}[h!]
%\centering
%\theverbbox
%\end{figure}
%\vspace{-0.28cm}
%
%The abstract syntax tree for \lstinlinep{map} is extracted using the Python standard  library packages  \lstinlinep{inspect} (to retrieve its source code) and \lstinlinep{ast} (to parse it into a syntax tree). 
%
%%\subsection{Metaprogramming in @\texttt{lang}}
%%Generic functions can be generated directly from ASTs as well, providing @\texttt{lang} with support for  straightforward metaprogramming. Listing \ref{metaprogramming} shows how to generate two more generic functions, \lstinlinep{scale} and \lstinlinep{negate}. The latter is derived from the former by using a library for manipulating Python syntax trees, \lstinlinep{astx}. In particular, the \lstinlinep{specialize} function replaces uses of the second argument of \lstinlinep{scale} with the literal \lstinlinep{-1} (and changes the function's name), leaving a function of one argument.
% 
%\subsection{Concrete Functions and Explicit Compilation}
%To compile a generic function to a particular \emph{concrete function}, a type must be provided for each argument, and typechecking and translation must then succeed. Listing \ref{compscript} shows how to explicitly provide type assignments to \lstinlinep{map} using the subscript operator (implemented using Python's operator overloading mechanism). We attempt to do so three times in Listing \ref{compscript}. The first, on line \ref{compscript}.7, fails due to a type error, which we handle so that the script can proceed. The error occurred  because the ordering of the argument types was incorrect. We provide a valid ordering on line \ref{compscript}.9 to generate the concrete function \lstinlinep{map_neg_f32}. We then provide a different type assignment to generate the concrete function \lstinlinep{map_neg_ci32}.
%Concrete functions are instances of \lstinlinep{ace.TypedFn}, consisting of an abstract syntax tree annotated with types and translations along with a reference to the original generic function. %The typing and translation process is mediated by the logic in the active base, types and target that have been provided, as we will describe in more detail below. 
%
%To produce an output file from an @\texttt{lang} ``compilation script'' like \lstinlinep{listing}\texttt{\ref{compscript}}\lstinlinep{.py}, the command \lstinlinep{acec} can be invoked from the shell, as shown in Listing \ref{mapc}. 
%\subsection{Types}
%\begin{codelisting}
%\lstinputlisting{listing4.py}
%\caption{[\texttt{listing\ref{compscript}.py}] The generic \texttt{map} function compiled to map the \texttt{negate} function over two  types of input.}
%\label{compscript}
%\end{codelisting}
%\begin{codelisting}
%\begin{lstlisting}[style=Bash]
%$ `acec listing3.py`
%Hello, compile-time world!
%[ace] TypeError in listing1.py (line 6, col 28): 
%      'GenericFnType(negate)' does not support [].
%[acec] listing3.cl successfully generated.
%\end{lstlisting}
%\caption{Compiling \texttt{listing\ref{compscript}.py} using the \texttt{acec} compiler.}
%\label{mapc}
%\end{codelisting}
%\begin{codelisting}
%\lstinputlisting[style=OpenCL]{listing5.cl}
%\caption{[\texttt{listing\ref{compscript}.cl}] The OpenCL file generated by Listing \ref{mapc}.}
%\label{mapout}
%\end{codelisting}
%%Lines \ref{compscript}.3-\ref{compscript}.5 construct the types assigned to the arguments of \lstinlinep{map} on lines \ref{compscript}.7-\ref{compscript}.10. In @\texttt{lang}, types are themselves values that can be manipulated at compile-time. This stands in contrast to other contemporary languages, where user-defined types (e.g. datatypes, classes, structs) are written declaratively at compile-time but cannot be constructed, inspected or passed around programmatically. More specifically, types are instances of a Python class that implements the \lstinlinep{ace.ActiveType} interface (see Sec. \ref{atypes}). 
%%As Python values, types can be assigned to variables when convenient (removing the need for  facilities like \lstinlinep{typedef} in C or \lstinlinep{type} in Haskell). Types, like all compile-time objects derived from @\texttt{lang} base classes, do not have visible state and operate in a referentially transparent manner (by constructor memoization, which we do not detail here).% These types are all implemented in the \lstinlinep{clx} library imported on line 1, none are built into @\texttt{lang} itself.
%
%The type named \lstinlinep{T1} on line \ref{compscript}.3 corresponds to the OpenCL type \lstinlinep{global float*}: a pointer to a 32-bit floating point number stored in the compute device's global memory (one of four address spaces defined by OpenCL \cite{opencl11}). It is constructed by applying \lstinlinep{clx.Ptr}, which is an @\texttt{lang} type constructor corresponding to pointer types, to a value representing the  address space, \lstinlinep{clx.global_}, and the type being pointed to. That type, \lstinlinep{clx.float}, is in turn the @\texttt{lang} type corresponding to \lstinlinep{float} in OpenCL (which, unlike C99, is always 32 bits). 
%The \lstinlinep{clx} library contains a full implementation of the OpenCL type system (including behaviors, like promotions, inherited from C99).
%@\texttt{lang} is \emph{unopinionated} about issues like memory safety and the wisdom of such promotions. We will discuss how to implement, as libraries, abstractions that are higher-level than raw pointers in Sec. \ref{examples}, but @\texttt{lang} does not prevent users from choosing a low level of abstraction or ``interesting'' semantics if the need arises (e.g. for compatibility with existing libraries; see the discussion in Sec. \ref{discussion}). We also note that we are being more verbose than necessary for the sake of pedagogy. The \lstinlinep{clx} library includes more concise shorthand for OpenCL's types: \lstinlinep{T1} is equal to \lstinlinep{clx.gp(clx.f32)}. %Similarly, the decorators in Listings \ref{map} and \ref{metaprogramming} could have been written \lstinlinep{clx.cl_fn}.\todo{move this back there probably}
%
%The type \lstinlinep{T2} on line \ref{compscript}.4 is a pointer to a \emph{complex integer} in global memory. It does not correspond direrctly to a type in OpenCL, because OpenCL does not include primitive support for complex numbers. Instead, it uses an active type constructor \lstinlinep{clx.Cplx}, which includes the necessary logic for typechecking operations on complex numbers and translating them to OpenCL (Sec. \ref{atypes}). This constructor is parameterized by the numeric type that should be used for the real and imaginary parts, here \lstinlinep{clx.int}, which corresponds to 32-bit OpenCL integers. Arithmetic operations with other complex numbers, as well as with plain numeric types (treated as if their imaginary part was zero), are supported. When targeting OpenCL, @\texttt{lang} expressions assigned type \lstinlinep{clx.Cplx(clx.int)} are compiled to OpenCL expressions of type \lstinlinep{int2}, a  \emph{vector type} of two 32-bit integers (a type that itself is not inherited from C99). This can be observed in several places on lines \ref{mapout}.14-\ref{mapout}.21. This choice is merely an implementation detail that can be kept private to \lstinlinep{clx}, however. An @\texttt{lang} value of type \lstinlinep{clx.int2} (that is, an actual OpenCL vector) \emph{cannot} be used when a \lstinlinep{clx.Cplx(clx.int)} is expected (and attempting to do so will result in a static type error): \lstinlinep{clx.Cplx} truly extends the type system, it is not a type alias.
%
%The type \lstinlinep{TF} on line \ref{compscript}.5 is extracted from the generic function \lstinlinep{negate} constructed in Listing \ref{metaprogramming}. Generic functions, according to Sec. \ref{genfn}, have not yet had a type assigned to them, so it may seem perplexing that we are nevertheless assigning a type to \lstinlinep{negate}. Although a conventional arrow type cannot be assigned to \lstinlinep{negate}, we can give it a \emph{singleton type}: a type that simply means ``this expression is the \emph{particular} generic function \lstinlinep{negate}''. This type could also have been explicitly written as \lstinlinep{ace.GenericFnType(listing2.negate)}. During typechecking and translation of \lstinlinep{map_neg_f32} and \lstinlinep{map_neg_ci32}, the call to \lstinlinep{f} on line \ref{map}.6 uses the type of the provided argument to compile the generic function that inhabits the singleton type of \lstinlinep{f} (\lstinlinep{negate} in both of these cases) to a concrete function. This is why there are two versions of \lstinlinep{negate} in the output in Listing \ref{mapout}. In other words, types \emph{propagate} into generic functions -- we didn't need to compile \lstinlinep{negate} explicitly. This also explains the error printed on line \ref{mapc}.3-\ref{mapc}.4: when this type was inadvertently assigned to the first argument \lstinlinep{input}, the indexing operation on line \ref{map}.6 resulted in an error. A generic function can only be \emph{statically} indexed by a list of types to turn it into a concrete function, not \emph{dynamically} indexed with a value of type \lstinlinep{clx.size_t} (the return type of the OpenCL primitive function \lstinlinep{get_global_id}).
%
%In effect, this scheme enables higher-order functions even when targeting languages, like OpenCL, that have no support for higher-order functions (OpenCL, unlike C99, does not support function pointers). Interestingly, because they have a singleton type, they are higher-order but not first-class functions. That is, the type system would prevent you from creating a heterogeneous list of generic functions. Concrete functions, on the other hand, can be given both a singleton type and a true function type. For example, \lstinlinep{listing2.negate[[clx.int]]} could be given type \lstinlinep{ace.Arrow(clx.int, clx.int)}. The base determines how to convert the @\texttt{lang} arrow type to an arrow type in the target language (e.g. a function pointer for C99, or an integer that indexes into a jump table constructed from knowledge of available functions of the appropriate type in OpenCL).
%
%Type assignment to generic functions is similar in some ways to template specialization in C++. In effect, both a template header and type parameters at call sites are being generated automatically by @\texttt{lang}. This simplifies a sophisticated feature of C++ and enables its use with other targets like OpenCL. %Other uses for C++ templates (and the preprocessor) are subsumed by the metaprogramming features discussed above\todo{cite template metaprogramming paper}.
%
%%\subsection{Within-Function Type Resolution}
%%\begin{codelisting}
%%\lstinputlisting{listing6.py}
%%\caption{\texttt{[listing6.py]} A function demonstrating whole-function type inference when multiple values with differing types are assigned to a single identifier, \texttt{y}.}
%%\label{inference}
%%\end{codelisting}
%%On line 5 in the generic \lstinlinep{map} function in Listing \ref{map}, the variable \lstinlinep{gid} is initialized with the result of calling the OpenCL primitive \lstinlinep{get_global_id}.  The type for \lstinlinep{gid} is never given explicitly. This is a simple case of @\texttt{lang}'s {\em within-function type resolution} strategy (we hesitate to call it \emph{type inference} because it does not, strictly speaking, take a constraint-solving approach). In this case, the type of \lstinlinep{gid} will resolve to \lstinlinep{size_t} because that is the return type of \lstinlinep{get_global_id} (as defined in the OpenCL specification, which the \lstinlinep{ace.OpenCL} module follows). The result can be observed on Lines 11 and 24 in Listing \ref{mapout}. 
%%
%%Inference is not restricted within single assignments, as in the \lstinlinep{map} example, however. Multiple assignments to the same identifier with values of differing types, or multiple return statements, can be combined if the types in each case are compatible with one another (e.g. by a subtyping relation or an implicit coercion). In Listing \ref{inference}, the \lstinlinep{threshold_scale} function assigns different values to \lstinlinep{y} in each branch of the conditional. In the first branch, the value \lstinlinep{0} is an \lstinlinep{int} literal. However, in the second branch of the loop, the type depends on the types of both arguments, \lstinlinep{x} and \lstinlinep{scale}. We show two choices for these types on Lines 11 and 12. Type inference correctly combines these two types according to OpenCL's C99-derived rules governing numeric types (defined by the user in the \lstinlinep{OpenCL} module, as we will describe in Section \ref{att}). We can verify this programmatically on Lines 12 and 13. Note that this example would also work correctly if the assignments to \lstinlinep{y} were replaced with \lstinlinep{return} statements (in other words, the return value of a function is treated as an assignable for the purpose of type inference).
%
%\subsection{Implicit Compilation and Interactive Execution}\label{compenv}\label{backend}\label{implicit}
%%Professional end-user programmers (e.g. scientists and engineers \cite{professional-end-users}) today generally use dynamically-typed high-level languages like MATLAB, Python, R or Perl for tasks that are not performance-sensitive, such as small-scale data analysis and plotting \cite{nguyen2010survey}. For portions of their analyses where the performance overhead of dynamic type checking and automatic memory management is too high, they will typically call into code written in a statically-typed, low-level language, most commonly C or Fortran, that uses low-level parallel abstractions like pthreads and MPI \cite{4222616,basili2008understanding}. Unfortunately, these low-level languages and abstractions are notoriously difficult to use and automatic verification is intractable in general.\todo{integrate this}
%
%
%A common workflow for \emph{professional end-user programmers} (e.g. scientists and engineers) is to use a simple scripting language for orchestration, small-scale data analysis and visualization and call into a low-level language for performance-critical sections. Python is both designed for this style of use and widely adopted for such tasks \cite{sanner1999python,nguyen2010survey}. Developers can call into native functions using Python's foreign function interface (FFI), for example. A more recent trend is to generate and compile code without leaving Python, using a Python wrapper around a compiler. For example, \lstinlinep{weave} works with C and C++, and \lstinlinep{pycuda} and \lstinlinep{pyopencl} work with CUDA and OpenCL, respectively \cite{klockner2011pycuda}. 
%\begin{codelisting}
%\lstinputlisting{listing8.py}
%\caption{[\texttt{listing\ref{py}.py}] A full OpenCL program using the \texttt{clx} Python bindings, including data transfer to and from a device and direct invocation of a generic function, \texttt{map}.}
%\label{py}
%\end{codelisting}
%The OpenCL language was designed for this workflow, exposing a retargetable compiler and data management  routines as an API, called the \emph{host API} \cite{opencl11}. The \lstinlinep{pyopencl} library exposes this API to Python and simplifies interoperation with \lstinlinep{numpy}, a popular package for manipulating contiguously-allocated numeric arrays in Python \cite{klockner2011pycuda}.
%
%@\texttt{lang} supports a refinement to this workflow, as an alternative to the \lstinlinep{acec} compiler described above, for targets that have wrappers like this available, including \lstinlinep{clx.opencl}. Listing \ref{py} shows an example of this workflow where the user chooses a compute device (line \ref{py}.3),  constructs a \lstinlinep{numpy} array (line \ref{py}.5), transfers it to the device (line \ref{py}.6), allocates an empty equal-sized buffer for the result of the computation (line \ref{py}.7), launches the {generic} kernel \lstinlinep{map} from Listing \ref{map} with these device arrays as well as the function \lstinlinep{negate} from Listing \ref{metaprogramming} (line \ref{py}.9) choosing a number of threads equal to the number of elements in the input array (line \ref{py}.10), and transfers the result back into main memory to check that the device computed the expected result (line \ref{py}.12).
%
%For developers experienced with the usual OpenCL or CUDA workflow, the fact that this can be accomplished in a total of 6 statements may be surprising. This simplicity is made possible by @\texttt{lang}'s implicit tracking of types throughout the code. First, \lstinlinep{numpy} keeps track of the type, shape and order of its arrays. The type of \lstinlinep{input}, for example, is \lstinlinep{numpy.float64} by default, its shape is \lstinlinep{(1024,1024)} and its order is row-major by default. The \lstinlinep{pyopencl} library, which the \lstinlinep{clx} mechanism is built upon, uses this metadata to automatically call the underlying OpenCL host API function for transferring byte arrays to the device without requiring the user to calculate the size. The \lstinlinep{clx} wrapper further retains this metadata in \lstinlinep{d_input} and \lstinlinep{d_output}, the Python wrappers around the allocated device arrays. The @\texttt{lang} active type of these wrappers is an instance of \lstinlinep{clx.NPArray} parameterized by this metadata. This type knows how to typecheck and translate operations, like indexing with a multi-dimensional thread index, automatically.
%
%By extracting the types from the arguments, we can call the generic function \lstinlinep{map} without first requiring an explicit type assignment, like we needed when using \lstinlinep{acec} above. In other words, dynamic types and other metadata can propagate from Python data structures into an @\texttt{lang} generic function as static type information, in the same manner as it propagated \emph{between} generic functions in the previous section. In both cases, typechecking and translation of \lstinlinep{map} happens the first time a particular type assignment is encountered and cached for subsequent use. When called from Python, the generated OpenCL source code is compiled for the device we selected using the OpenCL retargetable compilation infrastructure, and cached for subsequent calls. 
%
%The same program written using the OpenCL C API directly is an order of magnitude longer and significantly more difficult to comprehend. OpenCL does not support higher-order functions nor is there any way to write \lstinlinep{map} in a type-generic or shape-generic manner. If we instead use the \lstinlinep{pyopencl} library and apply the techniques described in \cite{klockner2011pycuda}, the program is still twice as large and less readable than this code. Both the \lstinlinep{map} and \lstinlinep{negate} functions must be explicitly specialized with the appropriate types using string manipulation techniques, and custom shapes and orders can be awkward to handle. Higher order functions are still not available, and must also be simulated by string manipulation. That approach also does not permit the use any of the language extensions that @\texttt{lang} enables (beyond the useful type for \lstinlinep{numpy} arrays just described; see Sec. \ref{examples} for more interesting possibilities).
%
%%Not shown are several additional conveniences, such as delegated kernel sizing and \lstinlinep{In} and \lstinlinep{Out} constructs that can reduce the size and improve the clarity of this code further; due to a lack of space, the reader is referred to the language documentation.

\section{Related Work}\label{related}

%\parindent0pt
%
%\subsection{Active Libraries}

%We have implemented the full OpenCL programming language as  primary example showed how to support low-level GPU programming from a high-level language. Other languages have supported similar workflows, e.g. Rust \cite{rustgpu}. These have not been fundamentally extensible and suffer from the ``bootstrapping'' problem described in Sec. \ref{intro}.

Early work proposing the inclusion of compile-time logic in libraries led to the phrase {\it active libraries} \cite{activelibraries}. We borrow the prefix ``active''. Our focus on type constructors and type system extension differs substantially from previous work, which focused on term rewriting for the purpose of optimization over a fixed semantics. Several contemporary projects have the same goals, e.g. LMS \cite{Rompf:2012:LMS}, which allows staged translation of well-typed Scala programs to other targets and supports composing optimizations. Macro systems can also be used to define optimizations and operations with interesting dynamic semantics. In both cases, the language's type system itself remains fixed. Our mechanism permits true type system extensions. New types are not merely aliases, nor must their rules be admissible in some base type system (e.g. \lstinlinep{string_in} would be difficult to define in Scala, particularly as directly as we can here). Note that in @\texttt{lang}, macros can be seen as a mode of use of the term constructor \verb|Call|, as we can give the macro a singleton type (like Python modules, discussed above) that performs the desired rewriting in \verb|ana_Call| and \verb|syn_Call|.

%Generic functions are a novel strategy for {\it function polymorphism} -- defining functions that operate over more than a single type. In @\texttt{lang}, generic functions are implicitly polymorphic and can be called with arguments of {\it any type that supports the operations used by the function}. This is related to structural polymorphism \cite{malayeri2009structural}. Structural types make explicit the structure required of an argument, unlike generic functions, which are only given singleton types because the structure may depend on the semantics of an active type. 
%Structural typing can be compared to the more \emph{ad hoc} approach taken by dynamically-typed languages like Python itself, sometimes called ``duck typing''. It is also comparable to the C++ template system, as discussed previously.

{\it Operator overloading} \cite{vanWijngaarden:Mailloux:Peck:Koster:Sintzoff:Lindsey:Meertens:Fisker:acta:1975} and {\it metaobject dispatch} \cite{Kiczales91} are run-time protocols that translate operator invocations into function calls. The function is typically selected according to the dynamic tag or value of one or more operands. These protocols share the notion of {\it inversion of control} with our strategy for targeted expressions. However, our strategy is a {\it compile-time} protocol. Note that we used Python's operator overloading and metaobject protocol for convenience in the static language. %Our dispatch protocol was designed to be conceptually similar to Python's operator overloading protocol \cite{python}.

Language-external mechanisms for creating and composing dialects (e.g. extensible compilers like Xoc \cite{conf/asplos/CoxBCKK08} or language workbenches \cite{erdweg2013state}) have ambiguity problems, and so they might benefit from the type constructor oriented view that we propose here as well. We argue that if ambiguities cannot occur and safety is guaranteed, there is no reason to leave the mechanism outside the language. Though we do not support syntax extension (other than via creative reuse of string literals), we argue that this is actually a benefit: we can use a variety of existing tools and avoid many facets of the expression problem \cite{wadler1998expression} in this way.

Typed Racket and other typed LISPs also add a type system to a fixed syntax atop a dynamically typed core \cite{TypedScheme2008}. However, these treat the semantics as a ``bag of rules'', so adding new rules can cause  ambiguities. Our work (particularly @$\lambda$, with its use of lists ubiquitously) provides a blueprint for a typed LISP oriented around type constructors, rather than rulesets. Type constructors are a natural organizational unit. For example, Harper's textbook, upon which we base our terminology and abstract syntax in this paper, identifies languages directly as a collection of type constructors \cite{pfpl}.

Bidirectional type systems are increasingly being used in practical settings, e.g. in Scala and C\#. They are useful in producing good error messages (which our mechanism shows can be customized) and help avoid the need for redundant type annotations while avoiding decidability limitations associated with whole-program type inference. Bidirectional techniques have also been used for adding refinement types to languages like ML and Twelf \cite{Lovas08abidirectional}. Refinement types can add stronger static checking over a fixed type system (analagous to formal verification), but cannot be used to add new operations directly to the language. Our use of a dynamic semantics for the static language relates to the notion of \emph{type-level computation}, being explored in a number of languages (e.g. Haskell) for reasons other than extensibility. %Combining the approaches may be a fruitful avenue for future work.

Recent work we have done on the Wyvern programming language uses bidirectional typechecking to control aspects of literal syntax in a manner similar to, but somewhat more flexible than, how we treat introductory forms \cite{TSLs}. Wyvern is its own language dialect and does not have an extensible type system, supporting only desugarings like SugarJ \cite{erdweg2011sugarj}. The Wyvern formalism guarantees hygiene, using an approach that is also likely applicable in the setting of this paper.
%\begin{figure*}[t]
%\small
%\vspace{-10pt}
%{\onecolumn
%\begin{longtable}{l l@{}l c@{}c c@{}c c@{}c c@{}c}
%%\toprule%
%
%&&  &&  {\bfseries Itself} && {\bfseries Extensible} && {\bfseries Extensible} && {\bfseries Unambiguous}\\
%
%{\bfseries Approach} && {\bfseries Examples} && {\bfseries Library} && {\bfseries Syntax} && {\bfseries Type System} && {\bfseries Composition}  \\
%
%\cmidrule(l){1-1} \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9} \cmidrule(l){10-11} 
%
%\endhead
%
%
%Active Types && && \CIRCLE && \Circle && \CIRCLE && \CIRCLE  \\
%\myrowcolour%
%Type-Specific Literals && \cite{Omar:2013:TWP:2489812.2489815} && \Circle && \CIRCLE && \Circle && \CIRCLE  \\
%Desugaring && \cite{erdweg2011sugarj} && \Circle && \CIRCLE && \Circle && \Circle\\
%\myrowcolour%
%Typed LISP && Qi, \cite{TypedScheme2008} &&  \CIRCLE && \Circle && \CIRCLE && \Circle \\
%Attribute Grammars && \cite{VanWyk:2010:SEA}, \cite{conf/asplos/CoxBCKK08} && \Circle && \CIRCLE && \CIRCLE && \Circle\\
%%\myrowcolour
%%Static macros /  && \cite{ScalaMacros2013}, \cite{MorphJ2011}, && \Circle && \Circle && \Circle && \CIRCLE && \Circle \\
%%\myrowcolour
%%Metaprogramming &&  \cite{Sheard:1999:UMS} && && && && && \\
%%
%%Cross-Compilation && \cite{Rompf:2012:LMS, Delite2011} && \CIRCLE && \Circle && \Circle && \Circle && \CIRCLE \\
%%
%\myrowcolour%
%%EDSL Frameworks && ? && \CIRCLE && \CIRCLE && \CIRCLE && \Circle && \Circle \\
%\end{longtable}}
%\caption{Comparison to related approaches to language extensibility.}\label{relatedtable}
%\end{figure*}
%
%TODO: mention this http://blog.codeclimate.com/blog/2014/05/06/gradual-type-checking-for-ruby/

\section{Discussion}\label{discussion}
This work aimed to show that one can add static typechecking to a language like Python as a library, without undue syntactic overhead using techniques available in a growing number of languages: reflection, quasiquotes and a form of open sum for representing type constructors (here, Python's classes). The type system is not fixed, but flexibly extensible in a natural and direct manner, without resorting to complex encodings and, crucially, without the possibility of ambiguities arising at link-time. 
%We achieve this by combining a bidrectional type system and an elaboration semantics and delegating to static functions associated with user-defined type constructors, rather than operating in the usual syntax-directed manner. 
%Our examples were not trivial, showing a prototypic object system, a powerful string tracking system and immutable records coexisting as libraries within a language that initially had no static typechecking at all.

Although we do not have space to detail it here, we have conducted a substantial case study using (an older version of) @\texttt{lang}: an implementation of the entirety of the OpenCL type system (a variant of C99), as well as several extensions to it, as a library. Our operations translate to Python's lower-level FFI with OpenCL but guarantee type safety at the interface between languages. We have built a neurobiological circuit simulation framework atop this library as well. Based on our examples in this paper and this admittedly anecdotal experience report, we submit that active type constructors deserve broad consideration by the research community as a foundational technique for structuring typed programming languages. We are currently refactoring these projects against the (new) bidirectional semantics described here. The language core of  has already been implemented.

There remain several promising avenues for future work, many of which we mentioned throughout this paper. From a practical perspective, extensible implicit coercions (i.e. subtyping) using a mechanism similar to our ``handle sets'' mechanism for binary expressions would be useful. An extensible mechanism supporting type index polymorphism would also be of substantial utility and theoretical interest. Debuggers and other tools that rely not just on Python's syntax but also its semantics cannot be used directly. We believe that active types can be used to control debugging and other tools, and plan to explore this in the future. We have also not yet evaluated the feasibility of implementing more advanced type systems (e.g. linear, dependent or flow-dependent type systems) and those that require a more ``global'' view (e.g. security-oriented types) using our framework.

From a theoretical perspective, the next step is to introduce a static semantics for the internal language and the static language, so that we can help avoid issues of extension correctness and guarantee extension safety (by borrowing techniques from the typed compilation literature). An even more interesting goal would be to guarantee that extensions are mutually conservative: that one cannot weaken any of the guarantees of the other. We believe that by enforcing strict abstraction barriers between extensions, we can approach this goal. Bootstrapping the @\texttt{lang} compiler would be an interesting direction to explore to enable this.
% Different dialects can be difficult to combine manually, as we discussed in the introduction. We have implemented the full OpenCL programming language (a variant of C99) as a library (though the details are left for future work), showing that such languages can instead be written as libraries, rather than as dialects.
% 
%\subsection{Compositional Reasoning}\label{safety}
%Every statement and expression in an @\texttt{lang} function is governed by exactly one active type, determined by the dispatch protocol, or by the single active base associated with the @\texttt{lang} function. The representational consistency check ensures that compilation is successful without requiring that types that abstract over other types (e.g. container types) know about their internal implementation details. Together, this permits compositional reasoning about the semantics of @\texttt{lang} functions even in the presence of many extensions. This stands on contrast to many prior approaches to extensibility, where extensions could insert themselves into the semantics in conflicting ways, making the order of imports matter (see Sec. \ref{related}).

%\begin{enumerate}
%\item We address the ``chicken-and-egg'' problem by designing @\texttt{lang}, an extensible language  embedded within Python \cite{Politz:2013:PFM:2509136.2509536,python} entirely as a library. The top-level of an @\texttt{lang} file is a \emph{compilation script} written in Python and Python serves as @\texttt{lang}'s \emph{type-level language}, as we will discuss.  @\texttt{lang} and Python thus share a common syntax and package system, so users of @\texttt{lang} can directly leverage established tools, infrastructure and coding standards. Functions marked by a decorator as under the control of @\texttt{lang} are, unlike Python functions, \emph{statically} typechecked. This makes it possible to rule out many potential issues ahead of execution, and thus  avoid  costly dynamic checks. %The semantics of  define run-time behavior.
%\item  % Instead, @\texttt{lang} has a statically-typed semantics that can be extended by users from within libraries. 
%There is \emph{no particular type system built into @\texttt{lang}}. Instead, a type is any  object implementing the \lstinlinep{ace.Type} interface constructed by the compilation script. We call these \emph{active types}. When statically assigning a type to a compound form, e.g. \lstinlinep{e.x}, the @\texttt{lang} {compiler} delegates to the type of a subexpression, e.g. to the type of \lstinlinep{e}, to determine how to assign a type to the expression as a whole. As we will show, this {type-directed protocol}, \emph{active typing}, permits the expression of a rich variety of type systems as libraries. Unlike syntax-directed protocols, there cannot be ambiguities when separately defined type sytems are combined (types control non-overlapping sets of expressions by construction). Base cases, e.g.  variables and certain statement forms, are handled on a {per-function basis} by a \emph{base semantics}, also a compile-time object implementing an interface, \lstinlinep{ace.Base}.
%\end{enumerate}
%% As a result, clients are able to import any combination of extensions with the confidence that link-time ambiguities cannot occur .
%
%The dynamic behavior of an @\texttt{lang} function is determined by translation to a \emph{target language}. We begin by  targeting Python, then discuss targeting OpenCL and CUDA, lower-level languages used to program many-core processors (e.g. GPUs). Active types and bases control translation by the same type-directed protocol that governs type assignment. We call this phase \emph{active translation}. A \emph{target} mediates this process and is also a user-defined compile-time object implementing an interface, \lstinlinep{ace.Target}. 
%%One base or type can support multiple targets. 
%When targeting a typed language, care must be taken to ensure that well-typed @\texttt{lang} expressions have well-typed translations, so we integrate a technique for compositionally reasoning about compiler correctness (developed for the TIL compiler for Standard ML \cite{tarditi+:til-OLD}) into the language.
%
%%TODO
%%
%%Static type systems are powerful tools for programming language design and implementation. By tracking the type of a value statically, a typechecker can verify the absence of many kinds of errors over all inputs. This simplifies and increases the performance of the run-time system, as errors need not be detected dynamically using tag checks and other kinds of assertions.
%
%It is legitimate to ask, however, why dynamically-typed languages are so widely-used in multiple domains. Although slow and difficult to reason about, these languages generally excel at satisfying the criteria of \textbf{ease-of-use}. More specifically, Cordy identified the principle of \emph{conciseness} as elimination of
%redundancy and the availability of reasonable defaults \cite{cordy1992hints}. Statically-typed languages, particularly those that professional end-users are exposed to, are often verbose, requiring explicit and often redundant type annotations on each function and variable declaration, separate header files, explicit template headers and  instantiation and other sorts of annotations.  Dynamically-typed languages, on the other hand, avoid most of this overhead by relying on support from the run-time system. @\texttt{lang} was first conceived to explore the question: \emph{does conciseness require run-time mechanisms, or can one develop a statically-typed language with the same usability profile?}
%
%
%Rather than designing a new syntax, or modifying the syntax of an existing language, we chose to utilize, \emph{without modification}, the syntax of an existing language, Python. This choice was not arbitrary, but rather a key means by which @\texttt{lang} satisfies extrinsic design criteria related to familiarity and tool support. Researchers often dismiss the importance of syntax. By repurposing a well-developed syntax, they no longer need to worry about the ``trivial'' task of implementing it.
%
%Most programming languages are {\em monolithic} -- a collection of primitives are given first-class treatment by the language implementation, and users can only creatively combine them to implement  abstractions of their design. Although highly-expressive general-purpose mechanisms have been developed (such as object systems or functional datatypes), these may not suffice when researchers or domain experts wish to evolve aspects of the type system, exert control over the representation of data, introduce specialized run-time mechanisms, if defining an abstraction in terms of existing mechanisms is unnatural or verbose, or if custom error messages are useful. In these situations, it would be desirable to have the ability to modularly extend existing systems with new compile-time logic and be assured that such extensions will never interfere with one another when used in the same program.
%
%Python does not truly prevent extensions from interfering with one another because it lacks, e.g., data hiding mechanisms. One type could change the implementation of another by replacing its methods, for example. But these kinds of conflicts do not occur ``innocently'', as conflicts between two libraries in SugarJ that use similar syntax might.
% 
%
%

Not all extensions will be useful. Indeed, some language designers worry that offering too much flexibility to users leads to abuse. This must be balanced with the possibilities made available by a vibrant ecosystem of competing statically-typed abstractions that can be developed and deployed as libraries, and thus more easily evaluated in the wild. With an appropriate community process, this could lead to faster adoption of ideas from the research community, and quicker abandonment of mistakes. %We also anticipate that coding guidelines and tools mandating the use of abstractions that are known to have certain desirable properties will replace designer-mandated enforcement.
% 
%This paper is presented as a language design paper. The theoretical foundations of this work lie in type-level computation. We have developed a simplified, type-theoretic formalism of active typechecking and translation, where we prove the safety properties we have outlined here, as well as several additional ones that are only possible to achieve using a typed metalanguage (currently in submission at ESOP 2014). The work described here  extends that mechanism in several directions: @\texttt{lang} supports a rich syntax, makes syntax trees available to extensions, supports a pluggable backend and per-function base semantics. It is implemented in an existing language and supports a wider range of practical extensions.

%\acks
%The author is grateful to Jonathan Aldrich, Robert Bocchino and anonymous referees for their useful suggestions. This work was funded by the DOE Computational Science Graduate Fellowship under grant number DE-FG02-97ER25308.
%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrv}

% The bibliography should be embedded for final submission.

\bibliography{../research}
%\softraggedright
%P. Q. Smith, and X. Y. Jones. ...reference text...
%
%\appendix
%\section{Source Code Emission}
%Portions of the \lstinlinep{Target} class related to source code emission are shown. If compilation occurred externally, the \lstinlinep{acec} library asks the targets used by the top-level typed functions to generate one or more files by deciding on a name and file extension on the basis of the name of the compilation script, and emitting code and \emph{snippets}, which are simply strings corresponding to functions, type declarations, imports and other top-level entities in the target language, each inserted at a \emph{location}. Each file being generated is a location, and locations might also be nested, depending on the language (e.g. the import block may be a separate location when generating Java, or the body of the \lstinlinep{<head>} tag when generating HTML+CSS+Javascript). Snippets are only added once to a particular location (so imports are not duplicated, for example).
%%We leave layering extensions to Python to created typed targets as future work.
%
%\begin{codelisting}
%\begin{lstlisting}
%class PyTarget(ace.Target):
%  class PyAST(object): """Base class for ASTs"""
%  target_type = ast
%  class Attribute(PyAST, ast.Attribute):
%    def emit(self, cg):
%      cg.append(self.value.emit(cg), '.', self.attr)
%  # ... 
%  class AnonModule(object): 
%    def __init__(self, name): 
%      self.name = name
%      self.guid = Guid()
%      
%    def emit(self, cg):
%      import_loc = cg['imports']
%      _cache = AnonModule._idx_cache[import_loc]
%      if self not in _cache:
%        anon_name = '__ace_' + str(len(_cache)+1)
%        imp_stmt = ('import ' + self.name + ' as ' + 
%          anon_name + cg.newline)
%        _cache[self] = (anon_name, imp_stmt)
%      else: 
%        anon_name, imp_stmt = _cache[self]
%      imports.add_snippet(imp_stmt)
%      cg.append(anon_name)
%      
%  def emit(self, script_name, cg):
%    file_name = '_' + script_name + '.py'
%    if file_name not in generator:
%      if 'imports' not in generator[file_name]:
%        generator[file_name].push_loc('imports')
%      if 'main' not in format_set[file_name]:
%        generator[file_name].push_loc('main')
%    translation.emit((file_name, 'main'))
%\end{lstlisting}
%\caption{Portions of the target showing how non-local code emission works.}
%\label{cg}
%\end{codelisting}
%
%\section{Kinding}
%\begin{figure}[t]
%\small
%%\vspace{-15pt}
%$\fbox{\inferrule{}{\tKindX{\tau}{\kappa}}}$
%%~~~$\chi ::= \cdot \pipe \checkmark$
%\begin{mathpar}
%\small\inferrule[k-var]{
%}{
%  \tKind{\tvarCtxX{t}{\kappa}}{\fvalCtx}{\tvar{t}}{\kappa}
%}
%~~~~
%\inferrule[k-arrow-i]{
%  \tKind{\tvarCtxX{t}{\kappa_1}}{\fvalCtx}{\tau}{\kappa_2}
%}{
%  \tKindX{\tlam{t}{\kappa_1}{\tau}}{\karrow{\kappa_1}{\kappa_2}}
%}
%~~~~
%\inferrule[k-arrow-e]{
%  \tKindX{\tau_1}{\karrow{\kappa_1}{\kappa_2}}\\\\
%  \tKindX{\tau_2}{\kappa_1}
%}{
%  \tKindX{\tapp{\tau_1}{\tau_2}}{\kappa_2}
%}
%
%\text{\color{gray} (kinding for integers, labels, lists, products and sums also standard)}
%
%%%
%%%\inferrule{ }{
%%% \tKindX{\tstr{str}}{\kstr}
%%%}(\text{str}^I_\tau)
%%%
%%%\inferrule{ }{
%%% \tKindX{\tunit}{\kunit}
%%%}(\text{1}^I_\tau)
%%%
%%%\inferrule{
%%% \tKindX{\tau_1}{\kappa_1}\\
%%% \tKindX{\tau_2}{\kappa_2}
%%%}{
%%% \tKindX{\tpair{\tau_1}{\tau_2}}{\kpair{\kappa_1}{\kappa_2}}
%%%}({\times}^I_\tau)
%%%
%%%\inferrule{
%%% \tKindX{\tau}{\kpair}{\kappa_1}{\kappa_2}
%%%}{
%%% \tKindX{\tfst{\tau}}{\kappa_1}
%%%}({\times}^{E1}_\tau)
%%%
%%%\inferrule{
%%% \tKindX{\tau}{\kpair}{\kappa_1}{\kappa_2}
%%%}{
%%% \tKindX{\tsnd{\tau}}{\kappa_2}
%%%}({\times}^{E2}_\tau)
%%%\inferrule{ }{
%%%\inferrule{
%%% \tKindX{\tau_{1}}{\kappa_{1}}\\
%%% \tKindX{\tau_{2}}{\kappa_{2}}
%%%}{
%%% \tKindX{\tpair{\tau_{1}}{\tau_{2}}}{\kpair{\kappa_{1}}{\kappa_{2}}}
%%%}~(\times_\tau)
%%%
%%%\inferrule{
%%% \tKindX{\tau}{\kpair{\kappa_{1}}{\kappa_{2}}}
%%%}{
%%% \tKindX{\tfst{\tau}}{\kappa_{1}}
%%%}~(\text{fst}_\tau)
%%%
%%%\inferrule{
%%% \tKindX{\tau}{\kpair{\kappa_{1}}{\kappa_{2}}}
%%%}{
%%% \tKindX{\tsnd{\tau}}{\kappa_{2}}
%%%}~(\text{snd}_\tau)
%%%
%%% TODO: list
%\inferrule[k-eq]{
% \kEq{\kappa}\\
% \tKindX{\tau_1}{\kappa}\\
% \tKindX{\tau_2}{\kappa}\\\\
% \tKindX{\tau_3}{\kappa'}\\
% \tKindX{\tau_4}{\kappa'}
%}{
% \tKindX{\tifeq{\tau_1}{\tau_2}{\kappa}{\tau_3}{\tau_4}}{\kappa'}
%}
%
%\inferrule[k-ty-i]{
% \fval{tycon}{\kappaidx}{-}{-} \in \fvalCtx\\\\
% \tKind{\tvarCtx}{\fvalCtx}{\tauidx}{\kappaidx}
%}{
% \tKindX{\ttype{tycon}{\tauidx}}{\kTypeBlur}
%}
%
%\inferrule[k-ty-e]{
% \tKindX{\tau}{\kTypeBlur}\\
% \fval{tycon}{\kappaidx}{-}{-} \in \fvalCtx\\\\
% \tKind{\tvarCtxX{x}{\kappaidx}}{\fvalCtx}{\tau_1}{\kappa}\\
% \tKindX{\tau_2}{\kappa}
%}{
% \tKind{\tvarCtx}{\fvalCtx}{\tfamcase{\tau}{tycon}{x}{\tau_1}{\tau_2}}{\kappa}
%}
%~~~~~~
%\inferrule[k-d-i]{
% \tKindX{\tau_1}{\kTypeBlur}\\\\
% \tKindX{\tau_2}{\kITerm}
%}{
% \tKindX{\tden{\iota}{\tau}}{\kDen}
%}
%
%%\inferrule[k-d-i-checked]{
%%  \tKind{\Delta}{\fvalCtx}{\tau}{\kTypeBlur}\\\\
%%  \isItermX{\ibar}
%%}{
%%  \tKind{\Delta}{\fvalCtx}{\tden{\ibar}{\tau}^\checkmark}{\kDen}
%%}
%%~~~~~~~
%\inferrule[k-d-e]{
% \tKindX{\tau}{\kDen}
%}{
% \tKindX{\ttypeof{\tau}}{\kTypeBlur}
%}
%%
%%\inferrule[iterm-intro]{
%%  \isIterm{\tvarCtx}{\emptyctx}{\fvalCtx}{\iota}
%%}{
%%  \tKindX{\titerm{\iota}}{\kITerm}
%%}
%
%\inferrule[k-reptype-i]{
% \isItype{\tvarCtx}{\fvalCtx}{\sbar}
%}{
% \tKindX{\titype{\sigma}}{\kIType}
%}
%\end{mathpar}
%%$\fbox{$\isItermX{\ibar}$}$
%%\begin{mathpar}
%%\inferrule[k-i-var]{ }{
%%  \isItermX{\evar{x}}
%%}
%%~~~~~~
%%\inferrule[k-i-fix]{
%%  \isItypeX{\sbar}\\
%%  \isItermX{\ibar}
%%}{
%%  \isItermX{\ifix{\evar{x}}{\sbar}{\ibar}}
%%}
%%~~~~~~
%%\inferrule[k-i-transof]{
%%  \tKindX{\tau}{\kDen}
%%}{
%%  \isItermX{\itransof{\tau}}
%%}
%%%\inferrule[i-lam-kinding]{
%%% \isItypeX{\sbar}\\\\
%%% \isItermX{\ibar}
%%%}{
%%% \isItermX{\ilam{\evar{x}}{\sbar}{\ibar}}
%%%}
%%
%%\text{\color{gray} (omitted rules are trivially recursive, like \textsc{k-i-fix})}
%%%\inferrule[iterm-dereify-kinding]{
%%% \tKindX{\tau}{\kITerm}
%%%}{
%%% \isItermX{\iup{\tau}}
%%%}
%%%
%%\end{mathpar}
%$\fbox{$\isItypeX{\sbar}$}$
%\begin{mathpar}
%%\inferrule[i-int-kinding]{ }{
%%  \isItypeX{\dint}
%%}
%%
%%\inferrule[i-prod-kinding]{
%%  \isItypeX{\sigma_1}\\\\
%%  \isItypeX{\sigma_2}
%%}{
%%  \isItypeX{\dpair{\sigma_1}{\sigma_2}}
%%}
%%
%\inferrule[k-s-unquote]{
% \tKindX{\tau}{\kIType}
%}{
% \isItypeX{\dup{\tau}}
%}
%
%%\inferrule[k-s-repof]{
%%  \tKindX{\tau}{\kTypeBlur}
%%}{
%%  \isItypeX{\trepof{\tau}}
%%}
%%
%\text{\color{gray} (omitted forms are trivially recursive)}
%\end{mathpar}
%\caption{\small Kinding for type-level terms. The kinding context $\tvarCtx$ maps type variables to kinds.}
%\label{tlkind}
%\end{figure} 
%Some of the kinding rules for type-level terms are shown.
%
%\section{Complex Numbers}
%We have implemented complex numbers (which have a logic similar to decimals, which we added late in the paper cycle). The API is somewhat dated, apologies for
%
%If \lstinlinep{c} is a complex number, then \lstinlinep{c.ni} and \lstinlinep{c.i} are its non-imaginary and imaginary components, respectively. These expressions are of the form \lstinlinep{Attribute}, so the typechecker calls \lstinlinep{type_Attribute} (line \ref{cplx}.7).
%This method receives the compilation context, \lstinlinep{context}, and the abstract syntax tree of the expression, \lstinlinep{node} and must return a type assignment for the node, or raise an \lstinlinep{ace.TypeError} if there is an error. In this case, a type assignment is possible if the attribute name is either \lstinlinep{"ni"} or \lstinlinep{"i"}, and an error is raised otherwise (lines \ref{cplx}.8-\ref{cplx}.10). We note that error messages are an important and sometimes overlooked facet of {ease-of-use} \cite{marceau2011measuring}. A common frustration with using general-purpose abstraction mechanisms to encode an abstraction is that they can produce  verbose and cryptic error messages that reflect the implementation details instead of the semantics. @\texttt{lang} supports custom error messages.%Active types permit direct encoding of semantics and customization of error messages.
%
%Complex numbers also support binary arithmetic operations partnered with both other complex numbers and with non-complex numbers, treating them as if their imaginary component is zero. The typechecking rules for this logic is implemented on lines \ref{cplx}.17-\ref{cplx}.29. Arithmetic operations are usually symmetric, so the dispatch protocol checks the types of both subexpressions for support. To ensure that the semantics remain deterministic in the case that both types support the binary operation, @\texttt{lang} asks the left first (via \lstinlinep{type_BinOp_left}), asking the right (via \lstinlinep{type_BinOp_right}) only if the left indicates an error. In either position, our implementation begins by recursively assigning a type to the other operand in the current context via the \lstinlinep{context.type} method (line \ref{cplx}.24). If supported, it applies the C99 rules for arithmetic operations to determine the resulting type (via \lstinlinep{c99_binop_t}, not shown). 
%
%Finally, a complex number can be constructed inside an @\texttt{lang} function using @\texttt{lang}'s special constructor form: \lstinlinep{[clx.Cplx](3,4)} represents $3+4i$, for example. The term within the braces is evaluated at \emph{compile-time}. Because \lstinlinep{clx.Cplx} evaluates not to an active type, but to a class, this form is assigned a type by handing control to the class object via the \emph{class method} \lstinlinep{type_New}. It operates as expected, extracting the types of the two arguments to construct an appropriate complex number type (lines \ref{cplx}.50-\ref{cplx}.57), raising a type error if the arguments cannot be promoted to a common type according to the rules of C99 or if two arguments were not provided.% (an exercise for the reader: modify this method to also allow a single argument for when the imaginary part is 0). 
%
%As seen in Listing \ref{mapout}, we are implementing complex numbers internally using OpenCL vector types, like \lstinlinep{int2}. Let us look first at \lstinlinep{trans_New} on lines \ref{cplx}.54-\ref{cplx}.60, where new complex numbers are translated to vector literals by invoking \lstinlinep{target.VecLit}. This will ultimately  generate the necessary OpenCL code, as a string, to complete compilation (these strings are not directly manipulated by extensions, however, to avoid problems with, e.g. precedence). For it to be possible to reason compositionally about the correctness of compilation, all complex numbers must translate to terms in the target language that have a consistent target type. The \lstinlinep{trans_type} method of the \lstinlinep{ace.ActiveType} associates a type in the target language, here a vector type like \lstinlinep{int2}, with the active type. @\texttt{lang} supports a mode where this \emph{representational consistency} is dynamically checked during compilation (requiring that the active target know how to assign types to terms in the target language, which can be done  for our OpenCL target as of this writing).
%
%The translation methods for attributes (line \ref{cplx}.12) and binary operations  (line \ref{cplx}.31) proceed in a straightforward manner. The context provides a method, \lstinlinep{trans}, for recursively determining the translation of subexpressions as needed. Of note is that the translation methods can assume that typechecking succeeded. For example, the implementation of \lstinlinep{trans_Attribute} assumes that if \lstinlinep{node.attr} is not \lstinlinep{'ni'} then it must have been \lstinlinep{'i'} on line \ref{cplx}.14, consistent with the implementation of \lstinlinep{type_Attribute} above it. Typechecking and translation are separated into two methods to emphasize that typechecking is not target-dependent, and to allow for more advanced uses, like type refinements and hypothetical typing judgements, that we do not describe here.
%
%\begin{codelisting}
%\lstinputlisting{listing9.py}
%\caption{\texttt{[}in \texttt{examples/clx.py]} The active type family \texttt{Ptr} implements the semantics of OpenCL pointer types.}
%\label{cplx}
%\end{codelisting}
%\section{Expressiveness}\label{examples}
%Thus far, we have focused mainly on the OpenCL target and shown examples of fairly low-level active types: those that implement OpenCL's primitives (e.g. \lstinlinep{clx.Ptr}), extend them in simple but convenient ways (e.g. \lstinlinep{clx.Cplx}) and those that make interactive execution across language boundaries safe and convenient (e.g. \lstinlinep{clx.NPArray}). 
%@\texttt{lang} was first conceived to answer the question: \emph{can we build a statically-typed language with the semantics of a C but the syntax and ease-of-use of a Python?} We submit that the work described above has answered this question in the affirmative. 
%
%But @\texttt{lang} has proven useful for more than low-level tasks like programming a GPU with OpenCL. We now describe several interesting extensions that implement the semantics of primitives drawn from a range of different language paradigms, to justify our claim that these mechanisms are highly expressive. 
%
%\subsection{Growing a Statically-Typed Python Inside an @\texttt{lang}}
%@\texttt{lang} comes with a target, base and type implementing Python itself: \lstinlinep{ace.python.python}, \lstinlinep{ace.python.base} and \lstinlinep{ace.python.dyn}. These can be supplemented by additional active types and used as the foundation for writing actively-typed Python functions. These functions can either be compiled ahead-of-time to an untyped Python file for execution, or be immediately executed with just-in-time compilation, just like the OpenCL examples we have shown. Many of the examples in the next section support this target in addition to the OpenCL target we have focused on thus far.
%
%\section{Recursive Labeled Sums}
%Listing 12 shows how recursive labeled sums (i.e. functional datatypes) would look.
%\begin{codelisting}
%\lstinputlisting{datatypes_t.py}
%\caption{\texttt{[datatypes\_t.py]} An example using statically-typed functional datatypes.}
%\label{datatypest}
%\end{codelisting}
%\begin{codelisting}
%\lstinputlisting{datatypes.py}
%\caption{\texttt{[datatypes.py]} The dynamically-typed Python code generated by running \texttt{acec datatypes\_t.py}.}
%\label{datatypes}
%\end{codelisting}
%Listing \ref{datatypest} shows an example of the use of statically-typed functional datatypes (a.k.a. recursive labeled sum types) together with the Python implementation just described. It shows two syntactic conveniences that were not mentioned previously: 1) if no target is provided to \lstinlinep{ace.fn}, the base can provide a default target (here, \lstinlinep{ace.python.python}); 2) a concrete function can be generated immediately by providing a type assignment after \lstinlinep{ace.fn} using braces. Listing \ref{datatypest} generates Listing \ref{datatypes}.
%
%Lines \ref{datatypest}.3-\ref{datatypest}.11 define a function that generates a recursive algebraic datatype representing a tree given a name and another @\texttt{lang} type for the data at the leaves. This type implemented by the active type family \lstinlinep{fp.Datatype}. A name for the datatype and the case names and types are provided programmatically on lines \ref{datatypest}.3-\ref{datatypest}.8. To support recursive datatypes, the case names are enclosed within a lambda term that will be passed a reference to the datatype itself. These lines also show two more active type families: units (the type containing just one value), and immutable pairs. Line \ref{datatypest}.13 calls this function with the @\texttt{lang} type for dynamic Python values, \lstinlinep{py.dyn}, to generate a type, aliased \lstinlinep{DT}. This type is implemented using class inheritance when the target is Python, as seen on lines \ref{datatypes}.1-\ref{datatypes}.13. For C-like targets, a \lstinlinep{union} type can be used (not shown).
%
%The generic function \lstinlinep{depth_gt_2} demonstrates two features. First, the base has been setup to treat the final expression in a top-level branch as the return value of the function, consistent with typical functional languages. Second, the \lstinlinep{case} ``method'' on line \ref{datatypest}.17 creatively reuses the syntax for dictionary literals to express a nested pattern matching construct. The patterns to the left of the colons are not treated as expressions (that is, a type and translation is never recursively assigned to them). Instead, the active type implements the standard algorithm for ensuring that the cases cover all possibilities and are not redundant \cite{pfpl}. If the final case were omitted, for example, the algorithm would statically indicate an error, just as in statically-typed functional languages like ML.
%
%%\subsection{Nested Data Parallelism}
%%Functional constructs have shown promise as the basis for a nested data-parallel parallel programming model. Many powerful parallel algorithms can be specified by composing primitives like \lstinlinep{map} and \lstinlinep{reduce} on persistent data structures like lists, trees and records. Data dependencies are directly encoded in these specifications, and automatic code generation techniques have shown promise in using this information to automatically execute these specifications on concurrent hardware and networks. The Copperhead programming language, for example, is based in Python as well and allows users to express functional programs over lists using common functional primitives, compiling them to CUDA code \cite{catanzaro2011copperhead}. By combining active types and the metaprogramming facilities described in Sec. 2 to implement optimizations, like fusion, on the typed syntax trees available from concrete functions, these same techniques can be implemented as an @\texttt{lang} library as well.
%
%\subsection{Product Types}
%Product types (types that group heterogeneous values together) like structs, unions, tuples, records and objects can all be implemented as @\texttt{lang} type families  parameterized by a dictionary mapping field names (indices, in the case of tuples) to types. Each family differs slightly in the semantics of the operations available. Structs support mutation, for example, while records do not. Object systems typically introduce additional logic for calling methods, binding special variables like \lstinlinep{this}, and accessing information through an inheritance hierarchy. Listing \ref{ooclxfppy} shows an example of each of these. The prototypic object system delegates to a backing record if a field is not available in the foreground. This example also demonstrates cross-compilation to C99, supported for a subset of operations in \lstinlinep{clx}. All three abstractions are implemented as simple \lstinlinep{struct}s, as can be seen in Listing \ref{ooclxfpc}. This example further demonstrates the use of argument annotations to generate concrete functions (line \ref{ooclxfppy}.9), which is a feature only available when using Python 3.3+. 
%
%\subsection{Distributed Programming}
%Many distributed programming abstractions can be understood as implementing object models with complex forms of dynamic dispatch. For example, in Charm++, dynamic dispatch involves message passing over a dynamically load-balanced network \cite{kale2009charm++}. While Charm++ is a sophisticated system, and we do not anticipate that implementing it as an @\texttt{lang} extension would be easy, it is possible to do so by targeting a system like Adaptive MPI, which exposes the Charm++ runtime system \cite{kale2009charm++}. %Erlang and other message-passing and actor-based systems can also be considered in this manner -- processes can be thought of as objects, channels as methods and messages as calls.
%
%%\subparagraph{Annotation and Extension Inference}
%%In addition to type annotations, OpenCL normally asks for additional annotations in a number of other situations.  Users can annotate functions that meet certain requirements  with the \lstinlinep{__kernel} attribute, indicating that they are callable from the host. The \lstinlinep{OpenCL} backend can check these and add this annotation automatically. Several types (notably, \lstinlinep{double}) and specialized primitive functions require that an OpenCL extension be enabled via a \lstinlinep{#pragma}. The OpenCL backend automatically detects many of these cases as well, adding the appropriate \lstinlinep{#pragma} declaration. An example of this can be seen in Listing \ref{mapout}, where the use of the \lstinlinep{double} type triggers the insertion of the \lstinlinep{cl_khr_f64} extension.
%%
%
%A number of recent languages designed for distributed programming on clusters use a partitioned global address space model, e.g. Chapel \cite{chapel}. These languages provide first-class support for accessing data transparently across a massively parallel cluster. Their type systems track information about \emph{data locality} so that the compiler can emit more efficient code. The extension mechanism can also track this information in a manner analogous to how address space information is tracked in the OpenCL example above, and target an existing portable run-time such as GASNet \cite{bonachea2002gasnet}.
%
%\subsubsection{Units of Measure}
%A number of domain-specific type systems can be implemented within @\texttt{lang} as well. For example, prior work has considered tracking units of measure (e.g. grams) statically to validate scientific code \cite{conf/cefp/Kennedy09}. This cannot easily be implemented using many existing abstraction mechanisms because this information should only be maintained statically to avoid excessive run-time overhead associated with tagging/boxing, and the typechecking logic is reasonably complex. The @\texttt{lang} extension mechanism allows this information to be tracked in the type system, but not included during translation, by a mechanism similar to how address space information is tracked with pointers. Because @\texttt{lang} extensions can use Python, the needed logic can be implemented directly.
%
%
%\subsection{OOFPCLX}
%\begin{codelisting}[t]
%\lstinputlisting{ooclxfp.py}
%\caption{\texttt{[ooclxfp.py]} An example combining structs and immutable records using a prototype-based object system, cross-compiled to C99. Uses Python 3 argument annotations.}
%\label{ooclxfppy}
%\end{codelisting}
%\begin{codelisting}[t]
%\lstinputlisting[style=OpenCL]{ooclxfp.c}
%\caption{\texttt{[ooclxfp.c]} The C99 code generated by running \texttt{acec ooclxfp.py}.}
%\label{ooclxfpc}
%\end{codelisting}

\end{document}