### 2014-08-05 sanitation-psp14/sanitation-psp14.tex
Added sec 3 back to the long version

\documentclass{acm_proc_article-sp}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{afterpage}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}

% Hack. Something's wrong with PLAS paper when using the ACM Proc docclass
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}

\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
    \makeatletter
\usepackage{enumitem}
\usepackage{enumerate}

% \btIfInRange{number}{range list}{TRUE}{FALSE}
%
% Test if int number <number> is element of a (comma separated) list of ranges
% (such as: {1,3-5,7,10-12,14}) and processes <TRUE> or <FALSE> respectively
%
        \newcount\bt@rangea
        \newcount\bt@rangeb

        \newcommand\btIfInRange[2]{%
            \global\let\bt@inrange\@secondoftwo%
            \edef\bt@rangelist{#2}%
            \foreach \range in \bt@rangelist {%
                \afterassignment\bt@getrangeb%
                \bt@rangea=0\range\relax%
                \pgfmathtruncatemacro\result{ ( #1 >= \bt@rangea) && (#1 <= \bt@rangeb) }%
                \ifnum\result=1\relax%
                    \breakforeach%
                    \global\let\bt@inrange\@firstoftwo%
                \fi%
            }%
            \bt@inrange%
        }

        \newcommand\bt@getrangeb{%
            \@ifnextchar\relax%
            {\bt@rangeb=\bt@rangea}%
            {\@getrangeb}%
        }

        \def\@getrangeb-#1\relax{%
            \ifx\relax#1\relax%
                \bt@rangeb=100000%   \maxdimen is too large for pgfmath
            \else%
                \bt@rangeb=#1\relax%
            \fi%
        }

%
% \btLstHL{range list}
%
        \newcommand{\btLstHL}[1]{%
            \btIfInRange{\value{lstnumber}}{#1}%
            {\color{black!10}}%
            {\def\lst@linebgrd}%
        }%

%
% \btInputEmph[listing options]{range list}{file name}
%
        \newcommand{\btLstInputEmph}[3][\empty]{%
            \lstset{%
                linebackgroundcolor=\btLstHL{#2}%
                \lstinputlisting{#3}%
            }% \only
        }

% Patch line number key to call line background macro
        \lst@Key{numbers}{none}{%
            \def\lst@PlaceNumber{\lst@linebgrd}%
            \lstKV@SwitchCases{#1}{%
                none&\\%
                left&\def\lst@PlaceNumber{\llap{\normalfont
                \lst@numberstyle{\thelstnumber}\kern\lst@numbersep}\lst@linebgrd}\\%
                right&\def\lst@PlaceNumber{\rlap{\normalfont
                \kern\linewidth \kern\lst@numbersep
                \lst@numberstyle{\thelstnumber}}\lst@linebgrd}%
            }{%
                \PackageError{Listings}{Numbers #1 unknown}\@ehc%
            }%
        }

% New keys
        \lst@Key{linebackgroundcolor}{}{%
            \def\lst@linebgrdcolor{#1}%
        }
        \lst@Key{linebackgroundsep}{0pt}{%
            \def\lst@linebgrdsep{#1}%
        }
        \lst@Key{linebackgroundwidth}{\linewidth}{%
            \def\lst@linebgrdwidth{#1}%
        }
        \lst@Key{linebackgroundheight}{\ht\strutbox}{%
            \def\lst@linebgrdheight{#1}%
        }
        \lst@Key{linebackgrounddepth}{\dp\strutbox}{%
            \def\lst@linebgrddepth{#1}%
        }
        \lst@Key{linebackgroundcmd}{\color@block}{%
            \def\lst@linebgrdcmd{#1}%
        }

% Line Background macro
        \newcommand{\lst@linebgrd}{%
            \ifx\lst@linebgrdcolor\empty\else
                \rlap{%
                    \lst@basicstyle
                    \color{-.}% By default use the opposite (`-`) of the current color (`.`) as background
                    \lst@linebgrdcolor{%
                        \kern-\dimexpr\lst@linebgrdsep\relax%
                        \lst@linebgrdcmd{\lst@linebgrdwidth}{\lst@linebgrdheight}{\lst@linebgrddepth}%
                    }%
                }%
            \fi
        }

 % Heather-added packages for the fancy table
 \usepackage{longtable}
 \usepackage{booktabs}
 \usepackage{pdflscape}
 \usepackage{colortbl}%
 \newcommand{\myrowcolour}{\rowcolor[gray]{0.925}}
 \usepackage{wasysym}

    \makeatother

\lstset{
  language=Python,
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  commentstyle=\itshape\color{light-gray},
  basicstyle=\ttfamily\scriptsize,
  morekeywords={lambda, self, assert, as, cls},
  numbers=left,
  numberstyle=\scriptsize\color{light-gray}\textsf,
  xleftmargin=2em,
  stringstyle=\color{mauve}
}
\lstdefinestyle{Bash}{
    language={}, 
    numbers=left,
    numberstyle=\scriptsize\color{light-gray}\textsf,
    moredelim=**[is][\color{blue}\bf\ttfamily]{`}{`},
}
\lstdefinestyle{OpenCL}{
	language=C++,
	morekeywords={kernel, __kernel, global, __global, size_t, get_global_id, sin, printf, int2}
}

\usepackage{float}
\floatstyle{ruled}
\newfloat{codelisting}{tp}{lop}
\floatname{codelisting}{Listing}
\setlength{\floatsep}{10pt}
\setlength{\textfloatsep}{10pt}

\usepackage{url}

\usepackage{todonotes}

\usepackage{placeins}

\usepackage{textpos}

\renewcommand\topfraction{0.85}
\renewcommand\bottomfraction{0.85}
\renewcommand\textfraction{0.1}
\renewcommand\floatpagefraction{0.85}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bussproofs}    % Gentzen-style deduction trees *with aligned sequents!*
\usepackage{mathpartir}	% For type-settting type checking rule figures
\usepackage{syntax}		% For type-setting formal grammars.

\usepackage{hyperref}		% For links in citations

\usepackage{float}			% Make figures float if [H] option is passed.

\iffalse
\usepackage{listings}		% For typesetting code listings
\usepackage{callout}
\usepackage{titlesec}
\usepackage[T1]{fontenc}
\usepackage{upquote}
\lstset{upquote=true}
\fi

\usepackage{textcomp}		% For \textquotesingle as used in introduction
\usepackage{color}			% for box colors, like in TAPL.

\usepackage{amsmath}		% Begin Carthage default packages
\usepackage{makeidx}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphics}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{epsfig}
\usepackage{csquotes}
\usepackage{enumitem}

\newtheorem{thm}{Theorem}                                                       
\newtheorem{cor}[thm]{Corollary}                                                
\newtheorem{lem}[thm]{Lemma}                                                    
\newtheorem{prop}[thm]{Proposition}                                             
\newtheorem{ax}[thm]{Axiom}                                                     
\theoremstyle{definition}                                                       
\newtheorem{defn}[thm]{Definition}                                              
\newtheorem{exam}[thm]{Example}                                                 
\newtheorem{rem}[thm]{Remark}

%
% For type setting inference rules with labels.
%
\newcommand{\inferlbl}[3]
			{\inferrule{#3}{#2}{\textsf{\footnotesize{\sc #1}}}}
\newcommand{\inferline}[3]
			{\inferrule{#3}{#2} & {\textsf{\footnotesize{\sc #1}}} \\ \\}

\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\lang}[1]{\Lagr\{#1\}}
\newcommand{\stru}[2]{ {\tt string\_union}(#1,#2)}

\newcommand{\dconvert}[2]{ {\tt dconvert}(#1,#2) }
\newcommand{\filter}[2]{ {\tt filter}(#1,#2) }
\newcommand{\ifilter}[2]{ {\tt ifilter}(#1,#2) }

\newcommand{\reduces}{ \Rightarrow }
\newcommand{\gvd}{\Gamma \vdash }
\newcommand{\ovd}{\Omega \vdash }

\newcommand{\trep}{{\tt rep}}

\newcommand{\tstrf}[1]{`#1\textrm'} %??
\newcommand{\strf}[1]{``#1"}

\newcommand{\iso}{\cong}
%%
%% Source and Target language definitions.
%%
\newcommand{\lambdas}{\lambda_S}
\newcommand{\lambdap}{\lambda_P}

% Source language terms.
\newcommand{\sisubst}[3]{{\sf rreplace}[#1](#2,#3)} \newcommand{\rreplace}[3]{{\sf rreplace}[#1](#2,#3)} % lots of legacy naming around these parts...

\newcommand{\rssreplace}[3]{\sisubst{#1}{#2}{#2}} % TODO-nrf fix this.

\newcommand{\coerce}[2]{ {\sf rcoerce}[#1](#2)}
\newcommand{\rcoerce}[2]{{\sf rcoerce}[#1](#2)}
\newcommand{\sistr}[1]{{\sf rstr}[#1]}   \newcommand{\rstr}[1]{{\sf rstr}[#1]} % Lots of legacy naming around these parts...

\newcommand{\strin}[1]{\sistr{#1}}
\newcommand{\rsconcat}[2]{{\sf rconcat}(#1,#2)} \newcommand{\rconcat}[2]{{\sf rconcat}(#1,#2)} % lots of legact naming around these parts..

% Source language types.
\newcommand{\stringin}[1]{{\sf stringin}[#1]}

% target language terms.
\newcommand{\tsubst}[3]{{\sf replace}(#1,#2,#3)} \newcommand{\metareplace}[3]{{\sf replace}(#1,#2,#3)} % TODO-nrf rename the commands. Lots of legacy naming around these parts...

\newcommand{\tcheck}[2]{{\sf check}(#1, #2)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\preplace}[3]{{\sf preplace}(#1,#2,#3)}
\newcommand{\tconcat}[2]{{\sf concat}(#1,#2)} \newcommand{\concat}[2]{{\sf concat}(#1,#2)} % lots of legacy naming around these parts...

\newcommand{\regext}[1]{ {\sf rx}[#1] } % TODO-nrf remove
\newcommand{\rx}[1]{ {\sf rx}[#1] }

% Target language types
\newcommand{\str}{{\sf string}}
\newcommand{\regex}{{\sf regex}}

% Meta-theoretic functions
\newcommand{\dosubst}[3]{{\tt replace}(#1,#2,#3)}
\newcommand{\lsubst}[3]{{\tt lsubst}(#1,#2,#3)} % This used to render lreplace(...) so there're probably mistkes wherever this command was used now.
\newcommand{\lreplace}[3]{{\tt lreplace}(#1, #2, #3)}

% Judgements
\newcommand{\err}{\ {\sf err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

% Relations
\newcommand{\treduces}{ \Downarrow }
\newcommand{\sreduces}{ \Downarrow }

%%
%% Constrain the size of full-page diagrams and rule lists
%%
%%\newcommand{\pagewidth}{5in}
%%\newcommand{\rulelistwidth}{3in}

% Names of type systems presented in paper
\newcommand{\lcs}{\lambda_{CS}}

\setlength{\grammarindent}{3em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%dep_docclass\conferenceinfo{-}{-} 
%dep_docclass\copyrightyear{-} 
%dep_docclass\copyrightdata{[to be supplied]}

%\titlebanner{{\tt \textcolor{Red}{{\small Under Review -- distribute within CMU only.}}}}        % These are ignored unless
%\preprintfooter{Distribute within CMU only.}   % 'preprint' option specified.

\newcommand{\Ace}{\textsf{Ace}}

\title{A Type System for String Sanitation\\Implemented Inside a Python}

%\authorinfo{~}{~}{~}
%dep_docclass\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
%dep_docclass           {School of Computer Science\\
%dep_docclass            Carnegie Mellon University}
%dep_docclass           {\{nathanfu, comar, aldrich\}@cs.cmu.edu}

\maketitle
\begin{abstract}
%Evidence suggests that programmers are reluctant to adopt new languages to
Security-oriented programming languages do not enjoy wide adoption. Conversely, library and framework-based approaches to security are pervasive. Unfortunately, libraries and frameworks are only effective solutions if they are implemented and used properly. In this paper, we propose that extensible programming languages provide a promising mechanism for increasing the amount of trust that can be place in libraries and frameworks. To demonstrate this approach, we present a conservative extension to the simply-typed lambda calculus for checking the correctness of input sanitation algorithms. We also demonstrate how this language's security guarantees are preserved under translation into an underlying language containing a regular expression library. We believe this approach -- complementing existing techniques with light-weight, composable type-based analyses -- constitutes a promising compromise between usability, potential for industry adoption, and theoretically grounded safety guarantees.
\end{abstract}

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Improper input sanitation is a leading cause of security vulnerabilities in web applications [OWASP]. Command injection attacks exploit improper input sanitation by inserting malicious code into an otherwise benign command. Modern web frameworks, libraries, and database abstraction layers attempt to ensure proper sanitation of user input. When these methods are unavailable or insufficient, developers implement custom sanitation techniques. Im both cases, sanitation algorithms are implemented using the language's regular expression capabilities and usually \emph{replace} potentially unsafe strings with equivalent escaped strings.

In this paper, we present a type system for implementing and statically checking input sanitation techniques. Our solution suggests a more general approach to the integration of security concerns into programming language design. This approach is characterized by \emph{composable} type system extensions which \emph{complement} existing and well-understood solutions with compile-time checks.

To demonstrate this approach, we present a simply typed lambda calculus with \emph{constrained strings}; that is, a set of string types parameterized by regular expressions. If $s : \stringin{r}$, then $s$ is a string matching the language $r$. Additionally, we include an operation $\rreplace{r}{s_1}{s_2}$ which corresponds to the replace mechanism available in most regular expression libraries; that is, any substring of $s_1$ matching $r$ is replaced with $s_2$. The type of this expression is the computed, and is likely ``smaller" or more constrained than the type of $s_1$. Libraries, frameworks or functions which construct and execute commands containing input can specify a safe subset $\stringin{r_{\sf spec}}$ of strings, and input sanitation algorithms can construct such a string using ${\sf rreplace}$ or, optionally, by coercion (in which case a runtime check is inserted). We also show how this simple can be translated into a host language containing a regular expression library such that the safety guarantee of the extended language is preserved.

Summarily, we present a simple type system extension which ensures the absence of input sanitation vulnerabilities by statically checking input sanitation algorithms which use an underlying regular expression library. This approach is \emph{composable} in the sense that it is a conservative extension. This approach is also \emph{complementary} to existing input sanitation techniques which use string replacement for input sanitation.

%%
% This is the paragraph which ties the paper back to the CFP.
%
% For refernce, here's a copy of the cfp:
%The importance of security and privacy in software engineering is now a pressing concern. 
%Over the last decade, many programmers have recognized the importance of including 
%security and privacy requirements at project start. This has led to efforts 
%in secure/security engineering and privacy engineering, which focus on guidelines and 
%best practices that can be used at the design stage to create safer code. 
%Unfortunately, these disciplines are not pervasive and still in their infancy.
%The goal of this workshop is to make security and privacy first class citizens in programming and programming languages.

%This forum aims to gather research and industry professionals to further the discussion on how to:
%   codify the principles from secure and privacy engineering into programming language constructs and or tools, 
%   create programming languages that have security and privacy as foundational tenets, and 
%   create/codify constructs or tools that enables secure and privacy-preserving (business) operations.
%   All submissions connected to the goals above, and their associated topics, are encouraged.
%%

\subsection{Related Work and Alternative Approaches}

The input sanitation problem is well-understood. There exist a large number of techniques and technologies, proposed by both practitioners and researchers, for preventing injection-style attacks. In this section, we explain how our approach to the input sanitation problem differs from each of these approaches. More important than these differences, however, is our more general assertion that language extensibility is a promising approach toward consideration of security goals in programming lanugage design.

Unlike \emph{frameworks and libraries} provided by languages such as Haskell and Ruby, our type system provides a \emph{static} guarantee that input is always properly sanitized before use. Doing so requires reasoning about the operations on regular languages corresponding to standard operations on strings; we are unaware of any production system which contains this form of reasoning. Therefore, even where frameworks and libraries provide a viable interface or wrapper around input sanitation, our approach is complementary because it ensurees the correctness of the framework or library itself. Furthermore, our approach is more general than database abstraction layers because our mechanism is applicable to all forms of command injection (e.g. shell injection or remote file inclusion).

A number of research languages provide static guarantees that a program is free of input sanitation vulnerabilities [Jif][Ur/Web]. Our work differes from these contributions in the ways following:

\begin{itemize}
  \item Our system is a light-weight solution to a single class of sanitation vulnerabilities (e.g. we do not address Cross-Site Scripting).
  \item Our system is defined as a library in terms of an extensible type system, as opposed to a stand-alone language. Instead of introducing new technologies and methodologies for addressing security problems, we provide a light-weight static analysis which complements approaches developers already understand well.
  \item Our implementation of the translation is implemented in Python and shares its grammar. Since Python is a popular programming language among web developers, the barrier between our research and adopted technologies is lower than for greenfield security-oriented languages.
\end{itemize}

We are also unaware of any extensible programming languages which emphasize applications to security concerns (TRUE?).

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{pierce} typechecks XML schemas using regular expressions. We differ from this and related work in at least two ways. First, our system  is defined within an extensible type system; second, and more importantly, we have demonstrated that regular expression types are applicable to the web security domain.

In conclusion, our system is novel in at least two ways:
\begin{itemize}
\item The safety guarantees provided by libraries and frameworks in popular languages
are not as (statically) justified as is often belived (or even claimed).
\item Our extension is the first major demonstration of how an extensible type
system may be used to provide light-weight, composable security analyses based upon
idiomatic code.
\end{itemize}

\subsection{Outline}

An outline of this paper follows:

\begin{itemize}
  \item In \S 2, we define the type system which is embedded in Ace. We include a type safety proof for the string
    segment of this language and prove the correctness of a translation to an underlying langguage $P$. In our theory,
    $P$ is a simply typed lambda calculus equipped with a minimal regular expression library; in an implementation, $P$
    stands in for Python or another underlying general-purpose programming language.
  \item In \S 3, we discuss our implemention of this translation as a type system extension  within the Ace programming language.
\end{itemize}

\section{A Type System for String Sanitation}

The $\lambdas$ language is characterized by a type of strings indexed by regular
expressions, together with operations on such strings which correspond to common
input sanitation patterns.
This section presents the grammar, typing rules and operational semantics for
$\lambdas$ as well as an underlying language $\lambdap$.

The system $\lambdas$ \ref{fig:lambdas} is the simply typed lambda calculus extended with \emph{regular expression types}, which are string
types ensuring a string belongs to a specified language. For instance, 
$S : \stringin{r}$ reads ``$s$ is a string matching $r$".
the system includes an operation for replacing all instances of a pattern $r$ in a string $s_1$ with another string $s_2$. 
Input sanitation algorithms -- as implemented by developers or within popular libraries and frameworks -- are often implemented in terms of this replace operation.
For instance, a developer might all potentially unsafe charapcters with excaped versions of the same character.
Regular expression types are used both to specify input sanitation algorithms, and at use sites as specifications.
Note that runtime error states ($S \err$) are introduced by coercion, not by replacement.

The language $\lambdap$ \ref{fig:lambdap} is a simple functional language extended with a minimal regular expression library.
Any general purpose programming language could stand in for $\lambdap$; for instance, SML has a regular expression library.
In an implementation, our correctness results are modulo the underlying language's correct implementation of regular expression matching (see P-E-Replace).

Finally, we define a translation from our type system $\lambdas$ into $\lambdap$.

%Note that we never insert an internal check where the type of the string implies
%that a check must succeed. Furthermore, expensive calculations (such as language
%inclusion or ${\sf rreplace}$) all occur at compile time.
%Since sanitation problems are generally expressible with smalle, simple expressions,
%we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
%support this assertion.

% Begin Grammar 
% Removes the < > from grammar productions.
\renewcommand{\grammarlabel}[2]{#1\hfill#2}

\begin{figure}[h]
\begin{grammar}
<r> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

\caption{Regular expressions over the alphabet $\Sigma$.}
\end{grammar}
\end{figure}

\begin{figure}[h]
\begin{grammar}

<$\psi$> ::=	...			\hfill	source types					\alt
$\stringin{r}$

<S> ::= ... \hfill source terms \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{S}{S}$ \alt
      $\sisubst{r}{S}{S}$ \alt
      $\coerce{r}{S}$
\caption{Syntax for the string sanitation fragment of our source language, $\lambdas$.}
\end{grammar}
\end{figure}

\begin{figure}[h]
\begin{grammar}

<$\theta$> ::= ... \hfill target types \alt
$\str$ \alt $\regex$

<P> ::= ... \hfill target terms \alt
  $\tstr{s}$ \alt
  $\rx{r}$ \alt
  $\tconcat{P}{P}$ \alt
  $\preplace{P}{P}{P}$ \alt
  $\tcheck{P}{P}$ 
  \end{grammar}
\caption{Syntax for the fragment of our target language, $\lambdap$, containing strings and statically constructed regular expressions.}
\label{fig:lcsSyntax}
\end{figure}

%The $\lcs$ language gives static semantics for common regular expression library
%functions. In this treatment, we include concatenation and filtering.
%The ${\tt filter}$ function removes all instances of a regular expression in a string,
%while concatenation ($+$) concatenates two strings.
%
%\subsection{Typing}
%
%The $\strin[r]$ type is parameterized by regular expressions; if $e:\strin[r]$,
%then $e \in r$. Mapping from an arbitrary $\str$ to a $\strin[r]$ requires
%defining an algorithm -- in terms of filter -- for converting a $\strin[.*]$
%into a $\strin[r]$. The static semantics of the language defines the types of
%operations on regular expressions in terms of well-understood properties about
%regular lanugages; we recall these properties in section 3.
%
%\subsection{Dynamics}
%
%There are two evaluation judgements: $e:T \reduces e'$ and $e:T \treduces i$.
%The $\reduces$ relation is between $\lcs$ expressions, while the $\treduces$
%relation is a mapping from $\lcs$ expressions into internal language expressions
%$i$ such that $i \ival$.
%
%Safety of the evaluation relation depends upon an injective mapping from $\lcs$ types info
%internal language types. This relation, $h$, is defined below.
%
%\subsection{Type Safety}
%
%The type safety proof relies upon some assumptions about the type system and
%dynamics of the internal language, as well as some properties of regular
%languages.
%
%There must exist a translation from $\lcs$ types to the types of the internal
%language. For the remainder of this paper, we call the type translation function
%$h$.
%
%\begin{defn}[Type Translation Function $h$]
%  The type translation function $h : Type \rightarrow IType$ is defined as follows:
%  \begin{itemize}
%    \item $\forall r. h(\strin[r]) = istr$
%    \item $h(\str) = istr$
%  \end{itemize}
%\end{defn}
%
%Additionally, we assume that the internal language contains an implementation of strings,
%together with operations for concatenation and filtering by regular expression.
%
%\begin{defn}[Types of internal values]
%  Let $\tstrf{s}$ range over string literals and $r$ over regular expressions.
%  Internal values are typed as follows:
%\begin{itemize}
%  \item If $e=\tstrf{s}$ then $e:istr$.
%  \item If $e=\ifilter{r}{\tstrf{s}}$ then $e:istr$.
%  \item If $e=\tstrf{s_1}+\tstrf{s_2}$ then $e:istr$.
%\end{itemize}
%\end{defn}
%
%For simplicity, we assume a fixed translation from
%$\lcs$ regular expressions to regular expressions recognizable by the internal
%language's regex library (in practice, a fixed translation is acceptable.)
%To summarize, we assume an internal language containing 
%a string type together with operations for string concatentation and filtering. We
%expect closure over strings for both operations.
%Finally, recall that ${\tt ifilter}$ is only needed for dynamic casts, which may be removed without
%descreasing the expressivity or even usability of the language.
%Finally, the semantics of the filter function are defined in terms of ${\tt rl\_filter}$,
%which is a static version of ${\tt ifilter}$.
%

% I can't achieve the correct latout here. We should have something like:
%       Lam_S Synax         Lam_P Syntax
%       Lam_S Sematics      LAm_P Semantics
%                Translation rules
% Or, alternatively, we can flip reflect things so that the entire page
% reads top-down (Lam_S syntax Lam_S smantics horizontally aligned w/ one
% another).

% Unfortunately, I can't get that to work. So, we're settling for something
% slightly ugly, where the fitures appear in the incorrect order and the 
% translation rules are on top.

% The translations rules *just won't* go to the bottom of the page, so 
% instead I am placing them at the top. Since the figures are out of order,
% it's even more important that we get the number gith. therefore, the counter
% is modified before and after this figure.
\addtocounter{figure}{4}
\begin{figure*}\label{fig:trans}

$\fbox{\inferrule{}{\trden{S} = P}}$
\begin{mathpar}

\inferrule[Tr-string]
{ \ }
{\trden{\strin{s}} = \tstr{s}}

\inferrule[Tr-Concat]
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2}
{\trden{\rsconcat{S_1}{S_2}} = \tconcat{P_1}{P_2}}

\inferrule[Tr-Subst]
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2 }
{\trden{ \sisubst{r}{S_1}{S_2} } = \tsubst{\rx{r}}{P_1}{P_2} }

\inferrule[Tr-Coerce-Ok]
{ S : \strin{r} \\ \lang{r'} \subseteq \lang{r}}
{\trden{ \coerce{r'}{S} } = \tstr{s}}

\inferrule[Tr-Coerce-NotOk]
{\trden{S} = P \\ S : \strin{r} \\ \lang{r'} \not \subseteq \lang{r}}
{\trden{\coerce{r'}{S}} = \tcheck{\rx{r'}}{P}}

\end{mathpar}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{figure*}
\addtocounter{figure}{-5}

%
% Begin typing/translation relation
%

\newcommand{\sctx}{\Psi} % Context for external typing
\newcommand{\tctx}{\Theta} % Context for internal typing

\newcommand{\ereduces}{\Downarrow}

\begin{figure}[t]\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash S : \psi}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \psi$
\begin{mathpar}
\inferrule[S-T-Stringin-I]
{s \in \lang{r}}
{\sctx \vdash \strin{s} : \stringin{r}}

\inferrule[S-T-Concat]
{\sctx \vdash S_1 : \stringin{r_1} \\ \sctx \vdash S_2 : \stringin{r_2}}
{\sctx \vdash \rsconcat{S_1}{S_2} : \stringin{r_1 \cdot r_2}}

\inferrule[S-T-replace]
{\sctx \vdash S_1 : \stringin{r_1} \\ \sctx \vdash S_2 : \stringin{r_2} \\\\ {\tt lsubst}(r,r_1,r_2) = r'}
{\sctx \vdash \sisubst{r}{S_1}{S_2} : \stringin{r'}}

\inferrule[S-T-coerce]
{\sctx \vdash S : \stringin{r'}}
{\sctx \vdash \coerce{r}{S} : \stringin{r}}

\end{mathpar}

\caption{Typing rules for our fragment of $\lambdas$. 
The typing context $\sctx$ is standard.}
\label{fig:etyping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{S \sreduces S}}$
$\fbox{\inferrule{}{S \err}}$
\begin{mathpar}
\inferrule[S-E-RStr]
{ \ }
{\strin{s} \sreduces \strin{s}}

\inferrule[S-E-Concat]
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2}} 
{\rsconcat{S_1}{S_2} \sreduces \strin{s_1 s_2}} % ???

\inferrule[S-E-Replace]
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2} \\ \dosubst{r}{s_1}{s_2} = s} 
{\sisubst{r}{S_1}{S_2} \sreduces \sistr{s}}

\inferrule[S-E-Coerce-OK]
{S \sreduces \strin{s} \\ s \in \lang{r}}
{\coerce{r}{S} \sreduces \strin{s}}

\inferrule[S-E-Coerce-Err]
{S \sreduces \strin{s} \\ s \not \in \lang{r}}
{\coerce{r}{S} \err}

\end{mathpar}
\caption{Big step semantics for our fragment of $\lambdas$.
Error propagation rules are omitted.}
\label{fig:eeval}
\end{figure}

\newpage

\begin{figure}[t]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash P : \theta}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \theta$

\begin{mathpar}
\inferrule[P-T-String]
{ \ }
{\tctx \vdash \tstr{s} : \str}

\inferrule[P-T-Regex]
{ \ }
{\tctx \vdash \rx{r} : \regex}

\inferrule[P-T-Concat]
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}
{\tctx \vdash \tconcat{P_1}{P_2} : \str}

\inferrule[P-T-Replace]
{\tctx \vdash P_1 : \regex \\ \tctx \vdash P_2 : \str \\ \tctx \vdash P_3 : \str }
{\tctx \vdash \preplace{P_1}{P_2}{P_3} : \str}

\inferrule[P-T-Check]
{\tctx \vdash P_1 : \regex \\ \tctx \vdash P_2 : \str \\ }
{\tctx \vdash \tcheck{P_1}{P_2} : \str}
\end{mathpar}
\caption{Typing rules for our fragment of $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:etyping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{P \treduces P}}$
$\fbox{\inferrule{}{P \err}}$

\begin{mathpar}
\inferrule[P-E-Str]
{ \ }
{\tstr{s} \treduces \tstr{s}}

\inferrule[P-E-Rx]
{ \ }
{\rx{r} \treduces \rx{r}}

\inferrule[P-E-Concat]
{P_1 \treduces \tstr{s_1} \\ P_2 \treduces \tstr{s_2}} 
{\tconcat{P_1}{P_2} \treduces \tstr{s_1 s_2}} % ???

\inferrule[P-E-Replace]
{P_1 \treduces \rx{r} \\ P_2 \treduces \tstr{s_2} \\ P_3 \treduces \tstr{s_3} \\ \dosubst{r}{s_2}{s_3} = s} 
{\preplace{P_1}{P_2}{P_3} \treduces \tstr{s}}

\inferrule[P-E-Check-OK]
{P_1 \treduces \rx{r} \\ P_2 \treduces \strin{s} \\ s \in \lang{r}}
{\tcheck{P_1}{P_2} \treduces \tstr{s}}

\inferrule[P-E-Check-Err]
{P_1 \treduces \rx{r} \\ P_2 \treduces \tstr{s} \\ s \not \in \lang{r}}
{\tcheck{P_1}{P_2} \err}
\end{mathpar}
\caption{Big step semantics for our fragment of $\lambdap$.
Error propagation rules are omitted.}
\label{fig:ieval}
\end{figure}

\newpage
\ 
\newpage
\ 
\newpage

\subsection{Properties of Regular Languages}

Our type safety proofs for languages S and P and our translation correctness
result all depend on some properties of regular languages. The crucial property
is a relationship between string substitution -- which is available in any regular
expression library -- and regular language substitution, which is a corresponding operation
on languages instead of strings \ref{thm:substcorrespondence}.
The decidability of language substitution is what enables static analysis of 
sanitation algorithms implemented in terms of string replacement

Throughout this section, we fix an alphabet $\Sigma$ over which strings $s$ and
regular expressions $r$ are defined. throughout the paper, $\lang{r}$ refers to the
language recognized by the expression $r$. This distinction between the expression
and its language -- typically elided in the literature -- makes our definition and
proofs about systems S and P more readable.

\begin{lem}{Properties of Regular Languages and Expressions.} \label{thm:regexprops}
The following are well-known properties of regular expressions which are necessary for our proofs:
\begin{description}
\item[(1):] If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1r_2}$
\item[(2):] For all strings $s$ and expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
\item[(3):] Regular languages are closed under complements and concatenation.
\item[(4):] The regular expressions correspond bijectively to the regular languages.
\end{description}
\end{lem}

\begin{defn}[$\tt{lsubst}$]
The function $\lsubst{r}{s_1}{s_2}$ produces a string in which all substrings of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

\begin{defn}[$\tt{lreplace}$]
The function $\lreplace{r}{r_1}{r_2}$ produces a regular expression in which any sublanguage $\lang{r_1'}$ of $\lang{r_1}$ satisfying the condition $\lang{r_1'} \subseteq{\lang{r}}$ is replaced with $\lang{r_2}$.
\end{defn}

\begin{lem}{Closure and Totality of Replacement.} \label{thm:total}
If $r, r_1$ and $r_2$ are regular expressions, then $\lreplace{r}{r_1}{r_2}$ is also a regular expression.
\end{lem}
\begin{proof}
By induction on $r$ and closure properties of regular expressions.
\end{proof}

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lreplace{r}{r_1}{r_2}$.
\end{lem}
\begin{proof}
The proof proceeds by structural induction on $r$.

\begin{description}
\item[case $r=\alpha$.]
If $s_1 = \alpha$ then $\alpha \in \lang{r_1}$ by assumption.
Therefore, $\lsubst{r}{s_1}{s_2} = \lsubst{\alpha}{\alpha}{s_2} = s_2$ and
$\lreplace{r}{r_1}{r_2} = \lreplace{\alpha}{r_1}{r_2}$. 
Since $s_1 = \alpha$ and $\alpha \in \lang{r_1}$, $r_1 \iso \alpha | r_1'$ for some $r_1'$.
Therefore, $\lreplace{\alpha}{r_1}{r_2} \iso \lreplace{\alpha}{\alpha|r_1'}{r_2}$ by Lemma X(1).
Finally, $s_2 \in \lang{r_2}$ which implies $s_1 \in \lang{r_2|r'}$.
If $s_1 \not = \alpha$ the $\lsubst{r}{s_1}{s_2} = s_1$ and $\lreplace{\alpha}{r_1}{r_2} = r_1$.

\item[case $r = a|b$].
Note that $\subst{a|b}{s_1}{s_2} = \lsubst{a}{\lsubst{b}{s_1}{s_2}}{s_2}$
and $\lsubst{b}{s_1}{s_2} \in \lreplace{b}{r_1}{r_2}$ by induction.
Therefore, $\lsubst{a|b}{s_1}{s_2}$ $\in \lreplace{a}{\lreplace{b}{r_1}{r_2}}{r_2}$ by induction and the definition of lreplace.
Finally, applying definitions once more, $\lsubst{a|b}{s_1}{s_2} \in \lreplace{a|b}{r_1}{r_2}$.

\item[case $r = ab$].
By a similar argument to the disjunctive case.

\item[case $r = a*$].
By considering the once unwinding of $a*$, noting that $s_1$ and $s_2$ are finite.
\end{description}
\end{proof}

% NRF The coercion lemma is no longer necessary, because it is a property of
% "strip"-style sanitation rather than the more general case of replacement 
% treated in this paper. However, we should mention this difference between my
% undergraduate thesis and this paper in the related work section if we are
% going tin clude a discussio of unpublished or quasi-published (e.g. src)
% results.
%\begin{lem}{Coercion Lemma.} \label{thm:coercionalemma}
%Suppose that $R$ and $L$ are regular expressions, and that $s \in R$ is a finite string.  
%Let $s' = \coerce(R,L,s)$ with all maximal substrings recognized by $L$ replaced with $\epsilon$.  Then $s'$ is recognized by $(R \backslash L) + \emptyset$ and the construction of $R \backslash L$ is decidable.
%\end{lem}
%\begin{proof}
%Let $F,G$ be FAs corresponding to $R$ and $L$, and let $G'$ be $G$ with its final states inverted (so that $G'$ is the complement of $L$).  Define an FA $H$ as a DFA corresponding to the NFA found by combining $F$ and $G'$ such that $H$ accepts only if $R$ and $L'$ accept or if $s$ is empty (this construction may result in an exponential blowup in state size.)  Clearly, $H$ corresponds to $R \backslash L + \emptyset$.  Thus, the construction of $R \backslash L + \emptyset$ is decidable.
%
%If $R \subset L$, $s' = \emptyset$.  If $L \subset R$, either $s' = \emptyset$, or $s' \in R$ and $s' \not \in L$. If $R$ and $L$ are not subsets of one another, then it may be the case that $L$ recognizes part of $R$.  Consider $L$ as the union of two languages, one which is a subset of $R$ and one which is disjoint.  The subset language is considered above and the disjoint language is inconsequential.
%\end{proof}

% Why is this necessary?
%\begin{lem}
%  For all $r$, if $s_1 \in r_1$ and $s_2 \in r_2$ then $\subst{r}{s_1}{s_2} \subseteq \lsubst{r}{r_1}{r_2}$.
%\end{lem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN PROOFS ABOUT S AND P
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Safety of the Source and Target Languages}

\begin{lem}
  If $\sctx \vdash S : \stringin{r}$ then $r$ is a well-formed regular expression. 
\end{lem}
\begin{proof}
  The only non-trivial case is S-T-Replace, which follows from \ref{thm:total}. 
\end{proof}

\begin{lem}
  If $\tctx \vdash P : \regex$ then $P \treduces \rx{r}$ such that $r$ is a well-formed regular expression. 
\end{lem}

We now prove safety for the string fragment of the source and target languages.

\begin{thm}{Safety for the String Fragment of P.}
  Let $S$ be a term in the source language. If $\sctx \vdash S : \stringin{r}$ 
  then $S \sreduces \rstr{s}$ and $\rstr{s} : \stringin{r}$, or else $S \err$.
\end{thm}
\begin{proof}
By induction on the derivation of $\sctx \vdash S : \psi$. The interesting case
is S-T-Replace, which requires Lemma C.

\begin{description}
\item[S-T-Stringin-I:]
If $S = \rsconcat{S_1}{S_2} : \stringin{r}$ then
$S \sreduces S$ by S-E-RStr, and $\sctx \vdash S : \psi$ by assumption.

\item[S-T-Concat:] 
Suppose $S = \rsconcat{S_1}{S_2} : \stringin{r_1 r_2}$.
By inversion, $\sctx \vdash S_1 : \stringin{r_1}$ and $\sctx \vdash S_1 : \stringin{r_2}$.
It follows by induction that either $S_1 \err$, $S_2 \err$, 
or $S_1 \sreduces \rstr{s_1}$ and $S_2 \sreduces \rstr{s_2}$ for some $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$.
In the latter case $S \reduces \rstr{s_1s_2}$ by S-E-Concat and $\sctx \vdash \rstr{s_1s_2} : \tstr{r_1r_2}$ by \ref{thm:regexprops}.
In the former cases, $S \err$.

\item[S-T-Replace:]
Suppose $S = \rreplace{r}{S_1}{S_2}$ and $\sctx \vdash S : \stringin{r'}$.
By inversion $\sctx \vdash S_1 : \stringin{r_1}$ and $\sctx \vdash S_2 : \stringin{r_2}$ such that $\lsubst{r}{r_1}{r_2} = r'$.
By induction, $S_1 \err$, $S_2 \err$ 
or $S_1 \sreduces \rstr{s_1}$ and $S_2 \sreduces \rstr{s_2}$ such that 
In the latter case, we know $\lreplace{r}{s_1}{s_2} \in \lang{\lsubst{r}{r_1}{r_2}}$ by Lemma C;
therefore by S-E-Replace, $S \sreduces \rstr{s}$ such that  $s \in \lang{\lsubst{r}{r_1}{r_2}} = \lang{r'}$.
So by S-T-String-I, $\rstr{s} : \stringin{r'}$.
In the former cases, $S \err$.

\item[S-T-Coerce:]
Suppose $S = \coerce{r}{S_1}$ and $S : \stringin{r}$.
By inversion, $\sctx \vdash S_1 : \stringin{r'}$.
By induction, $S_1 \err$ or $S_1 \sreduces \rstr{s}$.
In the former case $S \err$ by propagation rules.
In the latter case we have by property 2 of \ref{thm:regexprops} that $s \in \lang{r}$ or else
$s \not \in \lang{r}$.
If $s \in \lang{r}$ then $\rstr{s} : \stringin{r}$.
If $s \not \in \lang{r}$ then $S \err$. 
\end{description}
\end{proof}

\begin{thm}
  Let $P$ be a term in the target language. If $\tctx \vdash P : \theta$ 
  then $P \sreduces P'$ and $P' : \theta$.
\end{thm}
\begin{proof}
The proof proceeds by induction on the typing relation and is trivial give
and inversion lemma for the typing relation. \textbf{We can write up this proof if we end up having enough space...}
\end{proof}

\subsection{Translation Correctness}

We now present the main correctness result.

\begin{thm}
If $S : \strin{r}$ then there exists a P such that $\trden{s} = P$ and either: 
\begin{description}
  \item[(a)] $P \treduces \tstr{s}$ and $S \sreduces \strin{s}$, and $s \in lang{r}$.
  \item[(b)] $P \err$ and $S \err$.
  \end{description}
\end{thm}

\begin{proof}
The proof proceeds by induction on the typing relation for $S$.
Throughout the proof, properties from the closure lemma for regular languages
are necessary; for brevity, we elide these references.

\begin{description}
\item[S-T-String-I:] 
Let $S = \rstr{s}$ and suppose $\sctx \vdash \rstr{s} : \stringin{r}$.
Choose $T = \str{s}$ and note that $\trden{S} = P$ by Tr-String.
By P-E-String, $P \treduces \str{s}$ and by S-E-String $S \treduces \rstr{s}$.
Finally, by inversion of S-T-Stringin-I, $s \in \lang{r}$.

\item[S-T-Concat:]
Let $S = \rconcat{S_1}{S_2}$ and suppose $\sctx \vdash S : \stringin{r_1 r_2}$.
By inversion, $\sctx \vdash S_1 : \stringin{r_1}$.
It follows by induction that there exists a $P_1$ such that $\trden{S_1} = P_1$.
By a similar argument for $S_2$ and $r_2$, there exists a $P_2$ such that $\trden{S_2} = P_2$.
Choose $P = \concat{P_1}{P_2}$.

We first prove property (a).
Note that $S_1$ and $P_1$ are well typed (nrf ACTUALLY WE DON'T KNOW THAT $P_1$ IS WELL-TYPED!) and do not result in errors.
Therefore, $S_1 \sreduces \rstr{s_1}$ and $P_1 \treduces \str{s_1}$ for some $s_1 \in \lang{r_1}$ by theorems 8 and 9 respectively. % TODO-nrf make sure this numbering is correct
Similarly, $S_2 \sreduces \rstr{s_2}$ and $P_2 \treduces \str{s_2}$ for some $s_2 \in \lang{r_2}$.
Therefore, $S \sreduces \rstr{s_1 s_2}$ by S-E-Concat and $\concat{P_1}{P_2} \treduces \str{s_1 s_2}$ by P-E-Concat.
Finally, $s_1s_2 \in \lang{r_1}{r_2}$ by \ref{thm:regexprops}.

Consider property (b).
If $S_1 \err$ then $P_1 \err$ by induction, and it follows that $S \err$ and $P \err$ by respective error propagation rules.
Similarly, if $S_2 \err$ then $P_2 \err$ and it follow that $S \err$ and $P \err$ by induction and propagation.

\item[S-T-Replace:]
Let $S = \rreplace{r}{S_1}{S_2}$ and suppose $\sctx \vdash S : \stringin{r'}$ for some $s$.
By inversion of S-T-Replace, , $\sctx S_1 : \stringin{r_1}$ and $\sctx : \stringin{r_2}$ 
such that $\lsubst{r}{r_1}{r_2} = r'$.
By induction, there exists some $P_1, P_2$ such that $\trden{S_1} = P_1$, $\trden{S_2} = P_2$
and either (a) or (b) holds.

If (a) holds then $S_1 \sreduces \rstr{s_1}$ and $P_1 \treduces \str{s_1}$ for some $s_1 \in \lang{r_1}$,
and similarly for $S_2, P_2$ and some $s_2 \in \lang{r_2}$.
Therefore, by S-E-Replace, $S \sreduces \rstr{s}$ for some $s = \lreplace{r}{s_1}{s_2}$.
Choose $P = \preplace{r}{s_1}{s_2}$.
By a similar argument and P-E-Replace, $P \treduces \str{s}$ for some $s = \lreplace{r}{s_1}{s_2}$.
What remains to be shown is $\lreplace{r}{s_1}{s_2} \in \lang{\lsubst{r}{r_1}{r_2}}$,
which follows from Leamm D since $s_1 \in r_1$ and $s_2 \in r_2$.

If (b) holds for $S_1$ and $P_1$, then $S \err$ and $P \err$ by propagation rules.
Similarly, if (b) hols for $S_2$ and $P_2$ then $S \err$ and $P \err$ by propagation rules.

\item[S-T-Coerce:]
Let $S = \rcoerce{r}{S'}$ and suppose $\sctx \vdash \rcoerce{r}{S} : \stringin{r}$.
By inverstion $\Psi \vdash S' : \stringin{r'}$ for an arbitrary $r'$.
By induction there exists a $P'$ such that $\trden{S'} = P'$ and either (a) or (b)
holds for $S'$ and $P'$.

If (a) holds then $S' \sreduces \rstr{s'}$ and $P' \treduces \str{s'}$ for some $s' \in \lang{r'}$.
Note that either $s' \in \lang{r}$ or $s' \not \in \lang{r}$ by property 2 of \ref{thm:regexprops}.
Suppose $s' \in \lang{r}$. Then $\rcoerce{r}{S} \sreduces \rstr{S'}$ by S-E-Coerce.
Choose $P = \check{\rx{r}}{P'}$ and note that $P \treduces \str{s'}$ by P-E-Coerce.
Now suppose $s' \not \in \lang{r}$. Then $S \err$ and $P \err$ by P-E-Check-Err
and S-E-Coerce-Err.

Finally, if (b) holds then $S \err$ and $P \err$ by propagation.

\end{description}
\end{proof}

% NRF commented out section 2 in order to determine the length of section 1.
% Replace the pattern ^% with an emptty string in the selection starting at
% the first empty line of section 2 and ending just before \end{document}.

Papers that needs to be cited in this section:
\begin{itemize}
  \item Ur/Web OSDI paper
  \item Jif?
  \item OWASP
  \item XDuce and related papers. 
  \item src?
  \item Ace or Wyvern paper?
  \item hotsos?
  \item Haskell extension paper
  \item Maybe some popular FOSS libraries/frameworks that do input sanitation?
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% END SECTION TWO -- THIS IS THE CONFLICT LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\F}[1]{\textsf{#1}~}
\newcommand{\FF}[1]{\textsf{#1}}
\newcommand{\Q}{\FF{Arg}}
\newcommand{\xlA}[1]{\lfloor #1 \rfloor_{\lambda}}
\newcommand{\xA}[1]{$\lfloor #1 \rfloor_{\text{Ace}}$}
\newcommand{\rtotau}[1]{\lfloor #1 \rfloor}

\newcommand{\tremp}{{\tt tremp}}
\newcommand{\trdot}{{\tt trdot}}
\newcommand{\trchar}[1]{{\tt trchar}[#1]}
\newcommand{\trseq}[2]{{\tt trseq}(#1; #2)}
\newcommand{\tror}[2]{{\tt tror}(#1; #2)}
\newcommand{\tlstr}[1]{{\tt tstr}[#1]}
\begin{table*}[t]
\centering
\begin{tabular}{ l l l }
$r$ & $\xlA{r}$ & \xA{r}\\
\hline
$\epsilon$ & $\tremp$ & \verb|""|\\
$.$ & $\trdot$ & \verb|"."|\\
$a$ & $\trchar{a}$ & \verb|"|$a$\verb|"|\\
$r_1 \cdot r_2$ & $\trseq{\xlA{r_1}}{\xlA{r_2}}$ & \xA{r_1}\verb| + |\xA{r_2}\\
$r_1 + r_2$ & $\tror{\xlA{r_1}}{\xlA{r_2}}$ & \verb|"(" + |\xA{r_1}\verb! + "|" + !\xA{r_2}\verb| + ")"|\\
\\
$\psi$ & $\xlA{\psi}$ & \xA{\psi}\\
\hline
${\tt string\_in}[r]$ & $\fvar{stringin}[\xlA{r}]$ & \verb|string_in[|\xA{r}\verb|]|\\
\\
  $S$ & $\xlA{S}$ & \xA{S} \\
  \hline
  ${\tt str\_in}[s]$ & $\FF{intro}[\FF{str}[s]]()$ & \verb|"|$s$\verb|"|\\
  ${\tt concat}(S_1; S_2)$  & $\xlA{S_1}\cdot\FF{elim}[\tvar{concat}](\xlA{S_2})$ & \xA{S_1}\verb| + |\xA{S_2} \\
  ${\tt subst}[r](S_1; S_2)$  & $\xlA{S_1}\cdot\FF{elim}[\tvar{subst}~\xlA{r}](\xlA{S_2})$ & \xA{S_1}\verb|.subst(|\xA{r}\verb|, |\xA{S_2}\verb|)|\\
  ${\tt coerce}[r](S)$  & $\xlA{S}\cdot\FF{elim}[\tvar{coerce}~\xlA{r}]()$ & \xA{S}\verb|.coerce(|\xA{r}\verb|)|
\end{tabular}
\caption{Embeddings of the ${\tt string\_in}$ fragment into $\lamAce$ and Ace.}
\end{table*}
%
\newcommand{\atjsynX}[3]{\Gamma \vdash_\fvalCtx #1 \Rightarrow #2 \leadsto #3}
\newcommand{\atjanaX}[3]{\Gamma \vdash_\fvalCtx #1 \Leftarrow #2 \leadsto #3}
\newcommand{\atjerrX}[1]{\Gamma \vdash_\fvalCtx #1~ \mathtt{error}}
\begin{figure*}[t]
\small
$\fbox{\inferrule{}{\atjsynX{e}{\tau}{i}}}$~~~~
$\fbox{\inferrule{}{\atjanaX{e}{\tau}{i}}}$~~~~
%$\fbox{\inferrule{}{\atjerrX{e}}}$
\begin{mathpar}
\inferrule[att-flip]{
	\atjsynX{e}{\tau}{i}
}{
	\atjanaX{e}{\tau}{i}
}

%\inferrule[att-var]{
%	x \Rightarrow \tau \in \Gamma
%}{
%	\atjsynX{x}{\tau}{x}
%}
%
%\inferrule[att-asc]{
%    \tau \Downarrow_\fvalCtx \tau'\\
%	\atjanaX{e}{\tau'}{i}
%}{
%	\atjsynX{e : \tau}{\tau'}{i}
%}
%
%\inferrule[att-let-syn]{
%	\atjsynX{e_1}{\tau_1}{i_1}\\
%	\Gamma, x \Rightarrow \tau_1 \vdash_\fvalCtx e_2 \Rightarrow \tau_2 \leadsto i_2\\
%	\trepof{\tau_1} \Downarrow_\fvalCtx \titype{\sigma_1}
%}{
%	\atjsynX{\F{let}x = e_1~\F{in}e_2}{\tau_2}{(\ilam{x}{\sigma_1}{i_2})~i_1}
%}
%
%\inferrule[att-lam-ana]{
%	\Gamma, x \Rightarrow \tau_1 \vdash_\fvalCtx e \Leftarrow \tau_2 \leadsto i\\
%	\trepof{\tau_1} \Downarrow_\fvalCtx \titype{\sigma_1}
%}{
%	\atjanaX{\lambda x.e}{\ttype{arrow}{(\tau_1, \tau_2)}}{\ilam{x}{\sigma_1}{i}}
%}
%
%\inferrule[att-lam]{
%	\tau_1 \Downarrow_\fvalCtx \tau_1'\\
%		\trepof{\tau_1'} \Downarrow_\fvalCtx \titype{\sigma}\\\\
%	\Gamma, x \Rightarrow \tau_1' \vdash_\fvalCtx e \Rightarrow \tau_2 \leadsto i
%%	i \hookrightarrow_\fvalCtx i'\\
%%		\sigma \hookrightarrow_\fvalCtx \sigma'
%%	\ddbar{\fvar{Arrow}}{\fvalCtx}{\trepof{\tau_1'}}{\sbar_1}\\
%	%\delfromtau{$\Xi_0$}{\fvalCtx}{\tau_1'}{\sabs}\\\\
%}{
%	\atjsynX{\elam{x}{\tau_1}{e}}{\ttype{arrow}{(\tau_1', \tau_2)}}{\ilam{x}{\sigma'}{i}}
%}
%
\inferrule[att-intro-ana]{
	\vdash_\fvalCtx \FF{iana}(\fvar{tycon})=\taudef\\
	\taudef~\taut{opidx}~\taut{tyidx}~((\Gamma; e_1)? :: \ldots :: (\Gamma; e_n)? :: []) \Downarrow_\fvalCtx \titerm{i}\\\\
%	\trepof{\ttype{tycon}{\tauidx'}} \Downarrow_\fvalCtx \titype{\sigma}\\
		\trepof{\ttype{tycon}{\taut{tyidx}}} \Downarrow_\fvalCtx \titype{\sigma}\\
		\Gamma \vdash_\fvalCtx i : \sigma
}{
	\atjanaX{\FF{intro}[\taut{opidx}](e_1; \ldots; e_n)}{\ttype{tycon}{\taut{tyidx}}}{i}
}

%\inferrule[att-i-asc-ty]{
%	\atjanaX{I}{\tau}{i}
%}{
%	\atjsynX{I : \tau}{\tau}{i}
%}
%
%\inferrule[att-i-asc-tycon]{
%	\vdash_\fvalCtx \FF{isyn}(\fvar{tycon})=\taudef\\\\
%	\taudef~\tauidx~((\Gamma; e_1)? {::}{\ldots}{::}(\Gamma; e_n)? {::} []) \Downarrow_\fvalCtx \tden{\titerm{i}}{\ttype{tycon}{\tauidx'}}\\
%			\trepof{\tau_1'} \Downarrow_\fvalCtx \titype{\sigma}\\
%	\Gamma \vdash_\fvalCtx i : \sigma
%}{
%	\atjsynX{\FF{intro}[\tauidx](e_1; \ldots; e_n)] :: \fvar{tycon}}{\ttype{tycon}{\tauidx'}}{i'}
%}
%
\inferrule[att-elim-syn]{
	\atjsynX{e}{\ttype{tycon}{\taut{tyidx}}}{i}\\
	\vdash_\fvalCtx \FF{esyn}(\fvar{tycon})=\taudef\\\\
	\taudef~\taut{opidx}~\taut{tyidx}~\titerm{i}~((\Gamma; e)? :: (\Gamma; e_1)? :: \ldots :: (\Gamma; e_n)? :: []) \Downarrow_\fvalCtx (\tau, \titerm{i'})\\\\
			\trepof{\tau} \Downarrow_\fvalCtx \titype{\sigma}\\
	\Gamma \vdash_\fvalCtx i' : \sigma
}{
	\atjsynX{e\cdot\FF{elim}[\taut{opidx}](e_1; \ldots; e_n)}{\tau}{i'}
}
\end{mathpar}
%\vspace{-10px}
\caption{The bidirectional active typechecking and translation judgements.}
\label{atj}
\end{figure*}
\newcommand{\tlevalX}[2]{#1 \Downarrow_\fvalCtx #2}
\begin{figure*}[t]
\small
$\fbox{\inferrule{}{\tau \Downarrow_\fvalCtx \tau'}}$~~~~
\begin{mathpar}
\inferrule[trstr-eval]{ }{\tlevalX{\tlstr{s}}{\tlstr{s}}}

\inferrule[tremp-eval]{ }{\tremp \Downarrow_\fvalCtx \tremp}

\cdots

\inferrule[tror-eval]{
	\tlevalX{\tau_1}{\tau_1'}\\
	\tlevalX{\tau_2}{\tau_2'}
}{
	\tlevalX{\tror{\tau_1}{\tau_2}}{\tror{\tau_1'}{\tau_2'}}
}

trmatch

trlsubst

trsublang

%\inferrule[repof]{
%	\tau \Downarrow_\fvalCtx \ttype{tycon}{\tauidx}\\
%	\vdash_\fvalCtx \FF{rep}(\fvar{tycon}) = \taurep\\
%	\taurep~\tauidx \Downarrow_\fvalCtx \titype{\sigma}
%}{
%	\FF{repof}(\tau) \Downarrow_\fvalCtx \titype{\sigma}
%}
%
\inferrule[syn]{
    \tau \Downarrow_\fvalCtx (\Gamma; e)?\\
	\atjsynX{e}{\tau}{\iota}\\\\
    [\tau/\tvar{t}_{ty}, \titerm{\iota}/\tvar{t}_{trans}]\tau_2 \Downarrow_\fvalCtx \tau_2'
}{
	\FF{syn}(\tau_1; \tvar{t}_{ty}, \tvar{t}_{trans}.\tau_2) \Downarrow_\fvalCtx \tau_2'
}

\inferrule[ana]{
	\tau_1 \Downarrow_\fvalCtx (\Gamma; e)?\\
	\tau_2 \Downarrow_\fvalCtx \tau_2'\\
	\atjanaX{e}{\tau_2'}{\iota}\\\\
	[\titerm{\iota}/\tvar{t}_{trans}]\tau_3 \Downarrow_\fvalCtx \tau_3'
}{
	\FF{ana}(\tau_1; \tau_2; \tvar{t}_{trans}.\tau_3) \Downarrow_\fvalCtx \tau_3'
}

TODO: add rx, str stuff
\end{mathpar}
\caption{\small Normalization semantics for the type-level language. Missing rules (including error propagation rules and normalization of quoted internal terms and types) are unsurprising and will be given later.}
\label{tleval}
\end{figure*}
\begin{figure}[t]
\small\begin{flalign}
& \F{tycon}\fvar{stringin}~\F{of}\FF{R}~\{\\
& \quad \F{iana}\{\tlam{opidx}{\FF{String}}{
	\tlam{tyidx}{\FF{R}}{
	\tlam{a}{\klist{\Q}}{\\
& \quad\quad \tvar{arity0}~\tvar{a}~(\tvar{check}~\tvar{opidx}~\tvar{tyidx}~\titerm{\iup{\tvar{opidx}}})
	}}
}\}\\
& \quad \F{esyn}\{\tlam{opidx}{\kunit+(\FF{R}+\FF{R})}{
	\tlam{a}{\klist{\Q}}{\\
& \quad\quad \tsumcase{\tvar{opidx}}{\_}{\\
& \quad\quad\quad \tvar{arity2}~\tvar{a}~\tlam{a1}{\Q}{\tlam{a2}{\Q}{\\
& \quad\quad\quad\quad \tvar{rsyn}~\tvar{a1}~\tlam{r1}{\FF{R}}{\tlam{i1}{\kITerm}{~}}\\
& \quad\quad\quad\quad \tvar{rsyn}~\tvar{a2}~\tlam{r2}{\FF{R}}{\tlam{i2}{\kITerm}{~}}\\
& \quad\quad\quad\quad\quad (\ttype{stringin}{\FF{rseq}(\tvar{r1}; \tvar{r2})}, \\
& \quad\quad\quad\quad\quad\titerm{{\tt iconcat}(\iup{\tvar{i1}}; \iup{\tvar{i2}})})
}}\\
& \quad\quad}{opidx'}{d}}
}\}
\};\\
& \F{def}\tvar{concat} = \tinl{\kunit, \FF{R}+\FF{R}}{\tunit};\\
& \F{def}\tvar{subst} = \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinl{\FF{R},\FF{R}}{\tvar{r}}}};\\
& \F{def}\tvar{coerce} = \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinr{\FF{R},\FF{R}}{\tvar{r}}}}
\end{flalign}
\caption{Definition of $\phi_S$, which enables the embedding of fragment $S$ into $\lamAce$.}
\end{figure}
%\begin{figure}
%\[
%\begin{array}{lcl}
%% & & Definition & Kind\\
%\tvar{concat} & := & \tinl{\kunit, \FF{R}+\FF{R}}{\tunit}\\%& \kunit + (R + R)\\
%\tvar{replace} & := & \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinl{\FF{R},\FF{R}}{\tvar{r}}}}\\% & \karrow{a}{b}\\
%\tvar{coerce} & := & \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinr{\FF{R},\FF{R}}{\tvar{r}}}}\\
%\end{array}
%\]
%\caption{Definitions}
%\end{figure}

\subsection{Background: Ace}
TODO: Make a TR out of the OOPSLA submission.
\subsection{Explicit Conversions}
\subsection{Adding Subtyping to Ace}
\subsection{Theory}

\section{Related Work}

\section{Discussion}

\bibliographystyle{abbrv}

% The bibliography should be embedded for final submission.

\bibliography{../research}
%\softraggedright
%P. Q. Smith, and X. Y. Jones. ...reference text...

\end{document}