### 2014-08-14 sanitation-psp14/sanitation-psp14-short.tex
Layout

\documentclass[9pt]{sig-alternate}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{afterpage}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}

% Hack. Something's wrong with PLAS paper when using the ACM Proc docclass
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}

\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
    \makeatletter
\usepackage{enumitem}
\usepackage{enumerate}

% \btIfInRange{number}{range list}{TRUE}{FALSE}
%
% Test if int number <number> is element of a (comma separated) list of ranges
% (such as: {1,3-5,7,10-12,14}) and processes <TRUE> or <FALSE> respectively
%
        \newcount\bt@rangea
        \newcount\bt@rangeb

        \newcommand\btIfInRange[2]{%
            \global\let\bt@inrange\@secondoftwo%
            \edef\bt@rangelist{#2}%
            \foreach \range in \bt@rangelist {%
                \afterassignment\bt@getrangeb%
                \bt@rangea=0\range\relax%
                \pgfmathtruncatemacro\result{ ( #1 >= \bt@rangea) && (#1 <= \bt@rangeb) }%
                \ifnum\result=1\relax%
                    \breakforeach%
                    \global\let\bt@inrange\@firstoftwo%
                \fi%
            }%
            \bt@inrange%
        }

        \newcommand\bt@getrangeb{%
            \@ifnextchar\relax%
            {\bt@rangeb=\bt@rangea}%
            {\@getrangeb}%
        }

        \def\@getrangeb-#1\relax{%
            \ifx\relax#1\relax%
                \bt@rangeb=100000%   \maxdimen is too large for pgfmath
            \else%
                \bt@rangeb=#1\relax%
            \fi%
        }

%
% \btLstHL{range list}
%
        \newcommand{\btLstHL}[1]{%
            \btIfInRange{\value{lstnumber}}{#1}%
            {\color{black!10}}%
            {\def\lst@linebgrd}%
        }%

%
% \btInputEmph[listing options]{range list}{file name}
%
        \newcommand{\btLstInputEmph}[3][\empty]{%
            \lstset{%
                linebackgroundcolor=\btLstHL{#2}%
                \lstinputlisting{#3}%
            }% \only
        }

% Patch line number key to call line background macro
        \lst@Key{numbers}{none}{%
            \def\lst@PlaceNumber{\lst@linebgrd}%
            \lstKV@SwitchCases{#1}{%
                none&\\%
                left&\def\lst@PlaceNumber{\llap{\normalfont
                \lst@numberstyle{\thelstnumber}\kern\lst@numbersep}\lst@linebgrd}\\%
                right&\def\lst@PlaceNumber{\rlap{\normalfont
                \kern\linewidth \kern\lst@numbersep
                \lst@numberstyle{\thelstnumber}}\lst@linebgrd}%
            }{%
                \PackageError{Listings}{Numbers #1 unknown}\@ehc%
            }%
        }

% New keys
        \lst@Key{linebackgroundcolor}{}{%
            \def\lst@linebgrdcolor{#1}%
        }
        \lst@Key{linebackgroundsep}{0pt}{%
            \def\lst@linebgrdsep{#1}%
        }
        \lst@Key{linebackgroundwidth}{\linewidth}{%
            \def\lst@linebgrdwidth{#1}%
        }
        \lst@Key{linebackgroundheight}{\ht\strutbox}{%
            \def\lst@linebgrdheight{#1}%
        }
        \lst@Key{linebackgrounddepth}{\dp\strutbox}{%
            \def\lst@linebgrddepth{#1}%
        }
        \lst@Key{linebackgroundcmd}{\color@block}{%
            \def\lst@linebgrdcmd{#1}%
        }

% Line Background macro
        \newcommand{\lst@linebgrd}{%
            \ifx\lst@linebgrdcolor\empty\else
                \rlap{%
                    \lst@basicstyle
                    \color{-.}% By default use the opposite (`-`) of the current color (`.`) as background
                    \lst@linebgrdcolor{%
                        \kern-\dimexpr\lst@linebgrdsep\relax%
                        \lst@linebgrdcmd{\lst@linebgrdwidth}{\lst@linebgrdheight}{\lst@linebgrddepth}%
                    }%
                }%
            \fi
        }

 % Heather-added packages for the fancy table
 \usepackage{longtable}
 \usepackage{booktabs}
 \usepackage{pdflscape}
 \usepackage{colortbl}%
 \newcommand{\myrowcolour}{\rowcolor[gray]{0.925}}
 \usepackage{wasysym}

    \makeatother

\lstset{
  language=Python,
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  commentstyle=\itshape\color{light-gray},
  basicstyle=\ttfamily\scriptsize,
  morekeywords={lambda, self, assert, as, cls},
  numbers=left,
  numberstyle=\scriptsize\color{light-gray}\textsf,
  xleftmargin=2em,
  stringstyle=\color{mauve}
}
\lstdefinestyle{Bash}{
    language={}, 
    numbers=left,
    numberstyle=\scriptsize\color{light-gray}\textsf,
    moredelim=**[is][\color{blue}\bf\ttfamily]{`}{`},
}
\lstdefinestyle{OpenCL}{
	language=C++,
	morekeywords={kernel, __kernel, global, __global, size_t, get_global_id, sin, printf, int2}
}

\usepackage{float}
\floatstyle{ruled}
\newfloat{codelisting}{tp}{lop}
\floatname{codelisting}{Listing}
\setlength{\floatsep}{10pt}
\setlength{\textfloatsep}{10pt}

\usepackage{url}

\usepackage{todonotes}

\usepackage{placeins}

\usepackage{textpos}

\renewcommand\topfraction{0.85}
\renewcommand\bottomfraction{0.85}
\renewcommand\textfraction{0.1}
\renewcommand\floatpagefraction{0.85}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bussproofs}    % Gentzen-style deduction trees *with aligned sequents!*
\usepackage{mathpartir}	% For type-settting type checking rule figures
\usepackage{syntax}		% For type-setting formal grammars.

\usepackage{hyperref}		% For links in citations

\usepackage{float}			% Make figures float if [H] option is passed.

\iffalse
\usepackage{listings}		% For typesetting code listings
\usepackage{callout}
\usepackage{titlesec}
\usepackage[T1]{fontenc}
\usepackage{upquote}
\lstset{upquote=true}
\fi

\usepackage{textcomp}		% For \textquotesingle as used in introduction
\usepackage{color}			% for box colors, like in TAPL.

\usepackage{amsmath}		% Begin Carthage default packages
\usepackage{makeidx}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphics}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{epsfig}
\usepackage{csquotes}
\usepackage{enumitem}

\newtheorem{thm}{Theorem}                                                       
\newtheorem{cor}[thm]{Corollary}                                                
\newtheorem{lem}[thm]{Lemma}                                                    
\newtheorem{prop}[thm]{Proposition}                                             
\newtheorem{ax}[thm]{Axiom}                                                     
\theoremstyle{definition}                                                       
\newtheorem{defn}[thm]{Definition}                                              
\newtheorem{exam}[thm]{Example}                                                 
\newtheorem{rem}[thm]{Remark}

%
% For type setting inference rules with labels.
%
\newcommand{\inferlbl}[3]
			{\inferrule{#3}{#2}{\textsf{\footnotesize{\sc #1}}}}
\newcommand{\inferline}[3]
			{\inferrule{#3}{#2} & {\textsf{\footnotesize{\sc #1}}} \\ \\}

\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\lang}[1]{\Lagr\{#1\}}
\newcommand{\stru}[2]{ {\tt string\_union}(#1,#2)}

\newcommand{\dconvert}[2]{ {\tt dconvert}(#1,#2) }
\newcommand{\filter}[2]{ {\tt filter}(#1,#2) }
\newcommand{\ifilter}[2]{ {\tt ifilter}(#1,#2) }

\newcommand{\reduces}{ \Rightarrow }
\newcommand{\gvd}{\Gamma \vdash }
\newcommand{\ovd}{\Omega \vdash }

\newcommand{\trep}{{\tt rep}}

\newcommand{\tstrf}[1]{`#1\textrm'} %??
\newcommand{\strf}[1]{``#1"}

\newcommand{\iso}{\cong}
%%
%% Source and Target language definitions.
%%
\newcommand{\lambdas}{\lambda_{RS}}
\newcommand{\lambdap}{\lambda_P}

% Source language terms.
\newcommand{\sisubst}[3]{{\sf rreplace}[#1](#2;#3)} \newcommand{\rreplace}[3]{{\sf rreplace}[#1](#2;#3)} % lots of legacy naming around these parts...

\newcommand{\rssreplace}[3]{\sisubst{#1}{#2}{#2}} % TODO-nrf fix this.

\newcommand{\coerce}[2]{ {\sf rcoerce}[#1](#2)}
\newcommand{\rcoerce}[2]{{\sf rcoerce}[#1](#2)}
\newcommand{\sistr}[1]{{\sf rstr}[#1]}   \newcommand{\rstr}[1]{{\sf rstr}[#1]} % Lots of legacy naming around these parts...

\newcommand{\val}{{\sf val}}

\newcommand{\rcheck}[4]{ {\sf rcheck}[#1](#2;#3;#4) }

\newcommand{\strin}[1]{\sistr{#1}}
\newcommand{\rsconcat}[2]{{\sf rconcat}(#1;#2)} \newcommand{\rconcat}[2]{{\sf rconcat}(#1;#2)} % lots of legact naming around these parts..

% Source language types.
\newcommand{\stringin}[1]{{\sf stringin}[#1]}

% target language terms.
\newcommand{\tsubst}[3]{{\sf replace}(#1;#2;#3)} \newcommand{\metareplace}[3]{{\sf replace}(#1;#2;#3)} % TODO-nrf rename the commands. Lots of legacy naming around these parts...

\newcommand{\tcheck}[4]{{\sf check}(#1; #2; #3; #4)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\preplace}[3]{{\sf preplace}(#1;#2;#3)}
\newcommand{\tconcat}[2]{{\sf concat}(#1;#2)} \newcommand{\concat}[2]{{\sf concat}(#1;#2)} % lots of legacy naming around these parts...

\newcommand{\regext}[1]{ {\sf rx}[#1] } % TODO-nrf remove
\newcommand{\rx}[1]{ {\sf rx}[#1] }

% Target language types
\newcommand{\str}{{\sf string}}
\newcommand{\regex}{{\sf regex}}

% Meta-theoretic functions
\newcommand{\lsubst}[3]{{\tt subst}(#1;#2;#3)} % This used to renderlreplace(...) so there're probably mistkes wherever this command was used now.
\newcommand{\lreplace}[3]{{\tt lreplace}(#1; #2; #3)}

% Judgements
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

% Relations
\newcommand{\treduces}{ \Downarrow }
\newcommand{\sreduces}{ \Downarrow }

%%
%% Constrain the size of full-page diagrams and rule lists
%%
%%\newcommand{\pagewidth}{5in}
%%\newcommand{\rulelistwidth}{3in}

% Names of type systems presented in paper
\newcommand{\lcs}{\lambda_{S}}

\setlength{\grammarindent}{3em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%dep_docclass\conferenceinfo{-}{-} 
%dep_docclass\copyrightyear{-} 
%dep_docclass\copyrightdata{[to be supplied]}

%\titlebanner{{\tt \textcolor{Red}{{\small Under Review -- distribute within CMU only.}}}}        % These are ignored unless
%\preprintfooter{Distribute within CMU only.}   % 'preprint' option specified.

\newcommand{\Ace}{\textsf{Ace}}

\title{Statically Typed String Sanitation Inside a Python}
\numberofauthors{3}
\author{
  \alignauthor
  Nathan Fulton
  \alignauthor
  Cyrus Omar
  \alignauthor
  Jonathan Aldrich
  \and
  \affaddr{Carnegie Mellon University}\\
  \affaddr{Pittsburgh, PA}\\
  \email{\{nathanfu, comar, aldrich\}@cs.cmu.edu}
}
%\authorinfo{~}{~}{~}
%\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
 %          {School of Computer Science\\
  %          Carnegie Mellon University
   %        \{nathanfu, comar, aldrich\}@cs.cmu.edu}

\maketitle
\begin{abstract}
%Evidence suggests that programmers are reluctant to adopt new languages to
Web applications must ultimately command systems like web browsers and database engines using strings. Strings constructed using user input that has not been properly sanitized can thus cause command injection vulnerabilities. %Structured frameworks for constructing commands centralizes control over string sanitization, but verifying that the framework itself sanitizes user input correctly can still be tricky.
In this paper, we introduce \emph{regular string types}, which classify strings known statically to be in a specified regular language. These types come equipped with common operations like concatenation, substitution and coercion, so they  can be used to implement, in essentially a conventional manner, the parts of a web application or application framework that construct command strings. Simple type annotations at key interfaces can be used to statically verify that sanitization has been performed correctly without introducing redundant run-time checks. We specify this type system as a minimal typed lambda calculus, $\lambda_{RS}$.

To be practical, adopting a specialized type system like this should not require the adoption of a new programming language. Instead, we favor extensible type systems: new type system fragments like this should be distributed as libraries atop a mechanism that guarantees that they can be safely composed.  We support this by 1) specifying a  translation from $\lambda_{RS}$ to a language containing only strings and regular expressions, then, taking Python as such a language, 2) implement the type system together with the translation as a library using \texttt{atlang}, an extensible static type system for Python (being developed by the authors).% represents the most sophisticated application of this library to date.

% Security-oriented programming languages do not enjoy wide adoption. Conversely, library and framework-based approaches to security are pervasive. Unfortunately, libraries and frameworks are only effective solutions if they are implemented and used properly. In this paper, we propose that extensible programming languages provide a promising mechanism for increasing the amount of trust that can be place in libraries and frameworks. To demonstrate this approach, we present a conservative extension to the simply-typed lambda calculus for checking the correctness of input sanitation algorithms. We also demonstrate how this language's security guarantees are preserved under translation into an underlying language containing a regular expression library. We believe this approach -- complementing existing techniques with light-weight, composable type-based analyses -- constitutes a promising compromise between usability, potential for industry adoption, and theoretically grounded safety guarantees.
\end{abstract}

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Command injection vulnerabilities are among the most common and severe security vulnerabilities in modern web applications \cite{OWASP}. They arise because web applications, at their boundaries, control external systems by issuing  commands represented using  strings. For example, web browsers are controlled using HTML and Javascript sent from a server as a string, and database engines execute SQL queries also sent as strings. When these commands must include data derived from user input, care must be taken to ensure that the user cannot construct input that subverts the intended command. For example, a  SQL query constructed using string concatenation exposes a SQL injection vulnerability: 
\begin{lstlisting}[numbers=none]
'SELECT * FROM users WHERE name="' + name + '"'
\end{lstlisting}
If a malicious user enters the name \lstinline{'"; DROP TABLE users --'}, the entire database could be erased.

To avoid this problem, the program must \emph{sanitize} user input. For example, in this case, the developer (or, more often, a framework) might define a function \verb|sanitize| that prepends double quotes (and existing backslashes)  with a backslash, which SQL treats safely. Guaranteeing that user input has already been sanitized before it is used to construct a command is challenging. Note that this function is not idempotent, so it should only be called once.

We observe that most such sanitization techniques can be understood in terms of \emph{regular languages} \cite{cinderella}. For example, \verb|name| should be a string in the language described by the regular expression \verb!([^"\]|(\")|(\\))*! -- a sequence of characters other than quotation marks and backslashes; these can only appear escaped. This concrete pattern syntax can be understood to desugar, in a standard way, to the syntax for regular expressions shown in Figure \ref{fig:regex}, where $r \cdot r$ is sequencing and $r + r$ is  disjunction. We will work with this ``core'' for simplicity.

In this paper, we present a static type system that tracks the regular language a string belongs to. For example, the output of \verb|sanitize| will be in the regular language described by the regular expression above. We write this as $\mathcal{L}\{r\}$. We take advantage of closure and decidability properties of regular languages to support a number of useful operations on values of such \emph{regular string types}, including \emph{replacement} of substrings matching a given regular expression. These make it possible to implement sanitation protocols like the one just described in an essentially conventional manner. The result is a system where the fact that a string has been {correctly} sanitized is manifest in its type. Missing calls to sanitization functions are detected statically, and, importantly, so are \emph{incorrectly implemented sanitization functions} (i.e. these functions need not be trusted). These guarantees require run-time checks only when going from less precise to more precise types (e.g. at the edges of the system, where user input has not yet been validated). %implementing and statically checking input sanitation techniques. Our solution suggests a more general approach to the integration of security concerns into programming language design. This approach is characterized by \emph{composable} type system extensions which \emph{complement} existing and well-understood solutions with compile-time checks.

We will begin in Sec. \ref{calculus} by specifying this type system minimally, as a conservative extension of the simply typed lambda calculus called $\lambdas$. This allows us to specify the guarantees that the type system provides precisely. We also formally specify a translation from this calculus to a typed calculus with only standard strings and regular expressions, intending it as a guide to language implementors interested in building this feature into their own languages. This also demonstrates that no additional space overhead is required.

Waiting for a language designer to build this feature in is unsatisfying. Moreover, we would also face a ``chicken-and-egg problem'': justifying its inclusion into a commonly used language requires empirically demonstrating that it is useful, but this is difficult to do if developers have no way to use it in practice. As such, we take the position that a better path forward for the community is to work within a programming language where such type system fragments can be introduced modularly and orthogonally, as libraries.

In Sec. \ref{atlang}, we show how to implement the type system fragment we have specified  using \texttt{atlang}, an extensible static type system implemented as a library inside Python. \texttt{atlang} leverages local type inference to control the semantics of literal forms, so reguar string types can be introduced using string literals without any run-time overhead. Coercions that are known to be  safe due to a sublanguage relationship are performed implicitly, also without run-time overhead. This results in a \emph{usably secure} system: working with regular strings differs little from working with standard strings.

We conclude after discussing related work in Sec. \ref{related}.

\section{Regular String Types, Minimally}\label{calculus}
\renewcommand{\grammarlabel}[2]{#1\hfill#2}

\newcommand{\sctx}{\Psi} % Context for external typing
\newcommand{\tctx}{\Theta} % Context for internal typing

\newcommand{\ereduces}{\Downarrow}

\begin{figure}[h]
\begin{grammar}
<$r$> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

\caption{Regular expressions over the alphabet $\Sigma$.}
\label{fig:regex}
\end{grammar}
\end{figure}

\begin{figure}[h]
\begin{grammar}

<$\sigma$> ::=  $\sigma \rightarrow \sigma$     \hfill  source types          \alt
$\stringin{r}$

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x . e$ \alt
      $e~(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\rcoerce{r}{e}$ \alt
      $\rcheck{r}{e}{x.e}{e}$
\caption{Syntax of $\lambda_{RS}$.}\label{fig:lamrs}
\end{grammar}
\end{figure}

In this section, we define a minimal typed lambda calculus with regular string types called $\lambda_{RS}$. This will serve as the source language for our translation to a calculus with only standard strings and regular expressions, defined in Sec. \ref{translation}.

The syntax of $\lambda_{RS}$ is given in Figure \ref{fig:lamrs}, its static semantics are specified in Figure \ref{fig:etyping} and its evaluation semantics are specified in Figure \ref{fig:eeval}.

The system has regular expression types $\stringin{r}$ whre $r$ is a regular expression. 
Expressions of this type evaluate to string literals in the language described by $r$.
Operations on expressions of type $\stringin{r}$ preserve this property.

The premier operation for manipulating strings in $\lambdas$ is string substitution, 
which is a familiar operation to any programmer who has used regular expressions.
The replacement operation replaces all instances of a pattern in one string with
another string; for instance, $\lsubst{a|b}{a}{c} = c$.
In order to compute the type resulting from substitution, we also need to compute
the result of replacing one language with another inside a given language.
Finally, just for convenience, we provide a coerce operation. The introduction of
coercion requires handling of runtime errors.

The underlying language $\lambdap$ has only one type for strings. We prove that whenever
a term is translated from $\lambdas$ to $\lambdap$, correctness is preserved.
The only exception is in the case of unsafe casts in $\lambdas$, which are unnecessary but
are included to demonstrate that the regex library of $\lambdap$ may be used to insert
dynamic checks whenever even when developers are not careful about using statically checked
operations.

A brief outline of this section follows:
\begin{itemize}
  \item In \S 2.1-2.3, we define $\lambdas$, $\lambdap$ and the translation from one to the other.
  \item In \S 2.4 we state some properties about regular expressions which are needed in our correctness proofs.
   \item In \S 2.5 we prove type safety for $\lambdap$ as well as both type safety and correctness for $\lambdas$.
  \item In \S 2.6 we prove that translation preserves the correctness result about $\lambdas$.
\end{itemize}

%
% Begin typing/translation relation
%

% Begin Grammar 
% Removes the < > from grammar productions.
\renewcommand{\grammarlabel}[2]{#1\hfill#2}

% \begin{figure}[t]
% \begin{grammar}
% <r> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

% \caption{Regular expressions over the alphabet $\Sigma$.}
% \label{fig:regex}
% \end{grammar}
% \end{figure}

% \begin{figure}[h]
% \begin{grammar}

% <$\psi$> ::=	$\psi \rightarrow \psi$			\hfill	source types					\alt
% $\stringin{r}$

% <S> ::= 
%       $\lambda x . e$ \hfill source terms \alt
%       $e e$ \alt
%       $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
%       $\rsconcat{S}{S}$ \alt
%       $\sisubst{r}{S}{S}$ \alt
%       $\coerce{r}{S}$
% \caption{Syntax for the string sanitation fragment of our source language, $\lambdas$.}
% \end{grammar}
% \end{figure}

\subsection{Definition of $\lambdas$}

The $\lambdas$ system extends the simply-typed lambda calculus with regular
expression types; an explanation of significant typing rule follows:
\begin{itemize}
  \item Rule S-T-Concat states that the type of concatenated strings is 
    obtained by concatenating the regular expressions for each string.
  \item Rule S-T-Replace defines the type of $\textsf{rreplace}$. The expression $\textsf{rreplace}$
    evaluates to the result of substituting every string matching $r$ in $s_1$
    with $s_2$. This operation corresponds to the \textsf{str_replace} function of PHP.
    The type of these expressions is defined in terms of an extra-linguistic \textsf{lreplace}
    function, which is defined later in this section.
    The expression $\lreplace{r}{r_1}{r_2}$ is obtained by
    starting with $r_1$, and replacing any subexpression matching $r$ with $r_2$.
  \item Rule S-T-SafeCoerce allows only safe coercions between string types by
    expoiting the decidability of language inclusion.
  \item The check rules allow casts between strings that cannot be guaranteed at compile time, but might pass a runtime check.
\end{itemize}

\newpage

\begin{figure}[t]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash \iota : \tau}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \tau$

\begin{mathpar}
\inferrule[P-T-Var]
{ x:\tau \in \tctx }
{ \tctx \vdash x:\tau }

\inferrule[P-T-Abs]
{\tctx, x : \tau_1 \vdash \iota_2 : \tau_2}
{\tctx \vdash \lambda x.\iota_2 : \tau_1 \rightarrow \tau_2}

\inferrule[P-T-App]
{\tctx \vdash \iota_1 : \tau_2 \rightarrow \tau \\ \tctx \vdash \iota_2 : \tau_2}
{\tctx \vdash \iota_1(\iota_2) : \iota}

\inferrule[P-T-String]
{ \ }
{\tctx \vdash \tstr{s} : \str}

\inferrule[P-T-Regex]
{ \ }
{\tctx \vdash \rx{r} : \regex}

\inferrule[P-T-Concat]
{\tctx \vdash \iota_1 : \str \\ \tctx \vdash \iota_2 : \str}
{\tctx \vdash \tconcat{\iota_1}{\iota_2} : \str}

\inferrule[P-T-Replace]
{\tctx \vdash \iota_1 : \regex \\ \tctx \vdash \iota_2 : \str \\ \tctx \vdash \iota_3 : \str }
{\tctx \vdash \preplace{\iota_1}{\iota_2}{\iota_3} : \str}

\inferrule[P-T-Check]
{\tctx \vdash \iota_r : \regex \\ \tctx \vdash \iota_1 : \str \\ \tctx, x:\str \vdash \iota_2 : \sigma \\ \tctx \vdash \iota_3 : \sigma}
{\tctx \vdash \tcheck{\iota_r}{\iota_1}{x . \iota_2}{\iota_3} : \sigma}
\end{mathpar}
\caption{Typing rules for $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:ityping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\iota \treduces \iota}}$

\begin{mathpar}
\inferrule[P-E-Abs]
{ \ }
{\lambda x.e \sreduces \lambda x.e}

\inferrule[P-E-App]
{ \iota_1 \sreduces \iota_1' \\  \iota_2 \sreduces \iota_2' }
{ \iota_1(\iota_2) \sreduces \iota_1'(\iota_2')}

  \inferrule[P-E-Str]
{ \ }
{\tstr{s} \treduces \tstr{s}}

\inferrule[P-E-Rx]
{ \ }
{\rx{r} \treduces \rx{r}}

\inferrule[P-E-Concat]
{\iota_1 \treduces \tstr{s_1} \\ \iota_2 \treduces \tstr{s_2}} 
{\tconcat{\iota_1}{\iota_2} \treduces \tstr{s_1 s_2}} % ???

\inferrule[P-E-Replace]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \tstr{s_2} \\ \iota_3 \treduces \tstr{s_3} \\ \lsubst{r}{s_2}{s_3} = s} 
{\preplace{\iota_1}{\iota_2}{\iota_3} \treduces \tstr{s}}

\inferrule[P-E-Check-OK]
{\iota \treduces \tstr{s} \\ s \in \lang{r}}
{\tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} \treduces [\tstr{s}/x]\iota_1}

\inferrule[P-E-Check-NotOK]
{\iota \treduces \tstr{s} \\ s \not \in \lang{r}}
{\tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} \treduces \iota_2}
\end{mathpar}
\caption{Big step semantics for of $\lambdap$}.
\label{fig:ieval}
\end{figure}

\begin{figure}\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash x:\sigma}

\inferrule[S-T-Abs]
{\sctx, x : \sigma_1 \vdash e : \sigma_2}
{\sctx \vdash \lambda x.e : \sigma_1 \rightarrow \sigma_2}

\inferrule[S-T-App]
{\sctx \vdash e_1 : \sigma_2 \rightarrow \sigma \\ \sctx \vdash e_2 : \sigma_2}
{\sctx \vdash e_1(e_2) : \sigma}

\inferrule[S-T-Stringin-I]
{s \in \lang{r}}
{\sctx \vdash \strin{s} : \stringin{r}}

\inferrule[S-T-Concat]
{\sctx \vdash e_1 : \stringin{r_1} \\ \sctx \vdash e_2 : \stringin{r_2}}
{\sctx \vdash \rsconcat{e_1}{e_2} : \stringin{r_1 \cdot r_2}}

\inferrule[S-T-Replace]
{\sctx \vdash e_1 : \stringin{r_1} \\ \sctx \vdash e_2 : \stringin{r_2} \\\\ \lreplace{r}{r_1}{r_2} = r'}
{\sctx \vdash \sisubst{r}{e_1}{e_2} : \stringin{r'}}

\inferrule[S-T-SafeCoerce]
{\sctx \vdash e : \stringin{r'} \\ \lang{r'} \subseteq \lang{r}}
{\sctx \vdash \coerce{r}{e} : \stringin{r}}

\inferrule[S-T-Check]
{\sctx \vdash e_0 : \stringin{r_0} \\ \sctx, x:\stringin{r} \vdash e_1 : \sigma \\ \sctx \vdash e_2 : \sigma}
{\sctx \vdash \rcheck{r}{e_0}{x.e_1}{e_2} : \sigma}

\end{mathpar}

\caption{Typing rules for $\lambdas$. 
The typing context $\sctx$ is standard.}
\label{fig:etyping}
\end{figure}

\begin{figure}
\small
$\fbox{\inferrule{}{e \sreduces e}}$
\begin{mathpar}
\inferrule[S-E-Abs]                                                             
{ \ }                                                                           
{\lambda x.e \sreduces \lambda x.e}

\inferrule[S-E-App]
{ e_1 \sreduces e_1' \\ e_2 \sreduces e_2' }
{ e_1(e_2) \sreduces e_1'(e_2')}

\inferrule[S-E-RStr]
{ \ }
{\strin{s} \sreduces \strin{s}}

\inferrule[S-E-Concat]
{e_1 \sreduces \strin{s_1} \\ e_2 \sreduces \strin{s_2}} 
{\rsconcat{e_1}{e_2} \sreduces \strin{s_1 s_2}} % ???

\inferrule[S-E-Replace]
{e_1 \sreduces \strin{e_1} \\ e_2 \sreduces \strin{s_2} \\ \lsubst{r}{s_1}{s_2} = s} 
{\sisubst{r}{e_1}{e_2} \sreduces \sistr{s}}

\inferrule[S-E-SafeCoerce]
{e \sreduces \strin{s}}
{\coerce{r}{e} \sreduces \strin{s}}

\inferrule[S-E-Check-Ok]
{e \sreduces \strin{s} \\ s \in \lang{r}}
{
  \rcheck{r}{e}{x.e_1}{e_2} \sreduces [\strin{s} / x]e_1
}

\inferrule[S-E-Check-NotOk]
{
  e \sreduces \strin{s} \\ s \not \in \lang{r}
}
{
  \rcheck{r}{e}{x.e_1}{e_2} \sreduces e_2
}

\end{mathpar}
\caption{Big step semantics for $\lambdas$}.
\label{fig:eeval}
\end{figure}

\newpage
\ 
\newpage
\ 
\newpage

\subsection{Definition of $\lambdap$}

\begin{figure}[h]
\begin{grammar}

<$\theta$> ::= $\theta \rightarrow \theta$ \hfill target types \alt
$\str$ \alt $\regex$

<P> ::= $x$  \hfill target terms \alt
   $\lambda x . \iota$ \alt
   $\iota \iota$ \alt
  $\tstr{s}$ \alt
  $\rx{r}$ \alt
  $\tconcat{\iota}{\iota}$ \alt
  $\preplace{\iota}{\iota}{\iota}$ \alt
  $\tcheck{\rx{r}}{\iota}{\iota}{\iota}$ 
  \end{grammar}
\caption{Syntax for the target language, $\lambdap$, containing strings and statically constructed regular expressions.}
\label{fig:lcsSyntax}
\end{figure}

The system $\lambdap$ is a straight-forward extension of a simply typed lambda calculus
with a string type and a regular expression tpye. We include two operations which are available in the regular expression
library of any modern programming language. The check operation ensures that an expression
recognizes a string, and the replace operation is string replacement.

\subsection{Definition of Translation}

\begin{figure}[h]\label{fig:trans}

$\fbox{\inferrule{}{\sctx \vdash \trden{e} = \iota}}$
\begin{mathpar}

\inferrule[Tr-Abs]
{ \sctx \vdash \trden{e} = \iota }
{\sctx \vdash \trden{ \lambda x.e } = \lambda x . \iota}

\inferrule[Tr-App]
{\sctx \vdash \trden{e_1} = \iota_1 \\ \sctx \vdash \trden{e_2} = \iota_2}
{\sctx \vdash \trden{e_1(e_2)} = \iota_1(\iota_2)}

\inferrule[Tr-string]
{ \ }
{ \sctx \vdash  \trden{\strin{s}} = \tstr{s}}

\inferrule[Tr-Concat]
{ \sctx \vdash \trden{e_1} = \iota_1 \\ \sctx \vdash \trden{e_2} = \iota_2}
{ \sctx \vdash \trden{\rsconcat{e_1}{e_2}} = \tconcat{\iota_1}{\iota_2} }

\inferrule[Tr-Subst]
{ \sctx \vdash  \trden{e_1} = \iota_1 \\ \sctx \vdash  \trden{e_2} = \iota_2 }
{ \sctx \vdash  \trden{ \sisubst{r}{e_1}{e_2} } = \tsubst{\rx{r}}{\iota_1}{\iota_2} }

\inferrule[Tr-SafeCoerce]
{ \sctx \vdash \trden{e} = \iota }
{ \sctx \vdash  \trden{ \coerce{r'}{e} } = \iota }

\inferrule[Tr-Check]
{ \sctx \vdash \trden{e} = \iota_1 \\ \sctx \vdash \trden{e_1} = \iota_1?? \\ \sctx \vdash \trden{e_2} = \iota_2} 
{ \sctx \vdash \trden{ \rcheck{r}{e}{x.e_1}{e_2} } = \tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} }

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed.}
\end{figure}

The translation from $\lambdas$ to $\lambdap$ is defined in figure 5.
The coercion cases are most interesting. If the safety of coercion in manifest in the
types of the expressions, then no runtime check is inserted.
If the safety of coercion is not manifest in the types, then a check is inserted.

In practice, the type of a replacement rarely matches a specification.
Therefore, it is convienant in an implementation to always insert the appropriate
coercion, and then only raise type errors when an automatically inserted coercion
actually requires the insertion of a runtime check. Alternatively, this policy may 
be codified in the type system itself using subtyping \cite{fulton12}.

\subsection{Properties of Regular Languages}

Our type safety proof for language S replies on a relationship between
string substitution and language substitution given in lemma \ref{thm:substcorrespondence}.
We also rely upon several other properties of regular languages.
Throughout this section, we fix an alphabet $\Sigma$ over which strings $s$ and
regular expressions $r$ are defined. throughout the paper, $\lang{r}$ refers to the
language recognized by the regular expression $r$. This distinction between the regular expression
and its language -- typically elided in the literature -- makes our definition and
proofs about systems S and P more readable.

\begin{lem}{Properties of Regular Languages and Expressions.} \label{thm:regexprops}
The following are properties of regular expressions which are necessary for our proofs:
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1r_2}$.
For all strings $s$ and regular expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
Regular languages are closed under difference, right quotient, reversal, and string homomorphism.
\end{lem}

If any of these properties are unfamiliar, the reader may refer to a standard text on the subject \cite{cinderella}.

\begin{defn}[$\tt{subst}$]
The relation $\lsubst{r}{s_1}{s_2} = s$ produces a string $s$ in which all substrings of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

\begin{defn}[$\tt{lreplace}$]
  The relation $\lreplace{r}{r_1}{r_2} = r'$ relates $r, r_1,$ and $r_2$ to
  a language $r'$ containing all strings of $r_1$ except that any substring $s_{pre} s s_{post} \in \lang{r_1}$ where $s \in \lang{r}$
  is replaced by the set of strings $s_{pre} s_2 s_{post}$ for all $s_2 \in \lang{r_2}$ (the prefix and postfix positions may be empty).
\end{defn}

\begin{lem}{Closure.} \label{thm:total}
  If $\lang{r}, \lang{r_1}$ and $\lang{r_2}$ are regular expressions, then $\lang{\lreplace{r}{r_1}{r_2}}$ is also a regular language.
\end{lem}
\begin{proof}
The theorem follows from closure under difference, right quotient, reversal and string homomorphism.
\end{proof}

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of subst and lreplace; note that language substitutions over-approximate string substitutions.
\end{proof}

%\begin{description}
%\item[case $r=\alpha$.]
%If $s_1 = \alpha$ then $\alpha \in \lang{r_1}$ by assumption.
%Therefore, $\lsubst{r}{s_1}{s_2} = \lsubst{\alpha}{\alpha}{s_2} = s_2$ and
%$\lreplace{r}{r_1}{r_2} = \lreplace{\alpha}{r_1}{r_2}$. 
%Since $s_1 = \alpha$ and $\alpha \in \lang{r_1}$, $r_1 \iso \alpha | r_1'$ for some $r_1'$.
%Therefore, $\lreplace{\alpha}{r_1}{r_2} \iso \lreplace{\alpha}{\alpha|r_1'}{r_2}$ by Lemma X(1).
%Finally, $s_2 \in \lang{r_2}$ which implies $s_1 \in \lang{r_2|r'}$.
%If $s_1 \not = \alpha$ the $\lsubst{r}{s_1}{s_2} = s_1$ and $\lreplace{\alpha}{r_1}{r_2} = r_1$.
%
%\item[case $r = a|b$].
%Note that $\subst{a|b}{s_1}{s_2} = \lsubst{a}{\lsubst{b}{s_1}{s_2}}{s_2}$
%and $\lsubst{b}{s_1}{s_2} \in \lreplace{b}{r_1}{r_2}$ by induction.
%Therefore, $\lsubst{a|b}{s_1}{s_2} \in \lreplace{a}{\lreplace{b}{r_1}{r_2}}{r_2}$ by induction and the definition of lreplace.
%Finally, applying definitions once more, $\lsubst{a|b}{s_1}{s_2} \in \lreplace{a|b}{r_1}{r_2}$.
%
%\item[case $r = ab$].
%By a similar argument to the disjunctive case.
%
%\item[case $r = a*$].
%By considering the once unwinding of $a*$, noting that $s_1$ and $s_2$ are finite.
%\end{description}

% NRF The coercion lemma is no longer necessary, because it is a property of
% "strip"-style sanitation rather than the more general case of replacement 
% treated in this paper. However, we should mention this difference between my
% undergraduate thesis and this paper in the related work section if we are
% going tin clude a discussio of unpublished or quasi-published (e.g. src)
% results.
%\begin{lem}{Coercion Lemma.} \label{thm:coercionalemma}
%Suppose that $R$ and $L$ are regular expressions, and that $s \in R$ is a finite string.  
%Let $s' = \coerce(R,L,s)$ with all maximal substrings recognized by $L$ replaced with $\epsilon$.  Then $s'$ is recognized by $(R \backslash L) + \emptyset$ and the construction of $R \backslash L$ is decidable.
%\end{lem}
%\begin{proof}
%Let $F,G$ be FAs corresponding to $R$ and $L$, and let $G'$ be $G$ with its final states inverted (so that $G'$ is the complement of $L$).  Define an FA $H$ as a DFA corresponding to the NFA found by combining $F$ and $G'$ such that $H$ accepts only if $R$ and $L'$ accept or if $s$ is empty (this construction may result in an exponential blowup in state size.)  Clearly, $H$ corresponds to $R \backslash L + \emptyset$.  Thus, the construction of $R \backslash L + \emptyset$ is decidable.
%
%If $R \subset L$, $s' = \emptyset$.  If $L \subset R$, either $s' = \emptyset$, or $s' \in R$ and $s' \not \in L$. If $R$ and $L$ are not subsets of one another, then it may be the case that $L$ recognizes part of $R$.  Consider $L$ as the union of two languages, one which is a subset of $R$ and one which is disjoint.  The subset language is considered above and the disjoint language is inconsequential.
%\end{proof}

% Why is this necessary?
%\begin{lem}
%  For all $r$, if $s_1 \in r_1$ and $s_2 \in r_2$ then $\subst{r}{s_1}{s_2} \subseteq \lsubst{r}{r_1}{r_2}$.
%\end{lem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN PROOFS ABOUT S AND P
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Safety of the Source and Target Languages}

In this section, we establish type safety for the source ($\lambdas$) and
target ($\lambdap$) languages. In addition to type safety, we also prove
a stronger correctness property for $\lambdas$.

Our first two theorems establish that our rules preserve the well-formedness of
regular expressions. The standard lemmas required for safety of the simply
typed lambda claculus are also required; proofs for these do not differ substntially
from \cite{pfpl}.

\begin{lem}
  If $\sctx \vdash e : \stringin{r}$ then $r$ is a well-formed regular expression.
\end{lem}
\begin{proof}
  The only non-trivial case is S-T-Replace, which follows from lemma \ref{thm:total}. 
\end{proof}
\begin{lem}
If $\tctx \vdash \iota : \regex$ then $\iota \treduces \rx{r}$ such that $r$ is a well-formed regular expression. 
\end{lem}

We now prove safety for the string fragment of the source and target languages.
The easiest property is type safety of $\lambdap$, which follows almost directly from
type safety for the simply typed lambda calculus.

\begin{thm} 
  Let $\iota$ be a term in the target language. If $\tctx \vdash \iota : \tau$ 
  then $\iota \sreduces \iota'$ and $\tctx \vdash \iota' : \tau$.
\end{thm}

Safety for the string fragment of $\lambdas$ is more involved because it involves validating that
the type system's definition is justified by our theorems about regular languages.
We state the theorem for the string fragment only to demontrate this point;
full type safety for $\lambdas$ follows easily from safety for the string fagment.
Note that type safety does \emph{not} guarantee that strings of a type are in the
correct language. We isolate this property so that we may reason about its preservation
under translation to $\lambdap$, which does not have regular expression types.

\begin{thm}{Type Safety for the String Fragment of $\lambdas$.} \label{thm:scorrect}
  Let $S$ be a term in the source language. If $\sctx \vdash e : \stringin{r}$ 
  then $e \sreduces \rstr{s}$ and $\sctx \vdash \rstr{s} : \stringin{r}$.
\end{thm}
\begin{proof}
By induction on the typing relation.
The S-T-Concat case requires Lemma \ref{thm:regexprops} 
and the S-T-Replace case appeals to Lemma \ref{thm:substcorrespondence}.
\end{proof}

In addition to safety, $\lambdas$ requires a correctness result ensuring that well-typed terms of regular string type are in the language associated with their type.

\begin{thm}{Correctness of Input Sanitation for $\lambdas$.}\label{thm:scorrect}
  If  $\sctx \vdash e : \stringin{r}$ and $e \sreduces \rstr{s}$ then $s \in \lang{r}$.
\end{thm}
\begin{proof}
  Follows directly from type safety, canonical forms for $\lambdas$.
\end{proof}

\subsection{Translation Correctness}

The main theorem of this paper is Theorem \ref{thm:main},
which establishes that Theorem \ref{thm:scorrect} is preserved under translation
into the target language $\lambdap$.

Establishing this result requires an additional theorem establishing a relationship between canonical forms for the string fragments of $\lambdas$ and $\lambdap$.

\begin{thm}{Translation Correctness.} \label{thm:tcorrect}
If $\sctx \vdash e : \stringin{r}$ then there exists a $\iota$ such that $\trden{e} = \iota$, 
$\iota \treduces \tstr{s}$, and $e \sreduces \strin{s}$.
\end{thm}

\begin{proof}
The proof proceeds by induction on the typing relation for $e$ and an appropriate
choice of $\iota$; in each case, the choice is obvious and the rest of the proof
proceeds by our type safety theorems as well as an appeal to the induction hypothesis.
\end{proof}

Finally, our main theorem establishes that input sanitation correctness of $\lambdas$ is preserved under the translation into $\lambdap$.

\begin{thm}{Correctness of Input Sanitation for Translated Terms.}\label{thm:main}
  If $\trden{e} = \iota$ and $\sctx \vdash e : \stringin{r}$ then $\iota \sreduces \tstr{s}$
  for $s \in \lang{r}$.
\end{thm}
\begin{proof}
  By theorem \ref{thm:tcorrect}, $\iota \sreduces \tstr{s}$ implies that $e \sreduces \strin{s}$.
  By theorem \ref{thm:scorrect}, the above property together with the assumption that $e$ is well-typed implies that $s \in \lang{r}$.
\end{proof}
% PROOF OF TRANSLATION CORRECTNESS.
% This proof is slightly out of say. It's SUFFICIENT, but is actually more
% than sufficient because we prove \ref{scorrect} along with translation
% correctness....
%\begin{description}
%\item[S-T-String-I:] 
%Let $S = \rstr{s}$ and suppose $\sctx \vdash \rstr{s} : \stringin{r}$.
%Choose $T = \str{s}$ and note that $\trden{S} = P$ by Tr-String.
%By P-E-String, $P \treduces \str{s}$ and by S-E-String $S \treduces \rstr{s}$.
%Finally, by inversion of S-T-Stringin-I, $s \in \lang{r}$.
%
%\item[S-T-Concat:]
%Let $S = \rconcat{S_1}{S_2}$ and suppose $\sctx \vdash S : \stringin{r_1 r_2}$.
%By inversion, $\sctx \vdash S_1 : \stringin{r_1}$.
%It follows by induction that there exists a $P_1$ such that $\trden{S_1} = P_1$.
%By a similar argument for $S_2$ and $r_2$, there exists a $P_2$ such that $\trden{S_2} = P_2$.
%Choose $P = \concat{P_1}{P_2}$.
%
%We first prove property (a).
%Note that $S_1$ and $P_1$ do not result in errors and that $S_1$ is well typed.
%Therefore, $S_1 \sreduces \rstr{s_1}$ and $P_1 \treduces \str{s_1}$ for some $s_1 \in \lang{r_1}$ by theorems 8 and 9 respectively. % TODO-nrf make sure this numbering is correct
%Similarly, $S_2 \sreduces \rstr{s_2}$ and $P_2 \treduces \str{s_2}$ for some $s_2 \in \lang{r_2}$.
%Therefore, $S \sreduces \rstr{s_1 s_2}$ by S-E-Concat and $\concat{P_1}{P_2} \treduces \str{s_1 s_2}$ by P-E-Concat.
%Finally, $s_1s_2 \in \lang{r_1}{r_2}$ by \ref{thm:regexprops}.
%
%Consider property (b).
%If $S_1 \err$ then $P_1 \err$ by induction, and it follows that $S \err$ and $P \err$ by respective error propagation rules.
%Similarly, if $S_2 \err$ then $P_2 \err$ and it follow that $S \err$ and $P \err$ by induction and propagation.
%
%\item[S-T-Replace:]
%Let $S = \rreplace{r}{S_1}{S_2}$ and suppose $\sctx \vdash S : \stringin{r'}$ for some $s$.
%By inversion of S-T-Replace, , $\sctx S_1 : \stringin{r_1}$ and $\sctx : \stringin{r_2}$ 
%such that $\lsubst{r}{r_1}{r_2} = r'$.
%By induction, there exists some $P_1, P_2$ such that $\trden{S_1} = P_1$, $\trden{S_2} = P_2$
%and either (a) or (b) holds.
%
%If (a) holds then $S_1 \sreduces \rstr{s_1}$ and $P_1 \treduces \str{s_1}$ for some $s_1 \in \lang{r_1}$,
%and similarly for $S_2, P_2$ and some $s_2 \in \lang{r_2}$.
%Therefore, by S-E-Replace, $S \sreduces \rstr{s}$ for some $s = \lreplace{r}{s_1}{s_2}$.
%Choose $P = \preplace{r}{s_1}{s_2}$.
%By a similar argument and P-E-Replace, $P \treduces \str{s}$ for some $s = \lreplace{r}{s_1}{s_2}$.
%What remains to be shown is $\lreplace{r}{s_1}{s_2} \in \lang{\lsubst{r}{r_1}{r_2}}$,
%which follows from Leamm D since $s_1 \in r_1$ and $s_2 \in r_2$.
%
%If (b) holds for $S_1$ and $P_1$, then $S \err$ and $P \err$ by propagation rules.
%Similarly, if (b) hols for $S_2$ and $P_2$ then $S \err$ and $P \err$ by propagation rules.
%
%\item[S-T-Coerce:]
%Let $S = \rcoerce{r}{S'}$ and suppose $\sctx \vdash \rcoerce{r}{S} : \stringin{r}$.
%By inverstion $\Psi \vdash S' : \stringin{r'}$ for an arbitrary $r'$.
%By induction there exists a $P'$ such that $\trden{S'} = P'$ and either (a) or (b)
%holds for $S'$ and $P'$.
%
%If (a) holds then $S' \sreduces \rstr{s'}$ and $P' \treduces \str{s'}$ for some $s' \in \lang{r'}$.
%Note that either $s' \in \lang{r}$ or $s' \not \in \lang{r}$ by property 2 of \ref{thm:regexprops}.
%Suppose $s' \in \lang{r}$. Then $\rcoerce{r}{S} \sreduces \rstr{S'}$ by S-E-Coerce.
%Choose $P = \check{\rx{r}}{P'}$ and note that $P \treduces \str{s'}$ by P-E-Coerce.
%Now suppose $s' \not \in \lang{r}$. Then $S \err$ and $P \err$ by P-E-Check-Err
%and S-E-Coerce-Err.
%
%Finally, if (b) holds then $S \err$ and $P \err$ by propagation.
%
%\end{description}
\section{Implementation in {Atlang}}\label{atlang}
%\begin{figure}
%\begin{lstlisting}
%output of successful compilation
%\end{lstlisting}
%\caption{Output of successful compilation.}
%\end{figure}
%
%\begin{figure}
%\begin{lstlisting}
%output of failed compilation
%\end{lstlisting}
%\caption{Output of failed compilation.}
%\end{figure}
%

\subsection{An Example}\label{whatev}

\begin{figure}[h]\begin{lstlisting}
@fn
def sanitize(s : string_in[r'.*']):
  return (s.replace(r'"', '&quot;') 
           .replace(r'<', '&lt;')
           .replace(r'>', '&gt;'))

@fn
def results_query(s : string_in[r'[^"]*']):
  return 'SELECT * FROM users WHERE name="' + s + '"'

@fn
def results_div(s : string_in[r'[^<>]*']):
  return '<div>Results for ' + s + '</div>'

def main(db):
  input = sanitize(user_input())
  results = db.execute(results_query(input))
  return results_div(input) + format(results)
\end{lstlisting}
\vspace{-10px}
\caption{Regular string types in atlang, a library that enables static type checking for Python. }   \label{examplefig}

\end{figure}

Figure \ref{examplefig} demonstrates the use of string types in atlang.
The sanitize function takes an arbitrary string and returns a string without
double quotes or left and right brackets. In this example, we use HTML
escape sequences.

The main function receives user input and passes this input to a 
sanitize function, which replaces all double quotes and brackets with HTML
escape sequences.

The result of applying sanitize to input is appended to two functions which
construct a safe query and safe output.
The arguments to the result and output construction functions constitute \emph{specifications}.
In the case of \textsf{results_query}, this specification ensures that user input is always interpreted
as a string literal by the SQL server.
In the case of \textsf{results_div}, this specification ensures that user input does 
not contain any HTML tags, which is a conservative but effective policy for preventing
XSS attacks.

Note that ${\sf input}$ does not actually meet these specifications without additional machinery.
The type of ${\sf input}$ is quite large
and does not actually equal the specified domains of the query or output construction
methods. This mismatch is common -- in fact, nearly universal. Therefore, 
our implementation includes a simple subtyping relation between regular expression
types. 
This subtyping relation is justified theoretically by the fact that language inclusion is
decidable; see \cite{fulton12} for a formal definition of the subtyping relation.
Additionally, our extension remains composable because subtyping is defined on a type-by-type basis;
see \cite{fulton13} for a discussion of subtyping in Atlang (referred to there as Ace).

\subsection{Implementation of the Regular Expression Type}

\label{fig:impl}
\begin{figure}
\begin{lstlisting}
class string_in(atlang.Type):
  def __init__(self, rx):
    rx = rx_normalize(rx)
    atlang.Type.__init__(idx=rx)

  def ana_Str(self, ctx, node):
    if not in_lang(node.s, self.idx):
      raise atlang.TypeError("...", node)

  def trans_Str(self, ctx, node):
    return astx.copy(node)

  def syn_BinOp_Add(self, ctx, node):
    left_t = ctx.syn(node.left)
    right_t = ctx.syn(node.right)
    if isinstance(left_t, string_in):
      left_rx = left_t.idx
      if isinstance(right_t, string_in):
        right_rx = right_t.idx
        return string_in[lconcat(left_rx, right_rx)]
    raise atlang.TypeError("...", node)

  def trans_BinOp_Add(self, ctx, node):
    return astx.copy(node)

  def syn_Method_replace(self, ctx, node):
    [rx, exp] = node.args
    if not isinstance(rx, ast.Str):
      raise atlang.TypeError("...", node)
    rx = rx.s
    exp_t = ctx.syn(exp)
    if not isinstance(exp_t, string_in):
      raise atlang.TypeError("...", node)
    exp_rx = exp_t.idx
    return string_in[lreplace(self.idx, rx, exp_rx)]

  def trans_Method_replace(self, ctx, node):
    return astx.quote(
      """__import__(re); re.sub(%0, %1, %2)""",
      astx.Str(s=node.args[0]),
      astx.copy(node.func.value),
      astx.copy(node.args[1]))

  def syn_Method_check(self, ctx, node):
    [rx] = node.args
    if not isinstance(rx, ast.Str):
      raise atlang.TypeError("...", node)
    return string_in[rx.s]

  def trans_Method_check(self, ctx, node):
    return astx.quote(
      """__import__(string_in_helper);
      string_in_helper.coerce(%0, %1)""",
      astx.Str(s=other_t.idx),
      astx.copy(node))

  def check_Coerce(self, ctx, node, other_t):
    # coercions can only be defined between 
    # types with the same type constructor, 
    if rx_sublang(other_t.idx, self.idx):
      return other_t
    else: raise atlang.TypeError("...", node)
\end{lstlisting}
\vspace{-10px}
\caption{Implementation of the \texttt{string\_in} type constructor in atlang.}
\end{figure}

We implemented a variation on the type system presented in this paper. The only
significant difference is that we only support replacements where $s_2$ is the empty
string. Therefore, our implementation respects the system presented in section 2
only modulo the definition of \textsf{lreplace}.

Atlang translates programs using type definitions, which may extend both the
static and dynamic semantics of the language. New types are defined as Python
classes; figure \ref{fig:impl} contains the source code of our implementation.

The \textsf{string_in} type has an indexing regular expression \texttt{idx}.
Our translation is defined by the \textsf{trans_} methods while the \textsf{syn_}
methods define our type checker. Atlang defers type checking and translation
to these methods whenever an expression of type \textsf{string_in} is encountered.

\subsection{Related Work and Alternative Approaches}

The input sanitation problem is well-understood. There exist a large number of techniques and technologies, proposed by both practitioners and researchers, for preventing injection-style attacks. In this section, we explain how our approach to the input sanitation problem differs from each of these approaches. More important than these differences, however, is our more general assertion that language extensibility is a promising approach toward consideration of security goals in programming language design.

Unlike \emph{frameworks and libraries} provided by languages such as Haskell and Ruby, our type system provides a \emph{static} guarantee that input is always properly sanitized before use. Doing so requires reasoning about the operations on regular languages corresponding to standard operations on strings; we are unaware of any production system which contains this form of reasoning. Therefore, even where frameworks and libraries provide a viable interface or wrapper around input sanitation, our approach is complementary because it ensures the correctness of the framework or library itself. Furthermore, our approach is more general than database abstraction layers because our mechanism is applicable to all forms of command injection (e.g. shell injection or remote file inclusion).

A number of research languages provide static guarantees that a program is free of input sanitation vulnerabilities \cite{UrFlowOSDI10}. 
Unlike this work, our solution to the input sanitation problem has a very low barrier to adoption; for instance, our implementation conservatively extends Python -- a popular language among web developers.
We also believe our general approach is better-positioned for security, where continuously evolving threats might require frequent addition of new analyses; in these cases, the composability and generality of our approach is a substantial advantage.

%\begin{itemize}
%  \item Our system is a light-weight solution to a single class of sanitation vulnerabilities (e.g. we do not address Cross-Site Scripting).
%  \item Our system is defined as a library in terms of an extensible type system, as opposed to a stand-alone language. Instead of introducing new technologies and methodologies for addressing security problems, we provide a light-weight static analysis which complements approaches developers already understand well.
%  \item Our implementation of the translation is implemented in Python and shares its grammar. Since Python is a popular programming language among web developers, the barrier between our research and adopted technologies is lower than for greenfield security-oriented languages.
%\end{itemize}

The Wyvern programming language provides a general framework for composing language extensions \cite{tsl}\cite{wyvern}.
Our work identifies one particular extension, and is therefore complementary to Wyvern and related
work on extensible programming languages.
We are also unaware of any extensible programming languages which emphasize applications to security concerns.

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP,HosoyaPierce2002} checks XML documents against schema using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Our system is defined within an extensible type system.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML schema.
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
\end{itemize}

In conclusion, our contribution is a type system, implemented within an extensible type system, for checking the correctness of input sanitation algorithms.

\section{Conclusion}

Composable analyses which complement existing approaches constitute a promising approach toward the integration of security concerns into programming languages.
In this paper, we presented a system with both of these properties and defined a security-preserving transformation.
Unlike other approaches, our solution complements existing, familiar solutions while providing a strong guarantee that traditional library and framework-based approaches are implemented and utilized correctly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% END SECTION TWO -- THIS IS THE CONFLICT LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newcommand{\F}[1]{\textsf{#1}~}
%\newcommand{\FF}[1]{\textsf{#1}}
%\newcommand{\Q}{\FF{Arg}}
%\newcommand{\xlA}[1]{\lfloor #1 \rfloor_{\lambda}}
%\newcommand{\xA}[1]{$\lfloor #1 \rfloor_{\text{Ace}}$}
%\newcommand{\rtotau}[1]{\lfloor #1 \rfloor}
%
%\newcommand{\tremp}{{\tt tremp}}
%\newcommand{\trdot}{{\tt trdot}}
%\newcommand{\trchar}[1]{{\tt trchar}[#1]}
%\newcommand{\trseq}[2]{{\tt trseq}(#1; #2)}
%\newcommand{\tror}[2]{{\tt tror}(#1; #2)}
%\newcommand{\tlstr}[1]{{\tt tstr}[#1]}
%\begin{table*}[t]
%\centering
%\begin{tabular}{ l l l }
%$r$ & $\xlA{r}$ & \xA{r}\\
%\hline
%$\epsilon$ & $\tremp$ & \verb|""|\\
%$.$ & $\trdot$ & \verb|"."|\\
%$a$ & $\trchar{a}$ & \verb|"|$a$\verb|"|\\
%$r_1 \cdot r_2$ & $\trseq{\xlA{r_1}}{\xlA{r_2}}$ & \xA{r_1}\verb| + |\xA{r_2}\\
%$r_1 + r_2$ & $\tror{\xlA{r_1}}{\xlA{r_2}}$ & \verb|"(" + |\xA{r_1}\verb! + "|" + !\xA{r_2}\verb| + ")"|\\
%\\
%$\sigma$ & $\xlA{\sigma}$ & \xA{\sigma}\\
%\hline
%${\tt string\_in}[r]$ & $\fvar{stringin}[\xlA{r}]$ & \verb|string_in[|\xA{r}\verb|]|\\
%\\
%  $S$ & $\xlA{S}$ & \xA{S} \\
%  \hline
%  ${\tt str\_in}[s]$ & $\FF{intro}[\FF{str}[s]]()$ & \verb|"|$s$\verb|"|\\
%  ${\tt concat}(S_1; S_2)$  & $\xlA{S_1}\cdot\FF{elim}[\tvar{concat}](\xlA{S_2})$ & \xA{S_1}\verb| + |\xA{S_2} \\
%  ${\tt subst}[r](S_1; S_2)$  & $\xlA{S_1}\cdot\FF{elim}[\tvar{subst}~\xlA{r}](\xlA{S_2})$ & \xA{S_1}\verb|.subst(|\xA{r}\verb|, |\xA{S_2}\verb|)|\\
%  ${\tt coerce}[r](S)$  & $\xlA{S}\cdot\FF{elim}[\tvar{coerce}~\xlA{r}]()$ & \xA{S}\verb|.coerce(|\xA{r}\verb|)|
%\end{tabular}
%\caption{Embeddings of the ${\tt string\_in}$ fragment into $\lamAce$ and Ace.}
%\end{table*}
%%
%\newcommand{\atjsynX}[3]{\Gamma \vdash_\fvalCtx #1 \Rightarrow #2 \leadsto #3}
%\newcommand{\atjanaX}[3]{\Gamma \vdash_\fvalCtx #1 \Leftarrow #2 \leadsto #3}
%\newcommand{\atjerrX}[1]{\Gamma \vdash_\fvalCtx #1~ \mathtt{error}}
%\begin{figure*}[t]
%\small
%$\fbox{\inferrule{}{\atjsynX{e}{\tau}{i}}}$~~~~
%$\fbox{\inferrule{}{\atjanaX{e}{\tau}{i}}}$~~~~
%%$\fbox{\inferrule{}{\atjerrX{e}}}$
%\begin{mathpar}
%\inferrule[att-flip]{
%	\atjsynX{e}{\tau}{i}
%}{
%	\atjanaX{e}{\tau}{i}
%}
%
%%\inferrule[att-var]{
%%	x \Rightarrow \tau \in \Gamma
%%}{
%%	\atjsynX{x}{\tau}{x}
%%}
%%
%%\inferrule[att-asc]{
%%    \tau \Downarrow_\fvalCtx \tau'\\
%%	\atjanaX{e}{\tau'}{i}
%%}{
%%	\atjsynX{e : \tau}{\tau'}{i}
%%}
%%
%%\inferrule[att-let-syn]{
%%	\atjsynX{e_1}{\tau_1}{i_1}\\
%%	\Gamma, x \Rightarrow \tau_1 \vdash_\fvalCtx e_2 \Rightarrow \tau_2 \leadsto i_2\\
%%	\trepof{\tau_1} \Downarrow_\fvalCtx \titype{\sigma_1}
%%}{
%%	\atjsynX{\F{let}x = e_1~\F{in}e_2}{\tau_2}{(\ilam{x}{\sigma_1}{i_2})~i_1}
%%}
%%
%%\inferrule[att-lam-ana]{
%%	\Gamma, x \Rightarrow \tau_1 \vdash_\fvalCtx e \Leftarrow \tau_2 \leadsto i\\
%%	\trepof{\tau_1} \Downarrow_\fvalCtx \titype{\sigma_1}
%%}{
%%	\atjanaX{\lambda x.e}{\ttype{arrow}{(\tau_1, \tau_2)}}{\ilam{x}{\sigma_1}{i}}
%%}
%%
%%\inferrule[att-lam]{
%%	\tau_1 \Downarrow_\fvalCtx \tau_1'\\
%%		\trepof{\tau_1'} \Downarrow_\fvalCtx \titype{\sigma}\\\\
%%	\Gamma, x \Rightarrow \tau_1' \vdash_\fvalCtx e \Rightarrow \tau_2 \leadsto i
%%%	i \hookrightarrow_\fvalCtx i'\\
%%%		\sigma \hookrightarrow_\fvalCtx \sigma'
%%%	\ddbar{\fvar{Arrow}}{\fvalCtx}{\trepof{\tau_1'}}{\sbar_1}\\
%%	%\delfromtau{$\Xi_0$}{\fvalCtx}{\tau_1'}{\sabs}\\\\
%%}{
%%	\atjsynX{\elam{x}{\tau_1}{e}}{\ttype{arrow}{(\tau_1', \tau_2)}}{\ilam{x}{\sigma'}{i}}
%%}
%%
%\inferrule[att-intro-ana]{
%	\vdash_\fvalCtx \FF{iana}(\fvar{tycon})=\taudef\\
%	\taudef~\taut{opidx}~\taut{tyidx}~((\Gamma; e_1)? :: \ldots :: (\Gamma; e_n)? :: []) \Downarrow_\fvalCtx \titerm{i}\\\\
%%	\trepof{\ttype{tycon}{\tauidx'}} \Downarrow_\fvalCtx \titype{\sigma}\\
%		\trepof{\ttype{tycon}{\taut{tyidx}}} \Downarrow_\fvalCtx \titype{\sigma}\\
%		\Gamma \vdash_\fvalCtx i : \sigma
%}{
%	\atjanaX{\FF{intro}[\taut{opidx}](e_1; \ldots; e_n)}{\ttype{tycon}{\taut{tyidx}}}{i}
%}
%
%%\inferrule[att-i-asc-ty]{
%%	\atjanaX{I}{\tau}{i}
%%}{
%%	\atjsynX{I : \tau}{\tau}{i}
%%}
%%
%%\inferrule[att-i-asc-tycon]{
%%	\vdash_\fvalCtx \FF{isyn}(\fvar{tycon})=\taudef\\\\
%%	\taudef~\tauidx~((\Gamma; e_1)? {::}{\ldots}{::}(\Gamma; e_n)? {::} []) \Downarrow_\fvalCtx \tden{\titerm{i}}{\ttype{tycon}{\tauidx'}}\\
%%			\trepof{\tau_1'} \Downarrow_\fvalCtx \titype{\sigma}\\
%%	\Gamma \vdash_\fvalCtx i : \sigma
%%}{
%%	\atjsynX{\FF{intro}[\tauidx](e_1; \ldots; e_n)] :: \fvar{tycon}}{\ttype{tycon}{\tauidx'}}{i'}
%%}
%%
%\inferrule[att-elim-syn]{
%	\atjsynX{e}{\ttype{tycon}{\taut{tyidx}}}{i}\\
%	\vdash_\fvalCtx \FF{esyn}(\fvar{tycon})=\taudef\\\\
%	\taudef~\taut{opidx}~\taut{tyidx}~\titerm{i}~((\Gamma; e)? :: (\Gamma; e_1)? :: \ldots :: (\Gamma; e_n)? :: []) \Downarrow_\fvalCtx (\tau, \titerm{i'})\\\\
%			\trepof{\tau} \Downarrow_\fvalCtx \titype{\sigma}\\
%	\Gamma \vdash_\fvalCtx i' : \sigma
%}{
%	\atjsynX{e\cdot\FF{elim}[\taut{opidx}](e_1; \ldots; e_n)}{\tau}{i'}
%}
%\end{mathpar}
%%\vspace{-10px}
%\caption{The bidirectional active typechecking and translation judgements.}
%\label{atj}
%\end{figure*}
%\newcommand{\tlevalX}[2]{#1 \Downarrow_\fvalCtx #2}
%\begin{figure*}[t]
%\small
%$\fbox{\inferrule{}{\tau \Downarrow_\fvalCtx \tau'}}$~~~~
%\begin{mathpar}
%\inferrule[trstr-eval]{ }{\tlevalX{\tlstr{s}}{\tlstr{s}}}
%
%\inferrule[tremp-eval]{ }{\tremp \Downarrow_\fvalCtx \tremp}
%
%\cdots
%
%\inferrule[tror-eval]{
%	\tlevalX{\tau_1}{\tau_1'}\\
%	\tlevalX{\tau_2}{\tau_2'}
%}{
%	\tlevalX{\tror{\tau_1}{\tau_2}}{\tror{\tau_1'}{\tau_2'}}
%}
%
%trmatch
%
%trlsubst
%
%trsublang
%
%
%%\inferrule[repof]{
%%	\tau \Downarrow_\fvalCtx \ttype{tycon}{\tauidx}\\
%%	\vdash_\fvalCtx \FF{rep}(\fvar{tycon}) = \taurep\\
%%	\taurep~\tauidx \Downarrow_\fvalCtx \titype{\sigma}
%%}{
%%	\FF{repof}(\tau) \Downarrow_\fvalCtx \titype{\sigma}
%%}
%%
%\inferrule[syn]{
%    \tau \Downarrow_\fvalCtx (\Gamma; e)?\\
%	\atjsynX{e}{\tau}{\iota}\\\\
%    [\tau/\tvar{t}_{ty}, \titerm{\iota}/\tvar{t}_{trans}]\tau_2 \Downarrow_\fvalCtx \tau_2'
%}{
%	\FF{syn}(\tau_1; \tvar{t}_{ty}, \tvar{t}_{trans}.\tau_2) \Downarrow_\fvalCtx \tau_2'
%}
%
%\inferrule[ana]{
%	\tau_1 \Downarrow_\fvalCtx (\Gamma; e)?\\
%	\tau_2 \Downarrow_\fvalCtx \tau_2'\\
%	\atjanaX{e}{\tau_2'}{\iota}\\\\
%	[\titerm{\iota}/\tvar{t}_{trans}]\tau_3 \Downarrow_\fvalCtx \tau_3'
%}{
%	\FF{ana}(\tau_1; \tau_2; \tvar{t}_{trans}.\tau_3) \Downarrow_\fvalCtx \tau_3'
%}
%
%TODO: add rx, str stuff
%\end{mathpar}
%\caption{\small Normalization semantics for the type-level language. Missing rules (including error propagation rules and normalization of quoted internal terms and types) are unsurprising and will be given later.}
%\label{tleval}
%\end{figure*}
%\begin{figure}[t]
%\small\begin{flalign}
%& \F{tycon}\fvar{stringin}~\F{of}\FF{R}~\{\\
%& \quad \F{iana}\{\tlam{opidx}{\FF{String}}{
%	\tlam{tyidx}{\FF{R}}{
%	\tlam{a}{\klist{\Q}}{\\
%& \quad\quad \tvar{arity0}~\tvar{a}~(\tvar{check}~\tvar{opidx}~\tvar{tyidx}~\titerm{\iup{\tvar{opidx}}})
%	}}
%}\}\\
%& \quad \F{esyn}\{\tlam{opidx}{\kunit+(\FF{R}+\FF{R})}{
%	\tlam{a}{\klist{\Q}}{\\
%& \quad\quad \tsumcase{\tvar{opidx}}{\_}{\\
%& \quad\quad\quad \tvar{arity2}~\tvar{a}~\tlam{a1}{\Q}{\tlam{a2}{\Q}{\\
%& \quad\quad\quad\quad \tvar{rsyn}~\tvar{a1}~\tlam{r1}{\FF{R}}{\tlam{i1}{\kITerm}{~}}\\
%& \quad\quad\quad\quad \tvar{rsyn}~\tvar{a2}~\tlam{r2}{\FF{R}}{\tlam{i2}{\kITerm}{~}}\\
%& \quad\quad\quad\quad\quad (\ttype{stringin}{\FF{rseq}(\tvar{r1}; \tvar{r2})}, \\
%& \quad\quad\quad\quad\quad\titerm{{\tt iconcat}(\iup{\tvar{i1}}; \iup{\tvar{i2}})})
%}}\\
%& \quad\quad}{opidx'}{d}}
%}\}
%\};\\
%& \F{def}\tvar{concat} = \tinl{\kunit, \FF{R}+\FF{R}}{\tunit};\\
%& \F{def}\tvar{subst} = \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinl{\FF{R},\FF{R}}{\tvar{r}}}};\\
%& \F{def}\tvar{coerce} = \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinr{\FF{R},\FF{R}}{\tvar{r}}}}
%\end{flalign}
%\caption{Definition of $\phi_S$, which enables the embedding of fragment $S$ into $\lamAce$.}
%\end{figure}
%%\begin{figure}
%%\[
%%\begin{array}{lcl}
%%% & & Definition & Kind\\
%%\tvar{concat} & := & \tinl{\kunit, \FF{R}+\FF{R}}{\tunit}\\%& \kunit + (R + R)\\
%%\tvar{replace} & := & \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinl{\FF{R},\FF{R}}{\tvar{r}}}}\\% & \karrow{a}{b}\\
%%\tvar{coerce} & := & \tlam{r}{\FF{R}}{\tinr{\kunit, \FF{R}+\FF{R}}{\tinr{\FF{R},\FF{R}}{\tvar{r}}}}\\
%%\end{array}
%%\]
%%\caption{Definitions}
%%\end{figure}
%
%\subsection{Background: Ace}
%TODO: Make a TR out of the OOPSLA submission.
%\subsection{Explicit Conversions}
%\subsection{Adding Subtyping to Ace}
%\subsection{Theory}
%
%\section{Related Work}
%
%\section{Discussion}
%
%
\bibliographystyle{abbrv}

\bibliography{research}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}
%
%\bibitem{lamport94}
%Leslie Lamport,
%\emph{\LaTeX: a document preparation system}.
%Addison Wesley, Massachusetts,
%2nd edition,
%1994.
%
%\end{thebibliography}
\end{document}