### 2014-12-16 sanitation-psp14/tr/strings_chapter.tex
Updated TR

\documentclass[11pt,leqno]{article}
\usepackage{cmu-titlepage2}
\usepackage{fullpage}
\usepackage[small,bf]{caption2}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage{enumerate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{tabularx}
\usepackage{morefloats}
\usepackage{bussproofs}    % Gentzen-style deduction trees *with aligned sequents!*
\usepackage{mathpartir}    % For type-settting type checking rule figures
\usepackage{syntax}        % For type-setting formal grammars.
\usepackage{hyperref}       % For links in citations
\usepackage{enumerate}
\usepackage{enumitem}

\input{../../att-icfp14/macros-atlam}

\usepackage{todonotes}

\newtheorem{tr}{TR Item}
\renewcommand{\thetr}{\Alph{tr}} % "letter-numbered" theorems
\newtheorem{trexam}[tr]{Example}
\newtheorem{trthm}[tr]{Theorem}
\newtheorem{trlem}[tr]{Lemma}
\newtheorem{trdef}[tr]{Definition}
\newtheorem{trconj}[tr]{Conjecture}
\newtheorem{ass}[tr]{Assumption}
\newtheorem{condition}[tr]{Condition}

\newtheorem{thm}{Theorem}                                                       
\newtheorem{cor}[thm]{Corollary}                                                
\newtheorem{lem}[thm]{Lemma}                                                    
\newtheorem{prop}[thm]{Proposition}                                             
\newtheorem{ax}[thm]{Axiom}                                                     
\theoremstyle{definition}                                                       
\newtheorem{defn}[thm]{Definition}                                              
\newtheorem{exam}[thm]{Example}                                                 
\newtheorem{rem}[thm]{Remark} 
%\renewcommand*{\proofname}{Proof Sketch}

%
% For type setting inference rules with labels.
%
\newcommand{\inferlbl}[3]
            {\inferrule{#3}{#2}{\textsf{\footnotesize{\sc #1}}}}
\newcommand{\inferline}[3]
            {\inferrule{#3}{#2} & {\textsf{\footnotesize{\sc #1}}} \\ \\}

\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\lang}[1]{\Lagr\{#1\}}
\newcommand{\stru}[2]{ {\tt string\_union}(#1,#2)}

\newcommand{\dconvert}[2]{ {\tt dconvert}(#1,#2) }
\newcommand{\filter}[2]{ {\tt filter}(#1,#2) }
\newcommand{\ifilter}[2]{ {\tt ifilter}(#1,#2) }

\newcommand{\reduces}{ \Rightarrow }
\newcommand{\gvd}{\Gamma \vdash }
\newcommand{\ovd}{\Omega \vdash }

\newcommand{\trep}{{\tt rep}}

\newcommand{\tstrf}[1]{`#1\textrm'} %??
\newcommand{\strf}[1]{``#1"}

\newcommand{\iso}{\cong}
%%
%% Source and Target language definitions.
%%
\newcommand{\lambdas}{\lambda_{RS}}
\newcommand{\lambdap}{\lambda_P}

% Source language terms.
\newcommand{\sisubst}[3]{{\sf rreplace}[#1](#2;#3)} \newcommand{\rreplace}[3]{{\sf rreplace}[#1](#2;#3)} % lots of legacy naming around these parts...

\newcommand{\rssreplace}[3]{\sisubst{#1}{#2}{#2}} % TODO-nrf fix this.

\newcommand{\coerce}[2]{ {\sf rcoerce}[#1](#2)}
\newcommand{\rcoerce}[2]{{\sf rcoerce}[#1](#2)}
\newcommand{\sistr}[1]{{\sf rstr}[#1]}   \newcommand{\rstr}[1]{{\sf rstr}[#1]} % Lots of legacy naming around these parts...

\newcommand{\val}{{\sf val}}

\newcommand{\rcheck}[4]{ {\sf rcheck}[#1](#2;#3;#4) }

\newcommand{\strin}[1]{\sistr{#1}}
\newcommand{\rsconcat}[2]{{\sf rconcat}(#1;#2)} \newcommand{\rconcat}[2]{{\sf rconcat}(#1;#2)} % lots of legact naming around these parts..

% Source language types.
\newcommand{\stringin}[1]{{\sf stringin}[#1]}

% target language terms.
\newcommand{\tsubst}[3]{{\sf replace}(#1;#2;#3)} \newcommand{\metareplace}[3]{{\sf replace}(#1;#2;#3)} % TODO-nrf rename the commands. Lots of legacy naming around these parts...

\newcommand{\tcheck}[4]{{\sf check}(#1; #2; #3; #4)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\preplace}[3]{{\sf replace}(#1;#2;#3)}
\newcommand{\tconcat}[2]{{\sf concat}(#1;#2)} \newcommand{\concat}[2]{{\sf concat}(#1;#2)} % lots of legacy naming around these parts...

\newcommand{\regext}[1]{ {\sf rx}[#1] } % TODO-nrf remove
\newcommand{\rx}[1]{ {\sf rx}[#1] }

% Target language types
\newcommand{\str}{{\sf string}}
\newcommand{\regex}{{\sf regex}}

% Meta-theoretic functions
\newcommand{\lsubst}[3]{{\sf replace}(#1;#2;#3)} % This used to renderlreplace(...) so there're probably mistkes wherever this command was used now.
\newcommand{\lreplace}[3]{{\sf lreplace}(#1; #2; #3)}

\newcommand{\sctx}{\Psi} % Context for external typing
\newcommand{\tctx}{\Theta} % Context for internal typing
\newcommand{\ereduces}{\Downarrow}

\newcommand{\strcase}[3]{ {\sf rstrcase}(#1; #2; #3)}
\newcommand{\pstrcase}[3]{ {\sf strcase}(#1; #2; #3)}

\newcommand{\lhead}[1]{ {\sf lhead}(#1) }
\newcommand{\ltail}[1]{ {\sf ltail}(#1) }

% Judgements
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.
% Relations
\newcommand{\treduces}{ \Downarrow }
\newcommand{\sreduces}{ \Downarrow }

%%
%% Constrain the size of full-page diagrams and rule lists
%%
%%\newcommand{\pagewidth}{5in}
%%\newcommand{\rulelistwidth}{3in}

% Names of type systems presented in paper
\newcommand{\lcs}{\lambda_{S}}

\setlength{\grammarindent}{3em}

% Arguments to lreplace are as follows:
% lreplace(pattern to replace, target, replacement).
% 
% ie lreplace(a,b,c) reads "replace all instances of a in b with c" and
% correspondes to Python's b.replace(a,c). 
\renewcommand{\lreplace}[3]{{\tt lreplace}(#1, #2, #3)}
\newcommand{\freplace}[3]{{\tt freplace}(#1, #2, #3)}
\newcommand{\freplacep}[4]{{\tt freplace}(#1,#2,#3,#4)}

\newcommand{\eps}{\epsilon}

\title{Statically Typed String Sanitation Inside a Python (Technical Report)}
\author{Nathan Fulton \and Cyrus Omar \and Jonathan Aldrich}
\date{December 2014}
\trnumber{CMU-ISR-14-112}
\keywords{type systems; regular languages; input sanitation; string sanitation}
\support{This work was supported by the National Security Agency lablet contract \#H98230-14-C-0140.}
\abstract{This report contains supporting evidence for claims put forth and explained in the paper ``Statically Typed String Sanitation Inside a Python'' \cite{fulton2014}, including proofs of lemmas
and theorems asserted in the paper, examples, additional discussion of the paper's technical content, and errata.}
\begin{document}
\maketitle

\tableofcontents
\listoffigures
\clearpage

\section{Terminology and Notation}
Theorems and lemmas appearing in \cite{fulton2014} are numbered correspondingly, while supporting facts appearing only in the Technical Report are lettered.

\section{Regular Expressions}
The syntax of regular expressions over some alphabet $\Sigma$ is shown in Figure \ref{fig:regex}.

\begin{ass}[Regular Expression Congruences]
We assume regular expressions are implicitly identified up to the following congruences:
\begin{align*}
\epsilon\cdot r & \equiv r\\
r \cdot \epsilon & \equiv r\\
(r_1\cdot r_2) \cdot r_3 & \equiv r_1 \cdot (r_2 \cdot r_3)\\
r_1 + r_2 &\equiv r_2 + r_1\\
(r_1 + r_2) + r_3 &\equiv r_1 + (r_2 + r_3)\\
\epsilon^* & \equiv \epsilon
\end{align*}
\end{ass}

\begin{ass}[Properties of Regular Languages] \label{thm:regexprops}
We assume the following properties:
\begin{enumerate}

\item If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1\cdot r_2}$.
\item For all strings $s$ and regular expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
\item Regular languages are closed under reversal.
\end{enumerate}
\end{ass}

\section{$\lambdas$}
The syntax of $\lambdas$ is specified in Figure \ref{fig:glambdas}. The static semantics is specified in Figure \ref{fig:slambdas}.

\subsection{Head and Tail Operations}
The following correctness conditions must hold for any definition of $\lhead{r}$ and $\ltail{r}$.

\begin{condition}[Correctness of Head] \label{trlemmax}
If $c_1  s' \in \lang{r}$, then
$c_1 \in \lang{\lhead{r}}$.
\end{condition}

\begin{condition}[Correctness of Tail] \label{trlemma:tail-correctness}
If $c_1 s' \in \lang{r}$ then $s' \in \lang{\ltail{r}}$.
\end{condition}

For example, we conjecture (but do not here prove) that the definitions below satisfy these conditions. Note that these are slightly amended relative to the published paper.

\begin{defn}[Definition of $\lhead{r}$]\label{def:lhead}
We first define an auxiliary relation that determines the set of characters that the head might be, tracking the remainder of any sequences that appear:
\begin{align*}
\lhead{\epsilon, \epsilon} &= \emptyset\\
\lhead{\epsilon, r'}      &= \lhead{r', \epsilon} \\
\lhead{a, r'}             &= \{a\} \\
\lhead{r_1\cdot r_2, r'}  &= \lhead{r_1, r_2 \cdot r'} \\
\lhead{r_1 + r_2, r'}     &= \lhead{r_1, r'} \cup \lhead{r_2, r'} \\
\lhead{r^*, r'}           &= \lhead{r, \epsilon} \cup \lhead{r', \epsilon}
\end{align*}
We define $\lhead{r} = a_1 + a_2 + ... + a_i$ iff $\lhead{r, \epsilon} = \{a_1, a_2, ..., a_i\}$.
\end{defn}

% \begin{trlem}[Correctness of Head] \label{trlemmax}
% If $c_1  s' \in \lang{r}$, then
% $c_1 \in {\lhead{r,\epsilon}}$.
% \end{trlem}
% \begin{proof}
% By structural induction on $r$.

% The case for $r = \epsilon$ does not apply because $s$ is non-empty.

% If $r = a$ for some character $a$ then $s \in \lang{r}$ implies $c_1 = a$.
% By definition $\lhead{a, \epsilon} = \{a\}$.
% Finally, $a \in \{a\}$ by definition.

% If $r = r_1 \cdot r_2$ then we proceed by case analysis on $r_1$. 
% \begin{quote}
% We need not consider the case where $r_1=\epsilon$ by congruence.

% If $r_1=a$ then $c_1=a$ and $\lhead{a \cdot r_2, \epsilon}=\lhead{a, r_2}=a$ and $a \in \{a\}$ by definition.

% If $r_1=r_{11}\cdot r_{12}$, then we have that $\lhead{r_{11} \cdot r_{12} \cdot r_2, \epsilon}=\lhead{r_{11}, r_{12}\cdot r_2}$. By the IH we have that $c_1 \in {\lhead{r_{11}\cdot r_{12}, \epsilon}}$. By Lemma \ref{trlemma:head-concatenation-1}, we have that $c_1 \in \lhead{r_{11}\cdot r_{12} \cdot r_2, \epsilon}$.

% If $r_1=r_{11} + r_{12}$, then by definition we have that

% \end{quote}

% If $r = r_1 \cdot r_2$ then $s \in \lang{r}$ implies $s = c_1 s_1 s_2$
% such that $c_1 s_1 \in \lang{r_1}$. 
% By definition of lhead, $\lhead{r_1 \cdot r_2} = \lhead{r_1, r_2} =
% \cdots + \lhead{r_1, \epsilon} = \cdots + \lhead{r_1}$. Therefore, $r_1 \subseteq \lhead{r_1, r_2}$
% and so it suffices to show that $s = c_1 s_1 s_2$ such that $c_1 \in \lhead{r_1}$,
% \todo{this is why we break out the $c_1$ from the $s_1$, both here and going forward}
% which holds by induction because $c_1 s_1 \in \lang{r_1}$.

% If $r = r_1 + r_2$ then $s \in r$ implies either $s \in r_1$ or $s \in r_2$.
% Without loss of generality suppose $s \in r_1$.
% Then by induction $s = c_1 \cdots c_m$ such that $c_1 \in \lhead{r_1}$.
% By definition, $\lhead{r} = \lhead{r_1, \epsilon} + \lhead{r_2, \epsilon} = \lhead{r_1} + \cdots$.
% Therefore, $c_1 \in \lhead{r}$.

% If $r = r_1^*$ then $s \in r$ implies that $s = s_1 \cdots s_k$ where $k \ge 0$ and $s_i \in \lang{r}$.
% By definition of lhead, $\lhead{r_1^*} = \epsilon + \lhead{r_1}$.
% If $s = \epsilon$, then $s \in \lang{\epsilon} \subseteq \lang{\lhead{r_1}}$.
% If $s = s_1 \cdots s_k$, then $s_1 \in \lang{r}$ and so by induction $s_1 = c_1 \cdots c_k$ such that $c_1 \in \lang{\lhead{r}}$.
% \end{proof}

\begin{defn}[Brzozowski's Derivative]\label{def:derivative}
  The \emph{derivative of $r$ with respect to $s$} is denoted by $\delta_s(r)$
  and is $\delta_s(r) = \{t | st \in \lang{r}\}$.
\end{defn}

\begin{defn}[Definition of $\ltail{r}$]
If $\lhead{r, \epsilon} = \{a_1, a_2, ..., a_i\}$, then  
we define $\ltail{r} = \delta_{a_1}(r) + \delta_{a_2}(r) + ... + \delta_{a_i}(r)$.
\end{defn}

% Using these definitions of head and tail, we establish a correctness
% result upon which type soundness for the concatenation operator in
% $\lambdas$ depends.

% \begin{trthm}[Correctness of \textsf{ltail}]
% If $s \in \lang{r}$ then $s \in \lang{\lhead{r} \cdot \ltail{r}}$.
% \end{trthm}
% \begin{proof}
% The proof proceeds by structural induction on $r$.
% In each case, we identify sublanguages of the head and tail of $r$ such that 
% $s$ is in the concatenation of the the sublanguages.
% Throughout the proof, we use the fact that $\lang{r'} \subseteq \lang{r}$ and $\lang{q'} \subseteq \lang{q}$
% implies $\lang{r' \cdot q'} \subseteq \lang{r \cdot q}$.

% \begin{itemize}[label=$ $,itemsep=1ex]

% \item \textbf{Case $r = c$ for $c \in Sigma$}.
% By definition, if $s \in \lang{c}$ then $s = c$.
% By definition of lhead and ltail, $\lhead{c} = \lhead{c, \epsilon} = c$
% and $\lang{\epsilon} \subseteq \lang{\ltail{r}}$.
% Therefore, $c \in \lang{c \cdot \epsilon} \subseteq \lang{\lhead{r} + \ltail{r}}$.

% \item \textbf{Case $r = r_1 \cdot r_2$}.
% Suppose $s \in \lang{r}$.
% We may decompose $s$ as $s_1  \cdots s_n$ such that $s_1 = c_1 s_1' \in \lang{r_1}$.
% Note that $c_1 \in \lhead{r_1}$ by \ref{trlemmax}.
% Therefore, $c_1 \in \lhead{r_1 \cdot r_2}$ by definition of lhead.
% It suffices to show that $s_1' s_2 \in \lang{\delta_{c_1}(r)}$.
% Since $s = c_1 s_1' s_2\in \lang{r}$, it follows by the definition of derivative that $s_1' s_2 \in \delta_{c_1}(r)$.
% Therefore, $s = c_1 s_1' s_2 \in \lang{ c_1 \cdot \delta_{c_1} } \subseteq \lang{ \lhead{r} \cdot \ltail{r}}$.

% \item \textbf{Case $r = r_1 + r_2$}.
% Suppose $s \in \lang{r_1 + r_2}$
% so that $s \in \lang{r_1}$ or $s \in \lang{r_2}$.
% Consider without loss of generality the case where $s \in \lang{r_1}$.
% By induction, $s \in \lang{\lhead{r_1} \cdot \ltail{r_1}}$
% Note that $\lhead{r_1} \subseteq \lhead{r}$
% and $\ltail{r_1} \subseteq \ltail{r}$.
% Therefore, $\lhead{r_1} \cdot \ltail{r_2} \subseteq \lhead{r_1} \cdot \ltail{r_1}$.

% \item \textbf{Case $r = q^*$}.
% Either $s$ is the empty string or else the $k^{th}$ unwinding (i.e., $s = s_1 \cdots s_k$ where $k > 0$ and $s_i \in \lang{q}$).

% Suppose $s = \epsilon$.
% By definition of lhead, $\lhead{r} = \lhead{q^*} = \lhead{q^*, \epsilon} = \lhead{\epsilon, \epsilon} + \cdots = \epsilon + \cdots$.
% By definition of ltail, $\ltail{r} = \epsilon \cdots$.
% From these facts we conclude that $s = \epsilon \in \lang{\epsilon \cdot \epsilon} \subseteq \lang{\lhead{r} \cdot \ltail{r}}$.

% Suppose instead that $s = s_1 \cdots s_k$ for $k < 0$ where $s_i \in \lang{q}$.
% If $q = \epsilon$ then $s = \epsilon$ and the conclusion follows for the same reason as the epsilon subcase of the character case.
% Otherwise, $s$ is not the empty string and so $s = c_1 s_1 s_2$ where $s_1$ and $s_2$ might be empty.
% By the lemma above, $c_1 \in \lhead{s}$.
% By the definition of derivative, $s_1s_2 \in \delta_{c_1}(s)$.
% Therefore, $s \in \lang{\lhead{s} \cdot \delta_{c_1}(s)} \subseteq \lang{\lhead{r} \cdot \ltail{r}}$, where the set inclusion follows by the definition
% of lhead and ltail.
% \end{itemize}
% \end{proof}

% \begin{trexam}[Decomposition is not Reversible]\label{exam:headsandtails}
% $r \not = \lhead{r} \cdot \ltail{r}$.
% \end{trexam}
% \begin{proof}
% A simple counter-example is $ab + cd$. 
% Note that $\lhead{a\cdot b + c\cdot d} = a+c$ and $\ltail{a\cdot b + c\cdot d} = b + d$. 
% Therefore, $\{ad,bc\} \subset \lang{\lhead{a\cdot b + cd\cdot } \cdot \ltail{a\cdot b + c\cdot d}}$ even though neither of these is in $\lang{r}$.
% \end{proof}

% Example~\ref{exam:headsandtails} does not imply a counter-example to type soundness because $s \in \lang{r} \implies s \in \lhead{r}\cdot\ltail{r}$ is the
% property required for soundness.

% This is a general pattern in string operations: $\lambdas$ simulates -- within the type system -- common operations on strings. 
% If there is an operation for concatenating to strings, we define an operation for concatenating two regular expressions.
% If there is an operation for peeling off the first (n) characters of a string, then we define an operation for 
% converting a regular expression $r$ into a regular expression $r'$ which recognizes any $n^{th}$ suffix of a string in $r$.

% It is important to note, however, that the type system need not \emph{exactly} simulate the action of string operations.
% In the case of concatenation, we lose some information because more string values are possible -- according the types --
% than are actually possible in the dynamic semantics.
% Soundness is not lost because the types are conservative in their approximation.

% In the case of string replacement, there are \emph{trivial} definitions of substitution (on strings) and replacement (on languages) which over-approximate 
% the effect of a substitution. Closing these gaps in approximation is important, and motivates the string operations portion of this technical report.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % Substitution.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Replacement}
The following correctness condition must hold for any definition of $\lreplace{r}{r_1}{r_2}$.

\begin{condition}[Replacement Correctness] \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $$\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{r_1}{r_2}}$$
\end{condition}

We do not give a particular definition for $\lreplace{r}{r_1}{r_2}$ here.
%\begin{lem}
%  If $\emptyset \vdash e : \stringin{r}$ then $r$ is a well-formed regular expression.
%\end{lem}
%\begin{proof}
%  The proof proceeds by induction on the derivation of the premise.
%  The only non-trivial cases (those which require more than an appeal to inversion) are S-T-Case, S-T-Replace and S-T-Concat. 
%
%In the S-T-Case case, note that ${\sf lhead}$ and ${\sf ltail}$ are total functions for well-formed regular expressions to well-formed regular expressions.
%
%In the S-T-Concat case, note that Lemma~\ref{thm:total} implies that if $r_1$ and $r_2$ are regular expressions then so is $r_1 \cdot r_2$.
%
%In the S-T-Replace case, it suffices to show that $\lreplace{r}{r_1}{r_2}$ is a regular expression assuming (inductively) that $r, r_1$ and $r_2$ are all regular expressions.
%This follows from the Closure proposition.
%\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Type Safety and Security.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Small Step Semantics of $\lambdas$}
Figure~\ref{fig:sdlambdas} specifies a small-step operational semantics for $\lambdas$.

%\newcommand{\therel}{R_{\stringin{r}}}
%\begin{trdef}[Relation over $\stringin{r}$]
%$\therel(e)$ iff $e$ halts and $\emptyset \vdash e : \stringin{r}$.
%\end{trdef}

%\begin{trconj}
%if $\Gamma \vdash e : \sigma$ then $e \mapsto^* v$ such that $v \ \val$.
%\end{trconj}
%%\begin{trthm}[Termination of small step semantics]
%%\end{trthm}
%%\begin{proof}
%We do not develop the full proof here, but note that the simply typed lambda calculus terminates.
%For the string fragment, observe that the S-T- rules do not add any non-trivial binding structure because substitutions
%$[e/x]e'$ may only occur in the special case where $e = \rstr{s}$, so that the length of the term
%never increases and the number of free variables strictly decreases. Therefore, the standard normalization
%argument proceeds without complication after fixing an evaluation order for the compatibility rules 
%(all our other proofs are agnostic to evaluation order).
%\end{proof}

\begin{trlem}[Canonical Forms]\label{cfgen}
If $\emptyset \vdash v : \sigma$ then:
\begin{enumerate}
\item If $\sigma = \stringin{r}$ then $v = \rstr{s}$ and $s \in \lang{r}$.
\item If $\sigma = \sigma_1 \rightarrow \sigma_2$ then $v = \lambda x . e'$.
\end{enumerate}

\end{trlem}
\begin{proof}
By inspection of the static and dynamic semantics.
\end{proof}

\begin{trlem}[Progress]
If $\emptyset\vdash e : \sigma$ either $e = v$ for some $v$ or $e \mapsto e'$ for some $e'$.
\end{trlem}
\begin{proof}
The proof proceeds by rule induction on the derivation of $\emptyset\vdash e : \sigma$.
\begin{itemize}[label=$ $,itemsep=1ex]
\item \textbf{$\lambda$ fragment}. Cases SS-T-Var, SS-T-Abs, and SS-T-App are exactly as in a proof of progress for
the simply typed lambda calculus.

\item \textbf{S-T-Stringin-I}. Suppose $\emptyset\vdash \rstr{s} : \stringin{s}$.
Then $e=\rstr{s}$.

\item \textbf{S-T-Concat}. Suppose $\emptyset\vdash \rconcat{e_1}{e_2} : \stringin{r_1 \cdot r_2}$ and $\emptyset \vdash e_1 : \stringin{r_1}$ and $\emptyset \vdash e_2 : \stringin{r_2}$.
By induction, $e_1 \mapsto e_1'$ or $e_1 = v_1$ and similarly, $e_2 \mapsto e_2'$ or $e_2 = v_2$. 
If $e_1$ steps, then SS-E-Concat-Left applies and so
$\rconcat{e_1}{e_2} \mapsto \rconcat{e_1'}{e_2}$. Similarly, if $e_2$ steps
then $e$ steps by SS-E-Concat-Right.

In the remaining case, $e_1 =v_1$ and $e_2 =v_2$.
But then it follows by Canonical Forms that $e_1 = \rstr{s_1}$
and $e_2 = \rstr{s_2}$.
Finally, by SS-E-Concat, $\rconcat{\rstr{s_1}}{\rstr{s_2}} \mapsto \rstr{s_1s_2}$.

\def\mline#1{\par\hspace*{-\leftmargin}\parbox{\textwidth}{\[#1\]}}
\item \textbf{S-T-Case}.
Suppose $e = \strcase{e_1}{e_2}{x,y.e_3}$ and $\emptyset\vdash e_1 : \stringin{r}$.
%\begin{align}
%&\vdash e_1 : \stringin{r} \\
%&\vdash e_2 : \sigma \\
%x:\stringin{\lhead{r}}, y : \stringin{\ltail{r}} &\vdash e_3 : \sigma
%\end{align}
%Actually this isn't even necessary...
%By (2), $e_2 \mapsto e_2'$ or $e_2 \val$. 
%In the former case, $e \mapsto \strcase{e_1}{e_2'}{x,y.e_3}$.
By induction and Canonical Forms it follows that $e_1 \mapsto e_1'$ or $e_1 = \rstr{s}$.
In the former case, $e$ steps by S-E-Case-Left.
In the latter case, note that $s = \epsilon$ or $s = at$ for some string $t$.
If $s = \epsilon$ then $e$ steps by S-E-Case-$\epsilon$-Val,
and if $s = at$ the $e$ steps by S-E-Case-Concat.

\item \textbf{S-T-Replace}.
Suppose $e = \rreplace{r}{e_1}{e_2}$, $\emptyset\vdash e : \stringin{\lreplace{r}{r_1}{r_2}}$ and:
\begin{align}
&\emptyset\vdash e_1 : \stringin{r_1} \\
&\emptyset\vdash e_2 : \stringin{r_2}
\end{align}

By induction on (1), $e_1 \mapsto e_1'$ or $e_1 = v_1$  for some $e_1'$.
If $e_1 \mapsto e_1'$ then $e$ steps by SS-E-Replace-Left.
Similarly, if $e_2$ steps then $e$ steps by SS-E-Replace-Right.
The only remaining case is where $e_1 = v_1$ and also $e_2 = v_2$.
By Canonical Forms, $e_1 = \strin{s_1}$ and $e_2 = \strin{s_2}$.
Therefore, $e \mapsto \rstr{\lsubst{r}{s_1}{s_2}}$ by SS-E-Replace.

\item \textbf{S-T-SafeCoerce}.
Suppose that $\emptyset\vdash \rcoerce{r}{e_1} : \stringin{r}$.
and
%\begin{align}
$\emptyset\vdash e_1 : \stringin{r'}$ for $\lang{r'} \subseteq \lang{r}$.% \\
%&\lang{r'} \subseteq \lang{r}
%\end{align}
By induction, $e_1 =v_1$ or $e_1 \mapsto e_1'$ for some $e_1'$.
If $e_1 \mapsto e_1'$ then $e$ steps by SS-E-SafeCoerce-Step.
Otherwise, $e_1 =v$ and by Canonical Forms $e_1 = \rstr{s}$.
In this case, $e = \rcoerce{r}{\rstr{s}} \mapsto \rstr{s}$ by SS-E-SafeCoerce.

\item \textbf{S-T-Check}
Suppose that $\emptyset\vdash \rcheck{r}{e_0}{x.e_1}{e_2} : \stringin{r}$ and:
\begin{align}
\emptyset &\vdash e_0 : \stringin{r_0}  \\
\emptyset, x:\stringin{r} &\vdash e_1 : \sigma \\
\emptyset &\vdash e_2 : \sigma \label{align6}
\end{align}

By induction,  $e_0 \mapsto e_0'$ or $e_0 = v$.
In the former case $e$ steps by SS-E-Check-StepLeft.
Otherwise, $e_0 = \rstr{s}$ by Canonical Forms.
By Lemma~\ref{thm:regexprops} part 2, either $s \in \lang{r_0}$ or $s \not\in\lang{r_0}$.
In the former case $e$ takes a step by SS-E-Check-Ok.
In the latter case $e$ takes a step by SS-E-Check-NotOk.

\end{itemize}

\end{proof}

\begin{ass}[Substitution] \label{lem:subst}
If $\Psi, x : \sigma' \vdash e : \sigma$ and $\Psi \vdash e' : \sigma'$, then
$\Psi \vdash [e' / x] e : \sigma$.
\end{ass}

\begin{trlem}[Preservation for Small Step Semantics]
If $\emptyset \vdash e : \sigma$ and $e \mapsto e'$ then $\emptyset \vdash e' : \sigma$.
\end{trlem}
\begin{proof}

By induction on the derivation of $e \mapsto e'$ and $\emptyset \vdash e : \sigma$.

\begin{itemize}[label=$ $,itemsep=1ex]
\item \textbf{$\lambda$ fragment}. Cases SS-E-AppLeft, SS-E-AppRight, and SS-E-AppAbs are exactly as in a proof of type safety for
the simply typed lambda calculus.

\item \textbf{SS-E-Concat-Left}.
Suppose $e = \rconcat{e_1}{e_2} \mapsto \rconcat{e_1'}{e_2}$ and $e_1 \mapsto e_1'$.
The only rule that applies is S-T-Concat, so $\emptyset\vdash e_1 : \stringin{r_1}$ and $\emptyset\vdash e_2 : \stringin{r_2}$.
By induction, $\emptyset\vdash e_1' : \stringin{r_1}$.
Therefore, by S-T-Concat, $\emptyset\vdash \rconcat{e_1'}{e_2} : \stringin{r_1r_2}$.

\item {\textbf{SS-E-Concat-Right}}. 
Suppose $e = \rconcat{e_1}{e_2} \mapsto \rconcat{e_1}{e_2'}$ and $e_2 \mapsto e_2'$.
The only rule that applies is S-T-Concat, so $\emptyset\vdash e_1 : \stringin{r_1}$ and $\emptyset\vdash e_2 : \stringin{r_2}$.
By induction, $\emptyset\vdash e_2' : \stringin{r_2}$.
Therefore, by S-T-Concat, $\emptyset\vdash \rconcat{e_1}{e_2'} : \stringin{r_1r_2}$.

\item \textbf{SS-E-Concat}. 
Suppose $\rconcat{\rstr{s_1}}{\rstr{s_2}} \mapsto \rstr{s_1s_2}$.
The only applicable rule is S-T-Concat, so 
$\emptyset\vdash\rstr{s_1} : \stringin{r_1}$ 
and
$\emptyset\vdash\rstr{s_2} : \stringin{r_2}$ and $\emptyset \vdash \rconcat{\rstr{s_1}}{\rstr{s_2}} : \stringin{r_1 \cdot r_2}$. 
By Canonical Forms, $s_1 \in \lang{r_1}$ and $s_2 \in  \lang{r_2}$ from which it follows by Lemma~\ref{thm:regexprops}
that $s_1s_2 \in \lang{r_1\cdot r_2}$. Therefore, $\emptyset\vdash \rstr{s_1s_2} : \stringin{r_1\cdot r_2}$ by S-T-Rstr.

\item \textbf{S-E-Case-Left}.
Suppose 
$e \mapsto \strcase{e_1'}{e_2}{x,y.e_3}$
and $\emptyset \vdash e : \sigma$ and $e_1 \mapsto e_1'$.
The only rule that applies is S-T-Case, so:
\begin{align}
\emptyset&\vdash e_1 : \stringin{r} \label{align7} \\
\emptyset&\vdash e_2 : \sigma \label{align8} \\ 
\emptyset, x:\stringin{\lhead{r}}, y : \stringin{\ltail{r}} &\vdash e_3 : \sigma \label{align9}
\end{align}
By (\ref{align7}) and the assumption that $e_1 \mapsto e_1'$, it follows by induction that $\emptyset\vdash e_1' : \stringin{r}$.
This fact together with (\ref{align8}) and (\ref{align9}) implies by S-T-Case that
$\emptyset\vdash\strcase{e_1'}{e_2}{x,y.e_3} : \sigma$.

%\item \textbf{SS-E-Case-Right}. 
%We have that $e = \strcase{e_1}{e_2}{x,y.e_3}$, Suppose $e \mapsto \strcase{e_1}{e_2'}{x,y.e_3}$
%and $\emptyset \vdash e : \stringin{r}$.
%By inversion of S-T-Case:
%\begin{align}
%\emptyset&\vdash e_1 : \stringin{r} \label{align10}\\
%\emptyset&\vdash e_2 : \sigma \label{align11} \\ 
%x:\stringin{\lhead{r}}, y : \stringin{\ltail{r}} &\vdash e_3 : \sigma \label{align12}
%\end{align}
%By (\ref{align11}) and the assumption that $e_2 \mapsto e_2'$, it follows by induction that $\emptyset\vdash e_2' : \stringin{r}$.
%This fact together with (\ref{align10}) and (\ref{align12}) implies by S-T-Case that 
%$\emptyset\vdash\strcase{e_1}{e_2'}{x,y.e_3} : \sigma$.

\item \textbf{SS-E-Case-$\epsilon$-Val}.
Suppose $\strcase{e_0}{e_2}{x,y.e_3} \mapsto e_2$. 
The only rule that applies is S-T-Case, so $\emptyset \vdash e_2 : \sigma$.

\item \textbf{SS-E-Case-Concat}.
Suppose that $e = \strcase{\rstr{as}}{e_2}{x,y.e_3} \mapsto [\rstr{a}, \rstr{s} / x,y] e_3$
and that $\emptyset\vdash e : \sigma$.
The only rule that applies is S-T-Case so:
\begin{align}
\emptyset&\vdash \rstr{as} : \stringin{r} \label{align13}\\
\emptyset&\vdash e_2 : \sigma \label{align14} \\ 
\emptyset, x:\stringin{\lhead{r}}, y : \stringin{\ltail{r}} &\vdash e_3 : \sigma \label{align15}
\end{align}

We know that $as \in \lang{r}$ by Canonical Forms on (\ref{align13}) 
Therefore, $a \in \lang{\lhead{r}}$ by Condition \ref{trlemmax} and $s \in \lang{\ltail{r}}$ by Condition \ref{trlemma:tail-correctness}.

From these facts about $a$ and $s$ we know by S-T-Rstr that $\emptyset\vdash \rstr{a} : \stringin{\lhead{r}}$
and $\emptyset\vdash \rstr{s} : \stringin{\ltail{r}}$.
It follows by Assumption \ref{lem:subst} that  $\emptyset\vdash [\rstr{a}, \rstr{s} / x,y] e_3 : \sigma$.\todo{Cyrus stopped here}

\item \textbf{Cases SS-E-Replace-Left, SS-E-Replace-Right, SS-E-Check-StepLeft, SS-E-SafeCoerce-Step, SS-E-Check-StepRight}. \todo{todo}

\item \textbf{Case SS-E-Replace}.

Suppose $e = \rreplace{r}{\rstr{s_1}}{\rstr{s_2}} \mapsto \rstr{\lsubst{r}{s_1}{s_2}}$.
The only applicable rule is 
S-T-Replace, so
\begin{align*}
\emptyset &\vdash \rstr{s_1} : \stringin{r_1} \\
\emptyset &\vdash \rstr{s_2} : \stringin{r_2}
\end{align*}
By conanical forms,
$s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$.
Therefore, $\lreplace{r}{s_1}{s_2} \in \lang{\lreplace{r}{r_1}{r_2}}$ by Theorem~\ref{thm:substcorrespondence}.
It is finally derivable by S-T-Rstr that:

$\emptyset \vdash \rstr{\lreplace{r}{s_1}{s_2}} : \stringin{\lreplace{r}{r_1}{r_2}}$.

\item \textbf{Case SS-E-SafeCoerce}.
Suppose that $\rcoerce{r}{\rstr{s_1}} \mapsto \rstr{s_1}$.
The only applicable rule is S-T-SafeCoerce, so 
$\emptyset\vdash \rcoerce{r}{s_1} : \stringin{r}$.
By Canonical Forms, $s \in \lang{r}$. Therefore, $\emptyset\vdash \rstr{s} : \stringin{r}$.

\item \textbf{Case SS-E-Check-Ok}.
Suppose $\rcheck{r}{\rstr{s}}{x.e_1}{e_2} \mapsto \subst{\rstr{s}}{x}{e_1}$,
$s \in \lang{r}$, and $\emptyset \vdash \rcheck{r}{\rstr{s}}{x.e_1}{e_2} : \sigma$.
By inversion of S-T-Check, $x : \stringin{r} \vdash e_1 : \sigma$.
Note that $s \in \lang{r}$ implies that $s : \stringin{r}$ by S-T-RStr.
Therefore, $\emptyset \vdash \subst{\rstr{s}}{x}{e_1} : \sigma$.

\item \textbf{Case SS-E-Check-NotOk}.
Suppose $\rcheck{r}{\rstr{s}}{x.e_1}{e_2} \mapsto e_2$,
$s \not \in \lang{r}$, and $\emptyset \vdash \rcheck{r}{\rstr{s}}{x.e_1}{e_2} : \sigma$.
The only applicable rule is S-T-Check, so $\emptyset \vdash e_2 : \sigma$.

\end{itemize}
\end{proof}

\begin{trthm}[Type Safety for small step semantics.] \label{thm:sstypesafte}
If $\emptyset\vdash e : \sigma$ then either $e \ \val$ or $e \mapsto^* e'$ and $\emptyset\vdash e' : \sigma$.
\end{trthm}
\begin{proof}
Follows directly from progress and preservation.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BEGIN THE SEMANTIC CORRESPONDENCE ARGUMENT FOR LAMBDA_RS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Semantic Correspondence between Big and Small Step Semantics for $\lambdas$}

Before extending the previous theorem to the big step semantics, we first establish a correspondence between
the big step semantics in Figure~\ref{fig:sdlambdas} and the small step semantics in Figure~\ref{fig:dlambdas}.

% bookmark1
\begin{trlem} \label{lem:inductive}
If $e \sreduces v$ and $e \mapsto e'$ then $e' \sreduces v$. 
\end{trlem}
\begin{proof}
By induction on the structure of e.
\begin{itemize}[label=$ $,itemsep=1ex]

\item \textbf{Case $e = e_1(e_2)$}.
The only applicable rule is S-E-App, so $e_1 \sreduces \lambda x . e_3$
and $e_2 \sreduces v_2$ such that $[v_2/x] e_3 \sreduces v$.

The term $e = e_1(e_2)$ may step by three rules.

First, if $e$ steps by L-E-AppLeft then $e_1(e_2) \mapsto e_1'(e_2)$.
By induction, $e_1' \sreduces \lambda x . e_3$.
By S-E-App, $e_1'(e_2) \sreduces \lambda x . e_3$.

Second, if $e$ steps by L-E-AppRight then $e_1(e_2) \mapsto e_1(e_2')$.
By induction, $e_2' \sreduces v_2$.
By S-E-App, $e_1(e_2') \sreduces \lambda x . e_3$.

Third, if $e$ steps by L-E-AppAbs then $e = (\lambda x . e_3)(v_2) \mapsto [v_2 / x]e_3$.
By induction, $e = (\lambda x . e')v_2 \sreduces v$.
\end{itemize}
\end{proof}

\begin{trthm}[Semantic Correspondence for $\lambdas$ (Part I)]
If $e \sreduces v$ then $e \mapsto^* v$.  
\end{trthm}
\begin{proof}
We proceed by structural induction on $e$.

\begin{itemize}[label=$ $,itemsep=1ex]

% Abstraction case.
\item \textbf{Case $e = \lambda x . e_1$}. 
The only applicable rule is S-E-Abs, so $v = \lambda x . e_1$.
Note that $\lambda x . e_1 \mapsto^* \lambda x . e_1$ by RT-Refl.

% bookmark2
% Application case.
\item \textbf{Case $e=e_1(e_2)$}.
The only applicable rule is S-E-App. By inversion:
\begin{align*}
  e_1 &\sreduces \lambda x.e_1' \\
  e_2 &\sreduces v_2 \\
  [v_2/x]e_1' &\sreduces v
\end{align*}

From which it follows by induction that:
\begin{align*}
  e_1 &\mapsto^* \lambda x.e_1'  \\
  e_2 &\mapsto^* v_2 \\
  [v_2/x]e_1' &\mapsto^* v
\end{align*}

If $e_1 = \lambda x . e_1'$ and $e_2 = v_2$ (henceforth the reflexive case)
then $e \mapsto [v_2 / x] e_1'$ and the conclusion follows by RT-Trans.

If $e_1 \mapsto \lambda x . e_1'$ then $e_1(e_2) \mapsto (\lambda x . e_1')(e_2) \mapsto [e_2/x]e_1'$ and the conclusion follows by two applications
of RT-Trans.

If $e_1 \mapsto^k \lambda x . e_2'$ then $e_1 \mapsto e'$ so $e_1(e_2) \mapsto e'(e_2)$ and by Lemma~\ref{lem:inductive} $e'(e_2) \mapsto^* v$.
So $e_1(e_2) \mapsto e'(e_2) \mapsto^* v$; it follows by RT-Trans that $e_1(e_2) \mapsto v$.

\todo{not modified below this line...}

Note that the following rule is derivable by repeating applications of the left and right compatibility rules for application:
\begin{mathpar}
\inferrule[L*-App]
{ e_1 \mapsto^* e_1' \\ e_2 \mapsto^* e_2'  }
{ e_1(e_2) \mapsto^* e_1'(e_2') }
\end{mathpar}

From these facts and L-AppAbs, we may establish that $e_1(e_2) \mapsto^* ( \lambda x.e_2 )(v_2) \mapsto [v_2/x]e_2$.
Note that $[v_2/x]e_2 \mapsto^* v$, so by RT-Trans it follows that $e = e_1(e_2) \mapsto^* v$.

% String Values
\item \textbf{Case $e = \strin{s}$}. The only applicable rule is S-E-RStr, so $v = \strin{s}$.
By RT-Refl, $\strin{s} \mapsto^* \strin{s}$.

% String Concatenation.
\item \textbf{Case $e=\rconcat{e_1}{e_2}$}. 
The only applicable rule is S-E-Concat, so $v = \rstr{s_1s_2}$.
By inversion, $e_1 \sreduces \strin{s_1}$ and $e_2 \sreduces \strin{s_2}$.
By induction, $e_1 \mapsto^* \strin{s_1}$ and $e_2 \mapsto^* \strin{s_2}$.
Note that the rule following is derivable:
\begin{mathpar}
\inferrule[SS-E-Concat-LR*]
{ e_1 \mapsto^* e_1' \\ e_2 \mapsto^* e_2'  }
{ \rconcat{e_1}{e_2} \mapsto^* \rconcat{e_1'}{e_2'} }
\end{mathpar}

From these facts, it follows that $\rconcat{e_1}{e_2} \mapsto^* \rconcat{\strin{s_1}}{\strin{s_2}}$.
Finally, $\rconcat{\strin{s_1}}{\strin{s_2}} \mapsto \strin{s_1s_2}$ by SS-E-Concat.
By RT-Step, it follows that $\rconcat{e_1}{e_2} \mapsto^* \strin{s_1s_2}$.

%Strcase Case.
\item \textbf{Case $e=\strcase{e_1}{e_2}{x,y.e_3}$}.

There are two subcases. For the first, suppose $\strcase{e_1}{e_2}{x,y.e_3} \sreduces v$ was finally derived by S-E-Case-$\epsilon$.
By inversion: 
\begin{align*}
e_1 &\sreduces \strin{\epsilon} \\
e_2 &\sreduces v
\end{align*}
from which it follows by induction that:
\begin{align*}
e_1 &\mapsto^* \strin{\epsilon} \\
e_2 &\mapsto^* v
\end{align*}

Note that the following rule is derivable:

\begin{mathpar}
\inferrule[SS-E-Case-LR*]
{ e_1 \mapsto^* e_1' \\ e_2 \mapsto^* e_2'  }
{ \strcase{e_1}{e_2}{x,y.e_3} \mapsto^* \strcase{e_1'}{e_2'}{x,y.e_3} }
\end{mathpar}

From these facts is follows that $e \mapsto^* \strcase{\strin{\epsilon}}{v}{x,y.e_3}$.
By S-E-Case-$\epsilon$-Val and RT-Step it follows that $e \mapsto^* v$.

Now consider the other case where $\strcase{e_1}{e_2}{x,y.e_3} \sreduces v$ was finally derived by S-E-Case-Concat.
By inversion, 
$e_1 \sreduces \strin{as}$ and $[\strin{a}, \strin{s} / x, y]e_3 \sreduces v$. 
From these facts it follows by induction that 
$e_1 \mapsto^* \strin{as}$ and
$[\strin{a}, \strin{s} / x, y]e_3 \mapsto^* v$.

By the first of these facts, it is derivable via SS-E-Case-LR* that 
$e \mapsto^* \strcase{e_1'}{\strin{as}}{x,y.e_3}$.
SE-E-Case-Concat applies to this form, so by RT-Step we know $e \mapsto^* [\strin{a}, \strin{s} / x, y]e_3$.
Recall that $[\strin{a}, \strin{s} / x, y]e_3 \mapsto^* v$, so by RT-Trans we finally derive $e \mapsto^* v$.

% String Replacement.
\item \textbf{Case $e=\rreplace{r}{e_1}{e_2}$}.
There is only one applicable rule, so $v = \rstr{s}$ and by inversion it follows that:
\begin{align*}
e_1 &\sreduces \rstr{s_1} \\
e_2 &\sreduces \rstr{s_2}
\end{align*}

From which it follows by induction that:
\begin{align*}
e_1 &\mapsto^* \rstr{s_1} \\
e_2 &\mapsto^* \rstr{s_2}
\end{align*}

Furthermore, $\lsubst{r}{s_1}{s_2} = s$ by induction.
Note that the following rule is derivable:

\begin{mathpar}
\inferrule[SS-E-Replace-LR*]
{ e_1 \mapsto^* e_1' \\ e_2 \mapsto^* e_2'  }
{ \rreplace{r}{e_1}{e_2} \mapsto^* \rreplace{r}{e_1'}{e_2'} }
\end{mathpar}

From these facts, $\rreplace{r}{e_1}{e_2} \mapsto^* \rreplace{r}{\rstr{s_1}}{\rstr{s_2}}$.

Finally, $\rreplace{r}{\rstr{s_1}}{\rstr{s_2}} \mapsto \lsubst{r}{s_1}{s_2}$.

From these two facts we know via RT-Step that $\rreplace{r}{e_1}{e_2} \mapsto^* \rreplace{r}{e_1}{e_2}$.
Recall that $\lsubst{r}{s_1}{s_2} = s$, from which the conclusion follows.

\item \textbf{Case $e = \rcoerce{r}{e_1}$}.
In this case $e \sreduces v$ is only finally derivable via S-E-SafeCoerce.
Therefore, $v = \strin{s}$ and by inversion $e_1 \sreduces \strin{s}$.
By induction, $e_1 \mapsto^* \strin{s}$.

The following rule is derivable:

\begin{mathpar}
\inferrule[SS-E-SafeCoerce-Step]
{ e \mapsto^* e' }
{ \rcoerce{r}{e} \mapsto^* \rcoerce{r}{e'} }
\end{mathpar}

Applying this rule at $e_1 \mapsto^* \strin{s}$ derives $\rcoerce{r}{e_1} \mapsto^* \rcoerce{r}{\strin{s}}$.
In the final step, $\rcoerce{r}{\strin{s}} \mapsto \strin{s}$ by SS-E-SafeCoerce.
From this fact, we may derive via RT-Trans that $e \mapsto^* \strin{s}$ as required.

\item \textbf{Case $e = \rcheck{r}{e_1}{x.e_2}{e_3}$}.

Note that the rule following is derivable:

\begin{mathpar}
\inferrule[SS-E-Check-Step]
{ e_1 \mapsto^* e_1' \\ e_3 \mapsto^* e_3' }
{ \rcheck{r}{e_1}{x.e_2}{e_3} \mapsto^* \rcheck{r}{e_1'}{x.e_2}{e_3'} }
\end{mathpar}

There are two ways to finally derive $e \sreduces v$.
In both cases, $e_1 \sreduces \strin{s}$ by inversion.
Therefore, in both cases, $e_1 \mapsto^* \strin{s}$ by induction
and so $e \mapsto^* \rcheck{r}{\strin{s}}{x.e_2}{e_3}$ by SS-E-Check-Step.

Suppose $e \sreduces v$ is finally derived via SS-E-Check-Ok.
By the facts mentioned above and SS-E-Check-Step, $e \mapsto^* \rcheck{r}{\strin{s}}{x.e_2}{e_2}$.
Note that by inversion $s \in \lang{r}$. Therefore, SS-E-Check-Ok applies and so by RT-Trans $e \mapsto^* [\rstr{s}/x]e_1$.
By inversion, $[\strin{s}/x]e_1 \sreduces v$. Therefore, by induction and RT-Step $e \mapsto^* v$ as required.

Suppose that $e \sreduces v$ is instead finally derived via SS-E-Check-NotOk.
By inversion, $e_3 \sreduces v$ and by induction $e_3 \mapsto^* v$.
From these facts at SS-E-Check-Step, it is derivable that $e \mapsto^* \rcheck{r}{\strin{s}}{x.e_2}{v}$.

Also by inversion, $s \not \in \lang{r}$ and so SS-E-Check-NotOk applies.
Therefore, $\rcheck{r}{\strin{s}}{x.e_2}{v} \mapsto v$.

The conclusion $e \mapsto^* v$ follows from these facts by RT-Step.
\end{itemize}
\end{proof}
%% Abstraction case.
%\item \textbf{Case $e = \lambda x . e_1$}. 
%% String Values
%\item \textbf{Case $e = \strin{s}$}. 
%% String Concatenation.
%\item \textbf{Case $e=\rconcat{e_1}{e_2}$}. 
%%Strcase Case.
%\item \textbf{Case $e=\strcase{e_1}{e_1}{x,y.e_3}$}.
%% String Replacement.
%\item \textbf{Case $e=\rreplace{r}{e_1}{e_2}$}.
%% Coercion
%\item \textbf{Case $e = \rcoerce{r}{e_1}$}.
%% Replacement
%\item \textbf{Case $e = \rcheck{r}{e_1}{x.e_2}{e_3}$}.

%Establishing the other direction requires a minor lemma about the relationship between values and reflexivity in the big step semantics.
%
%\begin{trthm}[Small Step Values are Reflexive in the Big Step Semantics]
%If $v \ \val$ in the small step semantics then $v \sreduces v$.
%\end{trthm}
%\begin{proof}
%If $v \ \val$ is derived from SS-E-RStr then S-E-RStr establishes the result.
%Otherwise $v \ \val$ is derived from SS-E-Abs and so S-E-Abs establishes the result.
%\end{proof}

\begin{trthm}[Semantic Correspondence for $\lambdas$ (Part II)]
If $\emptyset\vdash e : \sigma$, $e \mapsto^* v$ and $v \ \val$ then $e \sreduces v$.  
\end{trthm}
\begin{proof}
The proof proceeds by structural induction on $e$.

\begin{itemize}[label=$ $,itemsep=1ex]

\item \textbf{Case $e = \concat{e_1}{e_2}$}.
By inversion, $\emptyset\vdash e_1 : \stringin{r_1}$.
By Type Safety, Canonical Forms and Termination it follows that $e_1 \mapsto^* \rstr{s_1}$ for some $s_1$.
By induction, $e_1 \sreduces \rstr{s_1}$.

Similarly, $e_2 \mapsto^* \rstr{s_2}$ and $e_2 \sreduces \rstr{s_2}$.

Note that $\concat{e_1}{e_2} \mapsto^* \concat{\rstr{s_1}}{\rstr{s_2}} \mapsto \rstr{s_1s_2}$ by SS-E-Concat-LR* and S-E-Concat.
Therefore, $e \mapsto^* \rstr{s_1s_2}$ by RT-Step. So it suffices to show that $e \sreduces \rstr{s_1s_2}$.

Finally, $e \sreduces \rstr{s_1s_2}$ follows via S-E-Concat from the facts that $e_1 \sreduces \rstr{s_1}$ and $e_2 \sreduces \rstr{s_2}$.
This completes the case.

\item \textbf{Case $e = \rreplace{r}{e_1}{e_2}$}.
By inversion of S-T-Replace, $\emptyset\vdash e_1 : \stringin{r_1}$ for some $r_1$.
It follows by Type Safety, Termination and Canonical Forms that $e_1 \mapsto^* \rstr{s_1}$.
By induction, $e_1 \sreduces \rstr{s_1}$.

Similarly, $e_2 \mapsto^* \rstr{s_2}$ and $e_2 \sreduces \rstr{s_2}$.

Note that $e \mapsto^* \rreplace{r}{\rstr{s_1}}{\rstr{s_2}} \mapsto \rstr{\lsubst{r}{s_1}{s_2}}$ by SS-Replace-LR* and SS-E-Replace.
Therefore $e \mapsto^* \rstr{\lsubst{r}{s_1}{s_2}}$ by RT-Step.

It suffices to show $e \sreduces \rstr{\lsubst{r}{s_1}{s_2}}$, which follows by S-E-Replace from the facts that $e_1 \sreduces \rstr{s_1}$ and
$e_2 \sreduces \rstr{s_2}$.

\item \textbf{Case $e = \strcase{e_1}{e_2}{x.y.e_3}$}.
By inversion, $\emptyset\vdash e_1 : \stringin{r}$ and $e_2 : \sigma$.
By Type Safety, Canonical Forms and Termination $e_1 \mapsto^* \stringin{s_1}$ and by induction $e_1 \sreduces \stringin{s_1}$.
Similarly, $e_2 \mapsto^* v_2$ and $\emptyset\vdash e_2 \sreduces v_2$.

By SS-E-Case-LR*, $\strcase{e_1}{e_2}{x,y.e_3} \mapsto^* \strcase{v_1}{v_2}{x,y.e_3}$.

Note that either $s_1 = \epsilon$ or $s_1 = as$ because we define strings as either empty or finite sequences of characters.
We proceed by cases.

If $s_1 = \epsilon$ then $\strcase{v}{v_2}{x,y.e_3} \mapsto v_2$ by SS-E-Case-$\epsilon$.
Therefore, by RT-Step, $e \mapsto^* v_2$.
Recall $e_1 \sreduces \strin{\epsilon}$ and $e_2 \sreduces v_2$, which is enough to establish by S-E-Case-$\epsilon$ that $e \sreduces v_2$.

If $s_1 = as$ instead, then $\strcase{\strin{s_1}}{v_2}{x,y.e_3} \mapsto [\strin{a},\strin{s} / x,y ] e_3$ by SS-E-Case-Concat.
Inversion of the typing relation satisfies the assumptions necessary to appeal to termination.
Therefore,

%\todo{Note that this substitution does not modify the syntactic structure of $e_3$ except by replacing free variables wit strings, and so} by induction,
\[ [\strin{a},\strin{s} / x,y ] e_3 \mapsto^* v \text{ for } v \ \val. \]

It follows by RT-Step that $e \mapsto^* v$.

Note that the substitution does not change the structure of $e_3$.
So by induction, $[\strin{a},\strin{s} / x,y ] e_3 \sreduces v$.
Recall that $e_1 \sreduces \strin{s_1}$ and so by S-E-Case it follows that $e \sreduces [a,s / x,y] e_3 \sreduces v$.
\end{itemize}

\todo{The cases for coercion and checking are straightforward.}
\end{proof}

\subsection{Extension of Safety for Small Step Semantics}

\begin{thm}[Type Safety] \label{thm:typesafety}
  If $\emptyset \vdash e : \sigma$ 
  and $e \sreduces e'$ then $\emptyset \vdash e' : \sigma$.
\end{thm}
\begin{proof}
If $\emptyset \vdash e : \sigma$ then $e \mapsto^* e'$.
Therefore, $e \sreduces e'$ by part 2 of the semantic correspondence theorem.

Since $\emptyset \vdash e : \sigma$ and $e \mapsto^* e'$, it follows that $\emptyset \vdash e' : \sigma$ by type safety for the small step semantics.
\end{proof}

\subsubsection{The Security Theorem}\label{sec:securitythm}

\begin{thm}[Correctness of Input Sanitation for $\lambdas$]\label{thm:scorrect}
  If  $\emptyset \vdash e : \stringin{r}$ and $e \sreduces \rstr{s}$ then $s \in \lang{r}$.
\end{thm}
\begin{proof}
If $\emptyset \vdash e : \stringin{r}$ and $e \sreduces \rstr{s}$ then $\emptyset \vdash \rstr{s} : \stringin{r}$ by Type Safety.
By Caonical Forms, $s \in \lang{r}$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LAMBDAP SECTION

\section{Proofs of Lemmas and Theorems About $\lambdap$}

This section follows the same structure as the safety proof for $\lambdas$ --
we prove type safety for a small-step semantics, prove a semantic correspondence,
and then transfer the safety result to the big-step semantics in the paper.

Our proofs here are more sketchy than in the $\lambdas$ section both because the
proof method is already demonstrated, and becaues $\lambdap$ is a fairly straightforward
extension of $\lambda$ for which these results are already estbalished.

\begin{lem}[Canonical Forms for Target Language]
~
\begin{itemize}
\item 
If $\emptyset \vdash \iota : \regex$ then $\iota \treduces \rx{r}$ such that $r$ is a well-formed regular expression. 
\item 
If $\emptyset \vdash \iota : \str$ then $\iota \treduces \tstr{s}$.
\end{itemize}
\end{lem}

\begin{thm}[Progress]
If $\emptyset\vdash \iota : \tau$ either $\iota = \dot{v}$ or $\iota \mapsto \iota'$ for some $\iota'$.
\end{thm}
\begin{proof}
The proof proceeds by induction on the typing assumption. We elide contexts where context has obviously not changed.

Consider only the string and regex (non-$\lambda$) fragments of $\lambdap$.

Regular expressions are always values.

For strings, the concatenation cases are trivial.

\begin{itemize}[label=$ $,itemsep=1ex]

\item \textbf{P-T-Case}.
Suppose $\emptyset \vdash \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3}$.
By inversion, $\iota_1 : \str$ and so either $\iota_1 \mapsto \iota_1'$
or by canonical forms, $\iota_1 = \tstr{s_1}$.
Similarly, $\iota_2 \mapsto \iota_2'$ or else $\iota_2 = \tstr{s_2}$.
In the former cases, progress occurs via the compatibility rules.
in the case where both are string values, progress occurs via the case concatenation rule.

\item \textbf{P-T-Replace}.
Suppose $\emptyset \vdash \preplace{\iota_1}{\iota_2}{\iota_3}$.
By inversion, $\iota_1 : \regex$ and so by canonical forms $\iota_1 = \rx{r}$.
By inversion, $\iota_2 : \str$ and so by induction either $\iota_2 \mapsto \iota_2'$
or else $\iota_2 = \tstr{s_2}$ for some string $s_2$.
Similarly, either $\iota_3$ steps or else $\iota_3 = \tstr{s_3}$.
In case any steps occur, progress occurs. In the remaining case,
PP-E-Replace applies and so progress occurs.

\item \textbf{P-T-Check}.
Finally, suppose $\emptyset \vdash \check{\iota_x}{\iota_1}{\iota_2}{\iota_3}$.
In case any of these step, then progress occurs.
In the remaining cases, applications of inversion and canonical forms for each
$\iota_x$ and $\iota_1$ implies that the term at hand equals $\check{\rx{r}}{\tstr{s}}{\iota_2}{\iota_3}$,
which evaluates to either $\iota_2$ or $\iota_3$.

\end{itemize}
\end{proof}

\begin{thm}[Preservation]
If $\emptyset \vdash \iota : \tau$ and $\iota \mapsto \iota'$ then $\emptyset \vdash \iota' : \tau$.
\end{thm}
\begin{proof}
\todo{Fill in if there's time. Otherwise there's nothing interesting going on.}
\end{proof}

\begin{thm}[Safety for Small-Step Semantics]
If $\iota \sreduces \dot{v}$ and $\iota \mapsto \iota'$ then $\iota' \sreduces \dot{v}$. 
\end{thm}
\begin{proof}
A direct result from progress and preservation.
\end{proof}

We only prove semantic correspondence in one direction; again, whereas $\lambdas$ proofs
were detailed, here we provide a less verbosy proof.

\begin{thm}[Semantic Correspondence]
If $\iota \mapsto^* \iota'$ then $\iota \sreduces \iota'$.
\end{thm}
\begin{proof}
By induction on the structure of $\iota$.

%\begin{itemize}[label=$ $,itemsep=1ex]
%\end{itemize}
\end{proof}

\begin{thm}[Safety for $\lambdap$] If $\emptyset \vdash \iota : \tau$ 
  then $\iota \sreduces \dot v$ and $\emptyset \vdash \dot v : \tau$.
\end{thm}
\begin{proof}
If $\emptyset \vdash \iota : \tau$ then $\iota \mapsto^* \iota'$.
Therefore, $\iota \sreduces \iota'$ by part 2 of the semantic correspondence theorem.

Since $\emptyset \vdash \iota : \tau$ and $\iota \mapsto^* \iota'$, it follows that $\emptyset \vdash \iota' : \tau$ by type safety for the small step semantics.

\end{proof}

We can also define canonical forms for regular expressions and strings in the usual way:

\section{Proofs and Lemmas and Theorems About Translation}

\begin{thm}[Translation Correctness]\label{thm:trcorrect}
  If $\tctx \vdash e : \sigma$ then 
  there exists an $\iota$ such that $\trden{e} = \iota$
  and $\trden{\tctx} \vdash \iota : \trden{\sigma}$.
  Furthermore, $e \sreduces v$ and
  $\iota \treduces \dot{v}$ such that
  $\trden{v} = \dot{v}$.
\end{thm}
\begin{proof}
We present a proof by induction on the derivation that $\tctx \vdash e : \sigma$.
we write $e \leadsto \iota$ as shorthand for the final property.

\begin{itemize}[label=$ $, itemsep=1ex]
\item \textbf{Case $e = \rstr{s}$}. Suppose $\tctx \vdash \rstr{s} : \sigma$.

By examination the syntactic structure of conclusions in the relation S-T, 
we know this is true just in case $\sigma = \stringin{r}$ for some $r$ such that
$s \in \lang{r}$; and of course, there is always such an $r$.

There are no free variables in $\rstr{s}$, so we might as well proceed from the
fact that $\emptyset \vdash \rstr{s} : \stringin{r}$.

By definition of the translation ($\trden{\cdot}$) the following statements hold:
\begin{align}
\trden{\rstr{s}} &= \str{s} \\
\trden{\stringin{r}} &= \str \\
\trden{\emptyset} &= \emptyset
\end{align}
Note that $\emptyset \vdash \str{s} : \str$ by P-T-Str.
Recall that contexts are standard and, in particular, can be weakened.
So since $\trden{\tctx}$ is either a weakening of $\emptyset$ or $\emptyset$ itself,
$\trden{\tctx} \vdash \tstr{s} : \str$ by weakening.

Summarily, $\str{s}$ is a term of $\lambdap$ such that $\trden{\tctx} \vdash \str{s} : \trden{\sigma}$

It remains to be shown that there exist $v, \dot{v}$ such that $\rstr{s} \sreduces v$,
$\str{s} \treduces \dot{v}$, and $\trden{v} = \dot{v}$. But this is immediate because
each term evaluates to itself and we have already established the equality.

\item \textbf{Case $e = \rconcat{e_1}{e_2}$}. This case is an obvious appeal to induction.

\item \textbf{Case $e = \strcase{e_1}{e_2}{x,y.e_3}$}. This case relies on our definition
of context translation.

Suppose $\sctx \vdash \strcase{e_1}{e_2}{x,y.e_3} : \sigma$.
By inversion of the typing relation it follows that
$\sctx \vdash e_1 : \stringin{r}$, $\sctx \vdash e_2 : \sigma$ and $\sctx, x : \stringin{\lhead{r}}, y : \stringin{\ltail{r}} \vdash e_3 : \sigma$.

By induction, there exists an $\iota_1$ such that $\trden{e_1} = \iota_1$,
$\trden{\sctx} \vdash \iota_1 : \trden{\sigma}$, and $e_1 \leadsto \iota_1$.
Similarly for $e_2$ and some $\iota_2$.

By canonical forms, $e_1 \sreduces \rstr{s}$ and so $\iota_1 \treduces \tstr{s}$ by $\leadsto$.

Choose $\iota = \tconcat{\iota_1}{\iota_2}{x,y.\iota_3}$ and note that by the properties established via induction,
$\trden{e} = \iota$ and $\trden{\sctx} \vdash \iota : \trden{\sigma}$.

Suppose $s = \epsilon$. Then $e \sreduces v$ where $e_2 \sreduces v$
and $\iota \treduces \dot{v}$ where $\iota_2 \treduces \dot{v}$.
But recall that $e_2 \leadsto v_2$ and so $\trden{v} = \dot{v}$.

Suppose otherwise that $s = at$ for some character $a$ and string $t$.
Then $e \sreduces v$ where $[a, t / x, y] e_3 \sreduces v$.
Similarly, $\iota \treduces \dot{v}$ where $[a, t / x, y] \iota_3 \treduces \dot{v}$

\end{itemize}
\end{proof}

\begin{thm}[Correctness of Input Sanitation for Translated Terms]\label{thm:main}
  If $\trden{e} = \iota$ and $\emptyset \vdash e : \stringin{r}$ then $\iota \sreduces \tstr{s}$
  for $s \in \lang{r}$.
\end{thm}
\begin{proof}
  By Theorem \ref{thm:trcorrect} and the rules given, $\iota \sreduces \tstr{s}$ implies that $e \sreduces \strin{s}$.
  Theorem \ref{thm:scorrect} together with the assumption that $e$ is well-typed implies that $s \in \lang{r}$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{research}

\clearpage
\renewcommand{\grammarlabel}[2]{#1\hfill#2}
\begin{figure}[b]
\small
  \begin{grammar}
<$r$> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

\caption{Regular expressions over the alphabet $\Sigma$.}
\label{fig:regex}
\end{grammar}
\end{figure}
\begin{figure}[t]
\small
  \begin{grammar}
<$\sigma$> ::=  $\sigma \rightarrow \sigma$ | $\stringin{r}$    \hfill  source types

<$e$> ::= 
      $x$ | $v$ \hfill source terms \alt 
      $\rsconcat{e}{e}$ | $\strcase{e}{e}{x,y.e}$ \hfill $s \in \Sigma^{*}$ \alt
      $\sisubst{r}{e}{e}$ | $\rcoerce{r}{e}$ | $\rcheck{r}{e}{x.e}{e}$

<$v$> ::= $\lambda x . e$ | $\strin{s}$ \hfill source values 
\caption{Syntax of $\lambda_{RS}$.}
\label{fig:glambdas}
\end{grammar}
\end{figure}

\renewcommand{\grammarlabel}[2]{#1\hfill#2}

\begin{figure}[t]
\small
  \begin{grammar}

<$\tau$> ::= $\tau \rightarrow \tau$ | $\str$ | $\regex$ \hfill target types

<$\iota$> ::= $x$ | $\dot{v}$ \hfill target terms \alt
$\tconcat{\iota}{\iota}$ | $\pstrcase{\iota}{\iota}{x,y.\iota}$ \alt
  $\rx{r}$ | $\preplace{\iota}{\iota}{\iota}$ | $\tcheck{\iota}{\iota}{\iota}{\iota}$

  <$\dot{v}$> ::= $\lambda x . \iota$ | $\tstr{s}$ | $\rx{r}$ \hfill target values

\end{grammar}
\caption{Syntax for the target language, $\lambdap$, containing strings and statically constructed regular expressions.}
\label{fig:lcsSyntax}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash x:\sigma}

\inferrule[S-T-Abs]
{\sctx, x : \sigma_1 \vdash e : \sigma_2}
{\sctx \vdash \lambda x.e : \sigma_1 \rightarrow \sigma_2}

\inferrule[S-T-App]
{\sctx \vdash e_1 : \sigma_2 \rightarrow \sigma \\ \sctx \vdash e_2 : \sigma_2}
{\sctx \vdash e_1(e_2) : \sigma}

\inferrule[S-T-Stringin-I]
{s \in \lang{r}}
{\sctx \vdash \strin{s} : \stringin{r}}

\inferrule[S-T-Concat]
{\sctx \vdash e_1 : \stringin{r_1} \\ \sctx \vdash e_2 : \stringin{r_2}}
{\sctx \vdash \rsconcat{e_1}{e_2} : \stringin{r_1 \cdot r_2}}

\inferrule[S-T-Case]
{ \sctx \vdash e_1 : \stringin{r} \\
  \sctx \vdash e_2 : \sigma \\
  \sctx, x : \stringin{\lhead{r}}, y : \stringin{\ltail{r}} \vdash e_3 : \sigma
}
{
  \sctx \vdash \strcase{e_1}{e_2}{x,y.e_3} : \sigma
}

\inferrule[S-T-Replace]
{\sctx \vdash e_1 : \stringin{r_1} \\ \sctx \vdash e_2 : \stringin{r_2} }
{\sctx \vdash \sisubst{r}{e_1}{e_2} : \stringin{\lreplace{r}{r_1}{r_2}}}

\inferrule[S-T-SafeCoerce]
{\sctx \vdash e : \stringin{r'} \\ \lang{r'} \subseteq \lang{r}}
{\sctx \vdash \coerce{r}{e} : \stringin{r}}

\inferrule[S-T-Check]
{\sctx \vdash e_0 : \stringin{r_0} \\ \sctx, x:\stringin{r} \vdash e_1 : \sigma \\ \sctx \vdash e_2 : \sigma}
{\sctx \vdash \rcheck{r}{e_0}{x.e_1}{e_2} : \sigma}
\end{mathpar}

\caption{Typing rules for $\lambdas$. 
The typing context $\sctx$ is standard.}
\label{fig:slambdas}
\end{figure}

\begin{figure}
  \small
$\fbox{\inferrule{}{e \sreduces v}}$
\begin{mathpar}
\inferrule[S-E-Abs]                                                             
{ \ }                                                                           
{\lambda x.e \sreduces \lambda x.e}

\inferrule[S-E-App]
{ e_1 \sreduces \lambda x . e_3 \\ e_2 \sreduces v_2 \\ [ v_2 / x ] e_3 \sreduces v }
{ e_1(e_2) \sreduces v }

\inferrule[S-E-RStr]
{ \ }
{\strin{s} \sreduces \strin{s}}

\inferrule[S-E-Concat]
{e_1 \sreduces \strin{s_1} \\ e_2 \sreduces \strin{s_2}} 
{\rsconcat{e_1}{e_2} \sreduces \strin{s_1 s_2}} % ???

\inferrule[S-E-Case-$\epsilon$]
{
  e_1 \sreduces \strin{\epsilon} \\
  e_2 \sreduces v_2
}
{
  \strcase{e_1}{e_2}{x,y.e_3} \sreduces v_2
}

\inferrule[S-E-Case-Concat]
{
  e_1 \sreduces \strin{a s} \\
  [\rstr{a},\rstr{s} / x,y]e_3 \sreduces v_3
}
{
  \strcase{e_1}{e_2}{x,y.e_3} \sreduces v_3
}

\inferrule[S-E-Replace]
{e_1 \sreduces \strin{s_1} \\ e_2 \sreduces \strin{s_2} \\ \lsubst{r}{s_1}{s_2} = s} 
{\sisubst{r}{e_1}{e_2} \sreduces \sistr{s}}

\inferrule[S-E-SafeCoerce]
{e \sreduces \strin{s}}
{\coerce{r}{e} \sreduces \strin{s}}

\inferrule[S-E-Check-Ok]
{e \sreduces \strin{s} \\ s \in \lang{r} \\ [\strin{s} / x]e_1 \sreduces v}
{
  \rcheck{r}{e}{x.e_1}{e_2} \sreduces v
}

\inferrule[S-E-Check-NotOk]
{
  e \sreduces \strin{s} \\ s \not \in \lang{r} \\
  e_2 \sreduces v
}
{
  \rcheck{r}{e}{x.e_1}{e_2} \sreduces v
}
\end{mathpar}
\caption{Big step semantics for $\lambdas$.}
\label{fig:dlambdas}
\end{figure}

\begin{figure}
  \small
$\fbox{\inferrule{}{e \mapsto e}}$

\begin{mathpar}

\inferrule[SS-E-AppLeft]
{e_1 \mapsto e_1'}
{e_1(e_2) \mapsto e_1'(e_2)}

\inferrule[SS-E-AppRight]
{e_2 \mapsto e_2'}
{v_1 \mapsto v_1}

\inferrule[SS-E-AppAbs]
{ \ }
{(\lambda x : \tau_{11} . t_{12}) v_2 \mapsto [v_2 / x]t_{12}}

\end{mathpar}

\fbox{\inferrule{}{e \mapsto^* e}}

\begin{mathpar}

\inferrule[RT-Refl]
{ \ }
{e \mapsto^* e}

\inferrule[RT-Trans]
{ e \mapsto^* e' \\ e' \mapsto e'' }
{ e \mapsto^* e'' }

\end{mathpar}

\label{fig:lambda}
\caption{Call-by-name small step Semantics for $\lambda$ and its reflexive, transitive closure.} \label{fig:lambda}
\end{figure}

\begin{figure}
  \small
$\fbox{\inferrule{}{e \mapsto e}}$ (Contiues figure 6)
%$\fbox{\inferrule{}{e \mapsto e}} \  \fbox{ \inferrule{}{v \ \val}}$
\begin{mathpar}
\inferrule[SS-E-Concat-Left]
{e_1 \mapsto e_1'}
{\rsconcat{e_1}{e_2} \mapsto \rsconcat{e_1'}{e_2}}

\inferrule[SS-E-Concat-Right]
{e_2 \mapsto e_2'}
{\rsconcat{v_1}{e_2} \mapsto \rsconcat{v_1}{e_2'}}

\inferrule[SS-E-Concat]
{ \ }
{\rsconcat{\strin{s_1}}{\strin{s_2}} \mapsto \strin{s_1 s_2}} % ???

\inferrule[SS-E-Case-Left]
{
  e_1 \mapsto e_1'
}
{
  \strcase{e_1}{e_2}{x,y.e_3} \mapsto \strcase{e_1'}{e_2}{x,y.e_3}
}

\inferrule[SS-E-Case-$\epsilon$-val]
{
  \ 
}
{
  \strcase{\strin{\epsilon}}{e_2}{x.y.e_3} \mapsto e_2
}

\inferrule[SS-E-Case-Concat]
{
  \ 
}
{
  \strcase{\strin{a s}}{e_2}{x,y.e_3} \mapsto [\rstr{a},\rstr{s} / x,y]e_3 
}

\inferrule[SS-E-Replace-Left]
{e_1 \mapsto e_1'}
{ \sisubst{r}{v_1}{e_2} \mapsto \sisubst{r}{v_1'}{e_2}}

\inferrule[SS-E-Replace-Right]
{e_2 \mapsto e_2'}
{ \sisubst{r}{e_1}{e_2} \mapsto \sisubst{r}{e_1}{e_2'}}

\inferrule[SS-E-Replace]
{ \ }
{\sisubst{r}{\strin{s_1}}{\strin{s_2}} \mapsto \sistr{\lsubst{r}{s_1}{s_2}}}

\inferrule[SS-E-SafeCoerce-Step]
{e \mapsto e'}
{\coerce{r}{e} \mapsto \coerce{r}{e'}}

\inferrule[SS-E-SafeCoerce]
{ \ }
{ \coerce{r}{\rstr{s}} \mapsto \rstr{s}}

%%% NRF-bookmark1
\inferrule[SS-E-Check-StepLeft]
{e \mapsto e'}
{\rcheck{r}{e}{x.e_1}{e_2} \mapsto \rcheck{r}{e'}{x.e_1}{e_2}}

\inferrule[SS-E-Check-Ok]
{s \in \lang{r}}
{\rcheck{r}{\strin{s}}{x.e_1}{e_2} \mapsto [\rstr{s} / x]e_1}

\inferrule[SS-E-Check-NotOk]
{s \not \in \lang{r}}
{\rcheck{r}{\strin{s}}{x.e_1}{e_2} \mapsto e_2}

\end{mathpar}
\caption{Small step semantics for $\lambdas$. Extends \ref{fig:lambda}.}
\label{fig:sdlambdas}
\end{figure}

\begin{figure}[t]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash \iota : \tau}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \tau$

\begin{mathpar}
\inferrule[P-T-Var]
{ x:\tau \in \tctx }
{ \tctx \vdash x:\tau }

\inferrule[P-T-Abs]
{\tctx, x : \tau_1 \vdash \iota_2 : \tau_2}
{\tctx \vdash \lambda x.\iota_2 : \tau_1 \rightarrow \tau_2}

\inferrule[P-T-App]
{\tctx \vdash \iota_1 : \tau_2 \rightarrow \tau \\ \tctx \vdash \iota_2 : \tau_2}
{\tctx \vdash \iota_1(\iota_2) : \iota}

\inferrule[P-T-String]
{ \ }
{\tctx \vdash \tstr{s} : \str}

\inferrule[P-T-Regex]
{ \ }
{\tctx \vdash \rx{r} : \regex}

\inferrule[P-T-Concat]
{\tctx \vdash \iota_1 : \str \\ \tctx \vdash \iota_2 : \str}
{\tctx \vdash \tconcat{\iota_1}{\iota_2} : \str}

\inferrule[P-T-Case]
{\tctx \vdash \iota_1 : \str \\ \tctx \vdash \iota_2 : \tau \\ \tctx, x:\str, y:\str \vdash \iota_3:\tau}
{\tctx \vdash \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} : \tau}

\inferrule[P-T-Replace]
{\tctx \vdash \iota_1 : \regex \\ \tctx \vdash \iota_2 : \str \\ \tctx \vdash \iota_3 : \str }
{\tctx \vdash \preplace{\iota_1}{\iota_2}{\iota_3} : \str}

\inferrule[P-T-Check]
{\tctx \vdash \iota_r : \regex \\ \tctx \vdash \iota_1 : \str \\ \tctx \vdash \iota_2 : \sigma \\ \tctx \vdash \iota_3 : \sigma}
{\tctx \vdash \tcheck{\iota_r}{\iota_1}{\iota_2}{\iota_3} : \sigma}
\end{mathpar}
\caption{Typing rules for $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:slambdap}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\iota \treduces \dot{v}}}$

\begin{mathpar}
\inferrule[P-E-Abs]
{ \ }
{\lambda x.e \sreduces \lambda x.e}

\inferrule[P-E-App]
{ \iota_1 \sreduces \lambda x . \iota_3 \\  \iota_2 \sreduces \dot{v}_2 \\ [\dot{v}_2 / x] \iota_3 \sreduces \dot{v}_3}
{ \iota_1(\iota_2) \sreduces \dot{v}_3}

\inferrule[P-E-Str]
{ \ }
{\tstr{s} \treduces \tstr{s}}

\inferrule[P-E-Rx]
{ \ }
{\rx{r} \treduces \rx{r}}

\inferrule[P-E-Concat]
{\iota_1 \treduces \tstr{s_1} \\ \iota_2 \treduces \tstr{s_2}} 
{\tconcat{\iota_1}{\iota_2} \treduces \tstr{s_1 s_2}} % ???

\inferrule[P-E-Case-$\epsilon$]
{
  \iota_1 \treduces \tstr{ \epsilon } \\
  \iota_2 \treduces \dot{v_2} \\
}
{
  \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} \treduces \dot{v_2}
}

\inferrule[P-E-Case-Concat]
{
  \iota_1 \treduces \tstr{as} \\
  [\tstr{a},\tstr{s}/x,y] \iota_3 \treduces \dot{v} 
}
{
  \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} \treduces \dot{v}
}

\inferrule[P-E-Replace]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \tstr{s_2} \\ \iota_3 \treduces \tstr{s_3} } 
{\preplace{\iota_1}{\iota_2}{\iota_3} \treduces \tstr{\lsubst{r}{s_2}{s_3} }}

\inferrule[P-E-Check-OK]
{
  \iota_x \treduces \rx{r} \\
  \iota \treduces \tstr{s} \\ 
s \in \lang{r} \\ 
\iota_1 \treduces \dot{v_1} 
}
{
  \tcheck{\iota_x}{\iota}{\iota_1}{\iota_2} \treduces \dot{v_1}
}

\inferrule[P-E-Check-NotOK]
{\iota_x \treduces \rx{r} \\ \iota \treduces \tstr{s} \\ s \not \in \lang{r} \\ \iota_2 \treduces \dot{v_2}}
{\tcheck{\iota_x}{\iota}{\iota_1}{\iota_2} \treduces \dot{v_2}}
\end{mathpar}
\caption{Big step semantics for $\lambdap$}.
\label{fig:dlambdap}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\iota \treduces \dot{v}}}$

\begin{mathpar}
\inferrule[PP-E-ConcatLeft]
{\iota_1 \mapsto \iota_1'}
{\tconcat{\iota_1}{\iota_2} \treduces \tconcat{\iota_1'}{\iota_2}}

\inferrule[PP-E-ConcatRight]
{\iota_2 \mapsto \iota_2'}
{\tconcat{\iota_1}{\iota_2} \treduces \tconcat{\iota_1}{\iota_2'}}

\inferrule[PP-E-Concat]
{\ }
{\tconcat{\tstr{s_1}}{\tstr{s_2}} \treduces \tstr{s_1s_2}}

\inferrule[PP-E-CaseLeft]
{ \iota_1 \mapsto \iota_1' }
{ \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} \mapsto \pstrcase{\iota_1'}{\iota_2}{x,y.\iota_3} }

\inferrule[PP-E-CaseRight]
{ \iota_2 \mapsto \iota_2' }
{ \pstrcase{\dot{v}}{\iota_2}{x,y.\iota_3} \mapsto \pstrcase{\dot{v}}{\iota_2'}{x,y.\iota_3} }

\inferrule[PP-E-Case]
{ }
{ \pstrcase{\tstr{s_1}}{\tstr{s_2}}{x,y.\iota_3} \mapsto \tstr{s_1s_2}}

\inferrule[PP-E-ReplaceLeft]
{\iota_1 \mapsto \iota_1'}
{ \preplace{\iota_1}{\iota_2}{\iota_3} \mapsto \preplace{\iota_1'}{\iota_2}{\iota_3} }

\inferrule[PP-E-ReplaceMid]
{\iota_2 \mapsto \iota_2'}
{ \preplace{\rx{r}}{\iota_2}{\iota_3} \mapsto \preplace{\rx{r}}{\iota_2'}{\iota_3} }

\inferrule[PP-E-ReplaceRight]
{\iota_3 \mapsto \iota_3'}
{ \preplace{\rx{r}}{\tstr{s_2}}{\iota_3} \mapsto \preplace{\rx{r}}{\tstr{s_2}}{\iota_3'} }

\inferrule[PP-E-Replace]
{\ }
{ \preplace{\rx{r}}{\tstr{s_2}}{\tstr{s_3}} \mapsto \tstr{\lsubst{r}{s_2}{s_3}} }

\inferrule[PP-E-Check-Left]
{ \iota_x \mapsto \iota_x' }
{ \check{\iota_x}{\iota}{\iota_1}{\iota_2} \mapsto \check{\iota_x}{\iota}{\iota_1}{\iota_2} }

\inferrule[PP-E-Check-Right]
{ \iota \mapsto \iota' }
{ \check{\rx{r}}{\iota}{\iota_1}{\iota_2} \mapsto \check{\rx{r}}{\iota'}{\iota_1}{\iota_2} }

\inferrule[PP-E-Check-OK]
{ s \in \lang{r} }
{ \check{\rx{r}}{\tstr{s}}{\iota_1}{\iota_2} \mapsto \iota_1}

\inferrule[PP-E-Check-OK]
{ s \not\in \lang{r} }
{ \check{\rx{r}}{\tstr{s}}{\iota_1}{\iota_2} \mapsto \iota_2}

\end{mathpar}
\caption{Small step semantics for $\lambdap$ (extends L-E rules)}.
\label{fig:sdlambdap}
\end{figure}

\begin{figure}[h]
\small
$\fbox{\inferrule{}{\trden{\sigma}=\tau}}$
\begin{mathpar}
\inferrule[Tr-T-String]
{ }
{ \trden{\stringin{r}} = \str }

\inferrule[Tr-T-Arrow]
{ \trden{\sigma_1} = \tau_1 \\ \trden{\sigma_2} = \tau_2 }
{ \trden{\sigma_1 \rightarrow \sigma_2} = \tau_1 \rightarrow \tau_2 }
\end{mathpar}
$\fbox{\inferrule{}{\trden{\Psi}=\Theta}}$
\begin{mathpar}
\inferrule[Tr-T-Context-Emp]
{ }{\trden{\emptyset} = \emptyset}

\inferrule[Tr-T-Context-Ext]
{\trden{\Psi}=\Theta\\ \trden{\sigma} = \tau}
{\trden{\Psi, x : \sigma} = \Theta, x : \tau}
\end{mathpar}
  $\fbox{\inferrule{}{ \trden{e} = \iota }}$
\begin{mathpar}
\inferrule[Tr-Var]
{ }
{ \trden{x} = x}

\inferrule[Tr-Abs]
{  \trden{e} = \iota }
{ \trden{ \lambda x.e } = \lambda x . \iota}

\inferrule[Tr-App]
{ \trden{e_1} = \iota_1 \\  \trden{e_2} = \iota_2}
{ \trden{e_1(e_2)} = \iota_1(\iota_2)}

\inferrule[Tr-Case]
{ \trden{e_1} = \iota_1 \\
   \trden{e_2} = \iota_2 \\
   \trden{e_3} = \iota_3
}
{ \trden{ \strcase{e_1}{e_2}{x,y.e_3} } = \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} }

\inferrule[Tr-string]
{ \ }
{   \trden{\strin{s}} = \tstr{s}}

\inferrule[Tr-Concat]
{  \trden{e_1} = \iota_1 \\  \trden{e_2} = \iota_2}
{  \trden{\rsconcat{e_1}{e_2}} = \tconcat{\iota_1}{\iota_2} }

\inferrule[Tr-Subst]
{   \trden{e_1} = \iota_1 \\   \trden{e_2} = \iota_2 }
{   \trden{ \sisubst{r}{e_1}{e_2} } = \tsubst{\rx{r}}{\iota_1}{\iota_2} }

\inferrule[Tr-SafeCoerce]
{  \trden{e} = \iota }
{   \trden{ \coerce{r'}{e} } = \iota }

\inferrule[Tr-Check]
{  
  \trden{e} = \iota \\  
   \trden{e_1} = \iota_1 \\  
   \trden{e_2} = \iota_2
} 
{ 
  \trden{ \rcheck{r}{e}{x.e_1}{e_2} } = \tcheck{\rx{r}}{\iota}{(\lambda x . \iota_1)(\iota)}{\iota_2} 
} 
\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).}
\label{fig:tr}
\end{figure}
\end{document}