### 2015-07-31 thesis/language-overview.tex

% !TEX root = omar-thesis.tex

\chapter{Language Overview}
\section{External Language}
\section{Internal Language}
\section{Static Language}
\section{Module Language}