### 2015-07-31 thesis/syntax-motivation.tex

% !TEX root = omar-thesis.tex

\chapter{Motivation}
\section{Motivating Examples}
\subsection{Lists}
\subsection{HTML}
\subsection{Regular Expressions}
\subsection{Monadic Commands}
\subsection{Quasiquotation}
\section{Existing Approaches}
\subsection{Dynamic String Parsing}
\subsection{Direct Syntax Extension}

Related work I haven't mentioned yet:
\begin{itemize}
\item Fan: http://zhanghongbo.me/fan/start.html
\item Well-Typed Islands Parse Faster: \\\url{http://www.ccs.neu.edu/home/ejs/papers/tfp12-island.pdf}
\end{itemize}
\subsection{Term Rewriting}